---
title: "07-NMF decomp LC for GSEA weights"
author: "Bernie Mulvey"
date: "2026-03-23"
output: html_document
---

```{r}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(SpatialExperiment)
library(RcppML)
library(data.table)
library(Matrix)

```

setup
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

## Drop Br5517 (unknown anatomy/orientation), Br5276 (LC has been depleted by QC), Br5712 (unknown anatomy/orientation). Keep 6297; no LC but anatomy is right in vicinity.
lc3 <- lc3[,!(lc3$brnum %in% paste0("Br",c(5517,5276,5712)))]

## subset to LC cluster
vdom <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
vdom <- vdom$HARMONYlmbna_HDG_SVG_2575
vdom <- vdom[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(vdom,2,"clusid")

vanno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
vdom <- merge.data.table(vdom,vanno[,.(clusid,anno)],by="clusid")
vdom <- DataFrame(vdom,row.names=vdom$rn)[colnames(lc3),]

colLabels(lc3) <- vdom$anno

lc3 <- lc3[,lc3$label=="LC"]
## norm
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

rm(vanno,vdom)
gc(full=T)
```

### cross-validation 1: steps of 5, 20-100 factors, 3 reps per k.
```{r}
options(RcppML.threads = 10)
setDTthreads(1,restore_after_fork = FALSE)
cvallc <- RcppML::crossValidate(logcounts(lc3),
    k=seq(20,100,5),
    tol = 1e-06, 
    maxit = 5000,
    reps = 3,
    verbose=T
)

cvallc <- as.data.table(cvallc)
# plot
library(ggplot2)
pdf("plots/10_subclustering/07thru10_ver2_53factorNMF/07-lcnmf_crossval3x_20to100_5step.pdf",width=6,height=4)
ggplot(cvallc,aes(x=k,y=value))+
    geom_point()+labs(y="reconstruction error",x="number of factors (k)",title="SRT LC NMF cross-validation")
dev.off()

## thats like impossibile to look at, so also plot just k up to 70
pdf("plots/10_subclustering/07thru10_ver2_53factorNMF/07-lcnmf_crossval3x_20-70only.pdf",width=6,height=4)
ggplot(cvallc[k<75],aes(x=k,y=value))+
    geom_point()+labs(y="reconstruction error",x="number of factors (k)",title="SRT LC NMF cross-validation")
dev.off()
```

### round two of crossval: looks like somewhere between 45 and 55 is a (subtle) elbow, at least with 3 reps and 5 steps. try 46-54 with the full (probably redundant) param settings from Erik's NMF reference code.
### cross-validation 1: steps of 5, 20-100 factors, 6 reps per k.
```{r}
rm(cvallc)
gc(full=T)

options(RcppML.threads = 10)
setDTthreads(1,restore_after_fork = FALSE)
cvallc <- RcppML::crossValidate(logcounts(lc3),
    k=seq(46,54,1),
    tol = 1e-04, 
    maxit = 5000,
    reps = 6,
    verbose=T,
    L1=0.1,
    diag=TRUE,
    mask_zeros = FALSE,
    nonneg=TRUE
)

cvallc <- as.data.table(cvallc)
# plot
pdf("plots/10_subclustering/07thru10_ver2_53factorNMF/07-lcnmf_crossval6x_46to54_1step.pdf",width=8,height=4)
ggplot(cvallc,aes(x=k,y=value))+
    geom_point(position=position_jitter(width=0.2))+labs(y="reconstruction error",x="number of factors (k)",title="SRT LC NMF cross-validation")+scale_x_continuous(breaks=seq(46,54,1))
dev.off()

```

### optimal k: 53

### get logcounts matrix, rownames, colnames, and rowdata
```{r}
lclct <- logcounts(lc3)
lcrn <- rownames(lclct)
lccn <- colnames(lclct)

# get gene metadata to add symbol info later
rddt <- as.data.table(rowData(lc3))

rm(lc3,cvallc)
gc(full=T)
```

#### WHERE VERSION 2 DIFFERS: k is crossval-determined to be 53. 

## run NMF using Erik Nelson's final parameter set for some of these args (1k max iters, L1 penalty 0.1, diag = T) https://github.com/LieberInstitute/spatial_hpc/blob/main/snRNAseq_hpc/code/nmf/nmf.R

but let's set tol lower for more precision
```{r}
setDTthreads(1,restore_after_fork = FALSE)

options(RcppML.threads=10)
lc.nmf <-nmf(lclct,
    k=53,
    tol = 1e-09, 
    maxit = 5000,
    verbose = T,
    L1 = 0.1,
    seed = 42,
    mask_zeros = FALSE,
    diag = TRUE,
    nonneg = TRUE
)

w <- as.data.frame(lc.nmf$w)
rownames(w) <- lcrn
colnames(w) <- paste0("LCnmf",c(1:53))

h <- as.data.frame(lc.nmf$h)
rownames(h) <- paste0("LCnmf",c(1:53))
colnames(h) <- colnames(lclct)

lc.nmf2 <- list(w=w,h=h)
saveRDS(lc.nmf2,"processed-data/10_subclustering/07thru10_ver2_53factorNMF/07-lcnmf53_tol1e-9.RDS")
```

### second run: same parameters, but dropping low/no detection genes. setup:
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

## Drop Br5517 (unknown anatomy/orientation), Br5276 (LC has been depleted by QC), Br5712 (unknown anatomy/orientation). Keep 6297; no LC but anatomy is right in vicinity.
lc3 <- lc3[,!(lc3$brnum %in% paste0("Br",c(5517,5276,5712)))]

## norm
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

clus <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
clus <- clus$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(clus,2,"clusid")
anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")

clus <- merge.data.table(clus,anno,by="clusid",all.x=T)
clus <- DataFrame(clus,row.names=clus$rn)[colnames(lc3),]

lc3$label <- clus$anno

lc3 <- lc3[,lc3$label=="LC"]
```

### filter
```{r}
check <- hist(log10(rowSums(logcounts(lc3))),breaks=seq(-2,8,0.25))
plot(check)
## looks like the minimum we want is somewhere around log10 1 to log10 2
View(cbind(check$breaks[2:length(check$breaks)],check$counts))
# 0.75-1: 1392; 1-1.25: 1254 
# <10^1.25 is the last major drop-off point, so let's set total gene logcounts 10^1.25 as the cutoff
dev.off()

## drop genes w less than 56 counts
keepg <- rownames(lc3)[rowSums(logcounts(lc3))>=10^1.25]
length(keepg) ## 15170
lc3 <- lc3[keepg,]

## clean up for second run
rm(h,lc.nmf2,w,lc.nmf,lclct,lcrn,lccn,check,keepg,rddt)
gc(full=T)

## renormalize
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)
## get new logcount matrix, rownames colnames and gene info
lclct <- logcounts(lc3)
lcrn <- rownames(lclct)
lccn <- colnames(lclct)

# get gene metadata to add symbol info later
rddt <- as.data.table(rowData(lc3)) 
```

run again
```{r}
rm(lc3,anno,clus)
gc(full=T)

setDTthreads(1,restore_after_fork = FALSE)
options("RcppML.threads"=11)
options("RcppML.verbose"=TRUE)

lc.nmf <-nmf(lclct,
    k=53,
    tol = 1e-09, 
    maxit = 5000,
    verbose = T,
    L1 = 0.1,
    seed = 42,
    mask_zeros = FALSE,
    diag = TRUE,
    nonneg = TRUE
)

w <- as.data.frame(lc.nmf$w)
rownames(w) <- lcrn
colnames(w) <- paste0("LCnmf",c(1:53))

h <- as.data.frame(lc.nmf$h)
rownames(h) <- paste0("LCnmf",c(1:53))
colnames(h) <- colnames(lclct)

lc.nmf2 <- list(w=w,h=h)
saveRDS(lc.nmf2,"processed-data/10_subclustering/07thru10_ver2_53factorNMF/07b-lcnmf53_tol1e-9_rowSumLogctFilt18.RDS")
```

repinfo
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] Matrix_1.7-4                data.table_1.18.2.1         RcppML_0.5.6               
 [4] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1        Seqinfo_1.0.0              
[10] IRanges_2.44.0              S4Vectors_0.48.0            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0       matrixStats_1.5.0          

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        beeswarm_0.4.0      rjson_0.2.23        xfun_0.56          
 [5] ggplot2_4.0.2       ggrepel_0.9.6       lattice_0.22-7      vctrs_0.7.1        
 [9] tools_4.5.2         parallel_4.5.2      tibble_3.3.1        pkgconfig_2.0.3    
[13] BiocNeighbors_2.4.0 RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5    
[17] compiler_4.5.2      farver_2.1.2        codetools_0.2-20    vipor_0.4.7        
[21] htmltools_0.5.9     yaml_2.3.12         pillar_1.11.1       BiocParallel_1.44.0
[25] DelayedArray_0.36.0 viridis_0.6.5       magick_2.9.0        abind_1.4-8        
[29] tidyselect_1.2.1    rsvd_1.0.5          digest_0.6.39       BiocSingular_1.26.0
[33] dplyr_1.2.0         rprojroot_2.1.1     fastmap_1.2.0       grid_4.5.2         
[37] here_1.0.2          cli_3.6.5           SparseArray_1.10.8  scater_1.38.0      
[41] magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1   scales_1.4.0       
[45] ggbeeswarm_0.7.3    rmarkdown_2.30      XVector_0.50.0      otel_0.2.0         
[49] gridExtra_2.3       ScaledMatrix_1.18.0 beachmat_2.26.0     evaluate_1.0.5     
[53] knitr_1.51          viridisLite_0.4.3   irlba_2.3.7         rlang_1.1.7.9000   
[57] Rcpp_1.1.1          glue_1.8.0          scuttle_1.20.0      rstudioapi_0.18.0  
[61] R6_2.6.1           
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-21
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0      2021-06-01 [2] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel           1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [2] CRAN (R 4.5.2)
 ggplot2                4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 gridExtra              2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 Matrix               * 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RcppML               * 0.5.6      2026-02-07 [2] Github (zdebruine/RcppML@5449a5b)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 rsvd                   1.0.5      2021-04-16 [2] CRAN (R 4.5.0)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scater                 1.38.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [2] CRAN (R 4.5.0)
 viridis                0.6.5      2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite            0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────
