---
title: "01-LC_nnsvg"
author: "Bernie Mulvey"
date: "2025-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(nnSVG)
library(scater)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallel)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallelly)
options(bphost="localhost")
library(BiocParallel)

```

### Setup: load data, apply cluster labels. LC is cluster X6.
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")

louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]

setnames(louvs,c("rn","cl"))
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$cl

```

### make subsetted SPE that only includes LC (X6)
```{r}
lconly <- lc3[,lc3$label=="X6"]
### 

### prep to iterate nnSVG over each tissue section
caps <- lapply(unique(lconly$sample_id),FUN=function(x){lconly[,lconly$sample_id==x]})
names(caps) <- unique(lconly$sample_id)

# length(caps)
# 77 tissue sections with 1+ LC spot

### remove samples with fewer than 25 LC spots
drops <- lapply(caps,FUN=function(x){ncol(x)<25})
## which(drops==T)
# V13M06-332_C1s1 V13M06-332_C1s2 V13M06-403_A1s1 V13M06-403_A1s2 V13M06-401_C1s2 V13B23-286_B1s1 V13B23-286_B1s2

caps <- caps[!(names(caps) %in% c("V13M06-332_C1s1","V13M06-332_C1s2","V13M06-403_A1s1","V13M06-403_A1s2","V13M06-401_C1s2","V13B23-286_B1s1","V13B23-286_B1s2"))]

# length(caps)
# 70 tissue sections with over 25 LC spots

## cleanup
rm(lc3,drops,louvs,lconly)
gc(full=T)
```

# run nnsvg: 1 section at a time x 10 threads

## test some parameters to see how many genes would even hypothetically be able to pass a xenium-nomination requirement of being nomsig in 50% of (remaining) samples, with different min counts and sample-specific minimum pct spots corresponding to S spots.

## (sapmle-specific threshold because there's a pretty big spread of how many LC spots in a sample -- ranges from 25 to ~250)
```{r}
# testres <- lapply(caps,FUN=function(x){
#   thresh <- round(100*(10/ncol(x)),digits = 1)
#   print(thresh)
#   filter_genes(x,filter_genes_ncounts=1,filter_genes_pcspots = thresh,filter_mito = FALSE)
# })
# 
# sampspergene <- as.data.frame(table(unlist(lapply(testres,rownames))))
# nrow(sampspergene[sampspergene$Freq>=35,])

# varying the values input to filter_genes above:
# 3814 genes with ≥ 2 counts in ≥4 LC spots in ≥35 samples 
# 2944 genes with " " " ≥ 6 " " " " "
# 6713 genes with ≥1 counts in ≥10 LC spots in ≥35 samples 

# rm(testres,sampspergene)

```

# let's see if nnsvg can run on these samples with 2 counts in at least 4 LC spots of a sample. casting the net any wider seems to just be crashing nnsvg. (if there's too few spots with counts of a gene for it to run, it'll just run perpetually, or inconsistently error out). 

```{r}
# prevent data.table from multithreading on parallel workers and thus overscheduling of processors
setDTthreads(threads = 1,restore_after_fork = FALSE)

# res_list2 <- list()
# i <- 1
res_list2[[7]] <- paste0(names(caps)[7]," failed")
# i<-8
res_list2[[43]] <- paste0(names(caps)[43]," failed")
# i <- 44
res_list2[[44]] <- paste0(names(caps)[44]," failed")
# i<-45
res_list2[[51]] <- paste0(names(caps)[51]," failed")
# i<-52
res_list2[[57]] <- paste0(names(caps)[57]," failed")
# i<-58
res_list2[[58]] <- paste0(names(caps)[58]," failed")
# i<-59
res_list2[[67]] <- paste0(names(caps)[67]," failed")
i<-68

# res_list2 <- lapply(caps,FUN=function(lc.samp){
# for (i in c(1:length(caps))){
# for (i in c(8:length(caps))){
# for (i in c(44:length(caps))){
# for (i in c(45:length(caps))){
# for (i in c(59:length(caps))){
for (i in c(68:length(caps))){
    print(Sys.time())
    
    lc.samp <- caps[[i]]
    
    thresh <- round(100*(4/ncol(lc.samp)),digits = 1)
    
    # BRISC (internally called thru nnsvg) REQUIRES at least 65 points for using the "AMMD" spatial-ordering scheme, which nnsvg uses by default. to avoid issues with samples having under 65 spots, check and set this parameter per sample)
    briscord <- ifelse(ncol(lc.samp)<65,yes = "Sum_coords",no="AMMD")
    
    ## filter, lognorm, and run nnSVG
    lc.samp <- filter_genes(lc.samp,filter_genes_ncounts = 2,filter_genes_pcspots = thresh,filter_mito = F)
    lc.samp <- computeLibraryFactors(lc.samp)
    lc.samp <- logNormCounts(lc.samp)
    
    out <- nnSVG(lc.samp,assay_name = "logcounts",BPPARAM = MulticoreParam(8), order = briscord)
    
    message(paste0("Completed sample ",unique(lc.samp$sample_id)))
    
    res_list2[[i]] <- as.data.table(rowData(out),keep.rownames=T)
    
    rm(lc.samp,out)
    gc(full=T)
    # return(ret)
#})
}

names(res_list2) <- names(caps)
```

####### Output messages (i=1 to i=7 (error)): ########
[1] "2025-02-02 18:26:44 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.4% (n = 5) of spatial locations
removed 25281 out of 29877 genes due to low expression
Completed sample V13M06-331_A1s2
[1] "2025-02-02 18:30:47 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.7% (n = 4) of spatial locations
removed 25929 out of 29877 genes due to low expression
Completed sample V13M06-331_A1s1
[1] "2025-02-02 18:34:16 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.3% (n = 4) of spatial locations
removed 26955 out of 29877 genes due to low expression
Completed sample V13M06-331_B1s2
[1] "2025-02-02 18:36:51 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.1% (n = 5) of spatial locations
removed 27518 out of 29877 genes due to low expression
Completed sample V13M06-331_B1s1
[1] "2025-02-02 18:38:56 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.9% (n = 4) of spatial locations
removed 22877 out of 29877 genes due to low expression
Completed sample V13M06-331_C1s2
[1] "2025-02-02 18:45:11 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.3% (n = 5) of spatial locations
removed 24134 out of 29877 genes due to low expression
Completed sample V13M06-331_C1s1
[1] "2025-02-02 18:50:16 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.4% (n = 5) of spatial locations
removed 24414 out of 29877 genes due to low expression
Timing stopped at: 0.029 0 0.029

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  0 remote errors, element index: 
  5463 unevaluated and other errors
  first remote error:


####### Output messages (i=8 to i=43 (error)) #######
[1] "2025-02-02 20:12:09 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.1% (n = 4) of spatial locations
removed 24306 out of 29877 genes due to low expression
Completed sample V13M06-331_D1s2
[1] "2025-02-02 20:17:07 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.6% (n = 5) of spatial locations
removed 25165 out of 29877 genes due to low expression
Completed sample V13M06-333_A1s1
[1] "2025-02-02 20:21:17 CST"
Gene filtering: retaining genes with at least 2 counts in at least 5.3% (n = 4) of spatial locations
removed 25554 out of 29877 genes due to low expression
Completed sample V13M06-333_A1s2
[1] "2025-02-02 20:25:05 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.5% (n = 4) of spatial locations
removed 23052 out of 29877 genes due to low expression
Completed sample V13M06-333_B1s2
[1] "2025-02-02 20:31:06 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4% (n = 4) of spatial locations
removed 24756 out of 29877 genes due to low expression
Completed sample V13M06-333_B1s1
[1] "2025-02-02 20:35:36 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.2% (n = 4) of spatial locations
removed 23277 out of 29877 genes due to low expression
Completed sample V13M06-333_C1s1
[1] "2025-02-02 20:41:26 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.3% (n = 5) of spatial locations
removed 24131 out of 29877 genes due to low expression
Completed sample V13M06-333_C1s2
[1] "2025-02-02 20:46:30 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.7% (n = 5) of spatial locations
removed 23439 out of 29877 genes due to low expression
Completed sample V13M06-333_D1s1
[1] "2025-02-02 20:52:14 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.5% (n = 4) of spatial locations
removed 22915 out of 29877 genes due to low expression
Completed sample V13M06-333_D1s2
[1] "2025-02-02 20:58:30 CST"
Gene filtering: retaining genes with at least 2 counts in at least 7.8% (n = 4) of spatial locations
removed 29316 out of 29877 genes due to low expression
Completed sample V13M06-332_A1s1
[1] "2025-02-02 20:59:01 CST"
Gene filtering: retaining genes with at least 2 counts in at least 6.3% (n = 4) of spatial locations
removed 27533 out of 29877 genes due to low expression
Completed sample V13M06-332_A1s2
[1] "2025-02-02 21:01:07 CST"
Gene filtering: retaining genes with at least 2 counts in at least 1.2% (n = 5) of spatial locations
removed 22884 out of 29877 genes due to low expression
Completed sample V13M06-332_B1s1
[1] "2025-02-02 21:07:45 CST"
Gene filtering: retaining genes with at least 2 counts in at least 1.7% (n = 4) of spatial locations
removed 23067 out of 29877 genes due to low expression
Completed sample V13M06-332_B1s2
[1] "2025-02-02 21:14:06 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.3% (n = 4) of spatial locations
removed 23742 out of 29877 genes due to low expression
Completed sample V13M06-332_D1s2
[1] "2025-02-02 21:19:58 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.2% (n = 5) of spatial locations
removed 23628 out of 29877 genes due to low expression
Completed sample V13M06-332_D1s1
[1] "2025-02-02 21:25:46 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.7% (n = 4) of spatial locations
removed 25045 out of 29877 genes due to low expression
Completed sample V13M06-404_A1s1
[1] "2025-02-02 21:30:15 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.4% (n = 5) of spatial locations
removed 25395 out of 29877 genes due to low expression
Completed sample V13M06-404_A1s2
[1] "2025-02-02 21:34:19 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3% (n = 4) of spatial locations
removed 25024 out of 29877 genes due to low expression
Completed sample V13M06-404_B1s2
[1] "2025-02-02 21:38:42 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.3% (n = 5) of spatial locations
removed 25199 out of 29877 genes due to low expression
Completed sample V13M06-404_B1s1
[1] "2025-02-02 21:42:55 CST"
Gene filtering: retaining genes with at least 2 counts in at least 13.8% (n = 5) of spatial locations
removed 29254 out of 29877 genes due to low expression
Completed sample V13M06-404_C1s1
[1] "2025-02-02 21:43:30 CST"
Gene filtering: retaining genes with at least 2 counts in at least 10.3% (n = 5) of spatial locations
removed 29006 out of 29877 genes due to low expression
Completed sample V13M06-404_C1s2
[1] "2025-02-02 21:44:18 CST"
Gene filtering: retaining genes with at least 2 counts in at least 6.2% (n = 5) of spatial locations
removed 28398 out of 29877 genes due to low expression
Completed sample V13M06-404_D1s1
[1] "2025-02-02 21:45:38 CST"
Gene filtering: retaining genes with at least 2 counts in at least 6.3% (n = 4) of spatial locations
removed 28384 out of 29877 genes due to low expression
Completed sample V13M06-404_D1s2
[1] "2025-02-02 21:46:58 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.8% (n = 4) of spatial locations
removed 26739 out of 29877 genes due to low expression
Completed sample V13M06-403_D1s2
[1] "2025-02-02 21:49:47 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.3% (n = 5) of spatial locations
removed 27940 out of 29877 genes due to low expression
Completed sample V13M06-403_D1s1
[1] "2025-02-02 21:51:34 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3% (n = 5) of spatial locations
removed 26668 out of 29877 genes due to low expression
Completed sample V13M06-402_A1s1
[1] "2025-02-02 21:54:30 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.2% (n = 5) of spatial locations
removed 26980 out of 29877 genes due to low expression
Completed sample V13M06-402_A1s2
[1] "2025-02-02 21:57:05 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.8% (n = 4) of spatial locations
removed 26446 out of 29877 genes due to low expression
Completed sample V13M06-402_B1s2
[1] "2025-02-02 22:00:09 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.8% (n = 4) of spatial locations
removed 24705 out of 29877 genes due to low expression
Completed sample V13M06-402_B1s1
[1] "2025-02-02 22:04:46 CST"
Gene filtering: retaining genes with at least 2 counts in at least 9.3% (n = 4) of spatial locations
removed 27467 out of 29877 genes due to low expression
Completed sample V13M06-402_C1s2
[1] "2025-02-02 22:06:54 CST"
Gene filtering: retaining genes with at least 2 counts in at least 12.9% (n = 4) of spatial locations
removed 28472 out of 29877 genes due to low expression
Completed sample V13M06-402_C1s1
[1] "2025-02-02 22:08:09 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.5% (n = 4) of spatial locations
removed 28172 out of 29877 genes due to low expression
Completed sample V13M06-402_D1s2
[1] "2025-02-02 22:09:41 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4% (n = 5) of spatial locations
removed 28697 out of 29877 genes due to low expression
Completed sample V13M06-402_D1s1
[1] "2025-02-02 22:10:45 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.1% (n = 4) of spatial locations
removed 25662 out of 29877 genes due to low expression
Completed sample V13M06-401_A1s2
[1] "2025-02-02 22:14:30 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.1% (n = 5) of spatial locations
removed 27484 out of 29877 genes due to low expression
Completed sample V13M06-401_A1s1
[1] "2025-02-02 22:16:37 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.6% (n = 4) of spatial locations
removed 23775 out of 29877 genes due to low expression
Timing stopped at: 0.006 0 0.006

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  0 remote errors, element index: 
  6102 unevaluated and other errors
  first remote error:
  
####### Output messages (i=44, error) #######
[1] "2025-02-02 22:35:41 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.4% (n = 4) of spatial locations
removed 24707 out of 29877 genes due to low expression
Timing stopped at: 0.014 0.001 0.013

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  1 remote errors, element index: 248
  4922 unevaluated and other errors
  first remote error:
Error in BRISC_estimation(coords = coords, y = y_i, x = X, cov.model = "exponential", : c++ error: dpotrf failed


####### Output messages (i=45 to i=51 (error)) #######
[1] "2025-02-02 22:40:22 CST"
Gene filtering: retaining genes with at least 2 counts in at least 16% (n = 4) of spatial locations
removed 28953 out of 29877 genes due to low expression
Completed sample V13M06-401_C1s1
[1] "2025-02-02 22:41:12 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.6% (n = 4) of spatial locations
removed 24631 out of 29877 genes due to low expression
Completed sample V13M06-401_D1s1
[1] "2025-02-02 22:45:59 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.1% (n = 4) of spatial locations
removed 23297 out of 29877 genes due to low expression
Completed sample V13M06-401_D1s2
[1] "2025-02-02 22:51:56 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.8% (n = 4) of spatial locations
removed 25551 out of 29877 genes due to low expression
Completed sample V13M06-386_A1s2
[1] "2025-02-02 22:55:51 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.3% (n = 5) of spatial locations
removed 26733 out of 29877 genes due to low expression
Completed sample V13M06-386_A1s1
[1] "2025-02-02 22:58:42 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.4% (n = 4) of spatial locations
removed 26083 out of 29877 genes due to low expression
Completed sample V13M06-386_C1s1
[1] "2025-02-02 23:02:08 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.7% (n = 5) of spatial locations
removed 25855 out of 29877 genes due to low expression
Timing stopped at: 0.014 0.001 0.014

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  0 remote errors, element index: 
  4022 unevaluated and other errors
  first remote error:

####### Output messages (i=52 to i=57 (error)) #######
[1] "2025-02-02 23:07:41 CST"
Gene filtering: retaining genes with at least 2 counts in at least 9.3% (n = 4) of spatial locations
removed 28224 out of 29877 genes due to low expression
Completed sample V13M06-386_D1s1
[1] "2025-02-02 23:09:11 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.7% (n = 5) of spatial locations
removed 28000 out of 29877 genes due to low expression
Completed sample V13M06-386_D1s2
[1] "2025-02-02 23:10:53 CST"
Gene filtering: retaining genes with at least 2 counts in at least 1.6% (n = 4) of spatial locations
removed 24864 out of 29877 genes due to low expression
Completed sample V13B23-366_A1s1
[1] "2025-02-02 23:15:31 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.3% (n = 4) of spatial locations
removed 26725 out of 29877 genes due to low expression
Completed sample V13B23-284_A1s1
[1] "2025-02-02 23:18:21 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.6% (n = 5) of spatial locations
removed 28407 out of 29877 genes due to low expression
Completed sample V13B23-284_A1s2
[1] "2025-02-02 23:19:41 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.2% (n = 4) of spatial locations
removed 25254 out of 29877 genes due to low expression
Timing stopped at: 0.019 0 0.02

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  1 remote errors, element index: 180
  4443 unevaluated and other errors
  first remote error:
Error in BRISC_estimation(coords = coords, y = y_i, x = X, cov.model = "exponential", : c++ error: dpotrf failed

####### Output messages (i=58 (error)) #######
[1] "2025-02-02 23:29:51 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.9% (n = 4) of spatial locations
removed 26395 out of 29877 genes due to low expression
Timing stopped at: 0.008 0.001 0.009

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  0 remote errors, element index: 
  3482 unevaluated and other errors
  first remote error:

####### Output messages (i=59 to i=67 (error)) #######
[1] "2025-02-02 23:32:43 CST"
Gene filtering: retaining genes with at least 2 counts in at least 5% (n = 4) of spatial locations
removed 27896 out of 29877 genes due to low expression
Completed sample V13B23-284_C1s2
[1] "2025-02-02 23:34:42 CST"
Gene filtering: retaining genes with at least 2 counts in at least 6.2% (n = 4) of spatial locations
removed 28478 out of 29877 genes due to low expression
Completed sample V13B23-284_C1s1
[1] "2025-02-02 23:35:58 CST"
Gene filtering: retaining genes with at least 2 counts in at least 12.5% (n = 4) of spatial locations
removed 28668 out of 29877 genes due to low expression
Completed sample V13B23-284_D1s2
[1] "2025-02-02 23:37:04 CST"
Gene filtering: retaining genes with at least 2 counts in at least 11.8% (n = 5) of spatial locations
removed 28973 out of 29877 genes due to low expression
Completed sample V13B23-284_D1s1
[1] "2025-02-02 23:37:54 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.4% (n = 5) of spatial locations
removed 25936 out of 29877 genes due to low expression
Completed sample V13F27-339_A1s1
[1] "2025-02-02 23:41:27 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4% (n = 4) of spatial locations
removed 26420 out of 29877 genes due to low expression
Completed sample V13F27-339_A1s2
[1] "2025-02-02 23:44:34 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4% (n = 4) of spatial locations
removed 26952 out of 29877 genes due to low expression
Completed sample V13F27-339_B1s2
[1] "2025-02-02 23:47:11 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.2% (n = 4) of spatial locations
removed 26833 out of 29877 genes due to low expression
Completed sample V13F27-339_B1s1
[1] "2025-02-02 23:49:56 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.8% (n = 5) of spatial locations
removed 25280 out of 29877 genes due to low expression
Timing stopped at: 0.016 0.001 0.016

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  0 remote errors, element index: 
  4597 unevaluated and other errors
  first remote error:

####### Output messages (i=68 to i=70) -- finished! #######

[1] "2025-02-03 11:11:55 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.8% (n = 5) of spatial locations
removed 24609 out of 29877 genes due to low expression
Completed sample V13F27-339_C1s1
[1] "2025-02-03 11:16:43 CST"
Gene filtering: retaining genes with at least 2 counts in at least 1.9% (n = 5) of spatial locations
removed 23215 out of 29877 genes due to low expression
Completed sample V13B23-286_A1s1
[1] "2025-02-03 11:22:53 CST"
Gene filtering: retaining genes with at least 2 counts in at least 1.9% (n = 4) of spatial locations
removed 23576 out of 29877 genes due to low expression
Completed sample V13B23-286_A1s2

## concatenate and save ## 
```{r}
# only get analyses that completed
res_list <- res_list2[unlist(lapply(res_list2,is.character))==F]

res_list <- lapply(res_list,FUN=function(x){
  # append nnsvg to the nnsvg analysis columns of the output rowDatas contained in res_list
  nongeneinfocols <- ncol(rowData(caps[[1]]))+1
  setnames(x,
           c(nongeneinfocols:ncol(x)),
           paste0("nnsvg_",names(x)[c(nongeneinfocols:ncol(x))])
  )
})

saveRDS(res_list,"processed-data/09_local_svgs/01a-LC_min-2count-4spot_prelim_fullout.RDS")
```

### tabulate significant genes per sample
```{r}
i <- 1
for (i in c(1:length(res_list))){
    tmp <- copy(res_list[[i]])
    tmp <- tmp[nnsvg_pval<0.05,.(gene_id,gene_name,nnsvg_rank)]
    setnames(tmp,"nnsvg_rank",names(res_list)[i])
    if(i==1){lcsvgranks <- copy(tmp)}
    else{lcsvgranks <- merge.data.table(lcsvgranks,tmp,by=c("gene_id","gene_name"),all.x=T,all.y=T)}
    rm(tmp)
}

lcsvgranks[,nasamps:=rowSums(is.na(lcsvgranks[,c(3:ncol(lcsvgranks)),with=F]))]
#
lcsvgranks[,nsigsamps:=(ncol(lcsvgranks)-3)-nasamps]
lcsvgranks[,meanrank:=rowMeans(lcsvgranks[,c(3:(ncol(lcsvgranks)-2)),with=F],na.rm=T)]
lcsvgranks[,nasamps:=NULL]

fwrite(lcsvgranks,"processed-data/09_local_svgs/01b-LC prelim nnSVG nomsig genes and ranks per sample.txt",sep='\t',quote=F)
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] BiocParallel_1.40.0         parallelly_1.42.0          
 [3] colorout_1.3-0.2            scater_1.34.0              
 [5] ggplot2_3.5.1               scuttle_1.16.0             
 [7] nnSVG_1.10.0                SpatialExperiment_1.16.0   
 [9] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
[11] Biobase_2.66.0              GenomicRanges_1.58.0       
[13] GenomeInfoDb_1.42.1         IRanges_2.40.1             
[15] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[17] MatrixGenerics_1.18.0       matrixStats_1.5.0          
[19] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] beeswarm_0.4.0          gtable_0.3.6            rjson_0.2.23           
 [4] xfun_0.50               ggrepel_0.9.6           lattice_0.22-6         
 [7] vctrs_0.6.5             tools_4.4.1             generics_0.1.3         
[10] tibble_3.2.1            BiocBaseUtils_1.8.0     BiocNeighbors_2.0.1    
[13] pkgconfig_2.0.3         Matrix_1.7-2            lifecycle_1.0.4        
[16] GenomeInfoDbData_1.2.13 compiler_4.4.1          munsell_0.5.1          
[19] codetools_0.2-20        vipor_0.4.7             snow_0.4-4             
[22] htmltools_0.5.8.1       yaml_2.3.10             pillar_1.10.1          
[25] crayon_1.5.3            DelayedArray_0.32.0     viridis_0.6.5          
[28] BRISC_1.0.6             magick_2.8.5            abind_1.4-8            
[31] rsvd_1.0.5              tidyselect_1.2.1        digest_0.6.37          
[34] BiocSingular_1.22.0     dplyr_1.1.4             rprojroot_2.0.4        
[37] fastmap_1.2.0           grid_4.4.1              here_1.0.1             
[40] colorspace_2.1-1        cli_3.6.3               SparseArray_1.6.0      
[43] magrittr_2.0.3          S4Arrays_1.6.0          withr_3.0.2            
[46] UCSC.utils_1.2.0        scales_1.3.0            ggbeeswarm_0.7.2       
[49] rmarkdown_2.29          XVector_0.46.0          httr_1.4.7             
[52] gridExtra_2.3           RANN_2.6.2              ScaledMatrix_1.14.0    
[55] beachmat_2.22.0         pbapply_1.7-2           evaluate_1.0.3         
[58] knitr_1.49              viridisLite_0.4.2       irlba_2.3.5.1          
[61] Rcpp_1.0.14             rdist_0.0.5             glue_1.8.0             
[64] rstudioapi_0.17.1       jsonlite_1.8.9          R6_2.5.1               
[67] zlibbioc_1.52.0        
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-03
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    beachmat               2.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocBaseUtils          1.8.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocNeighbors          2.0.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel         * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BRISC                  1.0.6   2024-10-02 [1] CRAN (R 4.4.1)
 P  cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
    colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1  2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6   2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra              2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
    jsonlite               1.8.9   2024-09-20 [1] CRAN (R 4.4.1)
    knitr                  1.49    2024-11-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magick                 2.8.5   2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    Matrix                 1.7-2   2025-01-23 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
    nnSVG                * 1.10.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    parallelly           * 1.42.0  2025-01-30 [1] CRAN (R 4.4.1)
    pbapply                1.7-2   2023-06-27 [1] CRAN (R 4.4.0)
    pillar                 1.10.1  2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
    RANN                   2.6.2   2024-08-25 [1] CRAN (R 4.4.1)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rdist                  0.0.5   2020-05-04 [1] CRAN (R 4.4.0)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4   2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ScaledMatrix           1.14.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
    scater               * 1.34.0  2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
    scuttle              * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    snow                   0.4-4   2021-10-27 [1] CRAN (R 4.4.0)
    SparseArray            1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.50    2025-01-07 [1] CRAN (R 4.4.1)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
