---
title: "01b-LC_nnsvg_rerun_error_samples"
author: "Bernie Mulvey"
date: "2025-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(nnSVG)
library(scater)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallel)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallelly)
options(bphost="localhost")
library(BiocParallel)

```

### Setup: load data, apply cluster labels. LC is cluster X6.
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")

louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]

setnames(louvs,c("rn","cl"))
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$cl

```

### make subsetted SPE that only includes LC (X6)
```{r}
lconly <- lc3[,lc3$label=="X6"]
### 

### prep to iterate nnSVG over each tissue section
caps <- lapply(unique(lconly$sample_id),FUN=function(x){lconly[,lconly$sample_id==x]})
names(caps) <- unique(lconly$sample_id)

# length(caps)
# 77 tissue sections with 1+ LC spot

### remove samples with fewer than 25 LC spots
drops <- lapply(caps,FUN=function(x){ncol(x)<25})
## which(drops==T)
# V13M06-332_C1s1 V13M06-332_C1s2 V13M06-403_A1s1 V13M06-403_A1s2 V13M06-401_C1s2 V13B23-286_B1s1 V13B23-286_B1s2

caps <- caps[!(names(caps) %in% c("V13M06-332_C1s1","V13M06-332_C1s2","V13M06-403_A1s1","V13M06-403_A1s2","V13M06-401_C1s2","V13B23-286_B1s1","V13B23-286_B1s2"))]

# length(caps)
# 70 tissue sections with over 25 LC spots

## cleanup
rm(lc3,drops,louvs,lconly)
gc(full=T)
```

## identify samples that run was attempted on but not in the initial output
```{r}
svgtmp <- fread("processed-data/09_local_svgs/01b-LC prelim nnSVG nomsig genes and ranks per sample.txt")

failt <- names(caps)[!(names(caps) %in% names(svgtmp))]

caps <- caps[failt]

rm(svgtmp)
```

# run nnsvg: 1 section at a time x 10 threads

# using let's see if nnsvg can run on these samples with 2 counts in at least 4 LC spots of a sample. casting the net any wider seems to just be crashing nnsvg. (if there's too few spots with counts of a gene for it to run, it'll just run perpetually, or inconsistently error out). 

```{r}
# prevent data.table from multithreading on parallel workers and thus overscheduling of processors
setDTthreads(threads = 1,restore_after_fork = FALSE)

res_list2 <- list()
# i <- 1
# i <- 4
# i<-5
i <- 6

res_list2[[4]] <- paste0(names(caps)[4]," failed twice")
res_list2[[5]] <- paste0(names(caps)[5]," failed twice")

# for (i in c(1:length(caps))){
# for (i in c(4:length(caps))){
for (i in c(6:length(caps))){

    print(Sys.time())
    
    lc.samp <- caps[[i]]
    
    thresh <- round(100*(4/ncol(lc.samp)),digits = 1)
    
    # BRISC (internally called thru nnsvg) REQUIRES at least 65 points for using the "AMMD" spatial-ordering scheme, which nnsvg uses by default. to avoid issues with samples having under 65 spots, check and set this parameter per sample)
    briscord <- ifelse(ncol(lc.samp)<65,yes = "Sum_coords",no="AMMD")
    
    ## filter, lognorm, and run nnSVG
    lc.samp <- filter_genes(lc.samp,filter_genes_ncounts = 2,filter_genes_pcspots = thresh,filter_mito = F)
    
    # per error message, drop post-filtering row (gene) #180 from sample 5
    if(i==5){
        lc.samp <- lc.samp[c(c(1:179),c(181:nrow(lc.samp))),]
    }
    
    lc.samp <- computeLibraryFactors(lc.samp)
    lc.samp <- logNormCounts(lc.samp)
    
    out <- tryCatch(nnSVG(lc.samp,assay_name = "logcounts",BPPARAM = MulticoreParam(8,stop.on.error = TRUE,log = T), order = briscord))
    
    message(paste0("Completed sample ",unique(lc.samp$sample_id)))
    
    res_list2[[i]] <- as.data.table(rowData(out),keep.rownames=T)
    
    rm(lc.samp,out)
    gc(full=T)
    # return(ret)
#})
}

names(res_list2) <- names(caps)
```


### output messages ###
Gene filtering: retaining genes with at least 2 counts in at least 2.4% (n = 5) of spatial locations
removed 24414 out of 29877 genes due to low expression
Completed sample V13M06-331_D1s1

[1] "2025-02-03 21:14:24 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.6% (n = 4) of spatial locations
removed 23775 out of 29877 genes due to low expression
Completed sample V13M06-401_B1s2

[1] "2025-02-03 21:19:14 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.4% (n = 4) of spatial locations
removed 24707 out of 29877 genes due to low expression
Completed sample V13M06-401_B1s1

### i=4, first try (error) ###
[1] "2025-02-03 21:23:17 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.7% (n = 5) of spatial locations
removed 25855 out of 29877 genes due to low expression
Timing stopped at: 0.014 0.001 0.014

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  0 remote errors, element index: 
  4022 unevaluated and other errors
  first remote error:
  
  ### trying again with i=4 ###
  [1] "2025-02-03 21:25:36 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.7% (n = 5) of spatial locations
removed 25855 out of 29877 genes due to low expression
Timing stopped at: 0.014 0 0.013

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  0 remote errors, element index: 
  4022 unevaluated and other errors
  first remote error:
  
## okay, so maybe V13M06-386_C1s2 just really isn't going to work.##
  
### proceeding to i=5 -- an error, but one that actually tells us a problem gene to kick--180: ###

  [1] "2025-02-03 21:28:52 CST"
Gene filtering: retaining genes with at least 2 counts in at least 3.2% (n = 4) of spatial locations
removed 25254 out of 29877 genes due to low expression
Timing stopped at: 0.019 0.001 0.021

Stop worker failed with the error: wrong args for environment subassignment
Error: BiocParallel errors
  1 remote errors, element index: 180
  4443 unevaluated and other errors
  first remote error:
Error in BRISC_estimation(coords = coords, y = y_i, x = X, cov.model = "exponential", : c++ error: dpotrf failed

### i=5 again with trycatch and BPPARAMs to get back potential debug info -- no further info on problem genes. this one might be borked too ###
############### LOG OUTPUT ###############
Task: 2
Node: 7
Timestamp: 2025-02-03 21:39:41.987561
Success: FALSE

Task duration:
   user  system elapsed 
 52.378   0.527  53.165 

Memory used:
           used  (Mb) gc trigger  (Mb) limit (Mb)  max used   (Mb)
Ncells  7941002 424.1   14833578 792.2         NA  14833578  792.2
Vcells 19871973 151.7   56347322 429.9     512000 279802649 2134.8

Log messages:
INFO [2025-02-03 21:38:48] loading futile.logger package
ERROR [2025-02-03 21:39:41] c++ error: dpotrf failed

stderr and stdout:
Timing stopped at: 0.021 0 0.021

Stop worker failed with the error: wrong args for environment subassignment

### skip and proceed to i=6 ### 
[1] "2025-02-03 21:43:11 CST"
Gene filtering: retaining genes with at least 2 counts in at least 4.9% (n = 4) of spatial locations
removed 26395 out of 29877 genes due to low expression
Completed sample V13B23-284_B1s1

[1] "2025-02-03 21:45:57 CST"
Gene filtering: retaining genes with at least 2 counts in at least 2.8% (n = 5) of spatial locations
removed 25280 out of 29877 genes due to low expression
Completed sample V13F27-339_C1s2

## concatenate and save ## 
```{r}
# only get analyses that completed
res_list <- res_list2[unlist(lapply(res_list2,is.character))==F]

res_list <- lapply(res_list,FUN=function(x){
  # append nnsvg to the nnsvg analysis columns of the output rowDatas contained in res_list
  nongeneinfocols <- ncol(rowData(caps[[1]]))+1
  setnames(x,
           c(nongeneinfocols:ncol(x)),
           paste0("nnsvg_",names(x)[c(nongeneinfocols:ncol(x))])
  )
})

saveRDS(res_list,"processed-data/09_local_svgs/01c-LC_min-2count-4spot_5of7_initialfails_fullout.RDS")
```

append these to the previous result table

### tabulate significant genes per sample
```{r}
svgtab <- fread("processed-data/09_local_svgs/01b-LC prelim nnSVG nomsig genes and ranks per sample.txt")

i <- 1
for (i in c(1:length(res_list))){
    tmp <- copy(res_list[[i]])
    tmp <- tmp[nnsvg_pval<0.05,.(gene_id,gene_name,nnsvg_rank)]
    setnames(tmp,"nnsvg_rank",names(res_list)[i])
    svgtab <- merge.data.table(svgtab,tmp,by=c("gene_id","gene_name"),all.x=T,all.y=T)
    rm(tmp)
}

# remove old summary columns
svgtab[,nsigsamps:=NULL]
svgtab[,meanrank:=NULL]

svgtab[,nasamps:=rowSums(is.na(svgtab[,c(3:ncol(svgtab)),with=F]))]
#
svgtab[,nsigsamps:=(ncol(svgtab)-3)-nasamps]
svgtab[,meanrank:=rowMeans(svgtab[,c(3:(ncol(svgtab)-2)),with=F],na.rm=T)]
svgtab[,nasamps:=NULL]

fwrite(svgtab,"processed-data/09_local_svgs/01d-LC prelim nnSVG 5addnl recovered samples nomsig genes and ranks per sample.txt",sep='\t',quote=F)
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] BiocParallel_1.40.0         parallelly_1.42.0          
 [3] colorout_1.3-0.2            scater_1.34.0              
 [5] ggplot2_3.5.1               scuttle_1.16.0             
 [7] nnSVG_1.10.0                SpatialExperiment_1.16.0   
 [9] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
[11] Biobase_2.66.0              GenomicRanges_1.58.0       
[13] GenomeInfoDb_1.42.1         IRanges_2.40.1             
[15] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[17] MatrixGenerics_1.18.0       matrixStats_1.5.0          
[19] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] beeswarm_0.4.0          gtable_0.3.6            rjson_0.2.23           
 [4] xfun_0.50               ggrepel_0.9.6           lattice_0.22-6         
 [7] vctrs_0.6.5             tools_4.4.1             generics_0.1.3         
[10] tibble_3.2.1            BiocNeighbors_2.0.1     pkgconfig_2.0.3        
[13] Matrix_1.7-2            lifecycle_1.0.4         GenomeInfoDbData_1.2.13
[16] compiler_4.4.1          munsell_0.5.1           codetools_0.2-20       
[19] vipor_0.4.7             htmltools_0.5.8.1       yaml_2.3.10            
[22] pillar_1.10.1           crayon_1.5.3            DelayedArray_0.32.0    
[25] sessioninfo_1.2.2       viridis_0.6.5           BRISC_1.0.6            
[28] magick_2.8.5            abind_1.4-8             rsvd_1.0.5             
[31] tidyselect_1.2.1        digest_0.6.37           BiocSingular_1.22.0    
[34] dplyr_1.1.4             rprojroot_2.0.4         fastmap_1.2.0          
[37] grid_4.4.1              here_1.0.1              colorspace_2.1-1       
[40] cli_3.6.3               SparseArray_1.6.0       magrittr_2.0.3         
[43] S4Arrays_1.6.0          withr_3.0.2             UCSC.utils_1.2.0       
[46] scales_1.3.0            ggbeeswarm_0.7.2        rmarkdown_2.29         
[49] XVector_0.46.0          httr_1.4.7              gridExtra_2.3          
[52] RANN_2.6.2              ScaledMatrix_1.14.0     beachmat_2.22.0        
[55] pbapply_1.7-2           evaluate_1.0.3          knitr_1.49             
[58] viridisLite_0.4.2       irlba_2.3.5.1           Rcpp_1.0.14            
[61] rdist_0.0.5             glue_1.8.0              rstudioapi_0.17.1      
[64] jsonlite_1.8.9          R6_2.5.1                zlibbioc_1.52.0        
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-03
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    beachmat               2.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocNeighbors          2.0.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel         * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BRISC                  1.0.6   2024-10-02 [1] CRAN (R 4.4.1)
 P  cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
    colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1  2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6   2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra              2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
    jsonlite               1.8.9   2024-09-20 [1] CRAN (R 4.4.1)
    knitr                  1.49    2024-11-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magick                 2.8.5   2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    Matrix                 1.7-2   2025-01-23 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
    nnSVG                * 1.10.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    parallelly           * 1.42.0  2025-01-30 [1] CRAN (R 4.4.1)
    pbapply                1.7-2   2023-06-27 [1] CRAN (R 4.4.0)
    pillar                 1.10.1  2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
    RANN                   2.6.2   2024-08-25 [1] CRAN (R 4.4.1)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rdist                  0.0.5   2020-05-04 [1] CRAN (R 4.4.0)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4   2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ScaledMatrix           1.14.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
    scater               * 1.34.0  2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
    scuttle              * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.50    2025-01-07 [1] CRAN (R 4.4.1)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
