---
title: "02sanity-pval_dists_per_DEtest"
author: "Bernie Mulvey"
date: "2025-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load de results from all-clusters-jointly DE analyses
```{r}
detabs <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")

detabs <- lapply(detabs,FUN=function(d){
    ret <- lapply(d,function(x){
        y <- x[[6]] # slot with DE tables
        z <- rbindlist(y,idcol="label_and_test")
        return(z)
    })
})

detabs <- lapply(detabs,function(x){
    rbindlist(x,idcol="specifiedDesign")
})

detabs <- rbindlist(detabs,idcol="ancestry_coding")

## tidy up any ugly contrast names
detabs[label_and_test %in% grep(label_and_test,pattern="_Sex",value=T),
       label:=gsub(label_and_test,pattern="^(.*)_Sex$",replacement="\\1")]
detabs[label_and_test %in% grep(label_and_test,pattern="_E4vE2",value=T),
       label:=gsub(label_and_test,pattern="^(.*)_E4vE2$",replacement="\\1")]
detabs[label_and_test %in% grep(label_and_test,pattern="E4.*_vs_.*",value=T),
       label:=gsub(label_and_test,pattern="^(.*)_E4.*vs.*$",replacement="\\1")]
detabs[label_and_test %in% grep(label_and_test,pattern="_E2.*_vs_.*",value=T),
       label:=gsub(label_and_test,pattern="^(.*)_E2.*vs.*$",replacement="\\1")]
detabs[label_and_test %in% grep(label_and_test,pattern="_E4.*_vs_.*",value=T),
       label:=gsub(label_and_test,pattern="^(.*)_E4.*vs.*$",replacement="\\1")]
detabs[label_and_test %in% grep(label_and_test,pattern="_by_YRI",value=T),
       label:=gsub(label_and_test,pattern="^(.*)_by_YRI$",replacement="\\1")]
detabs[label_and_test %in% grep(label_and_test,pattern="_YRI",value=T),
       label:=gsub(label_and_test,pattern="^(.*)_YRI$",replacement="\\1")]

detabs[,contrast:=apply(.SD,MARGIN=1,FUN=function(x){paste0(x[1],"_",x[2])}),.SDcols=c("label_and_test","ancestry_coding")]
## plot for main approach of interest (continuous ancestry), except for ancestry itself in which casae plot for continuous and cat
lapply(unique(detabs$specifiedDesign),function(d){
    subdat <- detabs[specifiedDesign == d,]
    p <- ggplot(subdat,(aes(x=P.Value)))+
        geom_histogram()+
        facet_wrap(contrast~label)
    pdf(paste0("plots/12_DEanalyses_removedsampsAndFinalNMseg/02b_Allclust_dropmixclusts_SexDEG_EDA/02sanity-pval_dists_",d,"Model.pdf"),width=20,height=14,onefile =FALSE)
    print(p)
    dev.off()
})
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.40.0 parallelly_1.43.0   colorout_1.3-0.2    gridExtra_2.3      
[5] ggtext_0.1.2        ggplot2_3.5.2       data.table_1.17.4   rlang_1.1.5        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6       dplyr_1.1.4        compiler_4.4.3     tidyselect_1.2.1  
 [5] Rcpp_1.0.14        xml2_1.3.8         parallel_4.4.3     dichromat_2.0-0.1 
 [9] scales_1.4.0       yaml_2.3.10        fastmap_1.2.0      here_1.0.1        
[13] R6_2.6.1           labeling_0.4.3     generics_0.1.4     knitr_1.50        
[17] tibble_3.2.1       rprojroot_2.0.4    pillar_1.10.2      RColorBrewer_1.1-3
[21] xfun_0.52          cli_3.6.4          withr_3.0.2        magrittr_2.0.3    
[25] digest_0.6.37      grid_4.4.3         gridtext_0.1.5     rstudioapi_0.17.1 
[29] lifecycle_1.0.4    vctrs_0.6.5        evaluate_1.0.3     glue_1.8.0        
[33] farver_2.1.2       codetools_0.2-20   rmarkdown_2.29     tools_4.4.3       
[37] pkgconfig_2.0.3    htmltools_0.5.8.1 
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-07-20
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────────────
 !  package      * version date (UTC) lib source
    BiocParallel * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli            3.6.4   2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.3)
    colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    data.table   * 1.17.4  2025-05-26 [1] CRAN (R 4.4.3)
    dichromat      2.0-0.1 2022-05-02 [1] CRAN (R 4.4.0)
    digest         0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate       1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver         2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics       0.1.4   2025-05-09 [1] CRAN (R 4.4.1)
    ggplot2      * 3.5.2   2025-04-09 [1] CRAN (R 4.4.1)
    ggtext       * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue           1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext       0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable         0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    knitr          1.50    2025-03-16 [1] CRAN (R 4.4.2)
    labeling       0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    parallelly   * 1.43.0  2025-03-24 [1] CRAN (R 4.4.1)
    pillar         1.10.2  2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6             2.6.1   2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer   1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp           1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.5   2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown      2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi     0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    scales         1.4.0   2025-04-24 [1] CRAN (R 4.4.1)
    sessioninfo    1.2.3   2025-02-05 [1] CRAN (R 4.4.1)
    tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    withr          3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun           0.52    2025-04-02 [1] CRAN (R 4.4.1)
    xml2           1.3.8   2025-03-14 [1] CRAN (R 4.4.1)
    yaml           2.3.10  2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────────────────
