---
title: "01-ID tissue section edges"
output: html_document
date: "2024-07-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(spdep)
library(scater)
library(dplyr)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load SPE
```{r}
lc <- readRDS("processed-data/02_build_spe/spe_raw.rds")

## make colnames globally unique
colnames(lc) <- lc$key
# ## we WANT out of tissue spots for edge detection.
# lc <- lc[,lc$in_tissue==TRUE]

# ## drop zero count spots, zero count genes
# lc2 <- lc[,colSums(counts(lc))!=0]
# lc2 <- lc2[rowSums(counts(lc))!=0,]
```

functions written to precede spotsweeper etc, code by Jacqui Thomson: https://github.com/LieberInstitute/spatialDLPFC_mdd_bpd/blob/devel/code/03_QC/01_revised_edge_detection.r 
```{r}

## initialize list of sample 
l1 = unique(lc$sample_id)
names(l1) = lapply(l1, function(x) unique(colData(lc)[lc$sample_id==x,"sample_id"]))

# extract colData per sample
l1 = lapply(l1, function(x) colData(lc[,lc$sample_id==x]))

## get nearest neighbors and determine whether they are all out of tissue, all in tissue, or mixed (ie edges):
newColData <- do.call(rbind, lapply(l1, function(x) {
	m1 = as.matrix(x[,c("array_row","array_col")])
	rownames(m1) = x$key
	nbors = dnearneigh(m1, d1=1, d2=2.1, bounds=c("GE","LE")) #d2 of 2.1 identifies closest 6 spots
	names(nbors) = x$key
	index= 1:nrow(x)
	off.tissue.index = index[x$in_tissue==FALSE]
	off.nbors = sapply(nbors, function(y) length(intersect(y, off.tissue.index)))
	x$keep = "keep"
	x[off.nbors>0,"keep"] = "tissue perimeter"
	x[off.tissue.index,"keep"] = "off tissue"

	east = max(x$array_row)
	west = min(x$array_row)
	north = max(x$array_col)
	south = min(x$array_col)
	cardinal.check = x$array_row %in% c(east, west) | x$array_col %in% c(north, south)
	x[x$keep!="off tissue" & cardinal.check,"keep"] = "image perimeter"
	return(x)
}))

### preview resulting colData
head(newColData)

###
newColData <- newColData[colnames(lc),]
colData(lc) <- newColData
```

```{r}
### plot check
lc$keep <- as.character(lc$keep)
pal <- c("pink","green","black","red")
names(pal) <- unique(lc$keep)

plts <- spatialLIBD::vis_grid_clus(lc,clustervar = "keep",return_plots = T,spatial = T,point_size = 1.5,alpha=0.75,image_id = "lowres",colors=c("pink","green","black","red"))

plts <- lapply(plts,FUN=function(x){x+theme(axis.text.x=element_blank(),axis.text.y=element_blank(),title=element_text(size=5),legend.text = element_text(size=4),legend.title = element_text(size=4))})

do.call("grid.arrange",plts)
dev.off()
```

This works very nicely, even for samples where there's an interior hole in the tissue or etc! Since the samples are tight packed against one another, and many have sort of ghostly-half-thickness H&E at the margins, we will definitely want to use this to get rid of edge spots. They could be cross-contaminated with neighbor spots from adjacent sections and have odd effects due to low gene/UMI counts in downstream analyses. The LC proper is never right on the edges here either, so no harm done removing the edge spots insofar as the main biological objective here.

Save (parts of) the colData with the in/out/margin classifications so we can bind them to the raw SPE prior to applying other QC steps.
```{r}
tmpcd <- as.data.table(newColData,keep.rownames=T)
tmpcd <- tmpcd[,.(rn,keep)]
setnames(tmpcd,2,"spot_type")
fwrite(tmpcd,"processed-data/04_QC/01-tissue margin spots.txt",sep='\t',quote=F)
```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] dplyr_1.1.4                 scater_1.32.0              
 [3] scuttle_1.14.0              spdep_1.3-5                
 [5] sf_1.0-16                   spData_2.3.1               
 [7] BiocParallel_1.38.0         parallelly_1.37.1          
 [9] colorout_1.3-0.2            spatialLIBD_1.16.2         
[11] SpatialExperiment_1.14.0    SingleCellExperiment_1.26.0
[13] SummarizedExperiment_1.34.0 Biobase_2.64.0             
[15] GenomicRanges_1.56.1        MatrixGenerics_1.16.0      
[17] matrixStats_1.3.0           gridExtra_2.3              
[19] ggplot2_3.5.1               Biostrings_2.72.1          
[21] GenomeInfoDb_1.40.1         XVector_0.44.0             
[23] IRanges_2.38.1              S4Vectors_0.42.1           
[25] BiocGenerics_0.50.0         data.table_1.15.4          
[27] rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0             bitops_1.0-7             
  [4] filelock_1.0.3            fields_16.2               tibble_3.2.1             
  [7] XML_3.99-0.17             lifecycle_1.0.4           edgeR_4.2.0              
 [10] doParallel_1.0.17         rprojroot_2.0.4           lattice_0.22-6           
 [13] magrittr_2.0.3            limma_3.60.3              plotly_4.10.4            
 [16] sass_0.4.9                rmarkdown_2.27            jquerylib_0.1.4          
 [19] yaml_2.3.9                httpuv_1.6.15             spam_2.10-0              
 [22] sp_2.1-4                  sessioninfo_1.2.2         cowplot_1.1.3            
 [25] DBI_1.2.3                 RColorBrewer_1.1-3        golem_0.4.1              
 [28] maps_3.4.2                abind_1.4-5               zlibbioc_1.50.0          
 [31] purrr_1.0.2               RCurl_1.98-1.14           rappdirs_0.3.3           
 [34] GenomeInfoDbData_1.2.12   ggrepel_0.9.5             irlba_2.3.5.1            
 [37] units_0.8-5               DelayedMatrixStats_1.26.0 codetools_0.2-20         
 [40] DelayedArray_0.30.1       DT_0.33                   tidyselect_1.2.1         
 [43] farver_2.1.2              UCSC.utils_1.0.0          ScaledMatrix_1.12.0      
 [46] viridis_0.6.5             shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [49] GenomicAlignments_1.40.0  jsonlite_1.8.8            BiocNeighbors_1.22.0     
 [52] e1071_1.7-14              iterators_1.0.14          dbscan_1.2-0             
 [55] foreach_1.5.2             tools_4.4.1               Rcpp_1.0.13              
 [58] glue_1.7.0                SparseArray_1.4.8         xfun_0.45                
 [61] here_1.0.1                withr_3.0.0               BiocManager_1.30.23      
 [64] fastmap_1.2.0             boot_1.3-30               fansi_1.0.6              
 [67] digest_0.6.36             rsvd_1.0.5                R6_2.5.1                 
 [70] mime_0.12                 wk_0.9.2                  colorspace_2.1-0         
 [73] RSQLite_2.3.7             config_0.3.2              utf8_1.2.4               
 [76] tidyr_1.3.1               generics_0.1.3            class_7.3-22             
 [79] rtracklayer_1.64.0        httr_1.4.7                htmlwidgets_1.6.4        
 [82] S4Arrays_1.4.1            pkgconfig_2.0.3           gtable_0.3.5             
 [85] blob_1.2.4                htmltools_0.5.8.1         dotCall64_1.1-1          
 [88] scales_1.3.0              png_0.1-8                 attempt_0.3.1            
 [91] knitr_1.48                rstudioapi_0.16.0         rjson_0.2.21             
 [94] curl_5.2.1                proxy_0.4-27              cachem_1.1.0             
 [97] KernSmooth_2.23-24        BiocVersion_3.19.1        parallel_4.4.1           
[100] vipor_0.4.7               AnnotationDbi_1.66.0      s2_1.1.6                 
[103] restfulr_0.0.15           pillar_1.9.0              grid_4.4.1               
[106] vctrs_0.6.5               promises_1.3.0            BiocSingular_1.20.0      
[109] dbplyr_2.5.0              beachmat_2.20.0           xtable_1.8-4             
[112] beeswarm_0.4.0            paletteer_1.6.0           evaluate_0.24.0          
[115] magick_2.8.3              cli_3.6.3                 locfit_1.5-9.10          
[118] compiler_4.4.1            Rsamtools_2.20.0          crayon_1.5.3             
[121] labeling_0.4.3            classInt_0.4-10           rematch2_2.1.2           
[124] ggbeeswarm_0.7.2          deldir_2.0-4              viridisLite_0.4.2        
[127] munsell_0.5.1             lazyeval_0.2.2            Matrix_1.7-0             
[130] ExperimentHub_2.12.0      benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[133] bit64_4.0.5               KEGGREST_1.44.1           statmod_1.5.0            
[136] shiny_1.8.1.1             AnnotationHub_3.12.0      memoise_2.0.1            
[139] bslib_0.7.0               benchmarkmeData_1.0.4     bit_4.0.5                
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-22
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular           1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings           * 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                   1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                  0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   class                  7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt               0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   dbscan                 1.2-0     2024-06-28 [1] CRAN (R 4.4.1)
   DelayedArray           0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                 2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                 0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64              1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                * 1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                  1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth             2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                  1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.37.1    2024-02-29 [1] CRAN (R 4.4.0)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises               1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                  0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13    2024-07-17 [1] CRAN (R 4.4.1)
   RCurl                  1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                     1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays               1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle              * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                   * 1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                     2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                   2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData               * 2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                * 1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                  0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                     0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                   0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector              * 0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
