---
title: "06-spotNMmetrics"
author: "Bernie Mulvey"
date: "2025-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## we appended the NM metrics in code/06b, so load the standalone colData that now has these for retained spots.
```{r}
cdtxt <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")

## get the 4 NM metric colnames
nmcols <- grep("NM",colnames(cdtxt),value=TRUE)
stopifnot(length(nmcols)!=4)

## get a histogram of each metric for each sample--exclude 0s

pltsets <- lapply(nmcols,FUN=function(n){
    ## set sample name to be rendered to indicate number of nonzero spots for metric
    tmp <- cdtxt[get(n)>0,.N,by="sample_id"]
    pltdat <- merge.data.table(copy(cdtxt),tmp,by="sample_id",all.x=T)
    pltdat[,sampid_title:=paste0(sample_id," ",N," nonzero")]
    
    p <- ggplot(pltdat[get(n)>0],aes(x=.data[[n]])) + 
        geom_histogram(bins=30) + 
        facet_wrap(~sampid_title,scales="free_y",ncol = 3)+
        ggtitle(n)
    return(p)
})

pdf("plots/06-NM_seg_hists_by_samp.pdf",height=length(unique(cdtxt$sample_id)),width=10,onefile=T)
i<-1
for(i in c(1:length(pltsets))){
    ## trying to get this to render as ~3x3in plots
    print(pltsets[[i]])
}
dev.off()

```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.2 (2024-10-31)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-17
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocParallel         * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 P  cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.2)
    colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    commonmark             1.9.2   2024-10-04 [1] CRAN (R 4.4.1)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1  2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
    ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    jsonlite               1.8.9   2024-09-20 [1] CRAN (R 4.4.1)
    knitr                  1.49    2024-11-08 [1] CRAN (R 4.4.1)
    labeling               0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.2)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magick                 2.8.5   2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    markdown               1.13    2024-06-04 [1] CRAN (R 4.4.0)
    Matrix                 1.7-1   2024-10-18 [1] CRAN (R 4.4.2)
    MatrixGenerics       * 1.18.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
    parallelly           * 1.42.0  2025-01-30 [1] CRAN (R 4.4.1)
    pillar                 1.10.1  2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4   2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
    sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    stringi                1.8.4   2024-05-06 [1] CRAN (R 4.4.0)
    stringr                1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    withr                  3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.50    2025-01-07 [1] CRAN (R 4.4.1)
    xml2                   1.3.6   2023-12-04 [1] CRAN (R 4.4.0)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
