---
title: "02-DEanalysis_minspots"
author: "Bernie Mulvey"
date: "2025-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgels <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/01a-DGElist_per_25HDG75SVGlouv1_cluster_normFilt_10spotmin.RDS")

### to avoid any issues, replace "Br6119(re-dis)" with "Br6119" in the samples tables
dgels <- lapply(dgels,FUN=function(d){
    d$samples$brnum[d$samples$brnum=="Br6119(re-dis)"] <- "Br6119"
    return(d)
})
```

## append global genetic ancestry proportion estimates (FLARE, 3-component CEU ancestry--the other two being YRI and CHB). this source file is intentionally not included in in github but is on JHPCE. The CHB proportions range 0.0001 to 0.01--so CEU is effectively the proportion that is not YRI, and we can just use CEU as the variable here.

```{r}
flares <- fread("raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")

dgels <- lapply(dgels,FUN=function(d){
    tmpsamp <- as.data.table(d$samples,keep.rownames=T)
    tmpsamp <- merge.data.table(tmpsamp,flares,by.x="brnum",by.y="ID")
    tmpsamp <- as.data.frame(tmpsamp,row.names=tmpsamp$rn)[rownames(d$samples),]
    tmpsamp$rn <- NULL
    d$samples <- tmpsamp
    return(d)
})
```

### DE with ancestry, without RIN
```{r}
setDTthreads(1,restore_after_fork = FALSE)
clus <- parallel::makeCluster(9,type = "PSOCK")
fulldgelres.norin <- parallel::clusterApply(x=dgels,cl = clus,fun=function(X){
    
    library(data.table)
    library(limma)
    library(edgeR)
    
    # prevent resource competition, though we aren't explicitly using DT at all here:
    setDTthreads(1,restore_after_fork=FALSE)
    dge <- X

    des <- model.matrix(~0+APOE+Sex+Age+YRI, data = dge$samples)
    des <- as.data.frame(des)
    
    # make these more readable
    colnames(des) <- gsub(colnames(des),pattern="APOE",replacement="")

    # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
    
    # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
    v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$brnum),adaptive.span = T,sample.weights = T)
    cont <- makeContrasts(apoe=(E4hom+E4het)-(E2hom+E2het),Sex=SexM,levels=des)
    v.swt.fit <- contrasts.fit(v.swt,contrasts=cont)
    v.swt.fit.e <- eBayes(v.swt.fit)
    v.swt.e.tt.apo <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
    
    v.swt.e.tt.sex <- topTable(v.swt.fit.e,coef = 2,number=Inf,adjust.method = "BH")
    
    
    retlist <- list(apo=v.swt.e.tt.apo,sex=v.swt.e.tt.sex)
    return(retlist)
})
parallel::stopCluster(clus)
```


```{r}
# make sure the output list has the corresponding cell cluster names
names(fulldgelres.norin) <- names(dgels)

# make data.tables
fulldgelres.norin <- lapply(fulldgelres.norin,function(x){
    lapply(x,as.data.table)
})

saveRDS(fulldgelres.norin,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02a-perCluster_DEtoptabs_10spotmin_noYRItest.RDS")
```

```{r}
rm(list=ls())
gc(full=T)
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats    
[3] graphics  grDevices
[5] utils     methods  
[7] base     

other attached packages:
[1] BiocParallel_1.40.0
[2] parallelly_1.43.0  
[3] gridExtra_2.3      
[4] ggplot2_3.5.2      
[5] colorout_1.3-0.2   
[6] edgeR_4.4.2        
[7] limma_3.62.2       
[8] data.table_1.17.0  
[9] rlang_1.1.5        

loaded via a namespace (and not attached):
 [1] vctrs_0.6.5       
 [2] cli_3.6.4         
 [3] knitr_1.50        
 [4] xfun_0.52         
 [5] pkgload_1.4.0     
 [6] generics_0.1.3    
 [7] glue_1.8.0        
 [8] statmod_1.5.0     
 [9] rprojroot_2.0.4   
[10] locfit_1.5-9.12   
[11] scales_1.4.0      
[12] grid_4.4.3        
[13] evaluate_1.0.3    
[14] tibble_3.2.1      
[15] lifecycle_1.0.4   
[16] compiler_4.4.3    
[17] codetools_0.2-20  
[18] dplyr_1.1.4       
[19] RColorBrewer_1.1-3
[20] pkgconfig_2.0.3   
[21] here_1.0.1        
[22] farver_2.1.2      
[23] lattice_0.22-6    
[24] R6_2.6.1          
[25] tidyselect_1.2.1  
[26] dichromat_2.0-0.1 
[27] pillar_1.10.2     
[28] magrittr_2.0.3    
[29] withr_3.0.2       
[30] tools_4.4.3       
[31] gtable_0.3.6      
> sessioninfo::session_info()
─ Session info ────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-05-07
 pandoc   NA
 quarto   1.6.40 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────
 !  package      * version  date (UTC) lib source
    BiocParallel * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli            3.6.4    2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools      0.2-20   2024-03-31 [1] CRAN (R 4.4.3)
    colorout     * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    data.table   * 1.17.0   2025-02-22 [1] CRAN (R 4.4.1)
    dichromat      2.0-0.1  2022-05-02 [1] CRAN (R 4.4.0)
    dplyr          1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
    edgeR        * 4.4.2    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate       1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
    farver         2.1.2    2024-05-13 [1] CRAN (R 4.4.0)
    generics       0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
    ggplot2      * 3.5.2    2025-04-09 [1] CRAN (R 4.4.1)
    glue           1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra    * 2.3      2017-09-09 [1] CRAN (R 4.4.0)
    gtable         0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
    here           1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
    knitr          1.50     2025-03-16 [1] CRAN (R 4.4.2)
    lattice        0.22-6   2024-03-20 [1] CRAN (R 4.4.3)
    lifecycle      1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
    limma        * 3.62.2   2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.4.1)
    magrittr       2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
    parallelly   * 1.43.0   2025-03-24 [1] CRAN (R 4.4.1)
    pillar         1.10.2   2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig      2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
    pkgload        1.4.0    2024-06-28 [1] CRAN (R 4.4.1)
    R6             2.6.1    2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer   1.1-3    2022-04-03 [1] CRAN (R 4.4.0)
 VP rlang        * 1.1.5    2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rprojroot      2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
    scales         1.4.0    2025-04-24 [1] CRAN (R 4.4.1)
    sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.4.1)
    statmod        1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
    tibble         3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect     1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
    vctrs          0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
    withr          3.0.2    2024-10-28 [1] CRAN (R 4.4.1)
    xfun           0.52     2025-04-02 [1] CRAN (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

───────────────────────────
>
