---
title: "02a-OneInteractionPerRun_libsizeadj"
author: "Bernie Mulvey"
date: "2025-05-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgel <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/01a-DGElist_allClust_10spotmin_NOfilt.RDS")

### to avoid any issues, replace "Br6119(re-dis)" with "Br6119" in the samples tables
dgel$samples$brnum[dgel$samples$brnum=="Br6119(re-dis)"] <- "Br6119"
```

### run low expression filtering separately for each model.

### model 1: sex*label + Age + APOE + YRI(continuous) + lib.size ###
### specify contrasts: Sex per cluster

```{r}
## make a copy of the dgelist for this analysis
dgelsex <- dgel
dgelsex$samples$label_sex <- paste0(dgelsex$samples$label,"_",dgel$samples$Sex)

## full design matrix with colnames matching clus geno group values
des <- model.matrix(~0 + label_sex + APOE + Age + YRI + lib.size, data = dgelsex$samples)

## fix names for readability
colnames(des) <- gsub(colnames(des),pattern="label_sex",replacement="")

## now filter expression. we DO want to get normalized lib sizes first, based on readme for the arguments for filterByExpr.DGEList (it will look for normalized lib sizes)
dgelsex <- edgeR::normLibSizes(dgelsex)

keep <- filterByExpr.DGEList(dgelsex,design = des)
dgelsex <- dgelsex[keep,,keep.lib.sizes=F]

## recalc lib sizes
dgelsex <- edgeR::normLibSizes(dgelsex)

## make sex contrasts for each cluster
i<-1
for (i in c(1:length(unique(dgelsex$samples$label)))){
    x <- unique(dgelsex$samples$label)[i]
    ret <- makeContrasts(
        str2expression(
            paste0(x,"_Sex=",paste0(x,"_M - ",x,"_F"))),
        levels=des)

    colnames(ret) <- paste0(x,"_Sex")

    if(i==1){
        trastmat <- ret
    } else{
        trastmat <- cbind(trastmat,ret)
    }
}
rm(i,x,ret)
```

### model 1: sex_label
run DE ###
```{r}
v.swt <- voomLmFit(dgelsex,design = des,block = as.factor(dgelsex$samples$brnum),adaptive.span = T,sample.weights = T)

## fit M-F contrasts for each label
v.swt.fit <- contrasts.fit(v.swt,contrasts=trastmat)
v.swt.fit.e <- eBayes(v.swt.fit,robust = T)

keeplist_sex_libsize <- list(voomLFt=v.swt,
                 trastsFit=v.swt.fit,
                 eBay=v.swt.fit.e,
                 dgelsex=dgelsex,
                 trasts=trastmat)

sexlabel_detabs <- lapply(colnames(v.swt.fit),FUN=function(t){
        restab <- as.data.table(topTable(v.swt.fit.e,coef = t,number=Inf,adjust.method = "BH"))
    return(restab)
})
names(sexlabel_detabs) <- colnames(trastmat)

rm(v.swt,v.swt.fit,v.swt.fit.e,dgelsex,des,trastmat)
gc(full=T)

keeplist_sex_libsize$sexlabel_detabs <- sexlabel_detabs
saveRDS(keeplist_sex_libsize,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_sexlabel.RDS")
```
############################

### model 2 E4 vs E2: APOE*label + Age + Sex + YRI + lib.size ###
### specify contrasts: E2 or E4 per cluster

```{r}
## make a copy of the dgelist for this analysis
dgelE2E4 <- dgel
dgelE2E4$samples$APOlargeGrp <- ifelse(dgelE2E4$samples$APOE %in% c("E2hom","E2het"),yes="E2",no="E4")
stopifnot(length(unique(dgelE2E4$samples$APOlargeGrp))==2)

dgelE2E4$samples$label_E2E4 <- paste0(dgelE2E4$samples$label,"_",dgelE2E4$samples$APOlargeGrp)

## full design matrix with colnames matching clus geno group values
des <- model.matrix(~0 + label_E2E4 + Sex + Age + YRI + lib.size, data = dgelE2E4$samples)

## fix names for readability
colnames(des) <- gsub(colnames(des),pattern="label_E2E4",replacement="")

## now filter expression. we DO want to get normalized lib sizes first, based on readme for the arguments for filterByExpr.DGEList (it will look for normalized lib sizes)
dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)

keep <- filterByExpr.DGEList(dgelE2E4,design = des)
dgelE2E4 <- dgelE2E4[keep,,keep.lib.sizes=F]

## recalc lib sizes
dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)

## make E4-E2 contrasts for each cluster
i<-1
for (i in c(1:length(unique(dgelE2E4$samples$label)))){
    x <- unique(dgelE2E4$samples$label)[i]
    ret <- makeContrasts(
        str2expression(
            paste0(x,"_E4vE2=",paste0(x,"_E4 - ",x,"_E2"))),
        levels=des)

    colnames(ret) <- paste0(x,"_E4vE2")

    if(i==1){
        trastmat <- ret
    } else{
        trastmat <- cbind(trastmat,ret)
    }
}
rm(i,x,ret,keep)
```

### model 2: E2orE4_label
run DE ###
```{r}
v.swt <- voomLmFit(dgelE2E4,design = des,block = as.factor(dgelE2E4$samples$brnum),adaptive.span = T,sample.weights = T)

## fit M-F contrasts for each label
v.swt.fit <- contrasts.fit(v.swt,contrasts=trastmat)
v.swt.fit.e <- eBayes(v.swt.fit,robust = T)

keeplist_E2E4_libsize <- list(voomLFt=v.swt,
                 trastsFit=v.swt.fit,
                 eBay=v.swt.fit.e,
                 dgelE2E4=dgelE2E4,
                 trasts=trastmat)

E2E4label_detabs <- lapply(colnames(v.swt.fit),FUN=function(t){
        restab <- as.data.table(topTable(v.swt.fit.e,coef = t,number=Inf,adjust.method = "BH"))
    return(restab)
})
names(E2E4label_detabs) <- colnames(trastmat)

keeplist_E2E4_libsize$E2E4label_detabs <- E2E4label_detabs
saveRDS(keeplist_E2E4_libsize,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_E2E4label.RDS")

rm(v.swt,v.swt.fit,v.swt.fit.e,dgelE2E4,des,trastmat)
gc(full=T)
```
##################################

### model 3/tests 3-6: E2hom vs. E4s, E2het vs. E4s, E4het vs E2s, E4hom vs E2s
## make a copy of the dgelist for this analysis
```{r}
dgelApoAlleles <- dgel
dgelApoAlleles$samples$label_apo <- paste0(dgelApoAlleles$samples$label,"_",dgel$samples$APOE)

## full design matrix with colnames matching clus geno group values
des <- model.matrix(~0 + label_apo + Sex + Age + YRI + lib.size, data = dgelApoAlleles$samples)

## fix names for readability
colnames(des) <- gsub(colnames(des),pattern="label_apo",replacement="")

## now filter expression. we DO want to get normalized lib sizes first, based on readme for the arguments for filterByExpr.DGEList (it will look for normalized lib sizes)
dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)

keep <- filterByExpr.DGEList(dgelApoAlleles,design = des)
dgelApoAlleles <- dgelApoAlleles[keep,,keep.lib.sizes=F]

## recalc lib sizes
dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)

## make sex contrasts for each cluster
i<-1
for (i in c(1:length(unique(dgelApoAlleles$samples$label)))){
    x <- unique(dgelApoAlleles$samples$label)[i]
    ret <- makeContrasts(
        str2expression(
            paste0(x,"_E2hom_vs_anyE4=",paste0(x,"_E2hom - (0.5*",x,"_E4het + 0.5*",x,"_E4hom)"))),
        str2expression(
            paste0(x,"_E2het_vs_anyE4=",paste0(x,"_E2het - (0.5*",x,"_E4het + 0.5*",x,"_E4hom)"))),
        str2expression(
            paste0(x,"_E4hom_vs_anyE2=",paste0(x,"_E4hom - (0.5*",x,"_E2het + 0.5*",x,"_E2hom)"))),
        str2expression(
            paste0(x,"_E4het_vs_anyE2=",paste0(x,"_E4het - (0.5*",x,"_E2het + 0.5*",x,"_E2hom)"))),
        levels=des)

    colnames(ret) <- paste0(x,c("_E2hom_vs_anyE4","_E2het_vs_anyE4","_E4hom_vs_anyE2","_E4het_vs_anyE2"))

    if(i==1){
        trastmat <- ret
    } else{
        trastmat <- cbind(trastmat,ret)
    }
}
rm(i,x,ret)
```

### model 3/tests 3-6: by E2 or E4 copies vs all in other allele group
run DE ###
```{r}
v.swt <- voomLmFit(dgelApoAlleles,design = des,block = as.factor(dgelApoAlleles$samples$brnum),adaptive.span = T,sample.weights = T)

## fit M-F contrasts for each label
v.swt.fit <- contrasts.fit(v.swt,contrasts=trastmat)
v.swt.fit.e <- eBayes(v.swt.fit,robust = T)

keeplist_ApoAlleles_libsize <- list(voomLFt=v.swt,
                 trastsFit=v.swt.fit,
                 eBay=v.swt.fit.e,
                 dgelApoAlleles=dgelApoAlleles,
                 trasts=trastmat)

ApoAllelesLabel_detabs <- lapply(colnames(v.swt.fit),FUN=function(t){
        restab <- as.data.table(topTable(v.swt.fit.e,coef = t,number=Inf,adjust.method = "BH"))
    return(restab)
})
names(ApoAllelesLabel_detabs) <- colnames(trastmat)

rm(v.swt,v.swt.fit,v.swt.fit.e,dgelApoAlleles,des,trastmat)
gc(full=T)

keeplist_ApoAlleles_libsize$ApoAllelesLabel_detabs <- ApoAllelesLabel_detabs
saveRDS(keeplist_ApoAlleles_libsize,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_ApoAllelesLabel.RDS")
```
############################

### 5/25: 
model 4/test 7: label*PREDOMINANT ancestry (categorical)
```{r}
## make a copy of the dgelist for this analysis
dgelYRI <- dgel

## make a predominant ancestry grouping and discard the 4 samples that are in the middle space
tmpsamp <- as.data.table(dgelYRI$samples,keep.rownames=T)
tmpsamp <- tmpsamp[YRI>0.75|YRI<0.25]
tmpsamp[,predomAncest:=ifelse(YRI>0.75,"YRI",
                              ifelse(YRI<0.25,"CEU","yougoofed"))]
stopifnot(nrow(tmpsamp[predomAncest=="yougoofed"])==0)

## make a label_predomancest grouping for the "interaction"
tmpsamp[,label_predomAncest:=paste0(label,"_",predomAncest)]

tmpsamp <- as.data.frame(tmpsamp,row.names=tmpsamp$rn)
tmpsamp$rn <- NULL

## get counts table and sample table back in sync in the dgelist
dgelYRI$counts <- dgelYRI$counts[,rownames(tmpsamp)]
stopifnot(nrow(tmpsamp)==ncol(dgelYRI$counts))
dgelYRI$samples <- tmpsamp

## full design matrix with colnames matching clus geno group values
des <- model.matrix(~0 + label_predomAncest +
                        Sex +
                        APOE +
                        Age + 
                        lib.size, 
                    data = dgelYRI$samples)

## make colnames more interpretable
colnames(des) <- gsub(colnames(des),pattern="label_predomAncest",replacement="")

## now filter expression. we DO want to get normalized lib sizes first, based on readme for the arguments for filterByExpr.DGEList (it will look for normalized lib sizes)
dgelYRI <- edgeR::normLibSizes(dgelYRI)

keep <- filterByExpr.DGEList(dgelYRI,design = des)
dgelYRI <- dgelYRI[keep,,keep.lib.sizes=F]

## recalc lib sizes
dgelYRI <- edgeR::normLibSizes(dgelYRI)

## make contrast matrix
for (lab in unique(dgelYRI$samples$label)){
    ret <- makeContrasts(
        str2expression(paste0(lab,"_YRI=",paste0(lab,"_YRI - ",lab,"_CEU"))),
        levels=des)

    colnames(ret) <- paste0(lab,c("_YRI"))

    if(!exists("trastmat")){
        trastmat <- ret
    } else{
        trastmat <- cbind(trastmat,ret)
    }
}

rm(tmpsamp,lab,ret,keep)
```

### 5/25: 
### model 4/test 7: label_predomAncest (CATEGORICAL, yri-derived)
run DE ###
```{r}
v.swt <- voomLmFit(dgelYRI,design = des,block = as.factor(dgelYRI$samples$brnum),adaptive.span = T,sample.weights = T)

## fit YRI effect for each label
v.swt.fit <- contrasts.fit(v.swt,trastmat)
v.swt.fit.e <- eBayes(v.swt.fit,robust = T)

keeplist_yrilabel_libsize <- list(voomLFt=v.swt,
                 trastsFit=v.swt.fit,
                 eBay=v.swt.fit.e,
                 dgelYRI=dgelYRI,
                 trasts=trastmat)

yrilabel_detabs <- lapply(colnames(v.swt.fit),FUN=function(t){
        restab <- as.data.table(topTable(v.swt.fit.e,coef = t,number=Inf,adjust.method = "BH"))
    return(restab)
})
names(yrilabel_detabs) <- colnames(v.swt.fit)

# rm(v.swt,v.swt.fit,v.swt.fit.e,dgelYRI,des,trastmat)
# gc(full=T)

keeplist_yrilabel_libsize$predomAncestLabel_detabs <- yrilabel_detabs
saveRDS(keeplist_yrilabel_libsize,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_predomAncestlabel.RDS")
```

## clean up, load the lists and concatenate into one giant result obj
```{r}
rm(list=ls())
gc(full=T)
```

## load the tmp RDSes
```{r}
sexlabel <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_sexlabel.RDS")
E2E4label <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_E2E4label.RDS")
ApoAllelesLabel <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_ApoAllelesLabel.RDS")
YRIlabel <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02aTMP_keeplist_predomAncestlabel.RDS")

## combine into one metalist
outlist <- list(SexByLabel=sexlabel,E2E4ByLabel=E2E4label,
                ApoAllelesByLabel=ApoAllelesLabel,predomAncestLabel=YRIlabel)

saveRDS(outlist,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02a-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabPredomAncest.RDS")
```
## can now delete 02aTMP RDSes.

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.40.0 parallelly_1.43.0   colorout_1.3-0.2    edgeR_4.4.2        
[5] limma_3.62.2        data.table_1.17.0   rlang_1.1.5        

loaded via a namespace (and not attached):
 [1] digest_0.6.37     codetools_0.2-20  fastmap_1.2.0     xfun_0.52        
 [5] lattice_0.22-6    knitr_1.50        htmltools_0.5.8.1 rmarkdown_2.29   
 [9] cli_3.6.4         grid_4.4.3        statmod_1.5.0     compiler_4.4.3   
[13] rprojroot_2.0.4   here_1.0.1        rstudioapi_0.17.1 tools_4.4.3      
[17] evaluate_1.0.3    yaml_2.3.10       locfit_1.5-9.12  
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-05-26
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 !  package      * version  date (UTC) lib source
    BiocParallel * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli            3.6.4    2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools      0.2-20   2024-03-31 [1] CRAN (R 4.4.3)
    colorout     * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    data.table   * 1.17.0   2025-02-22 [1] CRAN (R 4.4.1)
    digest         0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
    edgeR        * 4.4.2    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate       1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
    fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
    here           1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
    htmltools      0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
    knitr          1.50     2025-03-16 [1] CRAN (R 4.4.2)
    lattice        0.22-6   2024-03-20 [1] CRAN (R 4.4.3)
    limma        * 3.62.2   2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.4.1)
    parallelly   * 1.43.0   2025-03-24 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.5    2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown      2.29     2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot      2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi     0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
    sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.4.1)
    statmod        1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
    xfun           0.52     2025-04-02 [1] CRAN (R 4.4.1)
    yaml           2.3.10   2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
