---
title: "02b-OneInteractionPerRun_dropLowCountClusts_MODULARIZED"
author: "Bernie Mulvey"
date: "2025-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgel <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/01a-DGElist_allClust_10spotmin_NOfilt.RDS")

### to avoid any issues, replace "Br6119(re-dis)" with "Br6119" in the samples tables
dgel$samples$brnum[dgel$samples$brnum=="Br6119(re-dis)"] <- "Br6119"

### rather than adjust for lib.size, now we will drop the clusters making that necessary because they are of limited interest / mixed cell types: Vascular, Oligo_complex, Oligo, Astro_Oligo
newsamp <- dgel$samples
newsamp <- newsamp[!(newsamp$label %in% c("Vascular","Oligo_Complex","Oligo","Astro_Oligo")),]

newmat <- dgel$counts[,rownames(newsamp)]
newmat <- newmat[rowSums(newmat)>0,]
newg <- dgel$genes[rownames(newmat),]

## remove old dgelist
rm(dgel)

## make new dgelist from the subsetted data above
dgel <- DGEList(counts=newmat,samples=newsamp,genes=newg)
rm(newmat,newsamp,newg)

## edgeR's not going to like the decimals if the label names carried any, so remove them just in case
dgel$samples$label <-  gsub(x=dgel$samples$label,pattern="\\.",replacement="")
rownames(dgel$samples) <- gsub(x=rownames(dgel$samples),pattern="\\.",replacement="")
colnames(dgel$counts) <- gsub(x=colnames(dgel$counts),pattern="\\.",replacement="")
```

### run low expression filtering separately for each model. NO lib size cofactor--we're dropping the clusters that would necessitate it.

define a second dataset with categorical ancestry instead of continuous (comes out to N=26) and run all models/contrasts in both the full dataset (YRI continuous) and subsetted (YRI categorical).
```{r}
dgelYRI <- dgel

## make a label-predom ancestry grouping column
tmpsamp <- as.data.table(dgelYRI$samples,keep.rownames=T)
tmpsamp <- tmpsamp[YRI<0.25|YRI>0.75]
tmpsamp[,newYRI:=ifelse(YRI>0.75,"YRI","CEU")]
tmpsamp[,YRI:=NULL]
setnames(tmpsamp,"newYRI","YRI")

tmpsamp <- data.frame(tmpsamp,row.names=tmpsamp$rn)
tmpsamp$rn <- NULL

tmpcounts <- dgelYRI$counts
tmpcounts <- tmpcounts[,rownames(tmpsamp)]
tmpcounts <- as.data.frame(tmpcounts)

dgelYRI <- DGEList(counts=tmpcounts,samples=tmpsamp,genes=dgelYRI$genes[rownames(tmpcounts),])

rm(tmpsamp,tmpcounts)

dgels <- list(cont_ancest=dgel,cat_ancest=dgelYRI)
rm(dgel,dgelYRI)
```

##### In the next SEVERAL blocks, we will define a function that runs each DE analysis so that we can parallelize across dgelists (categorical or continuous YRI) and analysis parameters (defined functions). #####

##### MODEL 1: SEX DE (Sex_cluster "interaction" edgeR style #######
####################################################################
```{r}
sexfun <- function(dgel,n){
    dgelsex <- dgel
    dgelsex$samples$label_sex <- paste0(dgelsex$samples$label, "_", dgel$samples$Sex)
    
    # design mat
    des <- model.matrix(~0 + label_sex + APOE + Age + YRI, data = dgelsex$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_sex", replacement = "")
    
    
    # filter using design
    dgelsex <- edgeR::normLibSizes(dgelsex)
    keep <- filterByExpr.DGEList(dgelsex, design = des)
    dgelsex <- dgelsex[keep, , keep.lib.sizes = FALSE]
    dgelsex <- edgeR::normLibSizes(dgelsex)
    
    # define contrasts to pass to makeContrasts (we have to do it this way due to environment weirdness with parallelization)    contrasts_list <- list()
    contrasts_list <- list()
    for (current_label in unique(dgelsex$samples$label)) {
        contrast_name <- paste0(current_label, "_Sex")
        contrast_formula <- paste0(current_label, "_M - ", current_label, "_F")
        
        new_contrast <- list(contrast_formula)
        names(new_contrast) <- contrast_name
        contrasts_list <- c(contrasts_list, new_contrast)
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelsex, design = des, block = as.factor(dgelsex$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    sexlabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"))
    })
    names(sexlabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelsex, trasts = trastmat, detabs = sexlabel_detabs)
    
    saveRDS(keeplist, paste0("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT/02bTMP_keeplist_sexlabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#####################################################################
##### MODEL 2: E2/E4 carrier status "interaction" edgeR style #######
#####################################################################
```{r}
e2e4fun <- function(dgel,n){dgelE2E4 <- dgel
    dgelE2E4$samples$APOlargeGrp <- ifelse(dgelE2E4$samples$APOE %in% c("E2hom", "E2het"), yes = "E2", no = "E4")
    stopifnot(length(unique(dgelE2E4$samples$APOlargeGrp)) == 2)
    dgelE2E4$samples$label_E2E4 <- paste0(dgelE2E4$samples$label, "_", dgelE2E4$samples$APOlargeGrp)
    
    des <- model.matrix(~0 + label_E2E4 + Sex + Age + YRI, data = dgelE2E4$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_E2E4", replacement = "")

    # Filter using the design specified
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    keep <- filterByExpr.DGEList(dgelE2E4, design = des)
    dgelE2E4 <- dgelE2E4[keep, , keep.lib.sizes = FALSE]
    dgelE2E4 <- edgeR::normLibSizes(dgelE2E4)
    
    # define contrasts to pass to makeContrasts (we have to do it this way due to environment weirdness with parallelization)
    contrasts_list <- list()
    for (current_label in unique(dgelE2E4$samples$label)) {
        contrast_name <- paste0(current_label, "_E4vE2")
        contrast_formula <- paste0(current_label, "_E4 - ", current_label, "_E2")
        
        new_contrast <- list(contrast_formula)
        names(new_contrast) <- contrast_name
        contrasts_list <- c(contrasts_list, new_contrast)
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    ## fit and fit contrasts
    v.swt <- voomLmFit(dgelE2E4, design = des, block = as.factor(dgelE2E4$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    ## get top tables, save an RDS with the DE results and other parts
    E2E4label_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"))
    })
    names(E2E4label_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelE2E4, trasts = trastmat, detabs = E2E4label_detabs)
                     
    saveRDS(keeplist, paste0("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT/02bTMP_keeplist_E2E4label_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)}
```

##########################################################################
#### MODEL 3: APOE alleles (E2hom/het, 4hom/het) "intxns" edgeR style ####
##########################################################################
```{r}
apogenofun <- function(dgel,n){
    dgelApoAlleles <- dgel
    dgelApoAlleles$samples$label_apo <- paste0(dgelApoAlleles$samples$label, "_", dgelApoAlleles$samples$APOE)
    
    des <- model.matrix(~0 + label_apo + Sex + Age + YRI, data = dgelApoAlleles$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "label_apo", replacement = "")
    
    # filter using the design specified
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)
    keep <- filterByExpr.DGEList(dgelApoAlleles, design = des)
    dgelApoAlleles <- dgelApoAlleles[keep, , keep.lib.sizes = FALSE]
    dgelApoAlleles <- edgeR::normLibSizes(dgelApoAlleles)

    contrasts_list <- list()
    for (current_label in unique(dgelApoAlleles$samples$label)) {
        # Create a named vector of contrast formulas for the current label
        formulas <- c(
            paste0(current_label, "_E2hom - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
            paste0(current_label, "_E2het - (0.5*", current_label, "_E4het + 0.5*", current_label, "_E4hom)"),
            paste0(current_label, "_E4hom - (0.5*", current_label, "_E2het + 0.5*", current_label, "_E2hom)"),
            paste0(current_label, "_E4het - (0.5*", current_label, "_E2het + 0.5*", current_label, "_E2hom)"),
            paste0(current_label, "_E4hom - ", current_label, "_E4het"),
            paste0(current_label, "_E4hom - ((", current_label, "_E4het + ", current_label, "_E2het + ", current_label, "_E2hom)/3)")
        )
        names(formulas) <- paste0(current_label, c("_E2hom_vs_anyE4", "_E2het_vs_anyE4", "_E4hom_vs_anyE2", "_E4het_vs_anyE2", "_E4hom_vs_E4het", "_E4hom_vs_others"))
        contrasts_list <- c(contrasts_list, as.list(formulas))
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit and fit contrasts
    v.swt <- voomLmFit(dgelApoAlleles, design = des, block = as.factor(dgelApoAlleles$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs and other components, save
    ApoAllelesLabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"))
    })
    names(ApoAllelesLabel_detabs) <- colnames(trastmat)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelApoAlleles, trasts = trastmat, detabs = ApoAllelesLabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT/02bTMP_keeplist_ApoAllelesLabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

#############################################################################
#### MODEL 4: label-by-YRI (continuous or categorical) "intxns" edgeR style ####
#############################################################################
```{r}
ancesfun <- function(dgel,n){
    dgelances <- dgel
    
    # make design matrices based on YRI being continuous or categorical
    if (is.numeric(dgelances$samples$YRI)) {
        # Continuous YRI setup
        des <- model.matrix(~label:YRI + Sex + APOE + Age, data = dgelances$samples)
        colnames(des) <- gsub(x = colnames(des), pattern = "^label(.*):YRI$", replacement = "\\1_by_YRI")
        trastmat <- paste0(unique(dgelances$samples$label), "_by_YRI") # These are coefficients, not contrasts
        stopifnot(all(trastmat %in% colnames(des)))
    } else {
        # Categorical YRI setup
        dgelances$samples$labAnces <- paste0(dgelances$samples$label, "_", dgelances$samples$YRI)
        des <- model.matrix(~0 + labAnces + Sex + APOE + Age, data = dgelances$samples)
        
        contrasts_list <- list()
        for (current_label in unique(dgelances$samples$label)) {
            contrast_name <- paste0(current_label, "_YRI")
            contrast_formula <- paste0("labAnces", current_label, "_YRI - labAnces", current_label, "_CEU")
            
            new_contrast <- list(contrast_formula)
            names(new_contrast) <- contrast_name
            contrasts_list <- c(contrasts_list, new_contrast)
        }
        trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))
    }
    
    # Filtering using appropriate design from above
    dgelances <- edgeR::normLibSizes(dgelances)
    keep <- filterByExpr.DGEList(dgelances, design = des)
    dgelances <- dgelances[keep, , keep.lib.sizes = FALSE]
    dgelances <- edgeR::normLibSizes(dgelances)

    # Fit
    v.swt <- voomLmFit(dgelances, design = des, block = as.factor(dgelances$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    
    # Fit contrasts or coefficients depending on YRI type
    if (is.numeric(dgelances$samples$YRI)) {
        v.swt.fit <- contrasts.fit(v.swt, coefficients = trastmat)
    } else {
        v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    }
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)
    
    # Get results and other components, save
    yrilabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"))
    })
    names(yrilabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelances, trasts = trastmat, detabs = yrilabel_detabs)

    saveRDS(keeplist, paste0("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT/02bTMP_keeplist_AncesByLabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```


#############################################################################
####### MODEL 5: label*YRI*E4orE2 (cat-cat-cat) "intxns" edgeR style ########
#############################################################################
```{r}
ancesapofun <- function(dgel,n){
    if (is.numeric(dgel$samples$YRI)) {
        message("Model 5 is not run for continuous YRI, only categorical. Skipping for: ", n)
        return(NULL) # Exit the function
    }
    
    # for the categorical data, set it up and go
    dgelYRIAPO <- dgel
    dgelYRIAPO$samples$APOlargeGrp <- ifelse(dgelYRIAPO$samples$APOE %in% c("E2hom", "E2het"), yes = "E2", no = "E4")
    dgelYRIAPO$samples$lab_APOE_PredomAnces <- paste0(dgelYRIAPO$samples$label, "_", dgelYRIAPO$samples$APOlargeGrp, "_", dgelYRIAPO$samples$YRI)
    
    # design matrix
    des <- model.matrix(~0 + lab_APOE_PredomAnces + Sex + Age, data = dgelYRIAPO$samples)
    colnames(des) <- gsub(x = colnames(des), pattern = "^lab_APOE_PredomAnces(.*)$", replacement = "\\1")
    
    # filter using the design
    dgelYRIAPO <- edgeR::normLibSizes(dgelYRIAPO)
    keep <- filterByExpr.DGEList(dgelYRIAPO, design = des)
    dgelYRIAPO <- dgelYRIAPO[keep, , keep.lib.sizes = FALSE]
    dgelYRIAPO <- edgeR::normLibSizes(dgelYRIAPO)

    # define contrasts
    contrasts_list <- list()
    for (current_label in unique(dgelYRIAPO$samples$label)) {
        formulas <- c(
            paste0(current_label, "_E4_YRI - ", current_label, "_E4_CEU"),
            paste0(current_label, "_E2_YRI - ", current_label, "_E2_CEU"),
            paste0(current_label, "_E4_CEU - ", current_label, "_E2_CEU"),
            paste0(current_label, "_E4_YRI - ", current_label, "_E2_YRI")
        )
        names(formulas) <- paste0(current_label, c("_E4YRI_vs_E4CEU", "_E2YRI_vs_E2CEU", "_E4CEU_vs_E2CEU", "_E4YRI_vs_E2YRI"))
        contrasts_list <- c(contrasts_list, as.list(formulas))
    }
    trastmat <- do.call(makeContrasts, c(contrasts_list, list(levels = des)))

    # fit, fit contrasts
    v.swt <- voomLmFit(dgelYRIAPO, design = des, block = as.factor(dgelYRIAPO$samples$brnum), adaptive.span = TRUE, sample.weights = TRUE)
    v.swt.fit <- contrasts.fit(v.swt, contrasts = trastmat)
    v.swt.fit.e <- eBayes(v.swt.fit, robust = TRUE)

    # get toptabs, other results, and save
    yriapolabel_detabs <- lapply(colnames(v.swt.fit), FUN = function(t) {
        as.data.table(topTable(v.swt.fit.e, coef = t, number = Inf, adjust.method = "BH"))
    })
    names(yriapolabel_detabs) <- colnames(v.swt.fit)

    keeplist <- list(voomLFt = v.swt, trastsFit = v.swt.fit, eBay = v.swt.fit.e, 
                     dgel = dgelYRIAPO, trasts = trastmat, detabs = yriapolabel_detabs)
                     
    saveRDS(keeplist, paste0("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT/02bTMP_keeplist_YRIAPOlabel_noMixClusts_", n, ".RDS"))
    gc(full = TRUE)
}
```

## now run all of these simultaneously
```{r}
funs <- list(sexfun,e2e4fun,apogenofun,ancesfun,ancesapofun)

mapply(dgel=dgels,n=names(dgels),SIMPLIFY = FALSE,FUN=function(dgel,n){
    mclapply(X=funs,mc.cores=5,FUN=function(f){
        f(dgel,n)
    })
})
```

## clean up, load the lists and concatenate into one giant result obj
```{r}
rm(list=ls())
gc(full=T)
```

## load the tmp RDSes
## not included in concatenation because not run (properly, yet): YRI as continuous variable (model 5/test 8)
```{r}
categ_ances <- list.files("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT",pattern="_cat_ancest.RDS",full.names = TRUE)
cont_ances <- list.files("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT",pattern="_cont_ancest.RDS",full.names = TRUE)

## combine into one metalist
categ_ancesRDS <- lapply(categ_ances,readRDS)
names(categ_ancesRDS) <- gsub(categ_ances,pattern="processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT/02bTMP_keeplist_(.*)_noMix.*$",replacement="\\1")

cont_ancesRDS <- lapply(cont_ances,readRDS)
names(cont_ancesRDS) <- gsub(cont_ances,pattern="processed-data/12_DEanalyses_removedsampsAndFinalNMseg/MODULARIZEDTMP_OUT/02bTMP_keeplist_(.*)_noMix.*$",replacement="\\1")


outlist <- list(cont_ances=cont_ancesRDS,cat_ances=categ_ancesRDS)

saveRDS(outlist,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")
```
## can now delete 02bTMP RDSes.

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.40.0 parallelly_1.43.0   colorout_1.3-0.2    edgeR_4.4.2        
[5] limma_3.62.2        data.table_1.17.4   rlang_1.1.5        

loaded via a namespace (and not attached):
 [1] digest_0.6.37     codetools_0.2-20  fastmap_1.2.0     xfun_0.52         lattice_0.22-6   
 [6] knitr_1.50        htmltools_0.5.8.1 rmarkdown_2.29    cli_3.6.4         grid_4.4.3       
[11] statmod_1.5.0     compiler_4.4.3    rprojroot_2.0.4   here_1.0.1        rstudioapi_0.17.1
[16] tools_4.4.3       evaluate_1.0.3    yaml_2.3.10       locfit_1.5-9.12   datasets_4.4.3   
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-07-20
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────────────────────
 !  package      * version  date (UTC) lib source
    BiocParallel * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli            3.6.4    2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools      0.2-20   2024-03-31 [1] CRAN (R 4.4.3)
    colorout     * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    data.table   * 1.17.4   2025-05-26 [1] CRAN (R 4.4.3)
    digest         0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
    edgeR        * 4.4.2    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate       1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
    fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
    here           1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
    htmltools      0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
    knitr          1.50     2025-03-16 [1] CRAN (R 4.4.2)
    lattice        0.22-6   2024-03-20 [1] CRAN (R 4.4.3)
    limma        * 3.62.2   2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.4.1)
    parallelly   * 1.43.0   2025-03-24 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.5    2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown      2.29     2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot      2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi     0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
    sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.4.1)
    statmod        1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
    xfun           0.52     2025-04-02 [1] CRAN (R 4.4.1)
    yaml           2.3.10   2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────────────────
