---
title: "02c-Clusterwise_DEanalysis"
author: "Bernie Mulvey"
date: "2025-05-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgels <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/01b-Clusterwise_DGElist_per_25HDG75SVGlouv1_10spotmin_noFilt.RDS")

### to avoid any issues, replace "Br6119(re-dis)" with "Br6119" in the samples tables. should've been tackled in the DGElist making script but just in case.
dgels <- lapply(dgels,FUN=function(d){
    d$samples$brnum[d$samples$brnum=="Br6119(re-dis)"] <- "Br6119"
    return(d)
})
```

### DE with ancestry (YRI, CONTINUOUS), without RIN
```{r}
setDTthreads(1,restore_after_fork = FALSE)
clus <- parallel::makeCluster(9,type = "PSOCK")
fulldgelres.norin.yricont <- parallel::clusterApply(x=dgels,cl = clus,fun=function(X){
    
    library(data.table)
    library(limma)
    library(edgeR)
    
    # prevent resource competition, though we aren't explicitly using DT at all here:
    setDTthreads(1,restore_after_fork=FALSE)
    dge <- X

    des <- model.matrix(~0+APOE+Sex+Age+YRI, data = dge$samples)
    des <- as.data.frame(des)
    
    # filter low expression genes
    dge <- edgeR::calcNormFactors(dge)
    keep <- edgeR::filterByExpr.DGEList(dge,design=des)
    dge <- dge[keep,,keep.lib.sizes=FALSE]
    dge <- edgeR::calcNormFactors(dge)
    
    # make these more readable
    colnames(des) <- gsub(colnames(des),pattern="APOE",replacement="")

    # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
    v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$brnum),adaptive.span = T,sample.weights = T)
    cont <- makeContrasts(E4vsE2=(E4hom+E4het)-(E2hom+E2het),
                            E4hom_vs_E2=E4hom-(0.5*E2hom+0.5*E2het),
                            E4het_vs_E2=E4het-(0.5*E2hom+0.5*E2het),
                            E2hom_vs_E4=E2hom-(0.5*E4hom+0.5*E4het),
                            E2het_vs_E4=E2het-(0.5*E4hom+0.5*E4het),
                            Sex=SexM,
                            YRI=YRI,
                            levels=des)
    v.swt.fit <- contrasts.fit(v.swt,contrasts=cont)
    v.swt.fit.e <- eBayes(v.swt.fit)
    v.swt.e.tt.e4e2 <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e4hom <- topTable(v.swt.fit.e,coef = 2,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e4het <- topTable(v.swt.fit.e,coef = 3,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e2hom <- topTable(v.swt.fit.e,coef = 4,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e2het <- topTable(v.swt.fit.e,coef = 5,number=Inf,adjust.method = "BH")
    v.swt.e.tt.sex <- topTable(v.swt.fit.e,coef = 6,number=Inf,adjust.method = "BH")
    v.swt.e.tt.yri <- topTable(v.swt.fit.e,coef = 7,number=Inf,adjust.method = "BH")
    
    retlist <- list(E4vE2=v.swt.e.tt.e4e2,
                    E4hom_vs_E2=v.swt.e.tt.e4hom,
                    E4het_vs_E2=v.swt.e.tt.e4het,
                    E2hom_vs_E4=v.swt.e.tt.e2hom,
                    E2het_vs_E4=v.swt.e.tt.e2het,
                    Sex=v.swt.e.tt.sex,
                    YRI=v.swt.e.tt.yri)
    return(retlist)
})
parallel::stopCluster(clus)
```


```{r}
# make sure the output list has the corresponding cell cluster names
names(fulldgelres.norin.yricont) <- names(dgels)

# make data.tables
fulldgelres.norin.yricont <- lapply(fulldgelres.norin.yricont,function(x){
    lapply(x,as.data.table)
})

saveRDS(fulldgelres.norin.yricont,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02c-Clusterwise_DETopTabs_Sex_APO_YRIcontinuous.RDS")
```



######## model set 2 ########  predominant ancestry (drops 4 samples, so re-run other DE tests in this setup as well)

### DE with PREDOMINANT ancestry (categorical, YRI<0.25|>0.75 n=26 donors), without RIN
```{r}
dgels.pa <- lapply(dgels,function(x){
    tmpsamp <- as.data.table(x$samples,keep.rownames=T)
    tmpsamp <- tmpsamp[!is.na(predomAncest)]
    tmpsamp <- as.data.frame(tmpsamp,row.names=tmpsamp$rn)
    
    subdgel <- DGEList(counts=x$counts[,rownames(tmpsamp)],samples=tmpsamp,genes=x$genes[rownames(x$counts),])
    return(subdgel)
})

setDTthreads(1,restore_after_fork = FALSE)
clus <- parallel::makeCluster(9,type = "PSOCK")
fulldgelres.norin.predomancest <- parallel::clusterApply(x=dgels.pa,cl = clus,fun=function(X){
    
    library(data.table)
    library(limma)
    library(edgeR)
    
    # prevent resource competition, though we aren't explicitly using DT at all here:
    setDTthreads(1,restore_after_fork=FALSE)
    dge <- X

    des <- model.matrix(~0+APOE+Sex+Age+predomAncest, data = dge$samples)
    des <- as.data.frame(des)
    
    # filter low expression genes
    dge <- edgeR::calcNormFactors(dge)
    keep <- edgeR::filterByExpr.DGEList(dge,design=des)
    dge <- dge[keep,,keep.lib.sizes=FALSE]
    dge <- edgeR::calcNormFactors(dge)
    
    # make these more readable
    colnames(des) <- gsub(colnames(des),pattern="APOE",replacement="")

    # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
    v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$brnum),adaptive.span = T,sample.weights = T)
    cont <- makeContrasts(E4vsE2=(E4hom+E4het)-(E2hom+E2het),
                            E4hom_vs_E2=E4hom-(0.5*E2hom+0.5*E2het),
                            E4het_vs_E2=E4het-(0.5*E2hom+0.5*E2het),
                            E2hom_vs_E4=E2hom-(0.5*E4hom+0.5*E4het),
                            E2het_vs_E4=E2het-(0.5*E4hom+0.5*E4het),
                            Sex=SexM,
                            predomYRI=predomAncestYRI,
                            levels=des)
    v.swt.fit <- contrasts.fit(v.swt,contrasts=cont)
    v.swt.fit.e <- eBayes(v.swt.fit)
    v.swt.e.tt.e4e2 <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e4hom <- topTable(v.swt.fit.e,coef = 2,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e4het <- topTable(v.swt.fit.e,coef = 3,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e2hom <- topTable(v.swt.fit.e,coef = 4,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e2het <- topTable(v.swt.fit.e,coef = 5,number=Inf,adjust.method = "BH")
    v.swt.e.tt.sex <- topTable(v.swt.fit.e,coef = 6,number=Inf,adjust.method = "BH")
    v.swt.e.tt.pa <- topTable(v.swt.fit.e,coef = 7,number=Inf,adjust.method = "BH")
    
    retlist <- list(E4vE2=v.swt.e.tt.e4e2,
                    E4hom_vs_E2=v.swt.e.tt.e4hom,
                    E4het_vs_E2=v.swt.e.tt.e4het,
                    E2hom_vs_E4=v.swt.e.tt.e2hom,
                    E2het_vs_E4=v.swt.e.tt.e2het,
                    Sex=v.swt.e.tt.sex,
                    predomAncest=v.swt.e.tt.pa)
    return(retlist)
})
parallel::stopCluster(clus)
```


```{r}
# make sure the output list has the corresponding cell cluster names
names(fulldgelres.norin.predomancest) <- names(dgels.pa)

# make data.tables
fulldgelres.norin.predomancest <- lapply(fulldgelres.norin.predomancest,function(x){
    lapply(x,as.data.table)
})

saveRDS(fulldgelres.norin.predomancest,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02c-Clusterwise_DETopTabs_Sex_APO_predomAncest.RDS")
```

```{r}
rm(list=ls())
gc(full=T)
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.40.0 parallelly_1.43.0   colorout_1.3-0.2    edgeR_4.4.2        
[5] limma_3.62.2        data.table_1.17.0   rlang_1.1.5        

loaded via a namespace (and not attached):
 [1] digest_0.6.37     codetools_0.2-20  fastmap_1.2.0     xfun_0.52        
 [5] lattice_0.22-6    knitr_1.50        htmltools_0.5.8.1 rmarkdown_2.29   
 [9] cli_3.6.4         grid_4.4.3        statmod_1.5.0     compiler_4.4.3   
[13] rprojroot_2.0.4   here_1.0.1        rstudioapi_0.17.1 tools_4.4.3      
[17] evaluate_1.0.3    yaml_2.3.10       locfit_1.5-9.12  
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-05-26
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 !  package      * version  date (UTC) lib source
    BiocParallel * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli            3.6.4    2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools      0.2-20   2024-03-31 [1] CRAN (R 4.4.3)
    colorout     * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    data.table   * 1.17.0   2025-02-22 [1] CRAN (R 4.4.1)
    digest         0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
    edgeR        * 4.4.2    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate       1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
    fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
    here           1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
    htmltools      0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
    knitr          1.50     2025-03-16 [1] CRAN (R 4.4.2)
    lattice        0.22-6   2024-03-20 [1] CRAN (R 4.4.3)
    limma        * 3.62.2   2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.4.1)
    parallelly   * 1.43.0   2025-03-24 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.5    2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown      2.29     2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot      2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi     0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
    sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.4.1)
    statmod        1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
    xfun           0.52     2025-04-02 [1] CRAN (R 4.4.1)
    yaml           2.3.10   2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
