---
title: "02g-Clusterwise_DEanalysis"
author: "Bernie Mulvey"
date: "2025-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgels <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/01b-Clusterwise_DGElist_per_25HDG75SVGlouv1_10spotmin_edgeRdefaultLoExpFilt.RDS")

### to avoid any issues, replace "Br6119(re-dis)" with "Br6119" in the samples tables
dgels <- lapply(dgels,FUN=function(d){
    d$samples$brnum[d$samples$brnum=="Br6119(re-dis)"] <- "Br6119"
    return(d)
})
```


### DE with ancestry, without RIN
```{r}
setDTthreads(1,restore_after_fork = FALSE)
clus <- parallel::makeCluster(9,type = "PSOCK")
fulldgelres.norin <- parallel::clusterApply(x=dgels,cl = clus,fun=function(X){
    
    library(data.table)
    library(limma)
    library(edgeR)
    
    # prevent resource competition, though we aren't explicitly using DT at all here:
    setDTthreads(1,restore_after_fork=FALSE)
    dge <- X

    des <- model.matrix(~0+APOE+Sex+Age+YRI, data = dge$samples)
    des <- as.data.frame(des)
    
    # make these more readable
    colnames(des) <- gsub(colnames(des),pattern="APOE",replacement="")

    # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
    
    # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
    v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$brnum),adaptive.span = T,sample.weights = T)
    cont <- makeContrasts(E4vsE2=(E4hom+E4het)-(E2hom+E2het),
                            E4hom_vs_E2=E4hom-(0.5*E2hom+0.5*E2het),
                            E4het_vs_E2=E4het-(0.5*E2hom+0.5*E2het),
                            E2hom_vs_E4=E2hom-(0.5*E4hom+0.5*E4het),
                            E2het_vs_E4=E2het-(0.5*E4hom+0.5*E4het),                            
                            Sex=SexM,
                            YRI=YRI,
                            levels=des)
    v.swt.fit <- contrasts.fit(v.swt,contrasts=cont)
    v.swt.fit.e <- eBayes(v.swt.fit)
    v.swt.e.tt.e4e2 <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e4hom <- topTable(v.swt.fit.e,coef = 2,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e4het <- topTable(v.swt.fit.e,coef = 3,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e2hom <- topTable(v.swt.fit.e,coef = 4,number=Inf,adjust.method = "BH")
    v.swt.e.tt.e2het <- topTable(v.swt.fit.e,coef = 5,number=Inf,adjust.method = "BH")
    v.swt.e.tt.sex <- topTable(v.swt.fit.e,coef = 6,number=Inf,adjust.method = "BH")
    v.swt.e.tt.yri <- topTable(v.swt.fit.e,coef = 7,number=Inf,adjust.method = "BH")
    
    retlist <- list(E4vE2=v.swt.e.tt.e4e2,
                    E4hom_vs_E2=v.swt.e.tt.e4hom,
                    E4het_vs_E2=v.swt.e.tt.e4het,
                    E2hom_vs_E4=v.swt.e.tt.e2hom,
                    E2het_vs_E4=v.swt.e.tt.e2het,
                    Sex=v.swt.e.tt.sex,
                    YRI=v.swt.e.tt.yri)
    return(retlist)
})
parallel::stopCluster(clus)
```


```{r}
# make sure the output list has the corresponding cell cluster names
names(fulldgelres.norin) <- names(dgels)

# make data.tables
fulldgelres.norin <- lapply(fulldgelres.norin,function(x){
    lapply(x,as.data.table)
})

saveRDS(fulldgelres.norin,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02g-Clusterwise_DETopTabs_Sex_APO_YRI.RDS")
```

```{r}
rm(list=ls())
gc(full=T)
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    parallel  stats    
[4] graphics  grDevices utils    
[7] methods   base     

other attached packages:
 [1] gridExtra_2.3              
 [2] dreamlet_1.4.1             
 [3] variancePartition_1.36.3   
 [4] ggplot2_3.5.2              
 [5] SpatialExperiment_1.16.0   
 [6] SingleCellExperiment_1.28.1
 [7] SummarizedExperiment_1.36.0
 [8] Biobase_2.66.0             
 [9] GenomicRanges_1.58.0       
[10] GenomeInfoDb_1.42.3        
[11] IRanges_2.40.1             
[12] S4Vectors_0.44.0           
[13] BiocGenerics_0.52.0        
[14] MatrixGenerics_1.18.1      
[15] matrixStats_1.5.0          
[16] BiocParallel_1.40.0        
[17] parallelly_1.43.0          
[18] colorout_1.3-0.2           
[19] edgeR_4.4.2                
[20] limma_3.62.2               
[21] data.table_1.17.0          
[22] rlang_1.1.5                

loaded via a namespace (and not attached):
  [1] splines_4.4.3            
  [2] bitops_1.0-9             
  [3] tibble_3.2.1             
  [4] graph_1.84.1             
  [5] XML_3.99-0.18            
  [6] lifecycle_1.0.4          
  [7] Rdpack_2.6.3             
  [8] mixsqp_0.3-54            
  [9] rprojroot_2.0.4          
 [10] lattice_0.22-6           
 [11] MASS_7.3-65              
 [12] backports_1.5.0          
 [13] magrittr_2.0.3           
 [14] metafor_4.8-0            
 [15] DBI_1.2.3                
 [16] minqa_1.2.8              
 [17] RColorBrewer_1.1-3       
 [18] abind_1.4-8              
 [19] zlibbioc_1.52.0          
 [20] EnvStats_3.0.0           
 [21] purrr_1.0.4              
 [22] rmeta_3.0                
 [23] msigdbr_10.0.1           
 [24] RCurl_1.98-1.17          
 [25] GenomeInfoDbData_1.2.13  
 [26] ggrepel_0.9.6            
 [27] pbkrtest_0.5.3           
 [28] irlba_2.3.5.1            
 [29] annotate_1.84.0          
 [30] DelayedMatrixStats_1.28.1
 [31] codetools_0.2-20         
 [32] DelayedArray_0.32.0      
 [33] tidyselect_1.2.1         
 [34] UCSC.utils_1.2.0         
 [35] farver_2.1.2             
 [36] lme4_1.1-36              
 [37] mathjaxr_1.8-0           
 [38] jsonlite_2.0.0           
 [39] iterators_1.0.14         
 [40] tools_4.4.3              
 [41] progress_1.2.3           
 [42] Rcpp_1.0.14              
 [43] zenith_1.8.0             
 [44] glue_1.8.0               
 [45] SparseArray_1.6.2        
 [46] xfun_0.52                
 [47] here_1.0.1               
 [48] dplyr_1.1.4              
 [49] withr_3.0.2              
 [50] numDeriv_2016.8-1.1      
 [51] fastmap_1.2.0            
 [52] boot_1.3-31              
 [53] caTools_1.18.3           
 [54] digest_0.6.37            
 [55] truncnorm_1.0-9          
 [56] R6_2.6.1                 
 [57] colorspace_2.1-1         
 [58] scattermore_1.2          
 [59] gtools_3.9.5             
 [60] dichromat_2.0-0.1        
 [61] RSQLite_2.3.9            
 [62] RhpcBLASctl_0.23-42      
 [63] tidyr_1.3.1              
 [64] generics_0.1.3           
 [65] corpcor_1.6.10           
 [66] prettyunits_1.2.0        
 [67] httr_1.4.7               
 [68] S4Arrays_1.6.0           
 [69] pkgconfig_2.0.3          
 [70] gtable_0.3.6             
 [71] blob_1.2.4               
 [72] XVector_0.46.0           
 [73] remaCor_0.0.18           
 [74] zigg_0.0.2               
 [75] GSEABase_1.68.0          
 [76] scales_1.4.0             
 [77] png_0.1-8                
 [78] fANCOVA_0.6-1            
 [79] reformulas_0.4.0         
 [80] ashr_2.2-63              
 [81] knitr_1.50               
 [82] reshape2_1.4.4           
 [83] rjson_0.2.23             
 [84] nlme_3.1-167             
 [85] nloptr_2.2.1             
 [86] cachem_1.1.0             
 [87] stringr_1.5.1            
 [88] KernSmooth_2.23-26       
 [89] vipor_0.4.7              
 [90] metadat_1.4-0            
 [91] AnnotationDbi_1.68.0     
 [92] pillar_1.10.2            
 [93] grid_4.4.3               
 [94] vctrs_0.6.5              
 [95] gplots_3.2.0             
 [96] mashr_0.2.79             
 [97] xtable_1.8-4             
 [98] beeswarm_0.4.0           
 [99] Rgraphviz_2.50.0         
[100] evaluate_1.0.3           
[101] KEGGgraph_1.66.0         
[102] invgamma_1.1             
[103] magick_2.8.6             
[104] mvtnorm_1.3-3            
[105] cli_3.6.4                
[106] locfit_1.5-9.12          
[107] compiler_4.4.3           
[108] crayon_1.5.3             
[109] SQUAREM_2021.1           
[110] plyr_1.8.9               
[111] ggbeeswarm_0.7.2         
[112] stringi_1.8.7            
[113] assertthat_0.2.1         
[114] babelgene_22.9           
[115] lmerTest_3.1-3           
[116] Biostrings_2.74.1        
[117] aod_1.3.3                
[118] Matrix_1.7-3             
[119] hms_1.1.3                
[120] sparseMatrixStats_1.18.0 
[121] bit64_4.6.0-1            
[122] KEGGREST_1.46.0          
[123] statmod_1.5.0            
[124] rbibutils_2.3            
[125] Rfast_2.1.5.1            
[126] broom_1.0.7              
[127] memoise_2.0.1            
[128] RcppParallel_5.1.10      
[129] bit_4.6.0                
[130] EnrichmentBrowser_2.36.0 
> sessioninfo::session_info()
─ Session info ────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-05-13
 pandoc   NA
 quarto   1.6.40 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────
 !  package              * version    date (UTC) lib source
    abind                  1.4-8      2024-09-12 [1] CRAN (R 4.4.1)
    annotate               1.84.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    AnnotationDbi          1.68.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    aod                    1.3.3      2023-12-13 [1] CRAN (R 4.4.0)
    ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.4.0)
    assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.4.0)
    babelgene              22.9       2022-09-29 [1] CRAN (R 4.4.0)
    backports              1.5.0      2024-05-23 [1] CRAN (R 4.4.0)
    beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocParallel         * 1.40.0     2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    Biostrings             2.74.1     2024-12-16 [1] Bioconductor 3.20 (R 4.4.2)
    bit                    4.6.0      2025-03-06 [1] CRAN (R 4.4.1)
    bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.4.1)
    bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.4.1)
    blob                   1.2.4      2023-03-17 [1] CRAN (R 4.4.0)
    boot                   1.3-31     2024-08-28 [1] CRAN (R 4.4.3)
    broom                  1.0.7      2024-09-26 [1] CRAN (R 4.4.1)
    cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.4.0)
    caTools                1.18.3     2024-09-04 [1] CRAN (R 4.4.1)
 VP cli                    3.6.4      2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20     2024-03-31 [1] CRAN (R 4.4.3)
    colorout             * 1.3-0.2    2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1      2024-07-26 [1] CRAN (R 4.4.0)
    corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.4.0)
    crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.17.0     2025-02-22 [1] CRAN (R 4.4.1)
    DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.4.0)
    DelayedArray           0.32.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    DelayedMatrixStats     1.28.1     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37     2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4      2023-11-17 [1] CRAN (R 4.4.0)
    dreamlet             * 1.4.1      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    edgeR                * 4.4.2      2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    EnrichmentBrowser      2.36.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    EnvStats               3.0.0      2024-08-24 [1] CRAN (R 4.4.1)
    evaluate               1.0.3      2025-01-10 [1] CRAN (R 4.4.1)
    fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.4.0)
    farver                 2.1.2      2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.3      2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.3     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13     2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2      2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.2      2025-04-09 [1] CRAN (R 4.4.1)
    ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0      2024-09-30 [1] CRAN (R 4.4.1)
    gplots                 3.2.0      2024-10-05 [1] CRAN (R 4.4.1)
    graph                  1.84.1     2025-01-02 [1] Bioconductor 3.20 (R 4.4.2)
    gridExtra            * 2.3        2017-09-09 [1] CRAN (R 4.4.0)
    GSEABase               1.68.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.4.1)
    gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.4.0)
    here                   1.0.1      2020-12-13 [1] CRAN (R 4.4.0)
    hms                    1.1.3      2023-03-21 [1] CRAN (R 4.4.0)
    httr                   1.4.7      2023-08-15 [1] CRAN (R 4.4.0)
    invgamma               1.1        2017-05-07 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1     2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1    2022-10-03 [1] CRAN (R 4.4.0)
    iterators              1.0.14     2022-02-05 [1] CRAN (R 4.4.0)
    jsonlite               2.0.0      2025-03-27 [1] CRAN (R 4.4.1)
    KEGGgraph              1.66.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    KEGGREST               1.46.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.4.3)
    knitr                  1.50       2025-03-16 [1] CRAN (R 4.4.2)
    lattice                0.22-6     2024-03-20 [1] CRAN (R 4.4.3)
    lifecycle              1.0.4      2023-11-07 [1] CRAN (R 4.4.0)
    limma                * 3.62.2     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    lme4                   1.1-36     2025-01-11 [1] CRAN (R 4.4.1)
    lmerTest               3.1-3      2020-10-23 [1] CRAN (R 4.4.0)
    locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.4.1)
    magick                 2.8.6      2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3      2022-03-30 [1] CRAN (R 4.4.0)
    mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.4.0)
    MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.4.1)
    mathjaxr               1.8-0      2025-04-30 [1] CRAN (R 4.4.1)
    Matrix                 1.7-3      2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.4.1)
    memoise                2.0.1      2021-11-26 [1] CRAN (R 4.4.0)
    metadat                1.4-0      2025-02-04 [1] CRAN (R 4.4.1)
    metafor                4.8-0      2025-01-28 [1] CRAN (R 4.4.1)
    minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.4.0)
    mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.4.0)
    msigdbr                10.0.1     2025-03-19 [1] CRAN (R 4.4.3)
    mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.4.1)
    nlme                   3.1-167    2025-01-27 [1] CRAN (R 4.4.3)
    nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.4.1)
    numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.4.0)
    parallelly           * 1.43.0     2025-03-24 [1] CRAN (R 4.4.1)
    pbkrtest               0.5.3      2024-06-26 [1] CRAN (R 4.4.0)
    pillar                 1.10.2     2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.4.0)
    plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.4.0)
    png                    0.1-8      2022-11-29 [1] CRAN (R 4.4.0)
    prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.4.0)
    progress               1.2.3      2023-12-06 [1] CRAN (R 4.4.0)
    purrr                  1.0.4      2025-02-05 [1] CRAN (R 4.4.1)
    R6                     2.6.1      2025-02-15 [1] CRAN (R 4.4.1)
    rbibutils              2.3        2024-10-04 [1] CRAN (R 4.4.1)
    RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14     2025-01-12 [1] CRAN (R 4.4.1)
    RcppParallel           5.1.10     2025-01-24 [1] CRAN (R 4.4.1)
    RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.4.1)
    Rdpack                 2.6.3      2025-03-16 [1] CRAN (R 4.4.2)
    reformulas             0.4.0      2024-11-03 [1] CRAN (R 4.4.1)
    remaCor                0.0.18     2024-02-08 [1] CRAN (R 4.4.0)
    reshape2               1.4.4      2020-04-09 [1] CRAN (R 4.4.0)
    Rfast                  2.1.5.1    2025-03-14 [1] CRAN (R 4.4.1)
    Rgraphviz              2.50.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.4.0)
    rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5      2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmeta                  3.0        2018-03-20 [1] CRAN (R 4.4.0)
    rprojroot              2.0.4      2023-11-05 [1] CRAN (R 4.4.0)
    RSQLite                2.3.9      2024-12-03 [1] CRAN (R 4.4.1)
    S4Arrays               1.6.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0      2025-04-24 [1] CRAN (R 4.4.1)
    scattermore            1.2        2023-06-12 [1] CRAN (R 4.4.0)
    sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1     2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.2      2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    sparseMatrixStats      1.18.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.4.0)
    statmod                1.5.0      2023-01-06 [1] CRAN (R 4.4.0)
    stringi                1.8.7      2025-03-27 [1] CRAN (R 4.4.1)
    stringr                1.5.1      2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1      2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1      2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.4.0)
    truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    variancePartition    * 1.36.3     2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    vctrs                  0.6.5      2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.4.0)
    withr                  3.0.2      2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52       2025-04-02 [1] CRAN (R 4.4.1)
    XML                    3.99-0.18  2025-01-01 [1] CRAN (R 4.4.1)
    xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.4.0)
    XVector                0.46.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    zenith                 1.8.0      2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.4.1)
    zlibbioc               1.52.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

───────────────────────────────────
