---
title: "02-DEanalysis_minspots"
author: "Bernie Mulvey"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
dgel <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/01a-DGElist_per_25HDG75SVGlouv1_clusterSexAPO_normFilt_10spotmin.RDS")

### to avoid any issues, replace "Br6119(re-dis)" with "Br6119" in the samples tables
dgel$samples$brnum[dgel$samples$brnum=="Br6119(re-dis)"] <- "Br6119"
```

## append global genetic ancestry proportion estimates (FLARE, 3-component CEU ancestry--the other two being YRI and CHB). this source file is intentionally not included in in github but is on JHPCE. The CHB proportions range 0.0001 to 0.01--so CEU is effectively the proportion that is not YRI, and we can just use CEU as the variable here.

```{r}
flares <- fread("raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")

tmpsamp <- as.data.table(dgel$samples,keep.rownames=T)
tmpsamp <- merge.data.table(tmpsamp,flares,by.x="brnum",by.y="ID")
tmpsamp <- as.data.frame(tmpsamp,row.names=tmpsamp$rn)[rownames(dgel$samples),]
tmpsamp$rn <- NULL
dgel$samples <- tmpsamp
rm(flares)
```

### specify contrasts: Sex, E4 vs E2, E4 copy number vs E2, E2 copy number vs E4, per cluster
```{r}
## it's not going to like the LC.1, so change it to LC
dgel$samples$clus_geno_group <- gsub(dgel$samples$clus_geno_group,pattern="\\.",replacement="")
dgel$samples$label <-  gsub(dgel$samples$label,pattern="\\.",replacement="")

## full design matrix with colnames matching clus geno group values
des <- model.matrix(~0+clus_geno_group+Age+YRI, data = dgel$samples)
colnames(des) <- gsub(colnames(des),pattern="clus_geno_group",replacement="")

### apply default edgeR low count filtering
dgel2 <- filterByExpr.DGEList(dgel,design = des)
dgel2 <- dgel[dgel2,]
rm(dgel)

## make sex and e4s vs e2s contrasts for each cluster
i<-1
for (i in c(1:length(unique(dgel2$samples$label)))){
    x <- unique(dgel2$samples$label)[i]
# trasts <- lapply(unique(dgel$samples$label),function(x){
    ## get groups for sex DE
    
    malepat <- paste0(x,"_M_")
    malegrps <- unique(grep(dgel2$samples$clus_geno_group,pattern=malepat,value=T))
    fempat <- paste0(x,"_F_")
    femgrps <- unique(grep(dgel2$samples$clus_geno_group,pattern=fempat,value=T))
    
    if(length(femgrps)!=length(malegrps)){
        if(length(femgrps)<length(malegrps)){
            mult <- length(malegrps)/length(femgrps)
        } else {
            mult <- length(femgrps)/length(malegrps)
        }
    } else {mult <- 1}
    
    ### get groups for all E4 carriers vs all E2 carriers
    
    e4pat <- paste0("E4")
    e4grps <- unique(grep(dgel2$samples$clus_geno_group,pattern=e4pat,value=T))
    e4grps <- grep(e4grps,pattern=x,value=T)
    
    e2pat <- paste0("E2")
    e2grps <- unique(grep(dgel2$samples$clus_geno_group,pattern=e2pat,value=T))
    e2grps <- grep(e2grps,pattern=x,value=T)
    
    if(length(e4grps)!=length(e2grps)){
    if(length(e4grps)<length(e2grps)){
        e4e2mult <- length(e2grps)/length(e4grps)
    } else {
        e4e2mult <- length(e4grps)/length(e2grps)
    }
    } else {e4e2mult <- 1}
    
    ret <- makeContrasts(
        str2expression(
            paste0(x,"_Sex=",paste0("mult*(",paste0(malegrps,collapse="+"),")"),
               paste0("-(",paste0(femgrps,collapse="+"),")"))),
        
        str2expression(
            paste0(x,"_E4vE2=",paste0("e4e2mult*(",paste0(e4grps,collapse="+"),")"),
               paste0("-(",paste0(e2grps,collapse="+"),")"))),
    
        levels=des)
    
    colnames(ret) <- paste0(x,c("_Sex","_E4vsE2"))
    
    if(i==1){
        trastmat <- ret
    } else{ 
        trastmat <- cbind(trastmat,ret)
    }
}
rm(e2grps,e4grps,e2pat,e4pat,e4e2mult,femgrps,fempat,i,malegrps,malepat,mult,x,ret)
```

run DE ### LEFT OFF HERE probably needs to be jhpce'd because its going to take a while
```{r}
v.swt <- voomLmFit(dgel2,design = des,block = as.factor(dgel2$samples$brnum),adaptive.span = T,sample.weights = T)
v.swt.fit <- contrasts.fit(v.swt,contrasts=trastmat)
v.swt.fit.e <- eBayes(v.swt.fit)

keeplist <- list(voomLFt=v.swt,trastsFit=v.swt.fit,eBay=v.swt.fit.e,dgel=dgel2,trasts=trastmat)

saveRDS(keeplist,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02TESTa-allclust_sexAPOde_results.RDS")
rm(keeplist)

detabs <- lapply(c(1:length(colnames(trastmat))),FUN=function(t){
    restab <- as.data.table(topTable(v.swt.fit.e,coef = t,number=Inf,adjust.method = "BH"))
    return(restab)
})
names(detabs) <- colnames(trastmat)

saveRDS(detabs,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02TESTa-allclust_sexAPO_deTabs.RDS")
```

```{r}
rm(list=ls())
gc(full=T)
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.40.0 parallelly_1.43.0   colorout_1.3-0.2    edgeR_4.4.2        
[5] limma_3.62.2        data.table_1.17.0   rlang_1.1.5        

loaded via a namespace (and not attached):
 [1] digest_0.6.37     codetools_0.2-20  fastmap_1.2.0     xfun_0.51        
 [5] lattice_0.22-6    knitr_1.50        htmltools_0.5.8.1 rmarkdown_2.29   
 [9] cli_3.6.4         grid_4.4.3        statmod_1.5.0     compiler_4.4.3   
[13] rprojroot_2.0.4   here_1.0.1        rstudioapi_0.17.1 tools_4.4.3      
[17] evaluate_1.0.3    yaml_2.3.10       locfit_1.5-9.12  
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-04-28
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────
 ! package      * version  date (UTC) lib source
   BiocParallel * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 P cli            3.6.4    2025-02-13 [2] CRAN (R 4.4.1)
   codetools      0.2-20   2024-03-31 [1] CRAN (R 4.4.3)
   colorout     * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   data.table   * 1.17.0   2025-02-22 [1] CRAN (R 4.4.1)
   digest         0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
   edgeR        * 4.4.2    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
   evaluate       1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
   fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
   here           1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.50     2025-03-16 [1] CRAN (R 4.4.2)
   lattice        0.22-6   2024-03-20 [1] CRAN (R 4.4.3)
   limma        * 3.62.2   2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   locfit         1.5-9.12 2025-03-05 [1] CRAN (R 4.4.1)
   parallelly   * 1.43.0   2025-03-24 [1] CRAN (R 4.4.1)
 P rlang        * 1.1.5    2025-01-17 [2] CRAN (R 4.4.1)
   rmarkdown      2.29     2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot      2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi     0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
   sessioninfo    1.2.3    2025-02-05 [1] CRAN (R 4.4.1)
   statmod        1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
   xfun           0.51     2025-02-19 [1] CRAN (R 4.4.1)
   yaml           2.3.10   2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────────
