---
title: "03bii-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEAmulveyTFtargcollection"
author: "Bernie Mulvey"
date: "2025-07-20"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(clusterProfiler) # otherwise fgsea parallelization doesnt seem to work
library(fgsea)
library(SpatialExperiment)

require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallel)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallelly)
options(bphost="localhost")
library(BiocParallel)
```

### use TF-target sets compiled from enrichr sets on 12/18/23 (same RDS file as used for mulvey hypothalamus sex diffs). preformatted for fgsea.
```{r}
sets <- readRDS("raw-data/ref_data/00a-enrichrTFsets_forfgseainput_121823.RDS")
```


# PART 1: get signed DE t stat vectors per cluster per test
```{r}
de2b <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")

## we want continuous ancestry analyses, except for the ancestry-strat analysis
de2b2 <- list(sexlabel=de2b$cont_ances$sexlabel$detabs,E2E4label=de2b$cont_ances$E2E4label$detabs,ApoAllelesLabel=de2b$cont_ances$ApoAllelesLabel$detabs)
de2b2[[length(de2b2)+1]] <- de2b$cat_ances$AncesByLabel$detabs
names(de2b2)[length(de2b2)] <- "AncesByLabel"

de2b3 <- list()
i<-1
for (i in c(1:length(de2b2))){
    j<-1
    for (j in c(1:length(de2b2[[i]]))){
        de2b3[length(de2b3)+1] <- de2b2[[i]][j]
        names(de2b3)[length(de2b3)] <- names(de2b2[[i]])[j]
    }
    rm(j)
}
rm(i)

### drop sex chr genes from sex DE tests
de2b3[grep(names(de2b3),pattern="Sex",value=T)] <- lapply(de2b3[grep(names(de2b3),pattern="Sex",value=T)],FUN=function(x){
    x[!(chr %in% grep(chr,pattern="chrX|chrY",value=T))]})

### drop mito genes from all tests
de2b3 <- lapply(de2b3,FUN=function(x){
    x[!(chr %in% grep(chr,pattern="chrMT",value=T))]})

### get named, SIGNED vectors of DE t-stats from each test
### need to use gene symbols for compatibility with the TF-target sets
deres <- list()
i<-1
for (i in c(1:length(de2b3))){
    setnames(de2b3[[i]],"t","tstat")
    tmp <- copy(de2b3[[i]])[,.(gene_name,tstat)]
    
    dropg <- tmp[,.N,by="gene_name"][N>1,gene_name]
    tmp <- tmp[!(gene_name %in% dropg)]
    
    setorderv(tmp,"tstat",-1)
    ## get the stat column as a vector and give it the gene ids as names for fgsea
    ## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
    tmp <- as.data.frame(tmp)
    v <- tmp$tstat
    names(v) <- tmp$gene_name
    deres[[length(deres)+1]] <- v
    names(deres)[length(deres)] <- names(de2b3)[i]
    rm(v,tmp)
}
rm(i,de2b,de2b2,dropg)
gc(full=T)
```

### run fgsea-signed tests
### looping through each cluster-variable interaction (each run is very fast with 10 cpus in biocparallel, so just serially run 10cpu GSEA for each cluster)

```{r}
tfgseares <- mapply(X=deres,Y=names(deres),SIMPLIFY=FALSE,FUN=function(X,Y){
   register(MulticoreParam())
   
   # get cond1-up enriched results
   z1<-fgseaMultilevel(pathways=sets,stats = X,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")
   # coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
   z1$leadingEdge <- as.character(z1$leadingEdge)
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')
   
   # drop newlines from leadingEdge
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   # somehow, one run of this sometimes doesnt get them all so repeat for good meas
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   
   z1 <- as.data.table(z1)
   z1[,test:=Y]
   
   if(length(grep(Y,pattern="Sex",value=T))>0){
       z1[,dir:="male_up"]
   } else if(length(grep(Y,pattern="E4vE2",value=T))>0){
       z1[,dir:="E4carriers_up"]
   } else if(length(grep(Y,pattern="YRI",value=T))>0){
       z1[,dir:="YRI_up"]
   } else {
       z1[,dir:=gsub(Y,pattern="^.*_(E....)_vs_(.*)",replacement=paste0("\\1","_up_vs_","\\2"))]
   }
   
   # get cond2-up enriched results
   register(MulticoreParam())
   z2<-fgseaMultilevel(pathways=sets,stats = X,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "neg")
   # coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
   z2$leadingEdge <- as.character(z2$leadingEdge)
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern='"',replacement='')
   
   # drop newlines from leadingEdge
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")
   # somehow, one run of this sometimes doesnt get them all so repeat for good meas
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")
   
   z2 <- as.data.table(z2)
   z2[,test:=Y]
   
    if(length(grep(Y,pattern="Sex",value=T))>0){
       z2[,dir:="female_up"]
   } else if(length(grep(Y,pattern="E4vE2",value=T))>0){
       z2[,dir:="E4carriers_down"]
   } else if(length(grep(Y,pattern="YRI",value=T))>0){
       z2[,dir:="YRI_down"]
   } else {
       z2[,dir:=gsub(Y,pattern="^.*_(E....)_vs_(.*)",replacement=paste0("\\1","_down_vs_","\\2"))]
   }
   
   z<- as.data.table(rbind(z1,z2))
   return(z)
})
names(tfgseares) <- names(deres)

saveRDS(tfgseares,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/03bii-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEA_TFtargs.RDS")

rm(deres,tfgseares)
```


### PART 2: get UNsigned DE t stat vectors per cluster per test
```{r}
### get named, UNsigned vectors of DE t-stats from each test
deres <- list()
i<-1
for (i in c(1:length(de2b3))){
    tmp <- copy(de2b3[[i]])[,.(gene_name,abs(tstat))]

    dropg <- tmp[,.N,by="gene_name"][N>1,gene_name]
    tmp <- tmp[!(gene_name %in% dropg)]
    
    setnames(tmp,2,"tstat")
    setorderv(tmp,"tstat",-1)
    ## get the stat column as a vector and give it the gene ids as names for fgsea
    ## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
    tmp <- as.data.frame(tmp)
    v <- tmp$tstat
    names(v) <- tmp$gene_name
    deres[[length(deres)+1]] <- v
    names(deres)[length(deres)] <- names(de2b3)[i]
    rm(v,tmp)
}
rm(i)
```

run UNSIGNED fgsea (genes that tend to be affected, regardless of direction)
```{r}
tfgseares.unsign <- mapply(X=deres,Y=names(deres),SIMPLIFY=FALSE,FUN=function(X,Y){
   register(MulticoreParam())
   
   # get DE-enriched results
   z1<-fgseaMultilevel(pathways=sets,stats = X,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")
   
   # coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
   z1$leadingEdge <- as.character(z1$leadingEdge)
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')
   # drop newlines from leadingEdge
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   # belt and suspenders
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   
   z1 <- as.data.table(z1)
   z1[,test:=Y]
   z1[,dir:="either"]
   
   return(z1)
})

names(tfgseares.unsign) <- names(deres)

saveRDS(tfgseares.unsign,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/03bii-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEA_TFtargs_UNsigned.RDS")
```

In a subsequent script, we will determine TFs that are at least marginally expressed in a cluster and subset each result set for that cluster accordingly.

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] BiocParallel_1.40.0         parallelly_1.43.0           default_1.0.0              
 [4] colorout_1.3-0.2            SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1
 [7] SummarizedExperiment_1.36.0 Biobase_2.66.0              GenomicRanges_1.58.0       
[10] GenomeInfoDb_1.42.3         IRanges_2.40.1              S4Vectors_0.44.0           
[13] BiocGenerics_0.52.0         MatrixGenerics_1.18.1       matrixStats_1.5.0          
[16] fgsea_1.32.4                clusterProfiler_4.14.6      data.table_1.17.4          
[19] rlang_1.1.5                

loaded via a namespace (and not attached):
 [1] DBI_1.2.3               gson_0.1.0              magrittr_2.0.3         
 [4] DOSE_4.0.0              compiler_4.4.3          RSQLite_2.3.9          
 [7] png_0.1-8               vctrs_0.6.5             reshape2_1.4.4         
[10] stringr_1.5.1           pkgconfig_2.0.3         crayon_1.5.3           
[13] fastmap_1.2.0           magick_2.8.6            XVector_0.46.0         
[16] rmarkdown_2.29          enrichplot_1.26.6       UCSC.utils_1.2.0       
[19] purrr_1.0.4             bit_4.6.0               xfun_0.52              
[22] zlibbioc_1.52.0         cachem_1.1.0            aplot_0.2.5            
[25] jsonlite_2.0.0          blob_1.2.4              DelayedArray_0.32.0    
[28] R6_2.6.1                stringi_1.8.7           RColorBrewer_1.1-3     
[31] GOSemSim_2.32.0         Rcpp_1.0.14             knitr_1.50             
[34] ggtangle_0.0.6          R.utils_2.13.0          Matrix_1.7-3           
[37] splines_4.4.3           igraph_2.1.4            tidyselect_1.2.1       
[40] abind_1.4-8             qvalue_2.38.0           rstudioapi_0.17.1      
[43] dichromat_2.0-0.1       yaml_2.3.10             codetools_0.2-20       
[46] lattice_0.22-6          tibble_3.2.1            plyr_1.8.9             
[49] treeio_1.30.0           KEGGREST_1.46.0         evaluate_1.0.3         
[52] gridGraphics_0.5-1      Biostrings_2.74.1       pillar_1.10.2          
[55] ggtree_3.14.0           ggfun_0.1.8             generics_0.1.4         
[58] rprojroot_2.0.4         ggplot2_3.5.2           scales_1.4.0           
[61] tidytree_0.4.6          glue_1.8.0              lazyeval_0.2.2         
[64] tools_4.4.3             fs_1.6.6                fastmatch_1.1-6        
[67] cowplot_1.1.3           grid_4.4.3              tidyr_1.3.1            
[70] ape_5.8-1               AnnotationDbi_1.68.0    colorspace_2.1-1       
[73] nlme_3.1-167            GenomeInfoDbData_1.2.13 patchwork_1.3.0        
[76] cli_3.6.4               S4Arrays_1.6.0          dplyr_1.1.4            
[79] gtable_0.3.6            R.methodsS3_1.8.2       yulab.utils_0.2.0      
[82] digest_0.6.37           SparseArray_1.6.2       ggrepel_0.9.6          
[85] ggplotify_0.1.2         rjson_0.2.23            farver_2.1.2           
[88] memoise_2.0.1           htmltools_0.5.8.1       R.oo_1.27.0            
[91] lifecycle_1.0.4         httr_1.4.7              here_1.0.1             
[94] GO.db_3.20.0            bit64_4.6.0-1          
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-07-21
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    AnnotationDbi          1.68.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ape                    5.8-1   2024-12-16 [1] CRAN (R 4.4.1)
    aplot                  0.2.5   2025-02-27 [1] CRAN (R 4.4.1)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocParallel         * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    Biostrings             2.74.1  2024-12-16 [1] Bioconductor 3.20 (R 4.4.2)
    bit                    4.6.0   2025-03-06 [1] CRAN (R 4.4.1)
    bit64                  4.6.0-1 2025-01-16 [1] CRAN (R 4.4.1)
    blob                   1.2.4   2023-03-17 [1] CRAN (R 4.4.0)
    cachem                 1.1.0   2024-05-16 [1] CRAN (R 4.4.0)
 VP cli                    3.6.4   2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    clusterProfiler      * 4.14.6  2025-02-27 [1] Bioconductor 3.20 (R 4.4.2)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.3)
    colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    cowplot                1.1.3   2024-01-22 [1] CRAN (R 4.4.0)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.17.4  2025-05-26 [1] CRAN (R 4.4.3)
    DBI                    1.2.3   2024-06-02 [1] CRAN (R 4.4.0)
    default              * 1.0.0   2017-08-07 [1] CRAN (R 4.4.0)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    DOSE                   4.0.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    enrichplot             1.26.6  2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    fastmatch              1.1-6   2024-12-23 [1] CRAN (R 4.4.1)
    fgsea                * 1.32.4  2025-03-20 [1] Bioconductor 3.20 (R 4.4.3)
    fs                     1.6.6   2025-04-12 [1] CRAN (R 4.4.1)
    generics               0.1.4   2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3  2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggfun                  0.1.8   2024-12-03 [1] CRAN (R 4.4.1)
    ggplot2                3.5.2   2025-04-09 [1] CRAN (R 4.4.1)
    ggplotify              0.1.2   2023-08-09 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6   2024-09-07 [1] CRAN (R 4.4.1)
    ggtangle               0.0.6   2024-12-18 [1] CRAN (R 4.4.1)
    ggtree                 3.14.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    GO.db                  3.20.0  2024-12-12 [1] Bioconductor
    GOSemSim               2.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    gridGraphics           0.5-1   2020-12-13 [1] CRAN (R 4.4.0)
    gson                   0.1.0   2023-03-07 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    igraph                 2.1.4   2025-01-23 [1] CRAN (R 4.4.1)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    jsonlite               2.0.0   2025-03-27 [1] CRAN (R 4.4.1)
    KEGGREST               1.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    knitr                  1.50    2025-03-16 [1] CRAN (R 4.4.2)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.3)
    lazyeval               0.2.2   2019-03-15 [1] CRAN (R 4.4.0)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magick                 2.8.6   2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    Matrix                 1.7-3   2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1  2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    memoise                2.0.1   2021-11-26 [1] CRAN (R 4.4.0)
    nlme                   3.1-167 2025-01-27 [1] CRAN (R 4.4.3)
    parallelly           * 1.43.0  2025-03-24 [1] CRAN (R 4.4.1)
    patchwork              1.3.0   2024-09-16 [1] CRAN (R 4.4.1)
    pillar                 1.10.2  2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    plyr                   1.8.9   2023-10-02 [1] CRAN (R 4.4.0)
    png                    0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
    purrr                  1.0.4   2025-02-05 [1] CRAN (R 4.4.1)
    qvalue                 2.38.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    R.methodsS3            1.8.2   2022-06-13 [1] CRAN (R 4.4.0)
    R.oo                   1.27.0  2024-11-01 [1] CRAN (R 4.4.1)
    R.utils                2.13.0  2025-02-24 [1] CRAN (R 4.4.1)
    R6                     2.6.1   2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    reshape2               1.4.4   2020-04-09 [1] CRAN (R 4.4.0)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5   2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    RSQLite                2.3.9   2024-12-03 [1] CRAN (R 4.4.1)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0   2025-04-24 [1] CRAN (R 4.4.1)
    sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.2   2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    stringi                1.8.7   2025-03-27 [1] CRAN (R 4.4.1)
    stringr                1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1   2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    tidytree               0.4.6   2023-12-12 [1] CRAN (R 4.4.0)
    treeio                 1.30.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    xfun                   0.52    2025-04-02 [1] CRAN (R 4.4.1)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    yulab.utils            0.2.0   2025-01-29 [1] CRAN (R 4.4.1)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────────────────
