---
title: "11b-LMM_spotlogcounts_toNMmetrics_keepranfx"
author: "Bernie Mulvey"
date: "03-29-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(lme4)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

for plotting residualized values, we want to have the random effects as well -- which the original version of this code didn't include in the returned output. so we'll re-run, now outputting a random effect coef table and a main effect coef table per gene per APOE variable encoding and NM metric.

read in spe, append YRI genomic proportions, subset to NM+ cells in LC cluster
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

## calculate log counts
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

## read in nm subclusterings
lcnm <- fread("processed-data/10_subclustering/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnm <- lcnm[,.(rn,NMseg_subclus)]
lcnm <- DataFrame(lcnm,row.names=lcnm$rn)[colnames(lc3),]

## assign
colLabels(lc3) <- lcnm$NMseg_subclus

# subset to LC NM+, make rownames symbols
lcnm <- lc3[,lc3$label=="LC_NMpos"]

## Drop Br5517 (unknown anatomy/orientation), Br5276 (LC has been depleted by QC), Br5712 (unknown anatomy/orientation). Keep 6297; no LC but anatomy is right in vicinity.
lcnm <- lcnm[,!(lcnm$brnum %in% paste0("Br",c(5517,5276,5712)))]

rownames(lcnm) <- rowData(lcnm)$gene_name
rm(lc3)
gc()

## drop undetected genes
keepg <- rowSums(counts(lcnm))>0 #23744 genes detected at least once here.
lcnm <- lcnm[keepg,]

### get metadata -- only intensity and proportion for NM measures are continuous and appropriate for an LMM. (the N nm bodies and centroid signal count are both counts, and would require GLMMs). we can still inspect top genes from the continuous measures plotting across all four metrics with plots later.
flares <- fread("raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")
flares[ID=="Br6119",ID:="Br6119(re-dis)"]

lcnm.cd <- as.data.table(colData(lcnm),keep.rownames=T)
lcnm.cd <- lcnm.cd[,.(rn,Intensity_NM,Prop_NM,Age,Sex,APOE,brnum,capture_id,sample_id)]
setnames(lcnm.cd,"APOE","APOEgeno")

### append YRI
stopifnot(all(lcnm.cd$brnum%in%flares$ID))
lcnm.cd <- merge.data.table(lcnm.cd,flares[,.(ID,YRI)],by.x="brnum",by.y="ID")

## change APOE column to model-friendly genotype. we'll use three models with different approaches for APOE: E4+/-, individual genotypes, and an ordered factor of E4-,het,hom.
lcnm.cd[,E4carrier:="No"]
lcnm.cd[APOEgeno %in% c("E3/ E4","E4/ E4"),E4carrier:="Yes"]

lcnm.cd[,E4copies:="None"]
lcnm.cd[APOEgeno=="E3/ E4",E4copies:="Het"]
lcnm.cd[APOEgeno=="E4/ E4",E4copies:="Hom"]
lcnm.cd[,E4copies:=factor(E4copies,levels=c("None","Het","Hom"),ordered = TRUE)]

# recode APOE itself for model friendliess
lcnm.cd[APOEgeno=="E2/ E2",APOEgeno:="E2hom"]
lcnm.cd[APOEgeno=="E2/ E3",APOEgeno:="E2het"]
lcnm.cd[APOEgeno=="E3/ E4",APOEgeno:="E4het"]
lcnm.cd[APOEgeno=="E4/ E4",APOEgeno:="E4hom"]

lcnm.lct <- logcounts(lcnm)
lcnm.lct <- as.data.frame(as.matrix(lcnm.lct))
```

## clean up for modeling runs
```{r}
rm(lcnm,keepg,flares)
gc(full=T)
```

iterate over each gene and metric and APOE modeling specification
## note that, unlike modeling gene upon NM (expr ~ NM + covars), these models (NM ~ expr + covars) do NOT return "singular fit" warnings en masse

```{r}
setDTthreads(1,restore_after_fork=FALSE)

nmcor_apoes <- lapply(c("E4carrier","APOEgeno"),function(apo){
    
    nmcors <- lapply(c("Intensity_NM","Prop_NM"),function(nm){
        results <- mclapply(c(1:nrow(lcnm.lct)),mc.cores=8,FUN=function(i){

            setDTthreads(1,restore_after_fork=FALSE)
            xpr <- lcnm.lct[i,]
            xpr <- t(xpr)
            xpr <- as.data.table(xpr,keep.rownames=T)
            xpr <- merge.data.table(xpr,lcnm.cd,by="rn")
            
            curg <- colnames(xpr)[2]
            
            setnames(xpr,2,"val")
            setnames(xpr,nm,"NMvar")
            # fit lmms
            f <- formula(paste0("NMvar ~ val + Age + Sex + YRI +", apo," + (1|brnum) + (1|brnum:capture_id) + (1|brnum:capture_id:sample_id)"))
            f2 <- formula(paste0("NMvar ~ Age + Sex + YRI + ", apo," + (1|brnum) + (1|brnum:capture_id) + (1|brnum:capture_id:sample_id)")) # no term for gene
            
            lmmnm <- lmer(f,data=xpr)
            
            ## to perform LRT test, we shouldn't use REML, so run a second non REML
            lmmnm.compare <- lmer(f,data=xpr,REML = FALSE)
            lmmno <- lmer(f2,data=xpr,REML=FALSE)
            
            ## get LRT
            lrt <- stats::anova(lmmnm.compare,lmmno)
            
            ## get anova for model coeffs in the REML NM model
            an <- as.data.table(Anova(lmmnm),keep.rownames=T)
            
            ## get model coefficients
            fixfx <- as.data.table(fixef(lmmnm),keep.rownames=T)
            setnames(fixfx,c("variable","coef"))
            ## split Sex and APO into a variable column and group column for consistency with the anova table (e.g., SexM -> Sex , M) comes out as Sex here but "SexM" for the fixef. so to make sure we keep the sex coefficients, change SexM to Sex here
            fixfx[variable=="SexM",var_group:="M"]
            fixfx[variable=="SexM",variable:="Sex"]
            # pattern to fix the APOE fixed effects will begin with the variable we're testing
            fixpat <- paste0(apo,"(.*)")
            fixfx[variable %in% grep(variable,pattern=apo,value=T),var_group:=gsub(variable,pattern=fixpat,replacement="\\1")]
            fixfx[variable %in% grep(variable,pattern=apo,value=T),variable:=apo]
            
            ## make cleaner intercept rows
            fixfx[variable=="(Intercept)",variable:="Intercept"]
            ## append model coefficients
            
            an <- merge.data.table(fixfx,an,by.x="variable",by.y="rn")
            an[variable=="val",variable:=curg]
            
            ## append LRT test
            lrt <- as.data.table(lrt,keep.rownames=T)[rn=="lmmnm.compare",.(rn,Chisq,Df,`Pr(>Chisq)`,logLik)]
            ## change 'loglik' to coef; change rn to "LRTvs_no_gene_coef"
            lrt[,rn:="LRTvs_no_gene_coef"]
            setnames(lrt,"rn","variable")
            
            an <- rbind(an,lrt,fill=T)
            
            ## add the NMvar being fitted
            an[,NMvar:=nm]
            
            ## get the ranfx table
            ranfxtab <- as.data.table(ranef(lmmnm))
            ranfxtab[,gene_name:=curg]
            ranfxtab[,NMvar:=nm]
            
            retlist <- list(an,ranfxtab)
            names(retlist) <- c("fixed","ranfx")
            return(retlist)
        })
        names(results) <- rownames(lcnm.lct)
        return(results)
    })
    names(nmcors) <- c("Intensity_NM","Prop_NM")
    return(nmcors)

})
names(nmcor_apoes) <- c("E4carrierYN","PerGeno")


### a handful of these return error messages indicating one of the variables is constant. https://stats.stackexchange.com/questions/366784/how-to-get-around-the-glmer-warning-downdated-vtv-is-not-positive-definite
    
### only retain entries that are tabular results; concat those tabs for fixed and rand fx
nmcor_apoe_fixfxtabs <- lapply(nmcor_apoes,function(x){
    fixlist <- lapply(x,function(y){
        lapply(y, function(z){
            z[["fixed"]]
        })
    })
    fixlist2 <- lapply(fixlist,FUN=function(x){
        ret <- x[unlist(lapply(x,is.data.table))]
        ret <- rbindlist(ret,idcol="gene_name")
        return(ret)
    })
    return(fixlist2)
})

nmcor_apoe_ranfxtabs <- lapply(nmcor_apoes,function(x){
    ranlist <- lapply(x,function(y){
        lapply(y, function(z){
            z[["ranfx"]]
        })
    })
    ranlist2 <- lapply(fixlist,FUN=function(x){
        ret <- x[unlist(lapply(x,is.data.table))]
        ret <- rbindlist(ret,idcol="gene_name")
        return(ret)
    })
    return(ranlist2)
})


## spot check that we have random effects for all genes that we got fixed effect results for
stopifnot(all(unique(nmcor_apoe_fixfxtabs$E4carrierYN$Intensity_NM$gene_name) %in% unique(nmcor_apoe_ranfxtabs$E4carrierYN$Intensity_NM$gene_name)))

### final list should rearrange these back together now
finalout <- list(
    E4carrierYN = list(
        Intensity_NM = list(
            fixed = nmcor_apoe_fixfxtabs$E4carrierYN$Intensity_NM,
            ranfx = nmcor_apoe_ranfxtabs$E4carrierYN$Intensity_NM
        ),
        Prop_NM = list(
            fixed = nmcor_apoe_fixfxtabs$E4carrierYN$Prop_NM,
            ranfx = nmcor_apoe_ranfxtabs$E4carrierYN$Prop_NM
        )
    ),
    PerGeno = list(
        Intensity_NM = list(
            fixed = nmcor_apoe_fixfxtabs$PerGeno$Intensity_NM,
            ranfx = nmcor_apoe_ranfxtabs$PerGeno$Intensity_NM
        ),
        Prop_NM = list(
            fixed = nmcor_apoe_fixfxtabs$PerGeno$Prop_NM,
            ranfx = nmcor_apoe_ranfxtabs$PerGeno$Prop_NM
        )
))

# save it!
saveRDS(finalout,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/11-LCNMspots_NMmetric_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")
```

seshinf
```{r}
sessioninfo::session_info()
sessionInfo()
```
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-29
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [2] CRAN (R 4.5.0)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0      2021-06-01 [2] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 boot                   1.3-32     2025-08-29 [2] CRAN (R 4.5.2)
 broom                * 1.0.12     2026-01-27 [2] CRAN (R 4.5.2)
 car                  * 3.1-5      2026-02-03 [2] CRAN (R 4.5.2)
 carData              * 3.0-6      2026-01-30 [2] CRAN (R 4.5.2)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 Formula                1.2-5      2023-02-24 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [2] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 ggtext               * 0.1.2      2022-09-16 [2] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 gridExtra            * 2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gridtext               0.1.5      2022-09-16 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 labeling               0.4.3      2023-08-29 [2] CRAN (R 4.5.0)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 lme4                 * 1.1-38     2025-12-02 [2] CRAN (R 4.5.2)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 MASS                 * 7.3-65     2025-02-28 [2] CRAN (R 4.5.2)
 Matrix               * 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [2] CRAN (R 4.5.0)
 multcomp             * 1.4-29     2025-10-20 [2] CRAN (R 4.5.0)
 mvtnorm              * 1.3-3      2025-01-10 [2] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [2] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [2] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 parallelly           * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rbibutils              2.4.1      2026-01-21 [2] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 Rdpack                 2.6.5      2026-01-23 [2] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [2] CRAN (R 4.5.2)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 rsvd                   1.0.5      2021-04-16 [2] CRAN (R 4.5.0)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 sandwich               3.1-1      2024-09-15 [2] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scater                 1.38.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 survival             * 3.8-6      2026-01-16 [2] CRAN (R 4.5.2)
 TH.data              * 1.1-5      2025-11-17 [2] CRAN (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [2] CRAN (R 4.5.0)
 viridis                0.6.5      2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite            0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 withr                  3.0.2      2024-10-28 [2] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 xml2                   1.5.2      2026-01-17 [2] CRAN (R 4.5.2)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)
 zoo                    1.8-15     2025-12-15 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────
> sessionInfo()
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] lme4_1.1-38                 Matrix_1.7-4               
 [3] broom_1.0.12                car_3.1-5                  
 [5] carData_3.0-6               multcomp_1.4-29            
 [7] TH.data_1.1-5               MASS_7.3-65                
 [9] survival_3.8-6              mvtnorm_1.3-3              
[11] BiocParallel_1.44.0         parallelly_1.46.1          
[13] colorout_1.3-3              SpatialExperiment_1.20.0   
[15] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[17] Biobase_2.70.0              GenomicRanges_1.62.1       
[19] Seqinfo_1.0.0               IRanges_2.44.0             
[21] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[23] generics_0.1.4              MatrixGenerics_1.22.0      
[25] matrixStats_1.5.0           gridExtra_2.3              
[27] ggtext_0.1.2                ggplot2_4.0.2              
[29] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] Rdpack_2.6.5        sandwich_3.1-1      rlang_1.1.7.9000    magrittr_2.0.4     
 [5] scater_1.38.0       otel_0.2.0          compiler_4.5.2      vctrs_0.7.1        
 [9] pkgconfig_2.0.3     fastmap_1.2.0       backports_1.5.0     magick_2.9.0       
[13] XVector_0.50.0      scuttle_1.20.0      labeling_0.4.3      rmarkdown_2.30     
[17] sessioninfo_1.2.3   nloptr_2.2.1        ggbeeswarm_0.7.3    purrr_1.2.1        
[21] xfun_0.56           beachmat_2.26.0     DelayedArray_0.36.0 irlba_2.3.7        
[25] R6_2.6.1            RColorBrewer_1.1-3  boot_1.3-32         Rcpp_1.1.1         
[29] knitr_1.51          zoo_1.8-15          splines_4.5.2       tidyselect_1.2.1   
[33] rstudioapi_0.18.0   dichromat_2.0-0.1   abind_1.4-8         yaml_2.3.12        
[37] viridis_0.6.5       codetools_0.2-20    lattice_0.22-7      tibble_3.3.1       
[41] withr_3.0.2         S7_0.2.1            evaluate_1.0.5      xml2_1.5.2         
[45] pillar_1.11.1       reformulas_0.4.4    rprojroot_2.1.1     scales_1.4.0       
[49] minqa_1.2.8         glue_1.8.0          tools_4.5.2         BiocNeighbors_2.4.0
[53] ScaledMatrix_1.18.0 grid_4.5.2          tidyr_1.3.2         rbibutils_2.4.1    
[57] nlme_3.1-168        beeswarm_0.4.0      BiocSingular_1.26.0 vipor_0.4.7        
[61] Formula_1.2-5       cli_3.6.5           rsvd_1.0.5          S4Arrays_1.10.1    
[65] viridisLite_0.4.3   dplyr_1.2.0         gtable_0.3.6        digest_0.6.39      
[69] SparseArray_1.10.8  ggrepel_0.9.6       rjson_0.2.23        farver_2.1.2       
[73] htmltools_0.5.9     lifecycle_1.0.5     here_1.0.2          gridtext_0.1.5   
