---
title: "sharedAstroE4downreg_Enrichr"
author: "Bernie Mulvey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(enrichR)
library(biomaRt)
library(stringr)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
deres <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")

ancctrl <- deres$cat_ances$E2E4label$detabs

aa <- deres$cat_ances$YRIAPOlabel$detabs
ea <- aa[grep(names(aa),pattern="E4CEU_vs_E2CEU",value=T)]
aa <- aa[grep(names(aa),pattern="E4YRI_vs_E2YRI",value=T)]


ea.ast <- ea$Astro_E4CEU_vs_E2CEU
aa.ast <- aa$Astro_E4YRI_vs_E2YRI

setnames(ea.ast,c("logFC","adj.P.Val"),c("logFC.EA","FDR.EA"))
setnames(aa.ast,c("logFC","adj.P.Val"),c("logFC.AA","FDR.AA"))

shareg <- merge.data.table(ea.ast[lessOrEqual(FDR.EA,0.05),.(gene_id,gene_name,logFC.EA,FDR.EA)],aa.ast[lessOrEqual(FDR.AA,0.05),.(gene_id,gene_name,logFC.AA,FDR.AA)],by=c("gene_id","gene_name"))
shareg[,signmatch:=ifelse(greaterOrEqual(logFC.EA*logFC.AA,0),1,0)]
shareg <- shareg[signmatch==1]

# same DGE so same filtering was applied upstream for both EA and AA contrasts:
bkg <- ea.ast$gene_id
```

we need to be sure we have entrez symbols for enrichr gene matching

```{r}
biomart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL",dataset="hsapiens_gene_ensembl")
lut <- biomaRt::getBM(attributes=c("ensembl_gene_id",
                                   "entrezgene_id",
                                   "entrezgene_accession",
                                   "entrezgene_description"),
                      mart=biomart,
                      filter="ensembl_gene_id",
                      values=bkg) # uh ok "accession" = symbol

lut <- as.data.table(lut)
dropg <- lut[,.N,by="entrezgene_accession"][N>1,entrezgene_accession]
lut <- lut[!(entrezgene_accession %in% dropg)]
lut <- lut[!is.na(entrezgene_accession)]
## our background list (all genes tested for DE) which was used to make this lookup table should be updated to reflect genes where the symbols are not duplicated. make sure there's only one row per accession-id now:
stopifnot(length(unique(lut$entrezgene_accession))==nrow(lut))
bkg <- lut$entrezgene_accession

## merge shareg with the filterd lut, so only keeping genes with unique symbols after merge
shareg <- merge.data.table(shareg,lut[,.(ensembl_gene_id,entrezgene_accession)],by.x="gene_id",by.y="ensembl_gene_id") # 110 of 113 retained

## clean up
rm(biomart,aa,deres,ea,lut,dropg)
```

```{r}
### enrichr setup
## rummagenes don't work here unfortunately.
checkdbs <- enrichR::listEnrichrDbs()
enrichr_dbs <- grep(checkdbs$libraryName,pattern="Aging|Allen|ARCHS4|Azimuth|BioPlex|CellMarker|ChEA_2022|CORUM|DepMap_CRISPR_GeneDependency_CellLines_2023|Descartes|DGIdb_Drug_Targets_2024|Disease_Perturbations|Disease_Signatures|Drug_Perturbations|DrugMatrix|Elsevier_Pathway|ENCODE_Histone_Modifications_2015|ENCODE_TF_ChIP-seq_2015|Roadmap|FANTOM6|Gene_Perturbations|GO_Biological_Process_2025|GO_Cellular_Component_2025|GO_Molecular_Function_2025|HMS_LINCS|HuBMAP|huMAP|IDG_Drug_Targs_2022|Jensen_DISEASES_Curated_2025|KEGG_2021_Human|Kinase_Perturbations|KOMP2|L1000|Ligand_Perturbations|lncHUB|MAGNET|MGI_Mammalian_Phenotype_Level_4_2024|NIBR_DRUGseq_2025|PerturbAtlas|PFOCR_Pathways_2023|PPI_Hub_Proteins|Proteomics|Reactome_Pathways_2024|SILAC|SubCell|SynGO_2024|Tabula_Muris|Tabula_Sapiens|TF_Perturbations|TF-LOF_Expression|TG_GATES_2020|The_Kinase_Library_2024|Tissue_Protein_Expression|Transcription_Factor_PPIs|TRANSFAC_and_JASPAR|TRRUST|WikiPathways_2024_Human|^DSigDB",value=T)

## filter out the ARCHS4 predictions re: rare disease, NIH PIs--interesting but not useful here
enrichr_dbs <- grep(enrichr_dbs,pattern="AutoRIF|GeneRIF|from_GEO",value=T,invert=T)

rm(checkdbs)
```

# run enrichr for these dbs for downreg shared genes, upreg shared, and all shared genes
```{r}
sharegs <- list(shareup=shareg[greaterOrEqual(logFC.EA,0)&greaterOrEqual(logFC.AA,0),entrezgene_accession],
                sharedown=shareg[lessOrEqual(logFC.EA,0)&lessOrEqual(logFC.AA,0),entrezgene_accession],
                allshare=shareg$entrezgene_accession)
shareg.res <- list()
i <- 3
for (i in c(3:length(sharegs))){
    curhits <- sharegs[[i]]
    cursname <- names(sharegs)[i]
    cur_enrichr_dbs <- enrichr_dbs
    
    ## drop any dbs that throw weird factor-geneset-specific errors
    if(cursname=="sharedown"|cursname=="allshare"){
        cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="CellMarker_2024"]
        cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="PerturbAtlas_MouseGenePerturbationSigs"]
    } else {
      cur_enrichr_dbs <- enrichr_dbs
    }
    
    # run enrichr
    enrichreslist <- enrichr(genes=curhits,
                        background = bkg,
                        databases=cur_enrichr_dbs)
    
    eres.dt <- rbindlist(enrichreslist,idcol="Enrichr_database")
    eres.dt[,predomAncestWiseE4vsE2:=cursname]
    
    shareg.res[[i]] <- eres.dt
    names(shareg.res)[i] <- cursname
    
    rm(curhits,eres.dt,cur_enrichr_dbs)
    gc(full=T)
}

shareg.res <- rbindlist(shareg.res)
# tabulate how many genes were hit in the set
shareg.res[,N_genes:=apply(.SD,MARGIN=1,simplify=FALSE,FUN=function(x){
  max(c(1,length(str_split_fixed(x,";",n=Inf)))) # if no separator, 1 gene matched
}),.SDcols="Genes"]
shareg.res[,N_genes:=unlist(N_genes)] ## fix weirdness with apply within d.t.

# get rid of deprecated columns
shareg.res[,Old.P.value:=NULL]
shareg.res[,Old.Adjusted.P.value:=NULL]
shareg.res[,Rank:=as.integer(Rank)]

### save
saveRDS(shareg.res,"processed-data/12_DEanalyses_removedsampsAndFinalNMseg/14-AncestryShared_Astro_E4E2DEGs_Enrichr.RDS")
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] stringr_1.6.0       biomaRt_2.66.0      BiocParallel_1.44.0 parallelly_1.46.1   colorout_1.3-3     
[6] enrichR_3.4         data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] KEGGREST_1.50.0      gtable_0.3.6         rjson_0.2.23         xfun_0.56           
 [5] ggplot2_4.0.2        httr2_1.2.2          Biobase_2.70.0       lattice_0.22-7      
 [9] vctrs_0.7.1          tools_4.5.2          generics_0.1.4       stats4_4.5.2        
[13] curl_7.0.0           parallel_4.5.2       AnnotationDbi_1.72.0 tibble_3.3.1        
[17] RSQLite_2.4.5        blob_1.3.0           pkgconfig_2.0.3      dbplyr_2.5.1        
[21] RColorBrewer_1.1-3   S4Vectors_0.48.0     S7_0.2.1             lifecycle_1.0.5     
[25] compiler_4.5.2       farver_2.1.2         Biostrings_2.78.0    progress_1.2.3      
[29] statmod_1.5.1        Seqinfo_1.0.0        codetools_0.2-20     htmltools_0.5.9     
[33] yaml_2.3.12          pillar_1.11.1        crayon_1.5.3         rsconnect_1.7.0     
[37] cachem_1.1.0         limma_3.66.0         tidyselect_1.2.1     locfit_1.5-9.12     
[41] digest_0.6.39        stringi_1.8.7        purrr_1.2.1          dplyr_1.2.0         
[45] WriteXLS_6.8.0       rprojroot_2.1.1      fastmap_1.2.0        grid_4.5.2          
[49] here_1.0.2           cli_3.6.5            magrittr_2.0.4       dichromat_2.0-0.1   
[53] withr_3.0.2          edgeR_4.8.2          prettyunits_1.2.0    filelock_1.0.3      
[57] scales_1.4.0         rappdirs_0.3.4       bit64_4.6.0-1        XVector_0.50.0      
[61] rmarkdown_2.30       httr_1.4.7           bit_4.6.0            otel_0.2.0          
[65] png_0.1-8            datasets_4.5.2       hms_1.1.4            memoise_2.0.1       
[69] evaluate_1.0.5       knitr_1.51           IRanges_2.44.0       BiocFileCache_3.0.0 
[73] rlang_1.1.7.9000     glue_1.8.0           DBI_1.2.3            xml2_1.5.2          
[77] BiocGenerics_0.56.0  rstudioapi_0.18.0    R6_2.6.1            
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-04-28
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────────────────────
 package       * version    date (UTC) lib source
 AnnotationDbi   1.72.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Biobase         2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache   3.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics    0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel  * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 biomaRt       * 2.66.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Biostrings      2.78.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 bit             4.6.0      2025-03-06 [2] CRAN (R 4.5.0)
 bit64           4.6.0-1    2025-01-16 [2] CRAN (R 4.5.0)
 blob            1.3.0      2026-01-14 [2] CRAN (R 4.5.2)
 cachem          1.1.0      2024-05-16 [2] CRAN (R 4.5.0)
 cli             3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 codetools       0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout      * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 crayon          1.5.3      2024-06-20 [2] CRAN (R 4.5.0)
 curl            7.0.0      2025-08-19 [2] CRAN (R 4.5.0)
 data.table    * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DBI             1.2.3      2024-06-02 [2] CRAN (R 4.5.0)
 dbplyr          2.5.1      2025-09-10 [2] CRAN (R 4.5.0)
 dichromat       2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest          0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 dplyr           1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 edgeR           4.8.2      2025-12-23 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 enrichR       * 3.4        2025-02-02 [2] CRAN (R 4.5.0)
 evaluate        1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 farver          2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap         1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 filelock        1.0.3      2023-12-11 [2] CRAN (R 4.5.0)
 generics        0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 ggplot2         4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 glue            1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 gtable          0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here            1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 hms             1.1.4      2025-10-17 [2] CRAN (R 4.5.0)
 htmltools       0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 httr            1.4.7      2023-08-15 [2] CRAN (R 4.5.0)
 httr2           1.2.2      2025-12-08 [2] CRAN (R 4.5.2)
 IRanges         2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 KEGGREST        1.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 knitr           1.51       2025-12-20 [2] CRAN (R 4.5.2)
 lattice         0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lifecycle       1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 limma           3.66.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 locfit          1.5-9.12   2025-03-05 [2] CRAN (R 4.5.0)
 magrittr        2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 memoise         2.0.1      2021-11-26 [2] CRAN (R 4.5.0)
 otel            0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 parallelly    * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 pillar          1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig       2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 png             0.1-8      2022-11-29 [2] CRAN (R 4.5.0)
 prettyunits     1.2.0      2023-09-24 [2] CRAN (R 4.5.0)
 progress        1.2.3      2023-12-06 [2] CRAN (R 4.5.0)
 purrr           1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R6              2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rappdirs        0.3.4      2026-01-17 [2] CRAN (R 4.5.2)
 RColorBrewer    1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 rjson           0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang           1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown       2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot       2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 rsconnect       1.7.0      2025-12-06 [2] CRAN (R 4.5.2)
 RSQLite         2.4.5      2025-11-30 [2] CRAN (R 4.5.2)
 rstudioapi      0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 S4Vectors       0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7              0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 scales          1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 Seqinfo         1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo     1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 statmod         1.5.1      2025-10-09 [2] CRAN (R 4.5.0)
 stringi         1.8.7      2025-03-27 [2] CRAN (R 4.5.0)
 stringr       * 1.6.0      2025-11-04 [2] CRAN (R 4.5.0)
 tibble          3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyselect      1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 vctrs           0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 withr           3.0.2      2024-10-28 [2] CRAN (R 4.5.0)
 WriteXLS        6.8.0      2025-05-22 [2] CRAN (R 4.5.0)
 xfun            0.56       2026-01-18 [2] CRAN (R 4.5.2)
 xml2            1.5.2      2026-01-17 [2] CRAN (R 4.5.2)
 XVector         0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml            2.3.12     2025-12-10 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────────────────────
