---
title: "02-iSEE_pseudobulk_obj"
author: "Bernie Mulvey"
date: "2026-07-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)
library(SpatialExperiment)
library(spatialLIBD)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##
```


load unfiltered 9-domain pseudobulk for later (we need it to get ensgs in the meantime)
```{r}
unfilt <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/01a-DGElist_allClust_10spotmin_NOfilt.RDS")
```

## get all genes that passed model-specific expression filtering in any DE analysis
```{r}
deres <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/02b-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_JUL2025.RDS")

## for categorical ancestry, we only analyzed genotype-ancestry and across-ancestry:
deres$cat_ances <- deres$cat_ances[c("AncesByLabel","YRIAPOlabel")]

allens <- unique(unlist(lapply(deres,function(a){
    g <- unlist(lapply(a,(function(b){
        g2 <- unlist(lapply(b$detabs,function(t){
            return(t$gene_id)
        }))
        return(g2)
    })))
    return(unlist(g))
})))

rm(deres)
## also get all genes from marker analyses and NM DE
mks <- readRDS("processed-data/13_markers_and_registration_final/01-HDGSVG2575louv1_regWrapOut.RDS")
mks <- unique(mks$enrichment$ensembl)
allens <- unique(c(allens,mks))
rm(mks)

pairmks <- fread("processed-data/13_markers_and_registration_final/03-HDGSVG2575louv1_regWrapOut_Oligo_1vOthers_markers.txt")
allens <- unique(c(pairmks$ensembl,allens))
rm(pairmks)

nmde <- fread("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/04-LCNMpos_vs_LCNMneg_DE.txt")
allens <- unique(c(nmde$gene_id,allens))
rm(nmde)

## the filtering was a little more lenient in the (while unreported) preliminary runs of identifying markers between glial/achesert clusters--which only analyzed the domains in question, and with N=33--so grab those too
mks <- readRDS("processed-data/08_validitycheck_25hdg75svg_louv1/10-1vsOthers_markers_Oligos_Astros_ACHESERTS.RDS")
allens <- unique(c(allens,mks$oligos$ensembl,mks$astros$ensembl,mks$ache_serts$ensembl))
rm(mks)

## get all genes that were in the inhibitory_4 marker analysis since such genes are potentially of interest in here
inh4 <- fread("processed-data/10_subclustering/12-WebDiv_inhibitory_4_mks_vs_other_inhibitorys.txt")
allens <- unique(c(allens,inh4$ensembl))
# that adds 6k genes to the 8.6k we had up to here
rm(inh4)

# get all genes nominally associated with NM intensity
nmin <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/11-LCNMspots_NMmetric_LMMed_by_geneLogcounts_2apoeFactorizations_withRanEfFits.RDS")
# ugh only symbols in this one..
nmin <- nmin$E4carrierYN$Intensity_NM$fixed[`Pr(>Chisq)`<0.05,gene_name]
allens <- unique(c(allens,as.data.table(unfilt$genes)[gene_name %in% nmin,gene_id]))
rm(nmin)

# get TFs treated as expressed in relevant GSEAs--also symbols only
tfxpra <- fread("plots/manuscript/SuppTabs/TabDelims/ST-TFtargGSEA_TFin2pctOfDomSpotsOnAvgOrIn12Samps_E4vE2.txt")
tfxprb <- fread("plots/manuscript/SuppTabs/TabDelims/ST-TFtargGSEA_TFin2pctOfDomSpotsOnAvgOrIn12Samps_Sex.txt")

tfxpr <- unique(c(tfxpra$TF,tfxprb$TF))
allens <- unique(c(allens,as.data.table(unfilt$genes)[gene_name %in% tfxpr,gene_id]))
rm(tfxpr,tfxpra,tfxprb)
```

perform the minimal/generic filtering used elsewhere with the all-clusters object (we didn't actually use the 9 domain object because some were excluded from DE analyses. but for completeness..);  then calc normalized lib sizes and make a summarizedExperiment obj
```{r}
## subset
allens <- allens[allens %in% rownames(unfilt)]
filt <- unfilt[allens,]
### before we move forward, change "ACHE_SERT_LC_WM" to "ACHE_SERT_NE_WM" to avoid this cluster coming up when searching for "LC" in the pseudobulk browser!
colnames(filt) <- gsub(colnames(filt),pattern="ACHE_SERT_LC_WM",replacement="ACHE_SERT_NE_WM")
rownames(filt$samples) <- colnames(filt)
filt$samples$clus_sample_id <- gsub(filt$samples$clus_sample_id,
                                    pattern="ACHE_SERT_LC_WM",replacement="ACHE_SERT_NE_WM")
filt$samples$label <- gsub(filt$samples$label,
                           pattern="ACHE_SERT_LC_WM",replacement="ACHE_SERT_NE_WM")
filt$samples$clus_geno_group <- gsub(filt$samples$clus_geno_group,
                           pattern="ACHE_SERT_LC_WM",replacement="ACHE_SERT_NE_WM")


### now filter, normalize etc
filt <- edgeR::calcNormFactors(filt,method = "TMM")

subcoldat <- as.data.table(filt$samples,keep.rownames=T)
subcoldat <- subcoldat[,.(rn,sample_id,brnum,APOE,Ancestry,Sex,Age,hemisphere,tissueNMscore,YRI,label)]

subcoldat[,carriertype:=ifelse(APOE %in% c("E4het","E4hom"),"E4 carrier","E2 carrier")]
setnames(subcoldat,"carriertype","APOE Carrier Type")

setnames(subcoldat,"label","Domain")

subcoldat[APOE=="E2hom",APOE:="E2 Homozygous"]
subcoldat[APOE=="E2het",APOE:="E2 Heterozygous"]
subcoldat[APOE=="E4hom",APOE:="E4 Homozygous"]
subcoldat[APOE=="E4het",APOE:="E4 Heterozygous"]
setnames(subcoldat,"APOE","APOE Genotype")

setnames(subcoldat,"YRI", "Est. Pct African Genomic Ancestry")
subcoldat <- as.data.frame(subcoldat,row.names=subcoldat$rn)[colnames(filt$counts),]
subcoldat$rn <- NULL


## switch to symbols as rownames; need to drop dupe symbs
keepg <- as.data.table(filt$genes)[,.(gene_id,.N),by="gene_name"][N==1,gene_id]
filt <- filt[keepg,]
rownames(filt) <- filt$genes$gene_name

subrowdat <- as.data.table(filt$genes,keep.rownames=F)
subrowdat <- subrowdat[,.(gene_name,gene_id,chr)]
subrowdat <- as.data.frame(subrowdat,row.names=subrowdat$gene_name)[rownames(filt$counts),]
subrowdat$rn <- NULL

logcpmmat <- edgeR::cpm(filt,log=T,prior.count=0)

se2 <- SummarizedExperiment(assays=list(counts=filt$counts,logcpm=logcpmmat),rowData = subrowdat,colData = subcoldat)
```

test
```{r}
library(iSEE)
shiny::runApp(iSEE(se2))
```

save
```{r}
saveRDS(se2,"processed-data/14_apps_prep/02-iSEE_pbulkobj.RDS")
```
seshinf
```{r}
sessionInfo()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] shiny_1.12.1                iSEE_2.22.0                 colorout_1.3-3             
 [4] spatialLIBD_1.22.0          SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0              GenomicRanges_1.62.1       
[10] Seqinfo_1.0.0               IRanges_2.44.0              S4Vectors_0.48.1           
[13] BiocGenerics_0.56.0         generics_0.1.4              MatrixGenerics_1.22.0      
[16] matrixStats_1.5.0           edgeR_4.8.2                 limma_3.66.0               
[19] data.table_1.18.4          

loaded via a namespace (and not attached):
  [1] splines_4.5.2            later_1.4.5              BiocIO_1.20.0           
  [4] bitops_1.0-9             filelock_1.0.3           tibble_3.3.1            
  [7] XML_3.99-0.20            lifecycle_1.0.5          httr2_1.2.2             
 [10] doParallel_1.0.17        rprojroot_2.1.1          lattice_0.22-7          
 [13] crosstalk_1.2.2          magrittr_2.0.4           plotly_4.12.0           
 [16] sass_0.4.10              rmarkdown_2.30           jquerylib_0.1.4         
 [19] yaml_2.3.12              httpuv_1.6.16            otel_0.2.0              
 [22] sessioninfo_1.2.3        cowplot_1.2.0            DBI_1.2.3               
 [25] RColorBrewer_1.1-3       golem_0.5.1              abind_1.4-8             
 [28] purrr_1.2.1              RCurl_1.98-1.17          rappdirs_0.3.4          
 [31] circlize_0.4.17          ggrepel_0.9.6            irlba_2.3.7             
 [34] codetools_0.2-20         DelayedArray_0.36.1      DT_0.34.0               
 [37] scuttle_1.20.0           tidyselect_1.2.1         shape_1.4.6.1           
 [40] farver_2.1.2             ScaledMatrix_1.18.0      viridis_0.6.5           
 [43] shinyWidgets_0.9.0       BiocFileCache_3.0.0      GenomicAlignments_1.46.0
 [46] jsonlite_2.0.0           GetoptLong_1.1.0         BiocNeighbors_2.4.0     
 [49] scater_1.38.1            iterators_1.0.14         systemfonts_1.3.1       
 [52] foreach_1.5.2            tools_4.5.2              ragg_1.5.0              
 [55] Rcpp_1.1.1               glue_1.8.0               gridExtra_2.3           
 [58] SparseArray_1.10.8       mgcv_1.9-4               xfun_0.56               
 [61] here_1.0.2               dplyr_1.2.0              withr_3.0.2             
 [64] shinydashboard_0.7.3     BiocManager_1.30.27      fastmap_1.2.0           
 [67] shinyjs_2.1.1            digest_0.6.39            rsvd_1.0.5              
 [70] R6_2.6.1                 mime_0.13                textshaping_1.0.4       
 [73] colorspace_2.1-2         listviewer_4.0.0         Cairo_1.7-0             
 [76] dichromat_2.0-0.1        RSQLite_2.4.5            cigarillo_1.0.0         
 [79] config_0.3.2             tidyr_1.3.2              rtracklayer_1.70.1      
 [82] httr_1.4.7               htmlwidgets_1.6.4        S4Arrays_1.10.1         
 [85] pkgconfig_2.0.3          gtable_0.3.6             rsconnect_1.7.0         
 [88] blob_1.3.0               ComplexHeatmap_2.26.1    S7_0.2.1                
 [91] XVector_0.50.0           htmltools_0.5.9          rintrojs_0.3.4          
 [94] clue_0.3-66              scales_1.4.0             png_0.1-8               
 [97] attempt_0.3.1            knitr_1.51               rstudioapi_0.18.0       
[100] rjson_0.2.23             nlme_3.1-168             curl_7.0.0              
[103] shinyAce_0.4.4           cachem_1.1.0             GlobalOptions_0.1.3     
[106] BiocVersion_3.22.0       parallel_4.5.2           miniUI_0.1.2            
[109] vipor_0.4.7              AnnotationDbi_1.72.0     restfulr_0.0.16         
[112] pillar_1.11.1            grid_4.5.2               vctrs_0.7.1             
[115] promises_1.5.0           BiocSingular_1.26.0      dbplyr_2.5.1            
[118] beachmat_2.26.0          xtable_1.8-4             cluster_2.1.8.1         
[121] beeswarm_0.4.0           paletteer_1.7.0          evaluate_1.0.5          
[124] magick_2.9.0             cli_3.6.5                locfit_1.5-9.12         
[127] compiler_4.5.2           Rsamtools_2.26.0         rlang_1.1.7             
[130] crayon_1.5.3             labeling_0.4.3           rematch2_2.1.2          
[133] ggbeeswarm_0.7.3         viridisLite_0.4.3        BiocParallel_1.44.0     
[136] Biostrings_2.78.0        lazyeval_0.2.2           colourpicker_1.3.0      
[139] Matrix_1.7-4             ExperimentHub_3.0.0      benchmarkme_1.0.8       
[142] bit64_4.6.0-1            ggplot2_4.0.2            KEGGREST_1.50.0         
[145] statmod_1.5.1            AnnotationHub_4.0.0      fontawesome_0.5.3       
[148] igraph_2.2.1             memoise_2.0.1            bslib_0.10.0            
[151] benchmarkmeData_2.0.0    bit_4.6.0   
