---
title: "Fig2-Clusters, Markers, WeberDivecha registration, LC NMpos vs NMneg, LConly NMF-GSEA"
format: html
date: "05-19-2025"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(ggrastr)
library(ComplexHeatmap)
library(Cairo) # pdf device for ^ to properly save; also used for complexheatmap raster functions
library(circlize) #colorRamp2
library(viridis) # magma
library(ggrepel) # volcano point labeling

## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

## For the clustering spotplots, we want to use the H&E images and plot by capture area, so we will need to load the big daddy SPE
```{r}
lc2 <- readRDS("processed-data/04_QC/04-lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed.RDS")

# load clustering and the split-sections coldata, which has a column with pre-split spot ids we can use to match everything back to capture level
louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno,by="clusid")

splitcd <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")
splitcd <- splitcd[,.(rowname,original_spot_id)]

louvs <- merge.data.table(louvs,splitcd,by.x="rn",by.y="rowname") # split section ids

## also attach LCnm subsets because we will use them in this figure
lcnm <- fread("processed-data/10_subclustering/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnm <- lcnm[,.(rn,NMseg_subclus)]

louvs <- merge.data.table(louvs,lcnm,by="rn",all.x=T) 

louvs <- DataFrame(louvs,row.names=louvs$original_spot_id)[colnames(lc2),]
stopifnot(nrow(louvs)==ncol(lc2))

colLabels(lc2) <- louvs$anno
lc2$lcnm_subclus <- louvs$NMseg_subclus

rm(splitcd,lcnm)
gc(full=T)
```

## load cluster 9-color palette
```{r}
pal <- readRDS("plots/cluster_palette_narrowluminance_v2.RDS")
```

### custom function to extrct legend from one plot to extend to all ###
```{r}
extract_legend <- function(my_ggp) {
    step1 <- ggplot_gtable(ggplot_build(my_ggp))
    step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
    step3 <- step1$grobs[[step2]]
    return(step3)
}
```

## Fig 2A: 2x2 grid of plots with the clusters indicated
show the following captures: V13M06-331_A1, V13M06-333_B1, V13M06-404_B1, V13M06-403_D1. But first make a shared legend to extract
```{r}
tmpdat <- as.data.table(colData(lc2[,lc2$sample_id=="V13M06-331_A1"])[,c("label","array_row","array_col")])

tmpplt <- ggplot(tmpdat,aes(x=array_col,y=array_row,col=label))+
    geom_point()+
    scale_color_manual(values = pal)+
    theme(legend.key.height = unit(0.083,"in"),
          legend.key.width = unit(0.156,"in"),
          legend.text=element_text(size=7,margin=margin(0,0,0,-0.025,unit="in")),
          legend.title=element_blank(),
          legend.key=element_blank(),
          legend.background = element_blank())

leg.tmp <- extract_legend(tmpplt)
dev.off()
####

samps <- c("V13M06-331_A1","V13M06-333_B1","V13M06-404_B1","V13M06-403_D1")

plts <- lapply(samps,function(s){
    p <- vis_clus(lc2,sampleid=s,spatial = T,na_color = NA,auto_crop = T,point_size = 0.6,alpha=0.65,clustervar = "label",colors = pal,image_id = "lowres")+
        guides(color="none",fill="none")+
    #ggtitle("NPY")+
        theme(plot.title = element_text(size = 9, hjust = 0.5,margin=margin(0,0,0,0)),
          panel.border = element_blank(),
          plot.margin = margin(-0.1,0.025,-0.475,-0.45,unit="in"))
    p <- rasterize(p,layers="Points",dpi=300,dev="cairo_png")
    return(p)

})
### arrangeGrob() on the spotplot list --> 2x1 of plots from that list only, constituting the "first column" of the parent; leg.tmp2 is an additional thing being plotted in the second column of the parent.
pdf("plots/manuscript/Fig2/2A-Cluster_spotplots.pdf",height=3.5,width=4.25)
do.call("grid.arrange", list(arrangeGrob(grobs = plts, ncol = 2, nrow = 2), leg.tmp, ncol = 2, widths = c(2.8, 1.2)))
dev.off()

rm(plts,tmpdat,tmpplt,leg.tmp,samps)
gc(full=T)
```

Fig 2B: marker heatmap
```{r}
### get markers
globalmk <- fread("processed-data/13_markers_and_registration_final/01-HDGSVG2575louv1_regWrapOut_1vAll_markers.txt")

globalmk <- globalmk[fdr<0.05&logFC>0]
# exclude mtrnrs and mitochondrials
globalmk <- globalmk[!(gene %in% grep(gene,pattern="MTRNR|MT-",value=T))]
## get top 5 genes, by logFC, per cluster
mkg <- unique(globalmk[ , .SD[order(-logFC) , ][1:5] , by = "cluster"][,.(ensembl,cluster)])

## reorder the palette/clusters here so that the sort of gradations between glial clusters are apparent
pal2 <- pal[c(1:5,8,7,9,6)]

## extract these genes in the reverse order of the palette'd clusters so that they appear top corner to bottom corner in resulting heatmap
for (cl in names(pal2)){
    if(cl==names(pal2)[1]){
        mkg2 <- mkg[cluster==cl,ensembl]
    } else {
        mkg2 <- unique(c(mkg2,mkg[cluster==cl,ensembl]))
    }
}
mkg<-mkg2

rm(globalmk,mkg2,cl)

# global marker heatmap: top 5 genes by LFC per cluster x 9 clusters = 45 genes (max)

# get scaled expression data for global heatmap
## define function to z-scale expression
# z scale rows, from tony https://github.com/LieberInstitute/spatial_DG_lifespan/blob/9419eb91453cda1df494d93dc91156d819042c66/code/Pseudobulk/top_infant_DE_heatmap_enrichment.R#L106

scale_rows <- function(x) {
    m <- apply(x, 1, mean, na.rm = T)
    s <- apply(x, 1, sd, na.rm = T)
    return((x - m) / s)
}


## we reversed the order for the cluster-gene pairs, so use the original cluster order in pal for col ordering
clusord <- names(pal2)

lc2$label <- factor(lc2$label,levels=clusord)

### drop uncertain samples
lc2 <- lc2[,!(lc2$brnum %in% paste0("Br",c(5517,5276,5712)))]
    
# make an spe with reordering of the columns to be in the order of the clusters we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
i <- 1
for (i in c(1:length(levels(lc2$label)))) {
    curlev <- levels(lc2$label)[i]
    if (i == 1) {
        cnames <- colnames(lc2[, lc2$label == curlev])
    } else {
        cnames <- c(cnames, colnames(lc2[, lc2$label == curlev]))
    }
}
rm(i, curlev)

### extract log counts per marker, z-normalize, hang onto both

lc2 <- scater::computeLibraryFactors(lc2)
lc2 <- scater::logNormCounts(lc2)
pltdat <- SingleCellExperiment(lc2[mkg, cnames],rowData=rowData(lc2[mkg,])) 

# ^ gets rid of imagedata here but screws up assays somehow
assays(pltdat)$counts <- assays(lc2[mkg,cnames])$counts
assays(pltdat)$logcounts <- assays(lc2[mkg,cnames])$logcounts

# dump the screwed up assays in slot 1
assays(pltdat) <- assays(pltdat)[c(2:3)]

# change rownames to gene symbols
rownames(pltdat) <- rowData(pltdat)$gene_name
pltdat2 <- scale_rows(as.matrix(logcounts(pltdat)))

## labels also got bllown away here -_-
louvs2 <- louvs[colnames(pltdat),]
pltdat$label <- louvs2$anno
pltdat$label <- factor(pltdat$label,levels=clusord)
```

# 2B-continued make the plot.
```{r}
# z score 99 %ile is 3point something. So we can just plot a symmetric -3,0,3 to get most of the variation in the significant things  

htmap <- ComplexHeatmap::Heatmap(
    pltdat2,
    top_annotation = HeatmapAnnotation(
        `Visium Cluster` = pltdat$label,
        show_annotation_name = FALSE,
        gap = 0,
        col = list(`Visium Cluster`=pal2),
        gp = gpar(alpha = 1,cex=0.35)
    ),
    show_column_names = FALSE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    show_row_names = TRUE,
    col = colorRamp2(
        colors=c("blue","white","orange"),
        breaks = c(-3,0,3)
    ),
    cluster_columns = F, # handled by ordering the collumns of the spe above
    cluster_rows = F, #ibid
    heatmap_legend_param = list(
        title = "Z scored spot log counts",
        direction = "horizontal",
        title_position = "topcenter",
        at=c(-5,0,5) 
        # at argument forces the - region of the legend to show, it won't by default due to the long positive end 
    ),
    row_names_gp = gpar(fontsize = 5.5),
    raster_device = "CairoPNG",
    raster_quality = 10
)


### save plot; add title 

# unstack the legends in illustator; ital the gene names in illustrator.
CairoPDF("plots/manuscript/Fig2/2B-topSignifMarkersByLFC_heatmap.pdf", height = 4.5, width = 3.5)

draw(htmap,
    annotation_legend_side = "right",
    heatmap_legend_side = "bottom",
    legend_grouping = "original"
)
dev.off()

rm(pltdat,pltdat2,gns,cnames,clusts,pal2,clusord,mkg,htmap)
```

Fig 2C: WeberDivecha registration. we ran this when vetting different clustering/feature approaches.
```{r}
webdivreg <- readRDS("processed-data/13_markers_and_registration_final/02-LC3_registration_to_WeberDivecha23_final.RDS")

## get the registration values when using the max number of genes in both this dataset and the current webdiv dataset (slot 22 of the respective labeling approach for whichever data of theirs)
pltdat <- list()
i<-1
for (i in c(1:length(webdivreg))){
    out <- as.data.table(webdivreg[[i]][[22]])
    out[,webdivdat:=names(webdivreg)[i]]
    out <- melt.data.table(out,id.vars=c("LFF_clus","webdivdat"))
    setnames(out,"variable","WebDiv_label")
    pltdat[[i]] <- out
}
rm(i,out)

pltdat <- rbindlist(pltdat)
pltdat[,webdivdat:=as.character(webdivdat)]
pltdat[,LFF_clus:=as.character(LFF_clus)]

## subset to our main pure-ish clusters and their corresponding snuc/spatial types
pltdat <- pltdat[
    LFF_clus %in% c("LC","ACHE_SERT_LC_WM","ACHE_SERT_Npep","Astro","Oligo") 
    & 
    WebDiv_label %in% c("label_merged_x5HT","label_merged_xNE","label_merged_xastrocytes","label_merged_xinhibitory","label_merged_xexcitatory","label_merged_xoligodendrocytes","unsupervised_5HT_TRUE","supervised_NE_TRUE","annot_region_TRUE","annot_spot_TRUE")]

## prettify
pltdat[,WebDiv_label:=gsub(WebDiv_label,pattern="label_merged_x",replacement="snSeq ")]
pltdat[WebDiv_label=="snSeq inhibitory",WebDiv_label:="snSeq Inhibitory"]
pltdat[WebDiv_label=="snSeq excitatory",WebDiv_label:="snSeq Excitatory"]
pltdat[WebDiv_label=="snSeq astrocytes",WebDiv_label:="snSeq Astrocytes"]
pltdat[WebDiv_label=="snSeq oligodendrocytes",WebDiv_label:="snSeq Oligo"]

pltdat[WebDiv_label=="supervised_NE_TRUE",WebDiv_label:="snSeq Manual NE"]
pltdat[WebDiv_label=="unsupervised_5HT_TRUE",WebDiv_label:="snSeq Unsuperv. 5HT"]
pltdat[WebDiv_label=="annot_region_TRUE",WebDiv_label:="Vis. LC Region, Manual"]
pltdat[WebDiv_label=="annot_spot_TRUE",WebDiv_label:="Vis. LC Spot, Manual"]

pltdat[,LFF_clus:=gsub(LFF_clus,pattern="_",replacement=" ")]

pltdat[,WebDiv_label:=factor(WebDiv_label,levels=c(paste0("snSeq ",c("Oligo", "Astrocytes", "Unsuperv. 5HT", "5HT", "Inhibitory", "Excitatory", "NE","Manual NE")),"Vis. LC Spot, Manual","Vis. LC Region, Manual"))]
pltdat[,LFF_clus:=factor(LFF_clus,levels=c("Oligo","Astro","ACHE SERT Npep","ACHE SERT LC WM","LC"))]

## plot and save

pdf("plots/manuscript/Fig2/2C-WeberDivecha_registration.pdf",height=3.5,width=3.5)
ggplot(pltdat,aes(x=LFF_clus,y=WebDiv_label,fill=value))+
    geom_tile()+
    scale_x_discrete(expand = c(0,0)) +
    scale_y_discrete(expand = c(0,0)) +
    scale_fill_continuous(type="viridis",limits=c(min(pltdat$value),max(pltdat$value)))+
    labs(x="Cluster (This Dataset)",y="LC-NE snSeq/Visium\n(Weber, Divecha 2023)",fill="Spearman Cor.\n(Top ~4500 En-\nriched Genes)")+
    theme(
        axis.text.x=element_text(angle=45,hjust=1,size=8,colour = "#000000"),
        axis.text.y=element_text(size=8,colour = "#000000"),
        axis.title.x=element_text(size=9,colour = "#000000"),
        axis.title.y=element_text(size=9,colour = "#000000", margin=margin(0,0,0,-0.02,unit="in")),
        legend.title=element_text(size=8,hjust=0.5,colour = "#000000"),
        legend.text=element_text(size=7,colour = "#000000"),
        legend.background=element_blank(),
        legend.margin=margin(0,-0.08,0,-0.1,unit="in")
    )
dev.off()
```

Fig 2D: Show one of the previous samples, now with only LC NM+ and - labeled. V13M06-404_B1 looks like a good sample for illustrating
```{r}
fig2c <- lc2[,lc2$sample_id=="V13M06-404_B1"]
fig2c <- fig2c[,fig2c$lcnm_subclus %in% c("LC_NMpos","LC_NMneg")]
fig2c$lcnm_subclus <- gsub(fig2c$lcnm_subclus,pattern="_",replacement=" ")
fig2c$lcnm_subclus <- gsub(fig2c$lcnm_subclus,pattern="pos",replacement="+")
fig2c$lcnm_subclus <- gsub(fig2c$lcnm_subclus,pattern="neg",replacement="-")

lcnmpal <- c(pal[1],"#9a9a9a")
names(lcnmpal) <- c("LC NM+","LC NM-")

# point size of 0.9175 at this pdf size (2.4 x 3) and theme/legend parameter set gives spots nearly touching; the center-to-center distance between visium spots is 100µm (i.e., 100-(27.5µm radii of two spots)=45µm between spots). So we want to use approx (1-0.45)*0.9175 for point_size to accurately represent dead space (and NM overlap)

## that comes out a little small. use 0.55 straight up

## finally, spatialLIBD's autocrop is still including the vast majority of the image (whereas LC is a much smaller area of it, and despite us throwing out all the spots that aren't LC here. so some post-adjusting in illustrator will be required; triple all the parameters for sizes and themes besides legend so that we can rework it later).

plotfact <- 3
p <- vis_clus(fig2c,sampleid="V13M06-404_B1",spatial = T,na_color = NULL,auto_crop = T,point_size = 0.55*plotfact,alpha=0.45,clustervar = "lcnm_subclus",colors = lcnmpal,image_id = "hires")+
    theme(plot.title = element_text(size = 9, hjust = 0.5,margin=margin(0,0,0,0)),
          panel.border = element_blank(),
          plot.margin = margin(-0.15*plotfact,
                               0.0375*plotfact,
                               -0.7125*plotfact,
                               -0.675*plotfact,unit="in"),
          legend.key.height = unit(0.083,"in"),
          legend.key.width = unit(0.156,"in"),
          legend.text=element_text(size=7,margin=margin(0,0,0,-0.025,unit="in")),
          legend.title=element_text(size=8),
          legend.key=element_blank(),
          legend.background = element_blank())+
    scale_fill_manual(values=lcnmpal)+
    labs(fill="LC Spot Type")+
    guides(fill=guide_legend(override.aes=list(size=1.25,alpha=0.45)))

## also plot in higher res than would otherwise be necessary so we have good resolution of the illustrator-cropped out LC zoone
p <- rasterize(p,layers="Points",dpi=600,dev="cairo_png")

pdf("plots/manuscript/Fig2/Fig2D_needs_illust_postproc.pdf",height=plotfact*2.4,width=plotfact*3)
p
dev.off()

rm(fig2c,p,plotfact)
```

Fig 2E: LC NM+- DE volcano
```{r}
nmde <- fread("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/04-LCNMpos_vs_LCNMneg_DE.txt")

nmde[,yval:=-log10(adj.P.Val)]
nmde[,uptype:=ifelse(logFC>0,yes="LC NM+",no="LC NM-")]

## illustrate glial and inhibitory genes for NM- enrichment; LC markers and GPX3 (NM-assoc gene) for NM+ enrichment
nmde[gene_name %in% c("GAD1","GAD2","TCN","GJB6","VCAN","PHOX2B","SLC6A2","SLC18A2","TH","GPX3"),label:=gene_name]

p <- ggplot(nmde,aes(x=logFC,y=yval,col=uptype,alpha=ifelse(nmde$adj.P.Val<0.05,yes=1,no=0.4)))+
    geom_point(size=0.6,stroke = 0)+
    geom_hline(yintercept=-log10(0.05),linetype="dashed",col="red",linewidth=0.5)+
    labs(x="logFC\n(NM+ vs NM-)",y="-log10 (FDR)",col="LC\nSpot Type")+
    scale_color_manual(values=lcnmpal)+
        geom_text_repel(seed = 23942,aes(y = -log10(P.Value),
                x = logFC,
                label = ifelse(is.na(label), "", label)),
            min.segment.length = 0,
            size = 1.5,
            show.legend=FALSE,
            #nudge_x=1.25*sign(nudges$logFC),
            #seed=10403, 
            max.iter=100000,
            max.overlaps=10000,
            max.time=3,
            segment.size=0.25)+
    guides(alpha="none",color=guide_legend(override.aes=list(size=1.25,alpha=1)))+
    theme(axis.text.x=element_text(size=8),
          axis.text.y=element_text(size=8),
          axis.title.x=element_text(size=9),
          axis.title.y = element_text(size=9),
          legend.title=element_text(size=8,hjust=0.5),
          legend.text=element_text(size=7,margin=margin(0,0,0,-0.025,unit="in")),
          axis.ticks = element_line(linewidth=0.5),
          panel.grid.major = element_line(linewidth=0.5),
          panel.grid.minor = element_line(linewidth=0.25),
          legend.margin = margin(-0.25,0,0,-0.25,unit="in"),
          legend.background = element_blank(),
          legend.box.background = element_blank(),
          legend.key.height = unit(0.083,"in"),
          legend.key.width = unit(0.156,"in"),
          legend.key=element_blank())

p <- rasterize(p,layers="Points",dpi=300,dev="cairo_png")


## point labels are wildly screwed up and will need fixing in illustrator.
pdf("plots/manuscript/Fig2/2E-LCnmDE_volcano.pdf",height=2.5,width=3)
p
dev.off()

rm(p,nmde)
gc(full=T)
```

Fig 2F: LC NMF GSEA didn't really yield anything. instead, use the enrichr results on factor-specific genes (with background sets fed in as all genes on which the respective NMFs were performed).

Start by grabbing results for the following terms, which popped up in one or more factor-specific-gene set, then see how many we really have 
postsynaptic (GO)
presynaptic (GO)
ubiquitin (GO, other pathway/ontol dbs)
esr1
mitochondrial (GO, other pathway/ontol dbs)
calml3 (protein complexes)
znf365 (TF coexp, TF ppi)
LMTK3 (kinase coexp)
subceruleus (allen atlas regions)
COPS4 (protein complexes? or ppis? don't remember)
BOLA3 (" or " or TF coexp dont recall)
Sncg (mouse scseq type)
ZNF444 (TF coexp)
TCF25 (TF coexp)
Sensory (cell types/tissues)
Endothel (cell types/tissues)
Olig2 (TF coexpression)
```{r}
nmfgsea <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/06-LC_NMF_specificallyLoadedFactors_GSEA_msigdbAndTFtarg.RDS")

factspec <- fread("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/06-LCNMF_factors_and_factorSpecGenes_min25specGenes.txt")








```
