---
title: "Fig3-LC NMpos vs NMneg, LConly NMF-GSEA"
format: html
date: "07-09-2025"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(ggrastr)
library(Cairo) # pdf device for ^ to properly save; also used for complexheatmap raster functions
library(circlize) #colorRamp2
library(viridis) # magma
library(ggrepel) # volcano point labeling

## rstudio GUI tweaks
require(colorout)
ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```


## For the clustering spotplots, we want to use the H&E images and plot by capture area, so we will need to load the big daddy SPE
```{r}
lc2 <- readRDS("processed-data/04_QC/04-lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed.RDS")

# load clustering and the split-sections coldata, which has a column with pre-split spot ids we can use to match everything back to capture level
louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno,by="clusid")

splitcd <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")
splitcd <- splitcd[,.(rowname,original_spot_id)]

louvs <- merge.data.table(louvs,splitcd,by.x="rn",by.y="rowname") # split section ids

## also attach LCnm subsets because we will use them in this figure
lcnm <- fread("processed-data/10_subclustering/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnm <- lcnm[,.(rn,NMseg_subclus)]

louvs <- merge.data.table(louvs,lcnm,by="rn",all.x=T)
louvs$anno[louvs$anno=="Oligo_Complex"] <- "Oligo and Other Types Mixed"
louvs$anno[louvs$anno=="Astro_Oligo"] <- "Astro-Oligo Mixed"
louvs$anno[louvs$anno=="Oligo_Astro"] <- "Oligo-Astro Mixed"
louvs$anno[louvs$anno=="ACHE_SERT_LC_WM"] <- "*ACHE*, *SERT*, LC, Oligo Mixed"
louvs$anno[louvs$anno=="ACHE_SERT_Npep"] <- "*ACHE*, *SERT*, Neuropeptide Mixed"


louvs <- DataFrame(louvs,row.names=louvs$original_spot_id)[colnames(lc2),]
stopifnot(nrow(louvs)==ncol(lc2))

colLabels(lc2) <- louvs$anno
lc2$lcnm_subclus <- louvs$NMseg_subclus

rm(splitcd,lcnm)
gc(full=T)
```

## load cluster 9-color palette
```{r}
pal <- readRDS("plots/finalpal_somuchfornorainbows.RDS")

stopifnot(all(names(pal) %in% unique(lc2$label)))
```

### custom function to extrct legend from one plot to extend to all ###
```{r}
extract_legend <- function(my_ggp) {
    step1 <- ggplot_gtable(ggplot_build(my_ggp))
    step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
    step3 <- step1$grobs[[step2]]
    return(step3)
}
```

Fig 3A: Show one of the previous samples, now with only LC NM+ and - labeled. V13M06-404_B1 looks like a good sample for illustrating
```{r}
fig3a <- lc2[,lc2$sample_id=="V13M06-404_B1"]
fig3a <- fig3a[,fig3a$lcnm_subclus %in% c("LC_NMpos","LC_NMneg")]
fig3a$lcnm_subclus <- gsub(fig3a$lcnm_subclus,pattern="_",replacement=" ")
fig3a$lcnm_subclus <- gsub(fig3a$lcnm_subclus,pattern="pos",replacement="+")
fig3a$lcnm_subclus <- gsub(fig3a$lcnm_subclus,pattern="neg",replacement="-")

lcnmpal <- c(pal[1],"#9a9a9a")
names(lcnmpal) <- c("LC NM+","LC NM-")

# point size of 0.9175 at this pdf size (2.4 x 3) and theme/legend parameter set gives spots nearly touching; the center-to-center distance between visium spots is 100µm (i.e., 100-(27.5µm radii of two spots)=45µm between spots). So we want to use approx (1-0.45)*0.9175 for point_size to accurately represent dead space (and NM overlap)

## that comes out a little small. use 0.55 straight up

## finally, spatialLIBD's autocrop is still including the vast majority of the image (whereas LC is a much smaller area of it, and despite us throwing out all the spots that aren't LC here. so some post-adjusting in illustrator will be required; triple all the parameters for sizes and themes besides legend so that we can rework it later).

plotfact <- 3
p <- vis_clus(fig3a,sampleid="V13M06-404_B1",spatial = T,na_color = NULL,auto_crop = T,point_size = 0.55*plotfact,alpha=0.45,clustervar = "lcnm_subclus",colors = lcnmpal,image_id = "hires")+
    theme(plot.title = element_text(size = 9, hjust = 0.5,margin=margin(0,0,0,0,unit="in")),
          panel.border = element_blank(),
          plot.margin = margin(-0.15*plotfact,
                               0.0375*plotfact,
                               -0.7125*plotfact,
                               -0.675*plotfact,unit="in"),
          legend.key.height = unit(0.083,"in"),
          legend.key.width = unit(0.156,"in"),
          legend.text=element_text(size=7,margin=margin(0,0,0,-0.025,unit="in")),
          legend.title=element_text(size=8),
          legend.key=element_blank(),
          legend.background = element_blank())+
    scale_fill_manual(values=lcnmpal)+
    labs(fill="LC Spot Type")+
    guides(fill=guide_legend(override.aes=list(size=1.25,alpha=0.45)))

## also plot in higher res than would otherwise be necessary so we have good resolution of the illustrator-cropped out LC zoone
p <- rasterize(p,layers="Points",dpi=600,dev="cairo_png")

pdf("plots/manuscript/Fig3/Fig3A_needs_illust_postproc.pdf",height=plotfact*2.4,width=plotfact*3)
p
dev.off()

rm(fig3a,p,plotfact)
```

Fig 3B: LC NM+- DE volcano
```{r}
nmde <- fread("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/04-LCNMpos_vs_LCNMneg_DE.txt")

nmde[,yval:=-log10(adj.P.Val)]
nmde[,uptype:=ifelse(logFC>0,yes="LC NM+",no="LC NM-")]

## illustrate glial and inhibitory genes for NM- enrichment; LC markers and GPX3 (NM-assoc gene) for NM+ enrichment
nmde[gene_name %in% c("GAD1","STAT3","TCN","GJB6","VCAN","PHOX2B","SLC6A2","ATF4","TH","GPX3"),label:=gene_name]

p <- ggplot(nmde,aes(x=logFC,y=yval,col=uptype,alpha=ifelse(nmde$adj.P.Val<0.05,yes=1,no=0.4)))+
    geom_point(size=0.6,stroke = 0)+
    xlim(-max(abs(nmde$logFC)),max(abs(nmde$logFC)))+
    geom_hline(yintercept=-log10(0.05),linetype="dashed",col="red",linewidth=0.5)+
    labs(x="logFC\n(NM+ vs NM-)",y="-log10 (FDR)",col="LC\nSpot Type")+
    scale_color_manual(values=lcnmpal)+
        geom_text_repel(seed = 23942,aes(y = yval,
                x = logFC,
                label = ifelse(is.na(label), "", label)),
            min.segment.length = 0,
            size = 1.5,
            show.legend=FALSE,
            #nudge_x=1.25*sign(nudges$logFC),
            #seed=10403, 
            max.iter=100000,
            max.overlaps=10000,
            max.time=3)+
    guides(alpha="none",color=guide_legend(override.aes=list(size=1.25,alpha=1)))+
    theme(axis.text.x=element_text(size=8),
          axis.text.y=element_text(size=8),
          axis.title.x=element_text(size=9),
          axis.title.y = element_text(size=9),
          legend.title=element_text(size=8,hjust=0.5),
          legend.text=element_text(size=7,margin=margin(0,0,0,-0.025,unit="in")),
          axis.ticks = element_line(linewidth=0.5),
          panel.grid.major = element_line(linewidth=0.5),
          panel.grid.minor = element_line(linewidth=0.25),
          legend.margin = margin(-0.25,0,0,-0.25,unit="in"),
          legend.background = element_blank(),
          legend.box.background = element_blank(),
          legend.key.height = unit(0.083,"in"),
          legend.key.width = unit(0.156,"in"),
          legend.key=element_blank())

p <- rasterize(p,layers="Points",dpi=300,dev="cairo_png")


## point labels are wildly screwed up and will need fixing in illustrator. make it 2x size so that we can tell where labels and lines are supposed to be going (and then downsixze in illust.) 
pdf("plots/manuscript/Fig3/3B-LCnmDE_volcano.pdf",height=5,width=6)
p
dev.off()

rm(p,nmde)
gc(full=T)
```

Fig 3C: SCORPION GRN (cytoscape) using processed-data/12/12f

Fig 3D: NM DEG GSEA
```{r}
msignm <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/05-LCNMposVsNeg_msigdb_and_tftarg_GSEAs.RDS")
## ^ only grab the msigdb ones here; TF-targ enrichments were filtered beyond this point to those where TF is at least marginally expressed later
msignm <- msignm$msigdb_signed

tfnm <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/08-LCNM_APOESexDE_PredomAncesDE_LCnmf_LCnmfLct18Filt_TFtarg_GSEA_filterToClusAndDEgroupExpressed.RDS")
tfnm <- tfnm$LCNM_tftarg_signed

## also, the filtering allows for the possibility of transcriptional repressors (i.e., "targets" of the TF expressed enriched in cell group 1, TF expressed in cell group 2), which is going to be confusing for a figure. so make sure LCNM_neg_up TFs are expressed in LC_NMneg and vice versa for this
tfnm <- tfnm[(TF_LC_NMneg_Expr=="yes"&dir=="LCNM_neg_up")|(TF_LC_NMpos_Expr=="yes"&dir=="LCNM_pos_up")]
# that's the vast majority of them anyhow

#### NMpos TFs: 
# STAT3, 
# ATF4-upregulated (PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-upregulated_In_ATF4OE_txt-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors), 
# MEF2D-down
# EBF3-up (hum prot atlas shows this as being most highly expressed in human LC)

## NMpos-enriched TFtarg sets to keep:
# PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-downregulated_In_ATF4OE-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors
# PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-upregulated_In_ATF4OE_txt-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors
# EBF3_OE_SKNSH_HUMAN_GSE90682_RNASEQ_UP_TF_Perturbations_Followed_by_Expression
# MEF2D_KD_MOUSE_GSE47150_CREEDSID_GENE_956_DOWN_TF_Perturbations_Followed_by_Expression
# PMC7801680-41598_2020_79740_MOESM9_ESM.xlsx-STAT3_Target_Genes_In_AD1-STAT3_Target_Genes_Predicted_By_CHIP-seq_Overlap_With_Genes_In_AD_M1_STAT3_Rummagene_transcription_factors

#### NMneg TFs: ATF6-down, EBF3-down, MEF2D-UP, STAT3-UP, AR, ATF4-downregulated
## NMneg TFtarg sets to keep:
# PMC6345883-41598_2018_37397_MOESM10_ESM.xlsx-AR_D0_VS_D7_WO_B29_TRANS-Gene_Name_AR_Rummagene_transcription_factors
# PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-downregulated_In_ATF4OE-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors
# ATF6_ACTIVATION_HESC_HUMAN_GSE106847_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression
# EBF3_OE_SKNSH_HUMAN_GSE90682_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression
# MEF2D_KD_MOUSE_GSE47150_CREEDSID_GENE_956_UP_TF_Perturbations_Followed_by_Expression
# STAT3_KO_MOUSE_GSE38375_CREEDSID_GENE_2479_UP_TF_Perturbations_Followed_by_Expression

## this is very interesting--many bidirectional relationships here!

tfnm <- tfnm[pathway %in% c(
  "PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-downregulated_In_ATF4OE-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors",
  "PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-upregulated_In_ATF4OE_txt-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors",
  "EBF3_OE_SKNSH_HUMAN_GSE90682_RNASEQ_UP_TF_Perturbations_Followed_by_Expression",
  "MEF2D_KD_MOUSE_GSE47150_CREEDSID_GENE_956_DOWN_TF_Perturbations_Followed_by_Expression",
  "PMC7801680-41598_2020_79740_MOESM9_ESM.xlsx-STAT3_Target_Genes_In_AD1-STAT3_Target_Genes_Predicted_By_CHIP-seq_Overlap_With_Genes_In_AD_M1_STAT3_Rummagene_transcription_factors",
  "PMC6345883-41598_2018_37397_MOESM10_ESM.xlsx-AR_D0_VS_D7_WO_B29_TRANS-Gene_Name_AR_Rummagene_transcription_factors",
  "PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-downregulated_In_ATF4OE-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors",
  "ATF6_ACTIVATION_HESC_HUMAN_GSE106847_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression",
  "EBF3_OE_SKNSH_HUMAN_GSE90682_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression",
  "MEF2D_KD_MOUSE_GSE47150_CREEDSID_GENE_956_UP_TF_Perturbations_Followed_by_Expression",
  "STAT3_KO_MOUSE_GSE38375_CREEDSID_GENE_2479_UP_TF_Perturbations_Followed_by_Expression")]

### annotate TF with target regulation direction
tfnm[pathway=="PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-downregulated_In_ATF4OE-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors",tf:=paste0(tf," ","Downreg.")]
tfnm[pathway=="PMC8444148-NIHMS1529754-supplement-1529754_Tab_S2.xlsx-upregulated_In_ATF4OE_txt-RNA-seq_Of_C3_ATF4_OE_And_C3_Neurons_ATF4_Rummagene_transcription_factors",tf:=paste0(tf," ","Upreg.")]
tfnm[pathway=="EBF3_OE_SKNSH_HUMAN_GSE90682_RNASEQ_UP_TF_Perturbations_Followed_by_Expression",tf:=paste0(tf," ","Upreg.")]
tfnm[pathway=="MEF2D_KD_MOUSE_GSE47150_CREEDSID_GENE_956_DOWN_TF_Perturbations_Followed_by_Expression",tf:=paste0(tf," ","Upreg.")]
tfnm[pathway=="PMC7801680-41598_2020_79740_MOESM9_ESM.xlsx-STAT3_Target_Genes_In_AD1-STAT3_Target_Genes_Predicted_By_CHIP-seq_Overlap_With_Genes_In_AD_M1_STAT3_Rummagene_transcription_factors",tf:=paste0(tf," ","ChIP")]
tfnm[pathway=="PMC6345883-41598_2018_37397_MOESM10_ESM.xlsx-AR_D0_VS_D7_WO_B29_TRANS-Gene_Name_AR_Rummagene_transcription_factors",tf:=paste0(tf," ","Bidir.")]
tfnm[pathway=="ATF6_ACTIVATION_HESC_HUMAN_GSE106847_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression",tf:=paste0(tf," ","Downreg.")]
tfnm[pathway=="EBF3_OE_SKNSH_HUMAN_GSE90682_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression",tf:=paste0(tf," ","Downreg.")]
tfnm[pathway=="MEF2D_KD_MOUSE_GSE47150_CREEDSID_GENE_956_UP_TF_Perturbations_Followed_by_Expression",tf:=paste0(tf," ","Downreg.")]
tfnm[pathway=="STAT3_KO_MOUSE_GSE38375_CREEDSID_GENE_2479_UP_TF_Perturbations_Followed_by_Expression",tf:=paste0(tf," ","Downreg.")]

## how the shit to plot this..I think this is not getting the key point across, which is that NM+ neurons are both enriched in TF-upregulated genes AND depleted of TF-repressed genes


ggplot(tfnm,aes(x=-log10(pval),y=tf,color=NES,shape=`LC Spot Type\nEnriched for Set`,size=abs(NES)))+
  geom_point()+
  ylab("TF and effect on genes in geneset")+
  scale_color_gradient2(breaks=c(-2,0,2,4),low="#F15FFF",mid="gray",high = "darkorange4")+labs(color="Norm.\nEnrichment\nScore")
```

3E UPPER PORTION: LC NMF factors 3,4,6,24,45  (3, 6, 24, 45 strongest): LC
12: astrocytes/OPCs
14, 30: microglia/endomural

enrichr and projection of SPE NMF into WebDiv snseq

This plot will be laid out such that above the x axis is the enrichr results and below it are the projections onto webdiv. LC NMF loadings into GSEA didn't really yield anything. instead, use the enrichr results on factor-specific genes (with background sets fed in as all genes on which the respective NMFs were performed).

Start by grabbing results for the following terms, which popped up in one or more factor-specific-gene set, then see how many we really have 
postsynaptic (GO)
presynaptic (GO)
ubiquitin (GO, other pathway/ontol dbs)
esr1
mitochondrial (GO, other pathway/ontol dbs)
calml3 (protein complexes)
znf365 (TF coexp, TF ppi)
LMTK3 (kinase coexp)
subceruleus (allen atlas regions)
COPS4 (protein complexes? or ppis? don't remember)
BOLA3 (" or " or TF coexp dont recall)
Sncg (mouse scseq type)
ZNF444 (TF coexp)
TCF25 (TF coexp)
Sensory (cell types/tissues)
Endothel (cell types/tissues)
Olig2 (TF coexpression)
```{r}
nmfenrichr <- readRDS("processed-data/10_subclustering/09cV2-LCnmfFilt_factorSpecGenes_enrichr_results.RDS")
## fix something stupid where N_genes came back as a list
nmfenrichr[,N_genes:=as.numeric(N_genes)]

factspec <- fread("processed-data/10_subclustering/09bV2-LCNMFLogCt18Filt_factors_and_factorSpecGenes_min25specGenes.txt")

enrichrpltdat <- nmfenrichr[Term %in% grep(Term,pattern="postsynaptic|presynaptic|ubiquitin|esr1|mitochondr|calml3|znf365|LMTK3|subcoeruleus|COPS4|BOLA3|Sncg up|ZNF444|TCF25|HSF1 human tf|RPS6KB|FASTK|ADCK2|DAPK3|Sensory|Olig2|ATF3|ATF4|ATF6|XBP1",value=T,ignore.case=T)]
enrichrpltdat <- enrichrpltdat[LCnmf_filt_factor%in%paste0("LCnmf",c(3,6,24,45,12,14,30)),]

rm(nmfenrichr)

## subset to factors of interest and terms of interest first filtering for those signif after correction and matching terms of interest. term 4 has no notable enrichments.
keepterm <- unique(enrichrpltdat[Adjusted.P.value<0.2,max(N_genes),by=c("Enrichr_database","Term")])
setnames(keepterm,"V1","maxNgenes")

keepterm <- keepterm[Term %in% grep(Term,pattern="Ubiquitin|HSF1 human tf|RPS6KB|ubiquitin|FASTK|ADCK2|DAPK3|COPS4|BOLA3|ZNF365|ESR1|LMTK3|ATF3|ATF4|ATF6|XBP1|Olig2|Postsynaptic|Presynaptic|TCF25|subcoeruleus",value=T)|Term=="Mitochondrial Translation"|Term=="ESR1"|Term=="ATF3 A549 hg19"|(Term=="ATF6 Down"&Enrichr_database=="LINCS_L1000_CRISPR_KO_Consensus_Sigs")]


## drop a couple less important/redundant ones and ones driven by only one gene
keepterm <- keepterm[!(Term %in% grep(Term,pattern="MUT ESR1|AINV15 Mouse|FASTKD|HA1E|HCC515|ADCK2 IDG kinase|FASTK IDG|RPS6KB2 knockdown 96h HEPG2|RPS6KB2 knockdown 96h HT29|UTERUS",value=T,ignore.case=T))] # gene mutant screens; less relevant tissue; not intended gene (FASTK); cancer lines that parallel A549 (way more relevant); we care about the human kinase coexp; ibid; another one where we have A549 which is more relevant; ibid; irrel. tissue
keepterm <- keepterm[Term!="ZNF365 Up"] # weak/redundant
keepterm <- keepterm[maxNgenes>1]

## Now clean up/consolidate terms
keepterm[Term %in% grep(Term,pattern="subcoeruleus",value=T),pltterm:="Allen Ms r1/r2 Sub-<br>coeruleus nu."]
keepterm[,pltterm:=gsub(Term,pattern="^(.*) human tf ARCHS4 coexpression",replacement=paste0("*","\\1","*","Coexp. TFs"))]
keepterm[,pltterm:=gsub(Term,pattern="^(.*) human kinase ARCHS4 coexpression",replacement=paste0("*","\\1","*"," Coexp. Kinases"))]

keepterm[Term=="TCF25 Up",pltterm:="*TCF25* KD/KO Upreg."]
keepterm[Term=="LMTK3 Up",pltterm:="*LMTK3* KD/KO Upreg."]
keepterm[Term=="ATF3 Up",pltterm:="*ATF3* KD/KO Upreg."]
keepterm[Term=="ATF4 Up",pltterm:="*ATF4* KD/KO Upreg."]
keepterm[Term=="ATF3 Down",pltterm:="*ATF3* KD/KO Downreg."]
keepterm[Term=="ATF4 Down",pltterm:="*ATF4* KD/KO Downreg."]
keepterm[Term=="ATF6 Down",pltterm:="*ATF6* KD/KO Downreg."]

keepterm[Term=="OLIG2 23332759 ChIP-Seq OLIGODENDROCYTES Mouse",pltterm:="Ms Olig2 ChIP Targets"]
keepterm[Term=="Regulation Of Postsynaptic Membrane Neurotransmitter Receptor Levels (GO:0099072) BP",pltterm:="Regulation of Postsyn.<br>NT Recep. Levels"]
keepterm[Term=="Integral Component Of Presynaptic Active Zone Membrane (GO:0099059) CC",pltterm:="Presynaptic Active Zone<br>Memb. Integral Component"]
keepterm[Term %in% c("ESR1 knockdown 96h A549", "ESR1 up"),pltterm:="*ESR1* KD/KO Upreg."]
keepterm[Term=="Ubiquitin Ligase Activator Activity (GO:1990757)",pltterm:="Ubiquitin Ligase<br>Activator Activity"]
keepterm[Term=="Ubiquitin-Protein Transferase Activator Activity (GO:0097027)",pltterm:="Ubiquitin-Protein Transferase<br>Activator Activity"]
keepterm[Term=="ATF3 23680149 ChIP-Seq GBM1-GSC Human",pltterm:="ATF3 ChIP Targets"]

## drop one more
keepterm[Term!="Estrogens/ESR1 non-Genomic Signaling"] # v weak
## filter to terms we've renamed
keepterm <- keepterm[pltterm!=Term]

### merge these reduced terms back to the data we're keeping to plot
enrichrpltdat <- merge.data.table(enrichrpltdat[Term %in% keepterm$Term],keepterm,by=c("Enrichr_database","Term"))
stopifnot(all(keepterm$pltterm %in% enrichrpltdat$pltterm))

## make a new table with every factor and every term
pltdat <- unique(expand.grid(unique(enrichrpltdat[,.(LCnmf_filt_factor,pltterm)])))

pltdat <- merge(pltdat,enrichrpltdat[,.(LCnmf_filt_factor,pltterm,Adjusted.P.value,Odds.Ratio,N_genes)],by=c("LCnmf_filt_factor","pltterm"),all.x=T)
pltdat <- as.data.table(pltdat)
pltdat[,`-log10(FDR)`:=-log10(Adjusted.P.value)]
pltdat[Adjusted.P.value>0.2,Odds.Ratio:=NA]
pltdat[Adjusted.P.value>0.2,N_genes:=NA]

## attach percentage of factor-specific genes that are in the set as an alternative size scaling metric
factspecn <- factspec[,.N,by="factor"]
pltdat <- merge.data.table(pltdat,factspecn,by.x="LCnmf_filt_factor",by.y="factor")
pltdat[,propfactgenes:=100*(N_genes/N)]


pltdat[,pltterm:=factor(pltterm,levels=c(
  "Presynaptic Active Zone<br>Memb. Integral Component",
  "*ESR1* KD/KO Upreg.",
  "*ADCK2* Coexp. Kinases",
  "*DAPK3* Coexp. Kinases",
  "*FASTK* Coexp. Kinases",
  "*RPS6KB2* Coexp. Kinases",
  "Ubiquitin Ligase<br>Activator Activity",
  "Ubiquitin-Protein Transferase<br>Activator Activity",
  "*LMTK3* KD/KO Upreg.",
  "*ATF3* KD/KO Upreg.",
  "*ATF4* KD/KO Upreg.",
  "*ATF6* KD/KO Downreg.",
  "ATF3 ChIP Targets", 
  "*ATF3* KD/KO Downreg.",
  "*ATF4* KD/KO Downreg.",
  "*TCF25* KD/KO Upreg."))]
## reorder factor to match
## factor 45 has no significant hits left so remove
pltdat <- pltdat[LCnmf_filt_factor!="LCnmf45"]
pltdat[,LCnmf_filt_factor:=factor(LCnmf_filt_factor,levels=paste0("LCnmf",c(3,6,24,30,14,12)))]


pdf("plots/manuscript/Fig3/3Dupper-NMF53_Lowexpfilt_LCnmf_FactorSpecGene_Enrichr_facts3_6_12_14_24_30_maxAdjP0.2.pdf",height=4,width=4.35)
ggplot(pltdat,aes(x=LCnmf_filt_factor,y=pltterm,col=`-log10(FDR)`))+
    geom_point(aes(size=propfactgenes))+
    scale_color_viridis_c(option="magma")+
    #scale_x_discrete(expand = c(0,0)) +
    #scale_y_discrete(expand = c(0,0)) +
    labs(x="LC NMF Factor",y="Enrichr Term",size="% Factor Genes\nin Term")+
    theme(axis.text.x=element_text(angle=45,hjust=1,size=7,colour = "#000000"),
          axis.text.y=element_markdown(size=7,colour = "#000000"),
          axis.title.x=element_text(size=8,colour = "#000000"),
          axis.title.y=element_text(size=8,colour = "#000000", margin=margin(0,0,0,-0.02,unit="in")),
          legend.title=element_text(size=8,hjust=0.5,colour = "#000000"),
          legend.text=element_text(size=7,colour = "#000000"),
          legend.background=element_blank(),
          legend.margin=margin(0,-0.08,0,-0.1,unit="in"))
dev.off()
rm(pltdat,enrichrpltdat,keepterm)
gc(full=T)
```

Fig 3D lower is the NMF projection of these same factors onto webdiv LC snseq
```{r}
library(WeberDivechaLCdata)
lcsn <- WeberDivechaLCdata_singleNucleus()
colnames(colData(lcsn))
## several clusterings
# label-30, numerical
# label_inhibitory (14 subclusters)
# label_merged (10 clusters annotated)
# supervised_NE,unsupervised_NE,unsupervised_5HT (all T/F)
# we'll use label_merged here. label_inhibitory could be interesting for NMF on the other neuronal clusters some other time though.

## load filtered LC NMF output
lcnmf <- readRDS("processed-data/10_subclustering/07b-lcnmf53_tol1e-9_rowSumLogctFilt18.RDS")

## drop multimatches
dropg <- as.data.table(table(rownames(lcsn)))[N>1] # none
keepg <- rownames(lcnmf$w)[rownames(lcnmf$w) %in% rownames(lcsn)] 

## subset SCE and w accordingly
lcsn <- lcsn[keepg,]
lcnmf$w <- lcnmf$w[keepg,]

stopifnot(nrow(lcnmf$w)==nrow(lcsn))
```

## use rcppml::project (pretty straightforward but erik's script for posterity and the L2 parameter used). also, the argument order has changed for project relative to this code, i think. nonetheless:
https://github.com/LieberInstitute/spatial_hpc/blob/main/code/NMF/project_nmf_patterns.R

Note that Erik's code implies there are rownames on the nmf w matrix, which there aren't. so we also need to re-load the full mouse SPE to determine which genes were there to start with (i.e. are the rows of the w matrix) and which remained after ortholog filtering above.
```{r}
library(RcppML)
setDTthreads(1,restore_after_fork = FALSE)
options(RcppML.threads = 10)

set.seed(42)
lffprojinto.webdiv <- RcppML::project(data=logcounts(lcsn),w=as.matrix(lcnmf$w),L2=0.00005)
lffprojinto.webdiv <- t(lffprojinto.webdiv)


stopifnot(nrow(lffprojinto.webdiv)==ncol(lcsn))
rownames(lffprojinto.webdiv) <- colnames(lcsn)
colnames(lffprojinto.webdiv) <- paste0("V",c(1:53))
## append to sce
colData(lcsn) <- cbind(colData(lcsn),lffprojinto.webdiv)
colnames(colData(lcsn))[c(17:ncol(colData(lcsn)))] <- paste0("LCspeFiltNMF_",c(1:53))

usefulname <- colnames(colData(lcsn))[17:ncol(colData(lcsn))]

library(dplyr)
seed3 = as.matrix(colData(lcsn)[,usefulname])
seed3 = seed3>0
d3 = cbind.data.frame(webdivClus=as.data.frame(colData(lcsn))[,"label_merged"],
                      seed3) %>% 
  group_by(webdivClus) %>% add_tally(name="total") %>%
  group_by(webdivClus, total) %>%
  summarise_at(usefulname, sum) %>%
  tidyr::pivot_longer(usefulname, values_to="n", names_to="nmf") %>%
  mutate(prop=n/total)

seed4 = as.matrix(colData(lcsn)[,usefulname])
seed4 = apply(seed4, 2, scale)
d4 = cbind.data.frame(webdivClus=as.data.frame(colData(lcsn))[,"label_merged"],
                      seed4) %>% 
  group_by(webdivClus) %>%
  summarise_at(usefulname, mean) %>% 
  tidyr::pivot_longer(usefulname, values_to="scaled.avg", names_to="nmf")


pltdat <- left_join(d3[,c("webdivClus","nmf","prop")], 
                    d4[,c("webdivClus","nmf","scaled.avg")])

# to d.t.
pltdat <- as.data.table(pltdat)

rm(seed3,seed4,lcnmf,lcsn,lffprojinto.webdiv,d3,d4,dropg)
#reorder clusters to put LC/neurons first, then macrophages_microglia, then astrocytes
lcfirst <- c("NE","macrophages_microglia","astrocytes","OPCs","oligodendrocytes","endothelial_mural","5HT","excitatory","inhibitory","neurons_ambiguous")
pltdat[,webdivClus:=factor(webdivClus,rev(lcfirst))]


pltdat <- pltdat[!is.na(scaled.avg)]
range(pltdat$scaled.avg) # -1.72, 6.25
quantile(pltdat$scaled.avg,c(0.02,0.98)) # -1.12, 3.02

pltdat <- pltdat[nmf %in% paste0("LCspeFiltNMF_",c(3,6,24,30,14,12))]
pltdat[,nmf:=factor(nmf,levels=paste0("LCspeFiltNMF_",c(3,6,24,30,14,12)))]

pdf("plots/manuscript/Fig3/3D LOWER-K53_speLCnmfPrefilt_factors_projectToWeberDivechaSnseq.pdf",height=3,width=3.5)
ggplot(pltdat, aes(x=nmf, y=webdivClus, size=prop, color=scaled.avg))+
  geom_count()+
  theme_bw()+
  scale_size(range=c(0,3))+
  scale_color_viridis_c(option="F",end = 0.8)+
# scale_color_gradient2(low = "darkblue",mid="white",high="red",midpoint = 0,na.value = NA)+
  labs(y="Weber-Divecha Merged Cluster")+
  theme(axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        axis.text.y=element_markdown(size=7,colour = "#000000"),
        axis.title.y=element_text(size=8,colour = "#000000", margin=margin(0,0,0,-0.02,unit="in")),
        legend.title=element_text(size=8,hjust=0.5,colour = "#000000"),
        legend.text=element_text(size=7,colour = "#000000"),
        legend.background=element_blank(),
        legend.margin=margin(0,-0.08,0,-0.1,unit="in"))
dev.off()
```

repinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] dplyr_1.2.0                 RcppML_0.5.6                WeberDivechaLCdata_1.12.0  
 [4] ExperimentHub_3.0.0         AnnotationHub_4.0.0         BiocFileCache_3.0.0        
 [7] dbplyr_2.5.1                BiocParallel_1.44.0         parallelly_1.47.0          
[10] colorout_1.3-3              ggrepel_0.9.6               viridis_0.6.5              
[13] viridisLite_0.4.3           circlize_0.4.17             Cairo_1.7-0                
[16] ggrastr_1.0.2               spatialLIBD_1.22.0          SpatialExperiment_1.20.0   
[19] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0 Biobase_2.70.0             
[22] GenomicRanges_1.62.1        Seqinfo_1.0.0               IRanges_2.44.0             
[25] S4Vectors_0.48.1            BiocGenerics_0.56.0         generics_0.1.4             
[28] MatrixGenerics_1.22.0       matrixStats_1.5.0           gridExtra_2.3              
[31] ggtext_0.1.2                ggplot2_4.0.2               data.table_1.18.4          

loaded via a namespace (and not attached):
  [1] later_1.4.5              BiocIO_1.20.0            bitops_1.0-9            
  [4] filelock_1.0.3           tibble_3.3.1             XML_3.99-0.20           
  [7] lifecycle_1.0.5          httr2_1.2.2              edgeR_4.8.2             
 [10] doParallel_1.0.17        rprojroot_2.1.1          lattice_0.22-7          
 [13] magrittr_2.0.4           limma_3.66.0             plotly_4.12.0           
 [16] sass_0.4.10              rmarkdown_2.30           jquerylib_0.1.4         
 [19] yaml_2.3.12              httpuv_1.6.16            otel_0.2.0              
 [22] sessioninfo_1.2.3        cowplot_1.2.0            DBI_1.2.3               
 [25] RColorBrewer_1.1-3       golem_0.5.1              abind_1.4-8             
 [28] purrr_1.2.1              RCurl_1.98-1.17          rappdirs_0.3.4          
 [31] irlba_2.3.7              commonmark_2.0.0         codetools_0.2-20        
 [34] DelayedArray_0.36.1      DT_0.34.0                scuttle_1.20.0          
 [37] xml2_1.5.2               tidyselect_1.2.1         shape_1.4.6.1           
 [40] farver_2.1.2             ScaledMatrix_1.18.0      shinyWidgets_0.9.0      
 [43] GenomicAlignments_1.46.0 jsonlite_2.0.0           GetoptLong_1.1.0        
 [46] BiocNeighbors_2.4.0      scater_1.38.1            iterators_1.0.14        
 [49] foreach_1.5.2            tools_4.5.2              Rcpp_1.1.1              
 [52] glue_1.8.0               SparseArray_1.10.8       xfun_0.56               
 [55] here_1.0.2               withr_3.0.2              BiocManager_1.30.27     
 [58] fastmap_1.2.0            litedown_0.9             digest_0.6.39           
 [61] rsvd_1.0.5               R6_2.6.1                 mime_0.13               
 [64] colorspace_2.1-2         markdown_2.0             dichromat_2.0-0.1       
 [67] RSQLite_2.4.5            cigarillo_1.0.0          config_0.3.2            
 [70] tidyr_1.3.2              rtracklayer_1.70.1       httr_1.4.7              
 [73] htmlwidgets_1.6.4        S4Arrays_1.10.1          pkgconfig_2.0.3         
 [76] gtable_0.3.6             rsconnect_1.7.0          blob_1.3.0              
 [79] ComplexHeatmap_2.26.1    S7_0.2.1                 XVector_0.50.0          
 [82] htmltools_0.5.9          clue_0.3-66              scales_1.4.0            
 [85] png_0.1-8                attempt_0.3.1            knitr_1.51              
 [88] rstudioapi_0.18.0        rjson_0.2.23             curl_7.0.0              
 [91] cachem_1.1.0             GlobalOptions_0.1.3      stringr_1.6.0           
 [94] BiocVersion_3.22.0       parallel_4.5.2           vipor_0.4.7             
 [97] AnnotationDbi_1.72.0     restfulr_0.0.16          pillar_1.11.1           
[100] grid_4.5.2               vctrs_0.7.1              promises_1.5.0          
[103] BiocSingular_1.26.0      beachmat_2.26.0          xtable_1.8-4            
[106] cluster_2.1.8.1          beeswarm_0.4.0           paletteer_1.7.0         
[109] evaluate_1.0.5           magick_2.9.0             cli_3.6.5               
[112] locfit_1.5-9.12          compiler_4.5.2           Rsamtools_2.26.0        
[115] rlang_1.1.7              crayon_1.5.3             labeling_0.4.3          
[118] rematch2_2.1.2           ggbeeswarm_0.7.3         stringi_1.8.7           
[121] Biostrings_2.78.0        lazyeval_0.2.2           Matrix_1.7-4            
[124] benchmarkme_1.0.8        bit64_4.6.0-1            KEGGREST_1.50.0         
[127] statmod_1.5.1            shiny_1.12.1             gridtext_0.1.5          
[130] memoise_2.0.1            bslib_0.10.0             benchmarkmeData_2.0.0   
[133] bit_4.6.0               
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-07-06
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────────────
 package              * version   date (UTC) lib source
 abind                  1.4-8     2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub        * 4.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1     2020-05-03 [1] CRAN (R 4.5.0)
 beachmat               2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.5.0)
 benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0     2026-01-19 [1] CRAN (R 4.5.2)
 Biobase              * 2.70.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache        * 3.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27   2025-11-14 [1] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0    2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0     2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0     2026-01-14 [1] CRAN (R 4.5.2)
 bslib                  0.10.0    2026-01-26 [1] CRAN (R 4.5.2)
 cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.5.0)
 Cairo                * 1.7-0     2025-10-29 [1] CRAN (R 4.5.0)
 cigarillo              1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 circlize             * 0.4.17    2025-12-08 [1] CRAN (R 4.5.2)
 cli                    3.6.5     2025-04-23 [1] CRAN (R 4.5.0)
 clue                   0.3-66    2024-11-13 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1   2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20    2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3     2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-2     2025-09-22 [1] CRAN (R 4.5.0)
 commonmark             2.0.0     2025-07-07 [1] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1    2026-01-30 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2     2023-08-30 [1] CRAN (R 4.5.0)
 cowplot                1.2.0     2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0     2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.4    2026-05-06 [1] CRAN (R 4.5.2)
 DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr               * 2.5.1     2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.1    2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39    2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                * 1.2.0     2026-02-03 [1] CRAN (R 4.5.2)
 DT                     0.34.0    2025-09-02 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2     2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5     2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub        * 3.0.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 farver                 2.1.2     2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3     2023-12-11 [1] CRAN (R 4.5.0)
 foreach                1.5.2     2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4     2025-05-09 [1] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1    2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 GetoptLong             1.1.0     2025-11-28 [1] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3     2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2     2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr              * 1.0.2     2023-06-01 [1] CRAN (R 4.5.0)
 ggrepel              * 0.9.6     2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.3     2025-11-28 [1] CRAN (R 4.5.2)
 glue                   1.8.0     2024-09-30 [1] CRAN (R 4.5.0)
 golem                  0.5.1     2024-08-27 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2     2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9     2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.5.0)
 httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.5.0)
 httr                   1.4.7     2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2     2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7     2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14    2022-02-05 [1] CRAN (R 4.5.0)
 jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.5.0)
 jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51      2025-12-20 [1] CRAN (R 4.5.2)
 labeling               0.4.3     2023-08-29 [1] CRAN (R 4.5.0)
 later                  1.4.5     2026-01-08 [1] CRAN (R 4.5.2)
 lattice                0.22-7    2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5     2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 litedown               0.9       2025-12-18 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0     2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4     2025-09-12 [1] CRAN (R 4.5.0)
 markdown               2.0       2025-03-23 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4     2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1     2021-11-26 [1] CRAN (R 4.5.0)
 mime                   0.13      2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0     2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0     2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.47.0    2026-04-17 [1] CRAN (R 4.5.2)
 pillar                 1.11.1    2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.5.0)
 plotly                 4.12.0    2026-01-24 [1] CRAN (R 4.5.2)
 png                    0.1-8     2022-11-29 [1] CRAN (R 4.5.0)
 promises               1.5.0     2025-11-01 [1] CRAN (R 4.5.0)
 purrr                  1.2.1     2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1     2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4     2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1     2026-01-10 [1] CRAN (R 4.5.2)
 RcppML               * 0.5.6     2026-02-07 [1] Github (zdebruine/RcppML@5449a5b)
 RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.5.0)
 rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.5.0)
 restfulr               0.0.16    2025-06-27 [1] CRAN (R 4.5.0)
 rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7     2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30      2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1     2025-08-26 [1] CRAN (R 4.5.0)
 Rsamtools              2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 rsconnect              1.7.0     2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5     2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0    2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.5.0)
 rtracklayer            1.70.1    2025-12-20 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.1    2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 S7                     0.2.1     2025-11-14 [1] CRAN (R 4.5.2)
 sass                   0.4.10    2025-04-11 [1] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0     2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.1    2026-03-20 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0    2025-10-30 [1] Bioconductor 3.22 (R 4.5.2)
 Seqinfo              * 1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.5.0)
 shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.5.0)
 shiny                  1.12.1    2025-12-09 [1] CRAN (R 4.5.2)
 shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8    2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0    2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1     2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7     2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0     2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1     2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2     2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1     2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.5.0)
 viridis              * 0.6.5     2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite          * 0.4.3     2026-02-04 [1] CRAN (R 4.5.2)
 WeberDivechaLCdata   * 1.12.0    2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 withr                  3.0.2     2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56      2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20 2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2     2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12    2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────────────
