---
title: "04-Plot all removed spots per sample"
author: "Bernie Mulvey"
date: "2025-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


read in the raw SPE and the colData of the filtered SPE
```{r}
raw <- readRDS("processed-data/02_build_spe/spe_raw.rds")
# switch to the dataset-wide-specific-spotids used in all the other tables and spes
colnames(raw) <- raw$key

qccd <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")

## read annots of tissue margin spots, spotsweeper outlier labels
perimspot <- fread("processed-data/04_QC/01-tissue margin spots.txt")
swept <- fread("processed-data/04_QC/02-Spotsweeper_outliers_post-edge-rm.txt")
```


```{r}
tmpcd <- as.data.table(colData(raw),keep.rownames=T)

qccd[,QC:="Passed"]
qccd <- qccd[,.(original_spot_id,QC)]
setnames(qccd,1,"rn")

setnames(perimspot,"spot_type","QC")
perimspot <- perimspot[!(QC %in% c("keep","image perimeter"))]

swept <- swept[,.(rn,sum_umi_outliers_low,sum_gene_outliers_low,expr_chrM_ratio_outliers_hi)]
swept2 <- melt.data.table(swept,id.vars="rn")
swept2 <- swept2[value==T]
multiout <- swept2[,.N,by="rn"][N>1,rn]
swept2[rn %in% multiout,variable:="local_outlier_multiplemeasures"]
swept2 <- unique(swept2)[,.(rn,variable)]
setnames(swept2,c("rn","QC"))

initqcsteps <- rbind(perimspot,qccd,swept2)
stopifnot(nrow(initqcsteps[,.N,by="rn"][N>1])==0)

tmpcd <- tmpcd[,.(rn,sample_id,in_tissue,overlaps_tissue,sum_umi,sum_gene,array_row,array_col)]
tmpcd <- merge.data.table(tmpcd,initqcsteps,by="rn",all.x=T)

## annotate spots remaining after local outlier filtering and edge removal that were in the bottom 1%ile for gene count or UMI count. those thresholds were 17 and 23 respectively
tmpcd[is.na(QC)&sum_gene<=17,QC:="bottom_1pctile_sumgene"]
tmpcd[is.na(QC)&sum_umi<=23,QC:="bottom_1pctile_sumumi"]

stopifnot(nrow(tmpcd[is.na(QC)])==0)

## replace colData with tmpcd
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(raw),]
colData(raw) <- tmpcd
# clean up
rm(list=grep(ls(),pattern="raw",value=T,invert=T))
gc(full=T)
```


## define helper function to extract a single legend - from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

make plots
```{r}
# make a palette for the levels of QC and apply manually outside of spatiallibd
# unique(raw$QC)

pal <- c("green","darkgray","lightgray","yellow","yellow4","darkorange3","darkorange","orange","darkred")
names(pal) <- c("Passed","off tissue","tissue perimeter","bottom_1pctile_sumumi","bottom_1pctile_sumgene","sum_umi_outliers_low","sum_gene_outliers_low","expr_chrM_ratio_outliers_hi","local_outlier_multiplemeasures")

## make a plot for a sample containing all 9 QC values to extract the legend from and append globally later: V13M06-332_B1
tmpdat <- as.data.table(colData(raw))[sample_id=="V13M06-332_B1"]
leg.tmp <- ggplot(tmpdat,aes(x=array_col,y=array_row,color=QC))+geom_point()+scale_color_manual(values=pal)+theme(legend.text=element_text(size=7),legend.title=element_text(size=8))

leg.tmp <- extract_legend(leg.tmp)
dev.off()

plt <- vis_grid_clus(raw,clustervar="QC",auto_crop = T,point_size = 1.35,return_plots=T,height=4.25,width=4.25,spatial = T,image_id = "lowres",alpha=0.35)

plt <- lapply(plt,FUN=function(x){
    y <- x+guides(fill="none")+
        scale_fill_manual(values=pal)+
        theme(axis.text.x=element_blank(),axis.text.y=element_blank(),title=element_text(size=9))
    return(y)
})

### save--2 plots per row at 4 wide x 4 high + 3 for legend and spaces, 5in per row by 43 samples (22 rows) = 110in
pdf("plots/04_QC/04-Spots_by_QC_step.pdf",width=11,height=110,onefile=FALSE)
do.call("grid.arrange",list(arrangeGrob(grobs=plt,ncol=2),leg.tmp,ncol=2,widths=c(9,2)))
dev.off()
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] spatialLIBD_1.18.0          colorout_1.3-0.2           
 [3] escheR_1.6.0                SpatialExperiment_1.16.0   
 [5] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
 [7] Biobase_2.66.0              GenomicRanges_1.58.0       
 [9] GenomeInfoDb_1.42.1         IRanges_2.40.1             
[11] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[13] MatrixGenerics_1.18.0       matrixStats_1.5.0          
[15] gridExtra_2.3               ggplot2_3.5.1              
[17] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3       rstudioapi_0.17.1        jsonlite_1.8.9          
  [4] magrittr_2.0.3           ggbeeswarm_0.7.2         magick_2.8.5            
  [7] farver_2.1.2             rmarkdown_2.29           BiocIO_1.16.0           
 [10] fields_16.3              zlibbioc_1.52.0          vctrs_0.6.5             
 [13] memoise_2.0.1            config_0.3.2             Rsamtools_2.22.0        
 [16] paletteer_1.6.0          RCurl_1.98-1.16          benchmarkme_1.0.8       
 [19] htmltools_0.5.8.1        S4Arrays_1.6.0           AnnotationHub_3.14.0    
 [22] curl_6.2.0               BiocNeighbors_2.0.1      SparseArray_1.6.0       
 [25] sass_0.4.9               bslib_0.9.0              htmlwidgets_1.6.4       
 [28] plotly_4.10.4            cachem_1.1.0             GenomicAlignments_1.42.0
 [31] mime_0.12                lifecycle_1.0.4          iterators_1.0.14        
 [34] pkgconfig_2.0.3          rsvd_1.0.5               Matrix_1.7-2            
 [37] R6_2.5.1                 fastmap_1.2.0            GenomeInfoDbData_1.2.13 
 [40] shiny_1.10.0             digest_0.6.37            colorspace_2.1-1        
 [43] rematch2_2.1.2           AnnotationDbi_1.68.0     scater_1.34.0           
 [46] rprojroot_2.0.4          irlba_2.3.5.1            ExperimentHub_2.14.0    
 [49] RSQLite_2.3.9            beachmat_2.22.0          labeling_0.4.3          
 [52] filelock_1.0.3           httr_1.4.7               abind_1.4-8             
 [55] compiler_4.4.1           here_1.0.1               bit64_4.5.2             
 [58] withr_3.0.2              doParallel_1.0.17        attempt_0.3.1           
 [61] BiocParallel_1.40.0      viridis_0.6.5            DBI_1.2.3               
 [64] maps_3.4.2.1             sessioninfo_1.2.2        rappdirs_0.3.3          
 [67] DelayedArray_0.32.0      rjson_0.2.23             tools_4.4.1             
 [70] vipor_0.4.7              beeswarm_0.4.0           httpuv_1.6.15           
 [73] glue_1.8.0               restfulr_0.0.15          promises_1.3.2          
 [76] grid_4.4.1               generics_0.1.3           gtable_0.3.6            
 [79] tidyr_1.3.1              ScaledMatrix_1.14.0      BiocSingular_1.22.0     
 [82] XVector_0.46.0           ggrepel_0.9.6            BiocVersion_3.20.0      
 [85] foreach_1.5.2            pillar_1.10.1            spam_2.11-1             
 [88] limma_3.62.1             later_1.4.1              benchmarkmeData_1.0.4   
 [91] dplyr_1.1.4              BiocFileCache_2.14.0     lattice_0.22-6          
 [94] rtracklayer_1.66.0       bit_4.5.0.1              tidyselect_1.2.1        
 [97] locfit_1.5-9.10          scuttle_1.16.0           Biostrings_2.74.0       
[100] knitr_1.49               edgeR_4.4.1              xfun_0.50               
[103] statmod_1.5.0            DT_0.33                  UCSC.utils_1.2.0        
[106] lazyeval_0.2.2           yaml_2.3.10              shinyWidgets_0.8.7      
[109] evaluate_1.0.3           codetools_0.2-20         tibble_3.2.1            
[112] BiocManager_1.30.25      cli_3.6.3                xtable_1.8-4            
[115] jquerylib_0.1.4          munsell_0.5.1            golem_0.5.1             
[118] Rcpp_1.0.14              dbplyr_2.5.0             png_0.1-8               
[121] XML_3.99-0.17            parallel_4.4.1           blob_1.2.4              
[124] dotCall64_1.2            bitops_1.0-9             viridisLite_0.4.2       
[127] scales_1.3.0             purrr_1.0.2              crayon_1.5.3            
[130] cowplot_1.1.3            KEGGREST_1.46.0         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-03
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 !  package              * version   date (UTC) lib source
    abind                  1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
    AnnotationDbi          1.68.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    AnnotationHub          3.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
    beachmat               2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
    benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
    benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocFileCache          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocIO                 1.16.0    2024-12-12 [1] Github (Bioconductor/BiocIO@8e80624)
    BiocManager            1.30.25   2024-08-28 [1] CRAN (R 4.4.1)
    BiocNeighbors          2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel           1.40.0    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocVersion            3.20.0    2024-05-04 [1] Bioconductor 3.20 (R 4.4.0)
    Biostrings             2.74.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    bit                    4.5.0.1   2024-12-03 [1] CRAN (R 4.4.1)
    bit64                  4.5.2     2024-09-22 [1] CRAN (R 4.4.1)
    bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
    blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
    bslib                  0.9.0     2025-01-30 [1] CRAN (R 4.4.1)
    cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 P  cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
    codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
    colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1     2024-07-26 [1] CRAN (R 4.4.0)
    config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
    cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
    crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
    curl                   6.2.0     2025-01-23 [1] CRAN (R 4.4.1)
    data.table           * 1.16.4    2024-12-06 [1] CRAN (R 4.4.1)
    DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
    dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
    DelayedArray           0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    digest                 0.6.37    2024-08-19 [1] CRAN (R 4.4.1)
    doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
    dotCall64              1.2       2024-10-04 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
    DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
    edgeR                  4.4.1     2024-12-02 [1] Bioconductor 3.20 (R 4.4.2)
    escheR               * 1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    evaluate               1.0.3     2025-01-10 [1] CRAN (R 4.4.1)
    ExperimentHub          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
    fields                 16.3      2024-09-30 [1] CRAN (R 4.4.1)
    filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
    foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
    generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1    2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13    2024-12-12 [1] Bioconductor
    GenomicAlignments      1.42.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    GenomicRanges        * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
    golem                  0.5.1     2024-08-27 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
    gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
    htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
    httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
    httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
    iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
    jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
    jsonlite               1.8.9     2024-09-20 [1] CRAN (R 4.4.1)
    KEGGREST               1.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    knitr                  1.49      2024-11-08 [1] CRAN (R 4.4.1)
    labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
    later                  1.4.1     2024-11-27 [1] CRAN (R 4.4.1)
    lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
    lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
    lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
    limma                  3.62.1    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
    magick                 2.8.5     2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
    maps                   3.4.2.1   2024-11-10 [1] CRAN (R 4.4.1)
    Matrix                 1.7-2     2025-01-23 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.4.1)
    memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
    mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
    munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
    paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
    pillar                 1.10.1    2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
    plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
    png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
    promises               1.3.2     2024-11-28 [1] CRAN (R 4.4.1)
    purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
    R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
    rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
    RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14    2025-01-12 [1] CRAN (R 4.4.1)
    RCurl                  1.98-1.16 2024-07-11 [1] CRAN (R 4.4.0)
    rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
    restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
    rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4     2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown              2.29      2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
    Rsamtools              2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    RSQLite                2.3.9     2024-12-03 [1] CRAN (R 4.4.1)
    rstudioapi             0.17.1    2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
    rtracklayer            1.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Arrays               1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
    ScaledMatrix           1.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
    scater                 1.34.0    2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
    scuttle                1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
    shiny                  1.10.0    2024-12-14 [1] CRAN (R 4.4.1)
    shinyWidgets           0.8.7     2024-09-23 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    spam                   2.11-1    2025-01-20 [1] CRAN (R 4.4.1)
    SparseArray            1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    spatialLIBD          * 1.18.0    2024-11-07 [1] Bioconductor 3.20 (R 4.4.1)
    statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.50      2025-01-07 [1] CRAN (R 4.4.1)
    XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
    xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
    XVector                0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
