---
title: "01to03-Plot_initial_dimreds_by_covars"
author: "Bernie Mulvey"
date: "2024-01-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(Cairo)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load dimred mats and metadata
```{r}
metad <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")

dimr <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/02-PCA_harmo_UMAP_matrices_HiDevG_HVG_SVG_andmixed_featsets.RDS")

glmdimr <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/03-initial_HDG-SVG-HVG_sets_nullResid_to_approxGLMPCA.RDS")
```

covars to plot by: 
categorical sample_id, capture_id,  APOE, Ancestry, Sex, brnum
continuous sum_gene, sum_umi, expr_chrM_ratio, Age
```{r}
dimr <- lapply(dimr,as.data.table,keep.rownames=T)

covs <- c("sample_id", "capture_id", "APOE", "Ancestry", "Sex", "brnum", "sum_gene", "sum_umi", "expr_chrM_ratio", "Age")

bpmapply(d=dimr,f=names(dimr),SIMPLIFY=FALSE,BPPARAM = MulticoreParam(8),FUN=function(d,f){
    plotdr <- names(d)[c(1:3)] ## spot id and first 2 PCs/Harmos or 2 UMAP dims
    tmp <- copy(d)[,..plotdr]
    
    mdcol <- c("rowname",covs)
    
    tmp <- merge.data.table(tmp,metad[,..mdcol],by.x="rn",by.y="rowname")
    plts <- lapply(covs,function(x){
        if(x %in% c("sum_gene", "sum_umi", "expr_chrM_ratio", "Age")){
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=as.numeric(.data[[x]])))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("Dimred by ",x),color="none")
        } else {
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=.data[[x]]))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("Dimred by ",x))+
                guides(color="none")
        }
        
        
        q <- rasterize(p,layers="Points",dev="cairo_png",dpi=300)
        return(q)
    })
    Cairo(paste0("plots/07_featureSelection_dimred_harmony_clustering/01to03_PCA-UMAP-Harmony/stdPCA/",f,".png"),width=12,height=9,units="in",type="png",dpi=300)
    do.call("grid.arrange",c(plts,ncol=4))
    dev.off()
})

```

## same for glmpcas and descendants:
```{r}
glmdimr <- lapply(glmdimr,as.data.table,keep.rownames=T)

covs <- c("sample_id", "capture_id", "APOE", "Ancestry", "Sex", "brnum", "sum_gene", "sum_umi", "expr_chrM_ratio", "Age")

## fix names on the dimr list -- UMAPs don't indicate the upstream dimred type.
names(glmdimr)[seq(2,34,4)] <- gsub(names(glmdimr)[seq(2,34,4)],pattern="^UMAP",replacement="UMAP_glmPCA")

bpmapply(d=glmdimr,f=names(glmdimr),SIMPLIFY=FALSE,BPPARAM = MulticoreParam(8),FUN=function(d,f){
    plotdr <- names(d)[c(1:3)] ## spot id and first 2 PCs/Harmos or 2 UMAP dims

    tmp <- copy(d)[,..plotdr]
    
    mdcol <- c("rowname",covs)
    
    tmp <- merge.data.table(tmp,metad[,..mdcol],by.x="rn",by.y="rowname")
    plts <- lapply(covs,function(x){
        if(x %in% c("sum_gene", "sum_umi", "expr_chrM_ratio", "Age")){
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=as.numeric(.data[[x]])))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("Dimred by ",x),color="none")
        } else {
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=.data[[x]]))+
                geom_point(size=0.1,stroke = 0)+
                labs(title=paste0("Dimred by ",x))+
                guides(color="none")
        }
        
        
        q <- rasterize(p,layers="Points",dev="cairo_png",dpi=300)
        return(q)
    })
    Cairo(paste0("plots/07_featureSelection_dimred_harmony_clustering/01to03_PCA-UMAP-Harmony/glmPCA/",f,".png"),width=12,height=9,units="in",type="png",dpi=300)
    do.call("grid.arrange",c(plts,ncol=4))
    dev.off()
})

```

reprod
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0 parallelly_1.40.1   colorout_1.3-0.2   
 [4] Cairo_1.6-2         ggrastr_1.0.2       gridExtra_2.3      
 [7] ggtext_0.1.2        ggplot2_3.5.1       data.table_1.16.4  
[10] rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6      dplyr_1.1.4       compiler_4.4.1    tidyselect_1.2.1 
 [5] Rcpp_1.0.13-1     ggbeeswarm_0.7.2  xml2_1.3.6        parallel_4.4.1   
 [9] scales_1.3.0      yaml_2.3.10       fastmap_1.2.0     here_1.0.1       
[13] R6_2.5.1          generics_0.1.3    knitr_1.49        tibble_3.2.1     
[17] munsell_0.5.1     rprojroot_2.0.4   pillar_1.9.0      utf8_1.2.4       
[21] xfun_0.49         cli_3.6.3         withr_3.0.2       magrittr_2.0.3   
[25] digest_0.6.37     grid_4.4.1        gridtext_0.1.5    rstudioapi_0.17.1
[29] beeswarm_0.4.0    lifecycle_1.0.4   vipor_0.4.7       vctrs_0.6.5      
[33] evaluate_1.0.1    glue_1.8.0        sessioninfo_1.2.2 codetools_0.2-20 
[37] fansi_1.0.6       colorspace_2.1-1  rmarkdown_2.29    tools_4.4.1      
[41] pkgconfig_2.0.3   htmltools_0.5.8.1
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-12-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package      * version  date (UTC) lib source
   beeswarm       0.4.0    2021-06-01 [1] CRAN (R 4.4.0)
   BiocParallel * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   Cairo        * 1.6-2    2023-11-28 [1] CRAN (R 4.4.0)
 P cli            3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
   codetools      0.2-20   2024-03-31 [1] CRAN (R 4.4.1)
   colorout     * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace     2.1-1    2024-07-26 [1] CRAN (R 4.4.0)
   data.table   * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
   digest         0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
   dplyr          1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
   evaluate       1.0.1    2024-10-10 [1] CRAN (R 4.4.1)
   fansi          1.0.6    2023-12-08 [1] CRAN (R 4.4.0)
   fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
   generics       0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
   ggbeeswarm     0.7.2    2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2      * 3.5.1    2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr      * 1.0.2    2023-06-01 [1] CRAN (R 4.4.0)
   ggtext       * 0.1.2    2022-09-16 [1] CRAN (R 4.4.0)
   glue           1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
   gridExtra    * 2.3      2017-09-09 [1] CRAN (R 4.4.0)
   gridtext       0.1.5    2022-09-16 [1] CRAN (R 4.4.0)
   gtable         0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
   here           1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.49     2024-11-08 [1] CRAN (R 4.4.1)
   lifecycle      1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
   magrittr       2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
   munsell        0.5.1    2024-04-01 [1] CRAN (R 4.4.0)
   parallelly   * 1.40.1   2024-12-04 [1] CRAN (R 4.4.1)
   pillar         1.9.0    2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig      2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
   R6             2.5.1    2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp           1.0.13-1 2024-11-02 [1] CRAN (R 4.4.1)
 P rlang        * 1.1.4    2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.29     2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot      2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi     0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
   scales         1.3.0    2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo    1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
   tibble         3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect     1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
   utf8           1.2.4    2023-10-22 [1] CRAN (R 4.4.0)
   vctrs          0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
   vipor          0.4.7    2023-12-18 [1] CRAN (R 4.4.0)
   withr          3.0.2    2024-10-28 [1] CRAN (R 4.4.1)
   xfun           0.49     2024-10-31 [1] CRAN (R 4.4.1)
   xml2           1.3.6    2023-12-04 [1] CRAN (R 4.4.0)
   yaml           2.3.10   2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
