---
title: "04a-init_louvainClusters_purity"
output: html_document
date: "2024-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(ggbeeswarm)
library(Polychrome)
library(ggrastr)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
## 

library(parallel)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## plot these as demonstrated in OSCA advanced 5.2.3
```{r}
pur <- readRDS("processed-data/07_Clustering/02-SNNk10-15-20_louvain_allgeneGLMPCA_clustpurity.RDS")

## we also need the spot id-cluster label pairs for this one

SNNlouvlist <- readRDS("processed-data/07_Clustering/01-GLMPCAallgene_SNN10-15-20_leidenAndLouvain_res05-1-2.RDS")
snnlouvs <- lapply(SNNlouvlist,as.data.table)
snnlouv.dt <- merge.data.table(snnlouvs[[1]],snnlouvs[[2]],by="rn")
snnlouv.dt <- merge.data.table(snnlouv.dt,snnlouvs[[3]],by="rn")
keepclus <- c("rn",grep(names(snnlouv.dt),pattern="louv",value=T))

snnlouv.dt <- snnlouv.dt[,..keepclus]

mapply(pur,names(pur),SIMPLIFY=FALSE,FUN = function(p,n){
    p <- as.data.table(p,keep.rownames=T)
    getnames <- c("rn",n)
    clusassigns <- snnlouv.dt[,..getnames]
    setnames(clusassigns,2,"label")
    
    p <- merge.data.table(p,clusassigns,by="rn")
    
    nclus <- length(unique(p$label))
    pal <- palette36.colors(n = nclus+2)[c(3:(nclus+2))]
    names(pal) <- paste0("X",c(1:nclus))
    
    plt <- ggplot(p,aes(x=label,y=purity,col=maximum))+
        rasterize(geom_quasirandom(method="smiley"),dev="cairo_png")+
        scale_color_manual(values = pal)+
        theme(axis.text.x = element_text(angle = 90, hjust = 0.5))+
        ggtitle(n)

    pdf(paste0("plots/07_Clustering/04-Clusterpurity/",n,".pdf"),height=8,width=14)
    print(plt)
    dev.off()
})

```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] ggrastr_1.0.2       Polychrome_1.5.1    S4Vectors_0.42.1    BiocGenerics_0.50.0
 [5] colorout_1.3-0.2    ggbeeswarm_0.7.2    gridExtra_2.3       ggplot2_3.5.1      
 [9] data.table_1.15.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.5         dplyr_1.1.4          compiler_4.4.1      
 [4] tidyselect_1.2.1     png_0.1-8            scales_1.3.0        
 [7] yaml_2.3.9           fastmap_1.2.0        here_1.0.1          
[10] R6_2.5.1             labeling_0.4.3       generics_0.1.3      
[13] knitr_1.48           tibble_3.2.1         munsell_0.5.1       
[16] rprojroot_2.0.4      pillar_1.9.0         utf8_1.2.4          
[19] xfun_0.45            cli_3.6.3            withr_3.0.0         
[22] magrittr_2.0.3       digest_0.6.36        grid_4.4.1          
[25] rstudioapi_0.16.0    beeswarm_0.4.0       lifecycle_1.0.4     
[28] vipor_0.4.7          vctrs_0.6.5          scatterplot3d_0.3-44
[31] evaluate_0.24.0      glue_1.7.0           farver_2.1.2        
[34] sessioninfo_1.2.2    fansi_1.0.6          colorspace_2.1-0    
[37] rmarkdown_2.27       htmltools_0.5.8.1    tools_4.4.1         
[40] pkgconfig_2.0.3     
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-17
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package       * version date (UTC) lib source
   beeswarm        0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   BiocGenerics  * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli             3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout      * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace      2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   data.table    * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest          0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr           1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate        0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi           1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver          2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap         1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics        0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggbeeswarm    * 0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2       * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr       * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   glue            1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra     * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable          0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here            1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools       0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr           1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling        0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle       1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr        2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell         0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   pillar          1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig       2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   png             0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome    * 1.5.1   2022-05-03 [1] CRAN (R 4.4.0)
   R6              2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
 P rlang         * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown       2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot       2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi      0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Vectors     * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales          1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   scatterplot3d   0.3-44  2023-05-05 [1] CRAN (R 4.4.0)
   sessioninfo     1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   tibble          3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect      1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8            1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs           0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor           0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   withr           3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun            0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml            2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
