---
title: "04c-init_louvainCluster_RMSD"
output: html_document
date: "2025-01-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
rmsds <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/05c-init_SNNk10_louvain_clustRMSDs.RDS")

mapply(rmsds,names(rmsds),SIMPLIFY=FALSE,FUN=function(r,n){
    pdf(paste0("plots/07_featureSelection_dimred_harmony_clustering/04c_init_louv_RMSDs/",n,".pdf"),width=10,height=6)
    barplot(r,ylab="RMSD",xlab="Cluster",main=n)
    dev.off()
})

```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-0.2  data.table_1.16.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] compiler_4.4.1    here_1.0.1        fastmap_1.2.0     rprojroot_2.0.4  
 [5] cli_3.6.3         htmltools_0.5.8.1 tools_4.4.1       rstudioapi_0.17.1
 [9] yaml_2.3.10       rmarkdown_2.29    knitr_1.49        xfun_0.49        
[13] digest_0.6.37     evaluate_1.0.1   
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-01-02
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package     * version date (UTC) lib source
 P cli           3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout    * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   data.table  * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
   digest        0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
   evaluate      1.0.1   2024-10-10 [1] CRAN (R 4.4.1)
   fastmap       1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   here          1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools     0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr         1.49    2024-11-08 [1] CRAN (R 4.4.1)
 P rlang       * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown     2.29    2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot     2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi    0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
   sessioninfo   1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   xfun          0.49    2024-10-31 [1] CRAN (R 4.4.1)
   yaml          2.3.10  2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
