---
title: "01a-save_tempSingsamp_SPEs_for_nnSVG"
author: "Bernie Mulvey"
date: "2024-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(SpatialExperiment)
# library(SpatialFeatureExperiment)

```

# literally all we're doing here is saving a bunch of single-sample SPEs from which to run nnSVG without having to load the full SPE in every HPC batch job. (nnSVG runs on single samples, so we'd just be loading the entire SPE to throw 84/85ths of it back out or with 84/85ths of it occupying memory for no reason)

```{r}
# load sectionwise, filtered, lognormed SPE
lc3 <- readRDS("processed-data/06-QCed_SPE_split_to_tissSections.RDS")

caps <- unique(lc3$sample_id)
lsamps <- lapply(caps,FUN=function(x){lc3[,lc3$sample_id==x]})
names(lsamps) <- caps

mapply(s=lsamps,n=names(lsamps),SIMPLIFY=FALSE,FUN=function(s,n){saveRDS(s,paste0("code/07_featureSelection_dimred_harmony_clustering/01a_tmp_nnsvg_in/",n,".RDS"))})

write.table(names(lsamps),"code/07_featureSelection_dimred_harmony_clustering/01a_tmp_tiss_sec_sampids.txt",sep='\t',quote=F,row.names=F,col.names=F)
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0         parallelly_1.40.1          
 [3] colorout_1.3-0.2            scry_1.18.0                
 [5] nnSVG_1.10.0                scran_1.34.0               
 [7] scater_1.34.0               scuttle_1.16.0             
 [9] SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1
[11] SummarizedExperiment_1.36.0 Biobase_2.66.0             
[13] GenomicRanges_1.58.0        GenomeInfoDb_1.42.1        
[15] IRanges_2.40.1              S4Vectors_0.44.0           
[17] BiocGenerics_0.52.0         MatrixGenerics_1.18.0      
[19] matrixStats_1.4.1           gridExtra_2.3              
[21] ggtext_0.1.2                ggplot2_3.5.1              
[23] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] pbapply_1.7-2           magrittr_2.0.3          compiler_4.4.1         
 [4] vctrs_0.6.5             rdist_0.0.5             pkgconfig_2.0.3        
 [7] crayon_1.5.3            fastmap_1.2.0           magick_2.8.5           
[10] XVector_0.46.0          utf8_1.2.4              BRISC_1.0.6            
[13] rmarkdown_2.29          sessioninfo_1.2.2       UCSC.utils_1.2.0       
[16] ggbeeswarm_0.7.2        xfun_0.49               bluster_1.16.0         
[19] zlibbioc_1.52.0         beachmat_2.22.0         jsonlite_1.8.9         
[22] DelayedArray_0.32.0     irlba_2.3.5.1           parallel_4.4.1         
[25] cluster_2.1.8           R6_2.5.1                limma_3.62.1           
[28] Rcpp_1.0.13-1           knitr_1.49              Matrix_1.7-1           
[31] igraph_2.1.2            tidyselect_1.2.1        rstudioapi_0.17.1      
[34] abind_1.4-8             yaml_2.3.10             viridis_0.6.5          
[37] codetools_0.2-20        lattice_0.22-6          tibble_3.2.1           
[40] withr_3.0.2             evaluate_1.0.1          xml2_1.3.6             
[43] pillar_1.9.0            generics_0.1.3          rprojroot_2.0.4        
[46] munsell_0.5.1           scales_1.3.0            glue_1.8.0             
[49] metapod_1.14.0          tools_4.4.1             BiocNeighbors_2.0.1    
[52] ScaledMatrix_1.14.0     locfit_1.5-9.10         RANN_2.6.2             
[55] grid_4.4.1              edgeR_4.4.1             colorspace_2.1-1       
[58] GenomeInfoDbData_1.2.13 beeswarm_0.4.0          BiocSingular_1.22.0    
[61] vipor_0.4.7             cli_3.6.3               rsvd_1.0.5             
[64] fansi_1.0.6             S4Arrays_1.6.0          viridisLite_0.4.2      
[67] dplyr_1.1.4             gtable_0.3.6            digest_0.6.37          
[70] SparseArray_1.6.0       ggrepel_0.9.6           dqrng_0.4.1            
[73] rjson_0.2.23            htmltools_0.5.8.1       lifecycle_1.0.4        
[76] httr_1.4.7              here_1.0.1              statmod_1.5.0          
[79] gridtext_0.1.5         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-12-20
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version  date (UTC) lib source
   abind                  1.4-8    2024-09-12 [1] CRAN (R 4.4.1)
   beachmat               2.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.4.0)
   Biobase              * 2.66.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics         * 0.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocNeighbors          2.0.1    2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   BiocParallel         * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   BiocSingular           1.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   bluster                1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BRISC                  1.0.6    2024-10-02 [1] CRAN (R 4.4.1)
 P cli                    3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
   cluster                2.1.8    2024-12-11 [1] CRAN (R 4.4.1)
   codetools              0.2-20   2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-1    2024-07-26 [1] CRAN (R 4.4.0)
   crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
   DelayedArray           0.32.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   digest                 0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
   dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                  0.4.1    2024-05-28 [1] CRAN (R 4.4.0)
   edgeR                  4.4.1    2024-12-02 [1] Bioconductor 3.20 (R 4.4.2)
   evaluate               1.0.1    2024-10-10 [1] CRAN (R 4.4.1)
   fansi                  1.0.6    2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.42.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData       1.2.13   2024-12-12 [1] Bioconductor
   GenomicRanges        * 1.58.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ggbeeswarm             0.7.2    2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1    2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.4.1)
   ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.4.0)
   glue                   1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
   gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.4.0)
   gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.4.0)
   gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
   here                   1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7    2023-08-15 [1] CRAN (R 4.4.0)
   igraph                 2.1.2    2024-12-07 [1] CRAN (R 4.4.1)
   IRanges              * 2.40.1   2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   irlba                  2.3.5.1  2022-10-03 [1] CRAN (R 4.4.0)
   jsonlite               1.8.9    2024-09-20 [1] CRAN (R 4.4.1)
   knitr                  1.49     2024-11-08 [1] CRAN (R 4.4.1)
   lattice                0.22-6   2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.62.1   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   locfit                 1.5-9.10 2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.5    2024-09-20 [1] CRAN (R 4.4.1)
   magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                 1.7-1    2024-10-18 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   matrixStats          * 1.4.1    2024-09-08 [1] CRAN (R 4.4.1)
   metapod                1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   munsell                0.5.1    2024-04-01 [1] CRAN (R 4.4.0)
   nnSVG                * 1.10.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   parallelly           * 1.40.1   2024-12-04 [1] CRAN (R 4.4.1)
   pbapply                1.7-2    2023-06-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0    2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
   R6                     2.5.1    2021-08-19 [1] CRAN (R 4.4.0)
   RANN                   2.6.2    2024-08-25 [1] CRAN (R 4.4.1)
   Rcpp                   1.0.13-1 2024-11-02 [1] CRAN (R 4.4.1)
   rdist                  0.0.5    2020-05-04 [1] CRAN (R 4.4.0)
   rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                * 1.1.4    2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.29     2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot              2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
   rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.4.0)
   S4Arrays               1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors            * 0.44.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ScaledMatrix           1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scales                 1.3.0    2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.34.0   2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
   scran                * 1.34.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scry                 * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scuttle              * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sessioninfo            1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.28.1   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   SparseArray            1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SpatialExperiment    * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   statmod                1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.36.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   tibble                 3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.2.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   utf8                   1.2.4    2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5    2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.2    2024-10-28 [1] CRAN (R 4.4.1)
   xfun                   0.49     2024-10-31 [1] CRAN (R 4.4.1)
   xml2                   1.3.6    2023-12-04 [1] CRAN (R 4.4.0)
   XVector                0.46.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                   2.3.10   2024-07-26 [1] CRAN (R 4.4.0)
   zlibbioc               1.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
