---
title: "02-Initial_PCAs"
author: "Bernie Mulvey"
date: "2024-12-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(scran)
library(scater)
library(harmony)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### load SPE and feature sets
```{r}
lc3 <- readRDS("processed-data/06-QCed_SPE_split_to_tissSections.RDS")

feats <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/01d_HiDevG_HVG_SVG_andmixed_featsets.RDS")
```

run pca, HARMONY, and umaps thereupon; extract dimred mats and save separately
```{r}
i<-1
for (i in c(1:length(feats))){
    # standard PCAs
    set.seed(42)
    lc3 <- scater::runPCA(lc3,subset_row=feats[[i]],name=paste0("PCA_",names(feats)[i]))

    # standard UMAPs (which require PCA first)
    set.seed(42)
    lc3 <- scater::runUMAP(lc3,dimred=paste0("PCA_",names(feats)[i]),name=paste0("UMAP_",names(feats)[i]))
    
    
    # harmony runs -- we need to temporarily rename the PCA for the feature set to "PCA" for runharmony to accept it
    reducedDimNames(lc3)[which(reducedDimNames(lc3)==paste0("PCA_",names(feats)[i]))] <- "PCA"
    
    
    # data-based initialization of lambda (development ver of Harmony; requires explicit lambda=NULL)
    set.seed(42)
    lc3 <- RunHarmony(lc3,lambda=NULL,group.by.vars="sample_id",reduction.save=paste0("HARMONYlmbna_",names(feats)[i]))
    
    # restore the feature-set-specific name for the uncorrected PCA
    reducedDimNames(lc3)[which(reducedDimNames(lc3)=="PCA")] <- paste0("PCA_",names(feats)[i])
    
    
    # UMAP of data-based initialization of lambda
    set.seed(42)
    lc3 <- scater::runUMAP(lc3,dimred=paste0("HARMONYlmbna_",names(feats)[i]),name=paste0("HARMONYlmbna_",names(feats)[i],"_UMAP"))
    
}

## check that we have 8*length(feats) reduced dims on the object, plus the 3 crap 10x ones we forgot to remove beforehand
stopifnot(length(reducedDimNames(lc3))==4*length(feats))

```

### pull out and save the reducedDims; check prop variance explained by the top 50 PCs from each PCA entry (1, 5, 9...33)
```{r}
rdm <- reducedDims(lc3)
rdm <- as.list(rdm)
rdm <- rdm[grep(names(rdm),pattern="^10x",value=T,invert=T)]

ans <- lapply(seq(1,33,4),function(x){sum(attr(rdm[[x]],"percentVar"))})
names(ans) <- names(rdm)[seq(1,33,4)]
ans
# $PCA_HDG_1397
# [1] 26.14064
# 
# $PCA_nnSVG_1397
# [1] 32.82324
# 
# $PCA_HDG1391_SVGexclusive6
# [1] 26.16072
# 
# $PCA_HDG_SVG_2575
# [1] 32.79239
# 
# $PCA_HDG_SVG_7525
# [1] 28.7839
# 
# $PCA_HVG_1397
# [1] 29.46258
# 
# $PCA_HVG_SVG_5050
# [1] 28.88051
# 
# $PCA_HVG_SVG_2575
# [1] 30.42423
# 
# $PCA_HVG_SVG_7525
# [1] 30.64421

saveRDS(rdm,"processed-data/07_featureSelection_dimred_harmony_clustering/02-PCA_harmo_UMAP_matrices_HiDevG_HVG_SVG_andmixed_featsets.RDS")
```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] harmony_1.2.3               Rcpp_1.0.13-1              
 [3] BiocParallel_1.40.0         parallelly_1.40.1          
 [5] colorout_1.3-0.2            scater_1.34.0              
 [7] scran_1.34.0                scuttle_1.16.0             
 [9] SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1
[11] SummarizedExperiment_1.36.0 Biobase_2.66.0             
[13] GenomicRanges_1.58.0        GenomeInfoDb_1.42.1        
[15] IRanges_2.40.1              S4Vectors_0.44.0           
[17] BiocGenerics_0.52.0         MatrixGenerics_1.18.0      
[19] matrixStats_1.4.1           gridExtra_2.3              
[21] ggtext_0.1.2                ggplot2_3.5.1              
[23] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] magrittr_2.0.3          RcppAnnoy_0.0.22        compiler_4.4.1         
 [4] vctrs_0.6.5             pkgconfig_2.0.3         crayon_1.5.3           
 [7] fastmap_1.2.0           magick_2.8.5            XVector_0.46.0         
[10] utf8_1.2.4              rmarkdown_2.29          sessioninfo_1.2.2      
[13] UCSC.utils_1.2.0        ggbeeswarm_0.7.2        xfun_0.49              
[16] bluster_1.16.0          zlibbioc_1.52.0         beachmat_2.22.0        
[19] jsonlite_1.8.9          DelayedArray_0.32.0     irlba_2.3.5.1          
[22] parallel_4.4.1          cluster_2.1.8           R6_2.5.1               
[25] limma_3.62.1            knitr_1.49              Matrix_1.7-1           
[28] igraph_2.1.2            tidyselect_1.2.1        rstudioapi_0.17.1      
[31] abind_1.4-8             yaml_2.3.10             viridis_0.6.5          
[34] codetools_0.2-20        lattice_0.22-6          tibble_3.2.1           
[37] withr_3.0.2             evaluate_1.0.1          xml2_1.3.6             
[40] pillar_1.9.0            generics_0.1.3          rprojroot_2.0.4        
[43] munsell_0.5.1           scales_1.3.0            RhpcBLASctl_0.23-42    
[46] glue_1.8.0              metapod_1.14.0          tools_4.4.1            
[49] BiocNeighbors_2.0.1     ScaledMatrix_1.14.0     locfit_1.5-9.10        
[52] cowplot_1.1.3           grid_4.4.1              edgeR_4.4.1            
[55] colorspace_2.1-1        GenomeInfoDbData_1.2.13 beeswarm_0.4.0         
[58] BiocSingular_1.22.0     vipor_0.4.7             cli_3.6.3              
[61] rsvd_1.0.5              fansi_1.0.6             S4Arrays_1.6.0         
[64] viridisLite_0.4.2       dplyr_1.1.4             uwot_0.2.2             
[67] gtable_0.3.6            digest_0.6.37           SparseArray_1.6.0      
[70] ggrepel_0.9.6           dqrng_0.4.1             rjson_0.2.23           
[73] htmltools_0.5.8.1       lifecycle_1.0.4         httr_1.4.7             
[76] here_1.0.1              statmod_1.5.0           gridtext_0.1.5         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-12-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version  date (UTC) lib source
   abind                  1.4-8    2024-09-12 [1] CRAN (R 4.4.1)
   beachmat               2.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.4.0)
   Biobase              * 2.66.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics         * 0.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocNeighbors          2.0.1    2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   BiocParallel         * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   BiocSingular           1.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   bluster                1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 P cli                    3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
   cluster                2.1.8    2024-12-11 [1] CRAN (R 4.4.1)
   codetools              0.2-20   2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-1    2024-07-26 [1] CRAN (R 4.4.0)
   cowplot                1.1.3    2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
   DelayedArray           0.32.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   digest                 0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
   dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                  0.4.1    2024-05-28 [1] CRAN (R 4.4.0)
   edgeR                  4.4.1    2024-12-02 [1] Bioconductor 3.20 (R 4.4.2)
   evaluate               1.0.1    2024-10-10 [1] CRAN (R 4.4.1)
   fansi                  1.0.6    2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.42.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData       1.2.13   2024-12-12 [1] Bioconductor
   GenomicRanges        * 1.58.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ggbeeswarm             0.7.2    2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1    2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.4.1)
   ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.4.0)
   glue                   1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
   gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.4.0)
   gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.4.0)
   gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
   harmony              * 1.2.3    2024-11-27 [1] CRAN (R 4.4.1)
   here                   1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7    2023-08-15 [1] CRAN (R 4.4.0)
   igraph                 2.1.2    2024-12-07 [1] CRAN (R 4.4.1)
   IRanges              * 2.40.1   2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   irlba                  2.3.5.1  2022-10-03 [1] CRAN (R 4.4.0)
   jsonlite               1.8.9    2024-09-20 [1] CRAN (R 4.4.1)
   knitr                  1.49     2024-11-08 [1] CRAN (R 4.4.1)
   lattice                0.22-6   2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.62.1   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   locfit                 1.5-9.10 2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.5    2024-09-20 [1] CRAN (R 4.4.1)
   magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                 1.7-1    2024-10-18 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   matrixStats          * 1.4.1    2024-09-08 [1] CRAN (R 4.4.1)
   metapod                1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   munsell                0.5.1    2024-04-01 [1] CRAN (R 4.4.0)
   parallelly           * 1.40.1   2024-12-04 [1] CRAN (R 4.4.1)
   pillar                 1.9.0    2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
   R6                     2.5.1    2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                 * 1.0.13-1 2024-11-02 [1] CRAN (R 4.4.1)
   RcppAnnoy              0.0.22   2024-01-23 [1] CRAN (R 4.4.0)
   RhpcBLASctl            0.23-42  2023-02-11 [1] CRAN (R 4.4.0)
   rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                * 1.1.4    2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.29     2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot              2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
   rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.4.0)
   S4Arrays               1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors            * 0.44.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ScaledMatrix           1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scales                 1.3.0    2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.34.0   2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
   scran                * 1.34.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scuttle              * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sessioninfo            1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.28.1   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   SparseArray            1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SpatialExperiment    * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   statmod                1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.36.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   tibble                 3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.2.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   utf8                   1.2.4    2023-10-22 [1] CRAN (R 4.4.0)
   uwot                   0.2.2    2024-04-21 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5    2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.2    2024-10-28 [1] CRAN (R 4.4.1)
   xfun                   0.49     2024-10-31 [1] CRAN (R 4.4.1)
   xml2                   1.3.6    2023-12-04 [1] CRAN (R 4.4.0)
   XVector                0.46.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                   2.3.10   2024-07-26 [1] CRAN (R 4.4.0)
   zlibbioc               1.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
