---
title: "04-Initial_snnLeidenLouvain_clusterings"
author: "Bernie Mulvey"
date: "2024-12-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(scater)
library(scran)
library(bluster)
library(Matrix)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
lcdimrs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/02-PCA_harmo_UMAP_matrices_HiDevG_HVG_SVG_andmixed_featsets.RDS")

lcglmpca <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/03-initial_HDG-SVG-HVG_sets_nullResid_to_approxGLMPCA.RDS")

## in both of these outputs, the reduction with harmony correction for a feature set of interest is in the third of four slots for that feature set. so 3, 7, ...35.

clustdims <- lcdimrs[seq(3,35,4)]
clustdims <- c(clustdims,lcglmpca[seq(3,35,4)])

## we don't need the full objects now
rm(lcdimrs,lcglmpca)
gc(full=T)
```
### shared nearest neighbors with k=10
### leiden and louvain clustering for each k at resolutions of 0.5, 1, and 2 for each clustering approach
### need to use MAKE snngraph, not buildsnngraph for the initial neighbor graph; we can't tell the latter function we're giving it a reduced dimension matrix as a standalone object
```{r}
k <- 10

# parallelize the within-loop clustering calls, which are cpu heavy tasks
SNNlouvlist <- mapply(dr=clustdims,drn=names(clustdims),SIMPLIFY=FALSE,FUN=function(dr,drn){
    snn <- makeSNNGraph(dr, k=k,BPPARAM=MulticoreParam(6))
    set.seed(42)
    louv <- igraph::cluster_louvain(snn)
    set.seed(42)
    louv5 <- igraph::cluster_louvain(snn,resolution = 0.5)
    set.seed(42)
    louv2 <- igraph::cluster_louvain(snn,resolution=2)
    
    set.seed(42)
    leid <- igraph::cluster_leiden(snn)
    set.seed(42)
    leid5 <- igraph::cluster_leiden(snn,resolution = 0.5)
    set.seed(42)
    leid2 <- igraph::cluster_leiden(snn,resolution=2)
    
    rettab <- as.data.table(cbind(rownames(dr),
                            paste0("X",louv5$membership),
                            paste0("X",louv$membership),
                            paste0("X",louv2$membership),
                            paste0("X",leid5$membership),
                            paste0("X",leid$membership),
                            paste0("X",leid2$membership)))
    louvnames <- paste0(paste0("snn",drn,"_louv_res"),c("0.5","1","2"))
    leidnames <- paste0(paste0("snn",drn,"_leid_res"),c("0.5","1","2"))
    setnames(rettab,c("rn",louvnames,leidnames))
    return(rettab)
})

names(SNNlouvlist) <- names(clustdims)

saveRDS(SNNlouvlist,"processed-data/07_featureSelection_dimred_harmony_clustering//04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")

```

### repro inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0         parallelly_1.40.1          
 [3] colorout_1.3-0.2            Matrix_1.7-1               
 [5] bluster_1.16.0              scran_1.34.0               
 [7] scater_1.34.0               ggplot2_3.5.1              
 [9] scuttle_1.16.0              SpatialExperiment_1.16.0   
[11] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
[13] Biobase_2.66.0              GenomicRanges_1.58.0       
[15] GenomeInfoDb_1.42.1         IRanges_2.40.1             
[17] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[19] MatrixGenerics_1.18.0       matrixStats_1.4.1          
[21] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1        viridisLite_0.4.2       dplyr_1.1.4            
 [4] vipor_0.4.7             viridis_0.6.5           fastmap_1.2.0          
 [7] digest_0.6.37           rsvd_1.0.5              lifecycle_1.0.4        
[10] cluster_2.1.8           statmod_1.5.0           magrittr_2.0.3         
[13] compiler_4.4.1          tools_4.4.1             igraph_2.1.2           
[16] utf8_1.2.4              yaml_2.3.10             knitr_1.49             
[19] S4Arrays_1.6.0          dqrng_0.4.1             here_1.0.1             
[22] DelayedArray_0.32.0     abind_1.4-8             withr_3.0.2            
[25] grid_4.4.1              fansi_1.0.6             beachmat_2.22.0        
[28] colorspace_2.1-1        edgeR_4.4.1             scales_1.3.0           
[31] cli_3.6.3               rmarkdown_2.29          crayon_1.5.3           
[34] generics_0.1.3          metapod_1.14.0          rstudioapi_0.17.1      
[37] httr_1.4.7              rjson_0.2.23            ggbeeswarm_0.7.2       
[40] zlibbioc_1.52.0         XVector_0.46.0          vctrs_0.6.5            
[43] jsonlite_1.8.9          BiocSingular_1.22.0     BiocNeighbors_2.0.1    
[46] ggrepel_0.9.6           irlba_2.3.5.1           beeswarm_0.4.0         
[49] magick_2.8.5            locfit_1.5-9.10         limma_3.62.1           
[52] glue_1.8.0              codetools_0.2-20        gtable_0.3.6           
[55] UCSC.utils_1.2.0        ScaledMatrix_1.14.0     munsell_0.5.1          
[58] tibble_3.2.1            pillar_1.9.0            htmltools_0.5.8.1      
[61] GenomeInfoDbData_1.2.13 R6_2.5.1                rprojroot_2.0.4        
[64] evaluate_1.0.1          lattice_0.22-6          Rcpp_1.0.13-1          
[67] gridExtra_2.3           SparseArray_1.6.0       xfun_0.49              
[70] pkgconfig_2.0.3        
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-12-28
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version  date (UTC) lib source
   abind                  1.4-8    2024-09-12 [1] CRAN (R 4.4.1)
   beachmat               2.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.4.0)
   Biobase              * 2.66.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics         * 0.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocNeighbors          2.0.1    2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   BiocParallel         * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   BiocSingular           1.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   bluster              * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 P cli                    3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
   cluster                2.1.8    2024-12-11 [1] CRAN (R 4.4.1)
   codetools              0.2-20   2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-1    2024-07-26 [1] CRAN (R 4.4.0)
   crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
   DelayedArray           0.32.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   digest                 0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
   dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                  0.4.1    2024-05-28 [1] CRAN (R 4.4.0)
   edgeR                  4.4.1    2024-12-02 [1] Bioconductor 3.20 (R 4.4.2)
   evaluate               1.0.1    2024-10-10 [1] CRAN (R 4.4.1)
   fansi                  1.0.6    2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.42.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData       1.2.13   2024-12-12 [1] Bioconductor
   GenomicRanges        * 1.58.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ggbeeswarm             0.7.2    2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1    2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.4.1)
   glue                   1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
   gridExtra              2.3      2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
   here                   1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7    2023-08-15 [1] CRAN (R 4.4.0)
   igraph                 2.1.2    2024-12-07 [1] CRAN (R 4.4.1)
   IRanges              * 2.40.1   2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   irlba                  2.3.5.1  2022-10-03 [1] CRAN (R 4.4.0)
   jsonlite               1.8.9    2024-09-20 [1] CRAN (R 4.4.1)
   knitr                  1.49     2024-11-08 [1] CRAN (R 4.4.1)
   lattice                0.22-6   2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.62.1   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   locfit                 1.5-9.10 2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.5    2024-09-20 [1] CRAN (R 4.4.1)
   magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
   Matrix               * 1.7-1    2024-10-18 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   matrixStats          * 1.4.1    2024-09-08 [1] CRAN (R 4.4.1)
   metapod                1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   munsell                0.5.1    2024-04-01 [1] CRAN (R 4.4.0)
   parallelly           * 1.40.1   2024-12-04 [1] CRAN (R 4.4.1)
   pillar                 1.9.0    2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
   R6                     2.5.1    2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13-1 2024-11-02 [1] CRAN (R 4.4.1)
   rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                * 1.1.4    2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.29     2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot              2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
   rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.4.0)
   S4Arrays               1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors            * 0.44.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ScaledMatrix           1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scales                 1.3.0    2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.34.0   2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
   scran                * 1.34.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scuttle              * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sessioninfo            1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.28.1   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   SparseArray            1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SpatialExperiment    * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   statmod                1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.36.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   tibble                 3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.2.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   utf8                   1.2.4    2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5    2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.2    2024-10-28 [1] CRAN (R 4.4.1)
   xfun                   0.49     2024-10-31 [1] CRAN (R 4.4.1)
   XVector                0.46.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                   2.3.10   2024-07-26 [1] CRAN (R 4.4.0)
   zlibbioc               1.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
