---
title: "07-Manual_expression_based_LC_definition_for_comparison_to_louvain"
author: "Bernie Mulvey"
date: "2025-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
```

```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

### genes we want to use for determining cluster labels: DBH, TH, SLC6A2, PHOX2A, PHOX2B, DDC, SLC18A2

tmp <- lc3
rownames(tmp) <- rowData(tmp)$gene_name

tmp <- tmp[c("DBH","TH","SLC6A2","PHOX2A","PHOX2B","DDC","SLC18A2"),]

###
### hi stringency definition: ≥5 counts across DBH TH and 6A2 (all 3 required), and two counts among any others

histr <- apply(counts(tmp),MARGIN=2,simplify = FALSE,FUN = function(x){
  if (x["DBH"]>=1 & (x["TH"]>=1 | x["SLC6A2"]>=1) & sum(x["DBH"],x["TH"],x["SLC6A2"])>=5 & sum(x["PHOX2A"], x["PHOX2B"], x["DDC"], x["SLC18A2"])>=2){return("histr_LC")} else{
      return("nonLC")
  }
})
# sum(histr=="histr_LC")
# --> 6238 spots


###
### low stringency definition: ≥1 count DBH, 1 count of either TH or 6A2, and one count among any others

lostr <- apply(counts(tmp),MARGIN=2,simplify = FALSE,FUN = function(x){
  if (x["DBH"]>=1 & (x["TH"]>=1 | x["SLC6A2"]>=1) & (x["PHOX2A"]>=1 | x["PHOX2B"]>=1 | x["DDC"]>=1 | x["SLC18A2"]>=1)){return("lowstr_LC")} else{
      return("nonLC")
  }
})

### make into a table
histr <- as.data.table(histr)
histr <- t(histr)
histr <- as.data.table(histr,keep.rownames=T)

lostr <- as.data.table(lostr)
lostr <- t(lostr)
lostr <- as.data.table(lostr,keep.rownames=T)

setnames(histr,2,"histringency_LC")
setnames(lostr,2,"lostringency_LC")

manlcs <- merge.data.table(lostr,histr,by="rn")

#
fwrite(manlcs,"processed-data/07_featureSelection_dimred_harmony_clustering/07_manual_expression_defined_LC.txt",sep='\t',quote=F)

```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] colorout_1.3-0.2            SpatialExperiment_1.16.0   
 [3] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
 [5] Biobase_2.66.0              GenomicRanges_1.58.0       
 [7] GenomeInfoDb_1.42.1         IRanges_2.40.1             
 [9] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[11] MatrixGenerics_1.18.0       matrixStats_1.4.1          
[13] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] Matrix_1.7-1            jsonlite_1.8.9          rjson_0.2.23           
 [4] compiler_4.4.1          crayon_1.5.3            Rcpp_1.0.13-1          
 [7] magick_2.8.5            yaml_2.3.10             fastmap_1.2.0          
[10] lattice_0.22-6          here_1.0.1              R6_2.5.1               
[13] XVector_0.46.0          S4Arrays_1.6.0          knitr_1.49             
[16] DelayedArray_0.32.0     rprojroot_2.0.4         GenomeInfoDbData_1.2.13
[19] xfun_0.49               SparseArray_1.6.0       cli_3.6.3              
[22] magrittr_2.0.3          zlibbioc_1.52.0         digest_0.6.37          
[25] grid_4.4.1              rstudioapi_0.17.1       evaluate_1.0.1         
[28] sessioninfo_1.2.2       abind_1.4-8             rmarkdown_2.29         
[31] httr_1.4.7              tools_4.4.1             htmltools_0.5.8.1      
[34] UCSC.utils_1.2.0       
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-01-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version  date (UTC) lib source
   abind                  1.4-8    2024-09-12 [1] CRAN (R 4.4.1)
   Biobase              * 2.66.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics         * 0.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 P cli                    3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
   colorout             * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
   DelayedArray           0.32.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   digest                 0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
   evaluate               1.0.1    2024-10-10 [1] CRAN (R 4.4.1)
   fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.42.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData       1.2.13   2024-12-12 [1] Bioconductor
   GenomicRanges        * 1.58.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   here                   1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7    2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.40.1   2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   jsonlite               1.8.9    2024-09-20 [1] CRAN (R 4.4.1)
   knitr                  1.49     2024-11-08 [1] CRAN (R 4.4.1)
   lattice                0.22-6   2024-03-20 [1] CRAN (R 4.4.1)
   magick                 2.8.5    2024-09-20 [1] CRAN (R 4.4.1)
   magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                 1.7-1    2024-10-18 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   matrixStats          * 1.4.1    2024-09-08 [1] CRAN (R 4.4.1)
   R6                     2.5.1    2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13-1 2024-11-02 [1] CRAN (R 4.4.1)
   rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                * 1.1.4    2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.29     2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot              2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
   S4Arrays               1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors            * 0.44.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sessioninfo            1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.28.1   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   SparseArray            1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SpatialExperiment    * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SummarizedExperiment * 1.36.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   UCSC.utils             1.2.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   xfun                   0.49     2024-10-31 [1] CRAN (R 4.4.1)
   XVector                0.46.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                   2.3.10   2024-07-26 [1] CRAN (R 4.4.0)
   zlibbioc               1.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
