---
title: "03-Oligo_Astro_1vothers_markers"
format: html
date: "2026-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(spatialLIBD)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


## load SPE and assign hdg_svg_2575_louv_res1 clusterings as autoannotated (to start with), weber divecha lc datasets, 
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

louvs <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(louvs,2,"clusid")

anno <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots_prelim.txt")
anno <- anno[,.(clusid,anno)]

louvs <- merge.data.table(louvs,anno,by="clusid")
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$anno

rm(louvs,anno)
```

## make APOE columns tolerable as model variables; append genomwide YRI estimates
```{r}
tmpcd <- as.data.table(colData(lc3),keep.rownames=T)
tmpcd[APOE=="E4/ E4",APOE:="E4hom"]
tmpcd[APOE=="E3/ E4",APOE:="E4het"]
tmpcd[APOE=="E2/ E3",APOE:="E2het"]
tmpcd[APOE=="E2/ E2",APOE:="E2hom"]

flares<-fread("raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")
flares[ID=="Br6119",ID:="Br6119(re-dis)"]
stopifnot(sum(flares$ID %in% unique(tmpcd$brnum))==30)

tmpcd <- merge.data.table(tmpcd,flares[,.(ID,YRI)],by.x="brnum",by.y="ID")
stopifnot(nrow(tmpcd)==ncol(lc3))

tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lc3),]
tmpcd$rn <- NULL
colData(lc3) <- tmpcd

## make an oligo subobject and an astro subobject
oligs <- grep(unique(lc3$label),pattern="Olig",value=T)
lc3o <- lc3[,lc3$label %in% oligs]
asts <- grep(unique(lc3$label),pattern="Ast",value=T)
lc3a <- lc3[,lc3$label %in% asts]
rm(oligs,tmpcd,flares,asts)
```


## run registration wrapper for oligos
```{r}
regwrap.o <- registration_wrapper(lc3o,var_registration="label",var_sample_id = "sample_id",covars=c("Sex","APOE","Age","YRI"),gene_ensembl = "gene_id",gene_name="gene_name",min_ncells=10)

## save a table of one-vs-all markers
regwraptab <- as.data.table(regwrap.o$enrichment)
regwraptab <- melt.data.table(regwraptab,id.vars=c("ensembl","gene"))
regwraptab[,stat:=gsub(variable,pattern="^(logFC|fdr|p_value|t_stat)_.*$",replacement="\\1")]
regwraptab[,cluster:=gsub(variable,pattern="^(logFC|fdr|p_value|t_stat)_(.*)$",replacement="\\2")]

regwraptab[,variable:=NULL]
regwraptab2 <- dcast(regwraptab,cluster+gene+ensembl~stat)
fwrite(regwraptab2,"processed-data/13_2_markers_and_registration_final/03_2-HDGSVG2575louv1_regWrapOut_Oligo_1vOthers_markers.txt",sep="\t",quote=F,row.names=F,col.names=T)
rm(regwraptab,regwraptab2,regwrap.o)
```

and for astros
```{r}
regwrap.a <- registration_wrapper(lc3a,var_registration="label",var_sample_id = "sample_id",covars=c("Sex","APOE","Age","YRI"),gene_ensembl = "gene_id",gene_name="gene_name",min_ncells=10)

## save a table of one-vs-all markers
regwraptab <- as.data.table(regwrap.a$enrichment)
regwraptab <- melt.data.table(regwraptab,id.vars=c("ensembl","gene"))
regwraptab[,stat:=gsub(variable,pattern="^(logFC|fdr|p_value|t_stat)_.*$",replacement="\\1")]
regwraptab[,cluster:=gsub(variable,pattern="^(logFC|fdr|p_value|t_stat)_(.*)$",replacement="\\2")]

regwraptab[,variable:=NULL]
regwraptab2 <- dcast(regwraptab,cluster+gene+ensembl~stat)
fwrite(regwraptab2,"processed-data/13_2_markers_and_registration_final/03_2-HDGSVG2575louv1_regWrapOut_Astros_1vOthers_markers.txt",sep="\t",quote=F,row.names=F,col.names=T)
rm(regwraptab,regwraptab2,regwrap.a)
```

repinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] spatialLIBD_1.22.0          SpatialExperiment_1.20.0   
 [5] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0               IRanges_2.44.0             
[11] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3       rstudioapi_0.18.0        jsonlite_2.0.0          
  [4] shape_1.4.6.1            magrittr_2.0.4           ggbeeswarm_0.7.3        
  [7] magick_2.9.0             farver_2.1.2             GlobalOptions_0.1.3     
 [10] BiocIO_1.20.0            vctrs_0.7.1              memoise_2.0.1           
 [13] config_0.3.2             Rsamtools_2.26.0         paletteer_1.7.0         
 [16] RCurl_1.98-1.17          benchmarkme_1.0.8        htmltools_0.5.9         
 [19] S4Arrays_1.10.1          AnnotationHub_4.0.0      curl_7.0.0              
 [22] BiocNeighbors_2.4.0      SparseArray_1.10.8       sass_0.4.10             
 [25] bslib_0.10.0             htmlwidgets_1.6.4        httr2_1.2.2             
 [28] plotly_4.12.0            cachem_1.1.0             GenomicAlignments_1.46.0
 [31] mime_0.13                lifecycle_1.0.5          iterators_1.0.14        
 [34] pkgconfig_2.0.3          rsvd_1.0.5               Matrix_1.7-4            
 [37] R6_2.6.1                 fastmap_1.2.0            shiny_1.12.1            
 [40] clue_0.3-66              digest_0.6.39            colorspace_2.1-2        
 [43] rematch2_2.1.2           AnnotationDbi_1.72.0     scater_1.38.0           
 [46] rprojroot_2.1.1          irlba_2.3.7              ExperimentHub_3.0.0     
 [49] RSQLite_2.4.5            beachmat_2.26.0          filelock_1.0.3          
 [52] httr_1.4.7               abind_1.4-8              compiler_4.5.2          
 [55] here_1.0.2               bit64_4.6.0-1            doParallel_1.0.17       
 [58] attempt_0.3.1            S7_0.2.1                 viridis_0.6.5           
 [61] DBI_1.2.3                sessioninfo_1.2.3        rappdirs_0.3.4          
 [64] DelayedArray_0.36.0      rjson_0.2.23             tools_4.5.2             
 [67] vipor_0.4.7              otel_0.2.0               beeswarm_0.4.0          
 [70] httpuv_1.6.16            glue_1.8.0               restfulr_0.0.16         
 [73] promises_1.5.0           grid_4.5.2               rsconnect_1.7.0         
 [76] cluster_2.1.8.1          gtable_0.3.6             tidyr_1.3.2             
 [79] ScaledMatrix_1.18.0      BiocSingular_1.26.0      XVector_0.50.0          
 [82] ggrepel_0.9.6            BiocVersion_3.22.0       foreach_1.5.2           
 [85] pillar_1.11.1            limma_3.66.0             later_1.4.5             
 [88] circlize_0.4.17          benchmarkmeData_2.0.0    dplyr_1.2.0             
 [91] BiocFileCache_3.0.0      lattice_0.22-7           rtracklayer_1.70.1      
 [94] bit_4.6.0                tidyselect_1.2.1         ComplexHeatmap_2.26.1   
 [97] locfit_1.5-9.12          scuttle_1.20.0           Biostrings_2.78.0       
[100] knitr_1.51               gridExtra_2.3            edgeR_4.8.2             
[103] xfun_0.56                statmod_1.5.1            DT_0.34.0               
[106] lazyeval_0.2.2           yaml_2.3.12              shinyWidgets_0.9.0      
[109] cigarillo_1.0.0          evaluate_1.0.5           codetools_0.2-20        
[112] tibble_3.3.1             BiocManager_1.30.27      cli_3.6.5               
[115] xtable_1.8-4             jquerylib_0.1.4          golem_0.5.1             
[118] dichromat_2.0-0.1        Rcpp_1.1.1               dbplyr_2.5.1            
[121] png_0.1-8                XML_3.99-0.20            parallel_4.5.2          
[124] ggplot2_4.0.2            blob_1.3.0               bitops_1.0-9            
[127] viridisLite_0.4.2        scales_1.4.0             purrr_1.2.1             
[130] crayon_1.5.3             GetoptLong_1.1.0         rlang_1.1.7             
[133] cowplot_1.2.0            KEGGREST_1.50.0         
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-12
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   NA
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version   date (UTC) lib source
 abind                  1.4-8     2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub          4.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 attempt                0.3.1     2020-05-03 [1] CRAN (R 4.5.0)
 beachmat               2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.5.0)
 benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.5.0)
 benchmarkmeData        2.0.0     2026-01-19 [1] CRAN (R 4.5.2)
 Biobase              * 2.70.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocIO                 1.20.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27   2025-11-14 [1] CRAN (R 4.5.2)
 BiocNeighbors          2.4.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0    2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0     2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0     2026-01-14 [1] CRAN (R 4.5.2)
 bslib                  0.10.0    2026-01-26 [1] CRAN (R 4.5.2)
 cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.5.0)
 cigarillo              1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 circlize               0.4.17    2025-12-08 [1] CRAN (R 4.5.2)
 cli                    3.6.5     2025-04-23 [1] CRAN (R 4.5.0)
 clue                   0.3-66    2024-11-13 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1   2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20    2024-03-31 [1] CRAN (R 4.5.2)
 colorspace             2.1-2     2025-09-22 [1] CRAN (R 4.5.0)
 ComplexHeatmap         2.26.1    2026-01-30 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 config                 0.3.2     2023-08-30 [1] CRAN (R 4.5.0)
 cowplot                1.2.0     2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0     2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1  2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr                 2.5.1     2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39    2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0     2026-02-03 [1] CRAN (R 4.5.2)
 DT                     0.34.0    2025-09-02 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2     2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5     2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub          3.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 farver                 2.1.2     2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3     2023-12-11 [1] CRAN (R 4.5.0)
 foreach                1.5.2     2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4     2025-05-09 [1] CRAN (R 4.5.0)
 GenomicAlignments      1.46.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 GenomicRanges        * 1.62.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 GetoptLong             1.1.0     2025-11-28 [1] CRAN (R 4.5.2)
 ggbeeswarm             0.7.3     2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2                4.0.2     2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.3     2025-11-28 [1] CRAN (R 4.5.2)
 glue                   1.8.0     2024-09-30 [1] CRAN (R 4.5.0)
 golem                  0.5.1     2024-08-27 [1] CRAN (R 4.5.0)
 gridExtra              2.3       2017-09-09 [1] CRAN (R 4.5.0)
 gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2     2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9     2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.5.0)
 httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.5.0)
 httr                   1.4.7     2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2     2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7     2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14    2022-02-05 [1] CRAN (R 4.5.0)
 jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.5.0)
 jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51      2025-12-20 [1] CRAN (R 4.5.2)
 later                  1.4.5     2026-01-08 [1] CRAN (R 4.5.2)
 lattice                0.22-7    2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5     2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0     2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4     2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4     2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1     2021-11-26 [1] CRAN (R 4.5.0)
 mime                   0.13      2025-03-17 [1] CRAN (R 4.5.0)
 otel                   0.2.0     2025-08-29 [1] CRAN (R 4.5.0)
 paletteer              1.7.0     2026-01-08 [1] CRAN (R 4.5.2)
 parallelly           * 1.46.1    2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1    2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.5.0)
 plotly                 4.12.0    2026-01-24 [1] CRAN (R 4.5.2)
 png                    0.1-8     2022-11-29 [1] CRAN (R 4.5.0)
 promises               1.5.0     2025-11-01 [1] CRAN (R 4.5.0)
 purrr                  1.2.1     2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1     2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4     2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1     2026-01-10 [1] CRAN (R 4.5.2)
 RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.5.0)
 rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.5.0)
 restfulr               0.0.16    2025-06-27 [1] CRAN (R 4.5.0)
 rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7     2026-01-09 [1] CRAN (R 4.5.2)
 rprojroot              2.1.1     2025-08-26 [1] CRAN (R 4.5.0)
 Rsamtools              2.26.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 rsconnect              1.7.0     2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5     2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0    2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.5.0)
 rtracklayer            1.70.1    2025-12-20 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 S4Arrays               1.10.1    2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1     2025-11-14 [1] CRAN (R 4.5.2)
 sass                   0.4.10    2025-04-11 [1] CRAN (R 4.5.0)
 ScaledMatrix           1.18.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0     2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.5.0)
 shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.5.0)
 shiny                  1.12.1    2025-12-09 [1] CRAN (R 4.5.2)
 shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8    2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spatialLIBD          * 1.22.0    2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 statmod                1.5.1     2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0    2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1     2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2     2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1     2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5     2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.5.0)
 xfun                   0.56      2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20 2025-11-08 [1] CRAN (R 4.5.0)
 xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12    2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
