---
title: "01-Traditional_QC_Metrics"
output: html_document
date: "2024-06-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(ggrastr)
library(viridis)
library(ggtext)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

load dataset
```{r}
lc <- readRDS("processed-data/02_build_spe/spe_raw.rds")
```

spatial plots of unique # genes, unique # umis, mito %age
```{r}
samps <- unique(lc$sample_id)

## subset to tissue-overlapping spots
lc <- lc[,lc$in_tissue==TRUE]

# identify variables to plot
# colnames(colData(lc))

## we don't have logcounts yet, and these variables aren't even in an assay (these are colData()), but we need to specify one that exists to get an output.
plts <- spatialLIBD::vis_grid_gene(lc,geneid = "sum_umi",assayname = "counts",auto_crop = T, point_size = 0.575, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA,spatial=F)

plts <- lapply(plts,FUN=function(x){
    x <- x+theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4.5, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_markdown(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.375, 0, 0, 0, unit = "in"),
        panel.border=element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, 0.025, 0, -.4, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
    return(x)
})

pdf("plots/05_QC/01a-totalUMIs.pdf",height=14,width=9)
do.call("grid.arrange",c(plts,ncol=6))
dev.off()
```

## get global UMI quantiles and plot spots in the bottom 10%ile
```{r}
quantile(lc$sum_umi,probs = 0.1)
## <= 95 UMIs

# label low UMI spots
lc$low_umi <- ifelse(lc$sum_umi <= 95,yes="bottom_10pctile_UMIs",no="top_90pctile_UMIs")
lc$low_umi <- factor(lc$low_umi,levels=c("bottom_10pctile_UMIs","top_90pctile_UMIs"))

plts <- spatialLIBD::vis_grid_clus(lc,clustervar = "low_umi",auto_crop = T, point_size = 0.575, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA,spatial=F,sort_clust = F)

plts <- lapply(plts,FUN=function(x){
    x <- x+
        scale_fill_manual(labels=c("<=95 UMIs", ">95 UMIs"),values=c("red","palegreen"))+
    theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4.5, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_markdown(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.375, 0, 0, 0, unit = "in"),
        panel.border=element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, 0.025, 0, -.4, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
    return(x)
})

pdf("plots/05_QC/01b-95orfewerUMIs.pdf",height=14,width=9)
do.call("grid.arrange",c(plts,ncol=6))
dev.off()
```

### number of unique genes
```{r}
plts <- spatialLIBD::vis_grid_gene(lc,geneid = "sum_gene",assayname = "counts",auto_crop = T, point_size = 0.575, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA,spatial=F)

plts <- lapply(plts,FUN=function(x){
    x <- x+theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4.5, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_markdown(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.375, 0, 0, 0, unit = "in"),
        panel.border=element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, 0.025, 0, -.4, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
    return(x)
})

pdf("plots/05_QC/01c-uniqueGenes.pdf",height=14,width=9)
do.call("grid.arrange",c(plts,ncol=6))
dev.off()
```

### bottom 10%ile of unique genes
```{r}
quantile(lc$sum_gene,probs = 0.1)
## <= 59 UMIs

# label low UMI spots
lc$low_gene <- ifelse(lc$sum_gene <= 59,yes="bottom_10pctile_genes",no="top_90pctile_genes")
lc$low_gene <- factor(lc$low_gene,levels=c("bottom_10pctile_genes","top_90pctile_genes"))

plts <- spatialLIBD::vis_grid_clus(lc,clustervar = "low_gene",auto_crop = T, point_size = 0.575, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA,spatial=F,sort_clust = F)

plts <- lapply(plts,FUN=function(x){
    x <- x+
        scale_fill_manual(labels=c("<=59 Genes", ">59 genes"),values=c("red","palegreen"))+
    theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4.5, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_markdown(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.375, 0, 0, 0, unit = "in"),
        panel.border=element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, 0.025, 0, -.4, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
    return(x)
})

pdf("plots/05_QC/01d-59orfewerGenes.pdf",height=14,width=9)
do.call("grid.arrange",c(plts,ncol=6))
dev.off()
```

### % mito
```{r}
sum(is.na(lc$expr_chrM_ratio))

## convert NAs to 0s
lc$expr_chrM_ratio[is.na(lc$expr_chrM_ratio)] <- 0

## throw spots that are 100% mito
lc <- lc[,lc$expr_chrM_ratio!=1]

plts <- spatialLIBD::vis_grid_gene(lc,geneid = "expr_chrM_ratio",assayname = "counts",auto_crop = T, point_size = 0.575, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA,spatial=F)

plts <- lapply(plts,FUN=function(x){
    x <- x+theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4.5, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_markdown(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.375, 0, 0, 0, unit = "in"),
        panel.border=element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, 0.025, 0, -.4, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
    return(x)
})

pdf("plots/05_QC/01e-pctMito.pdf",height=14,width=9)
do.call("grid.arrange",c(plts,ncol=6))
dev.off()
```

# reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] ggtext_0.1.2                viridis_0.6.5              
 [3] viridisLite_0.4.2           BiocParallel_1.38.0        
 [5] parallelly_1.37.1           colorout_1.3-0.2           
 [7] ggrastr_1.0.2               spatialLIBD_1.16.2         
 [9] SpatialExperiment_1.14.0    SingleCellExperiment_1.26.0
[11] SummarizedExperiment_1.34.0 Biobase_2.64.0             
[13] GenomicRanges_1.56.1        MatrixGenerics_1.16.0      
[15] matrixStats_1.3.0           gridExtra_2.3              
[17] ggplot2_3.5.1               Biostrings_2.72.1          
[19] GenomeInfoDb_1.40.1         XVector_0.44.0             
[21] IRanges_2.38.0              S4Vectors_0.42.0           
[23] BiocGenerics_0.50.0         data.table_1.15.4          
[25] rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0            
  [3] bitops_1.0-7              filelock_1.0.3           
  [5] fields_15.2               tibble_3.2.1             
  [7] XML_3.99-0.16.1           lifecycle_1.0.4          
  [9] edgeR_4.2.0               doParallel_1.0.17        
 [11] rprojroot_2.0.4           lattice_0.22-6           
 [13] magrittr_2.0.3            limma_3.60.3             
 [15] plotly_4.10.4             sass_0.4.9               
 [17] rmarkdown_2.27            jquerylib_0.1.4          
 [19] yaml_2.3.8                httpuv_1.6.15            
 [21] spam_2.10-0               sessioninfo_1.2.2        
 [23] cowplot_1.1.3             DBI_1.2.3                
 [25] RColorBrewer_1.1-3        golem_0.4.1              
 [27] maps_3.4.2                abind_1.4-5              
 [29] zlibbioc_1.50.0           purrr_1.0.2              
 [31] RCurl_1.98-1.14           rappdirs_0.3.3           
 [33] GenomeInfoDbData_1.2.12   ggrepel_0.9.5            
 [35] irlba_2.3.5.1             commonmark_1.9.1         
 [37] DelayedMatrixStats_1.26.0 codetools_0.2-20         
 [39] DelayedArray_0.30.1       xml2_1.3.6               
 [41] DT_0.33                   scuttle_1.14.0           
 [43] tidyselect_1.2.1          UCSC.utils_1.0.0         
 [45] farver_2.1.2              ScaledMatrix_1.12.0      
 [47] shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [49] GenomicAlignments_1.40.0  jsonlite_1.8.8           
 [51] BiocNeighbors_1.22.0      scater_1.32.0            
 [53] iterators_1.0.14          foreach_1.5.2            
 [55] tools_4.4.1               Rcpp_1.0.12              
 [57] glue_1.7.0                SparseArray_1.4.8        
 [59] xfun_0.45                 here_1.0.1               
 [61] dplyr_1.1.4               withr_3.0.0              
 [63] BiocManager_1.30.23       fastmap_1.2.0            
 [65] fansi_1.0.6               digest_0.6.35            
 [67] rsvd_1.0.5                R6_2.5.1                 
 [69] mime_0.12                 colorspace_2.1-0         
 [71] markdown_1.13             RSQLite_2.3.7            
 [73] config_0.3.2              utf8_1.2.4               
 [75] tidyr_1.3.1               generics_0.1.3           
 [77] rtracklayer_1.64.0        httr_1.4.7               
 [79] htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [81] pkgconfig_2.0.3           gtable_0.3.5             
 [83] blob_1.2.4                htmltools_0.5.8.1        
 [85] dotCall64_1.1-1           scales_1.3.0             
 [87] png_0.1-8                 attempt_0.3.1            
 [89] knitr_1.47                rstudioapi_0.16.0        
 [91] rjson_0.2.21              curl_5.2.1               
 [93] cachem_1.1.0              stringr_1.5.1            
 [95] BiocVersion_3.19.1        parallel_4.4.1           
 [97] vipor_0.4.7               AnnotationDbi_1.66.0     
 [99] restfulr_0.0.15           pillar_1.9.0             
[101] grid_4.4.1                vctrs_0.6.5              
[103] promises_1.3.0            BiocSingular_1.20.0      
[105] dbplyr_2.5.0              beachmat_2.20.0          
[107] xtable_1.8-4              beeswarm_0.4.0           
[109] paletteer_1.6.0           evaluate_0.24.0          
[111] magick_2.8.3              cli_3.6.2                
[113] locfit_1.5-9.9            compiler_4.4.1           
[115] Rsamtools_2.20.0          crayon_1.5.3             
[117] labeling_0.4.3            rematch2_2.1.2           
[119] ggbeeswarm_0.7.2          stringi_1.8.4            
[121] munsell_0.5.1             lazyeval_0.2.2           
[123] Matrix_1.7-0              ExperimentHub_2.12.0     
[125] benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[127] bit64_4.0.5               KEGGREST_1.44.1          
[129] statmod_1.5.0             shiny_1.8.1.1            
[131] AnnotationHub_3.12.0      gridtext_0.1.5           
[133] memoise_2.0.1             bslib_0.7.0              
[135] benchmarkmeData_1.0.4     bit_4.0.5                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-24
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package              * version     date (UTC) lib source
   abind                  1.4-5       2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0      2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1       2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0       2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8       2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4       2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache          2.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23     2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular           1.20.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1      2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings           * 2.72.1      2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5       2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5       2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7       2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4       2023-03-17 [1] CRAN (R 4.4.0)
   bslib                  0.7.0       2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0       2024-05-16 [1] CRAN (R 4.4.0)
 P cli                    3.6.2       2023-12-11 [2] CRAN (R 4.4.0)
   codetools              0.2-20      2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2     2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0       2023-01-23 [1] CRAN (R 4.4.0)
   commonmark             1.9.1       2024-01-30 [1] CRAN (R 4.4.0)
   config                 0.3.2       2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3       2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3       2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1       2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4      2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3       2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0       2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1      2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.35      2024-03-11 [1] CRAN (R 4.4.0)
   doParallel             1.0.17      2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64              1.1-1       2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4       2023-11-17 [1] CRAN (R 4.4.0)
   DT                     0.33        2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0       2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0      2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6       2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2       2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0       2024-05-15 [1] CRAN (R 4.4.0)
   fields                 15.2        2023-08-17 [1] CRAN (R 4.4.0)
   filelock               1.0.3       2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2       2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3       2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1      2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12      2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1      2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2       2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1       2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr              * 1.0.2       2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5       2024-01-10 [1] CRAN (R 4.4.0)
   ggtext               * 0.1.2       2022-09-16 [1] CRAN (R 4.4.0)
   glue                   1.7.0       2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1       2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3         2017-09-09 [1] CRAN (R 4.4.0)
   gridtext               0.1.5       2022-09-16 [1] CRAN (R 4.4.0)
   gtable                 0.3.5       2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1       2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1     2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4       2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15      2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7       2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   irlba                  2.3.5.1     2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14      2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4       2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8       2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1      2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   knitr                  1.47        2024-05-29 [1] CRAN (R 4.4.0)
   labeling               0.4.3       2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.3.2       2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6      2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2       2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4       2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3      2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.9     2024-03-01 [1] CRAN (R 4.4.0)
   magick                 2.8.3       2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3       2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2       2023-12-15 [1] CRAN (R 4.4.0)
   markdown               1.13        2024-06-04 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0       2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0       2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1       2021-11-26 [1] CRAN (R 4.4.0)
   mime                   0.12        2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1       2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0       2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.37.1      2024-02-29 [1] CRAN (R 4.4.0)
   pillar                 1.9.0       2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3       2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4      2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8       2022-11-29 [1] CRAN (R 4.4.0)
   promises               1.3.0       2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2       2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1       2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3       2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3       2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.12      2024-01-09 [1] CRAN (R 4.4.0)
   RCurl                  1.98-1.14   2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2       2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15      2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21      2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4       2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27        2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4       2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7       2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0      2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5       2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   S4Arrays               1.4.1       2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sass                   0.4.9       2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0       2023-11-28 [1] CRAN (R 4.4.0)
   scater                 1.32.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2       2021-12-06 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1     2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6       2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spam                   2.10-0      2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8       2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2      2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   statmod                1.5.0       2023-01-06 [1] CRAN (R 4.4.0)
   stringi                1.8.4       2024-05-06 [1] CRAN (R 4.4.0)
   stringr                1.5.1       2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1       2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1       2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1       2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0       2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4       2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5       2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7       2023-12-18 [1] CRAN (R 4.4.0)
   viridis              * 0.6.5       2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite          * 0.4.2       2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0       2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45        2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.16.1 2024-01-22 [1] CRAN (R 4.4.0)
   xml2                   1.3.6       2023-12-04 [1] CRAN (R 4.4.0)
   xtable                 1.8-4       2019-04-21 [1] CRAN (R 4.4.0)
   XVector              * 0.44.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.8       2023-12-11 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
