---
title: "04_2-NMseg_subclustering"
author: "Bernie Mulvey"
date: "2026-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## load SPE (NM segmentation results are already embedded); set (prop_NM<=0.003 & prop_NM>0) to 0. values in the 0.002-0.003 range are the most frequent spot value for prop_NM in all clusters, so anything at or below this is noise or background.

```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

tmpcd <- as.data.table(colData(lc3),keep.rownames=T)
tmpcd[Prop_NM>0&Prop_NM<=0.003,Prop_NM:=0]
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lc3),]
# DO retain the rownames (copied to column rn above) for ease of retrieving neighbor spots downstream

colData(lc3) <- tmpcd

rm(tmpcd)
```


assign hdg_svg_2575_louv_res1 clusterings as annotated
```{r}
louvs <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(louvs,2,"clusid")

annots <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")

louvs <- merge.data.table(louvs,annots,by="clusid")
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$anno

rm(louvs,annots)
```

### for the lc cluster, make an NM+ and NM- designation
```{r}
colData(lc3) <- cbind(colData(lc3),spatialCoords(lc3))

tmpcd <- as.data.table(colData(lc3)) # rownames are already in a column now, no need to keep propagating them
tmpcd[label=="LC"&Prop_NM>0,NMseg_subclus:="LC_NMpos"]
tmpcd[label=="LC"&Prop_NM==0,NMseg_subclus:="LC_NMneg"]

## append
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lc3),]
colData(lc3) <- tmpcd
rm(tmpcd)
```


for non-LC clusters--astrocyte/vascular especially--make NM-adjacent and non-adjacent and LC adjacent/nonadjacent label sets.
### adapt a block from Michael Totty's spotsweeper package to identify immediately neighboring spots in glial clusters. this uses spatial coordinates in full pixel resolution, so append those to the coldata so we can do this without lugging around the whole SPE.
```{r}
unique_sample_ids <- unique(lc3$sample_id)
setDTthreads(1,restore_after_fork=FALSE)

## break apart coldata per sample, find neighbors, annotate accordingly, return the sample coldata with the additional columns
newcds <- lapply(unique_sample_ids,function(s){
    spe_subset <- lc3[, lc3$sample_id == s]
    dnn <- BiocNeighbors::findKNN(spatialCoords(spe_subset), 
        k = 6, warn.ties = FALSE)$index
    rownames(dnn) <- colnames(spe_subset)
    neighborhoods <- mclapply(seq_len(nrow(dnn)),mc.cores=8,FUN=function(i) {
        setDTthreads(1,restore_after_fork=FALSE)
        indices <- dnn[i, ]
        indices <- indices[indices != 0]
        as.data.table(colData(spe_subset)[indices, c("label","rn")]) # return the neighbors' spot ids (as "rn") and cluster labels
    })
    ## add the center spot id as the names of the list output
    names(neighborhoods) <- rownames(dnn)
    
    ## collapse into a data table with the center spot's id propagated in
    neighborhoods <- rbindlist(neighborhoods,idcol="centerspot_rn")

    neighborhoods[,centerspot_type:=colData(spe_subset)[neighborhoods$centerspot_rn,"NMseg_subclus"]]
    
    ## annotate any form of LC-adjacency for the non-LC spots
    neighborhoods[label!="LC" & centerspot_type %in% c("LC_NMpos","LC_NMneg"),LCadjacent:=paste0(label,"_LCadj")]
    
    neighborhoods[label!="LC" & !(centerspot_type %in% c("LC_NMpos","LC_NMneg")),LCadjacent:=paste0(label,"_LCnonadj")]
    
    ## further annotate NM+ LC adjacency/nonadjacency 
    neighborhoods[!is.na(LCadjacent) & centerspot_type=="LC_NMpos",nonLC_NMseg_subclus:=paste0(label,"_LCNMadj")]
    neighborhoods[!is.na(LCadjacent) & centerspot_type=="LC_NMneg",nonLC_NMseg_subclus:=paste0(label,"_LCnonNMadj")]
    
    ## among spots with >1 LC neighbor, there could be NM+ and NM- neighbors. if all are NM-, keep that label. if there are both, keep the NM+ label.
    
    newcdout <- unique(
        neighborhoods[LCadjacent %in% grep(LCadjacent,pattern="_LCadj",value=T) &
                          nonLC_NMseg_subclus %in% grep(nonLC_NMseg_subclus,pattern="_LCNMadj",value=T),
                      .(rn,LCadjacent,nonLC_NMseg_subclus)])
    
    stopifnot(nrow(newcdout)==length(unique(newcdout$rn)))
    
    ## get remaining lc adjacents
    newcdout <- rbind(newcdout,
                      unique(neighborhoods[LCadjacent %in% grep(LCadjacent,pattern="_LCadj",value=T) & !(rn %in% newcdout$rn),
                                           .(rn,LCadjacent,nonLC_NMseg_subclus)]))
    
    stopifnot(nrow(newcdout)==length(unique(newcdout$rn)))

    
    ## get everything else
    newcdout <- rbind(newcdout,
                      unique(neighborhoods[!(rn %in% newcdout$rn),
                                           .(rn,LCadjacent,nonLC_NMseg_subclus)]))
    
    ## finally, set the nonLC_NMseg_subclus to be blah_LCnonadj for totally LC-non-neighboring spots
    newcdout[is.na(nonLC_NMseg_subclus),nonLC_NMseg_subclus:=LCadjacent]
    
    stopifnot(nrow(newcdout)==length(unique(newcdout$rn)))

    return(newcdout)
})

## collapse it all back together and check for any neighborless spots that weren't accounted for here ^
newcdout <- rbindlist(newcds)

stopifnot(nrow(newcdout)==ncol(lc3))
# nrow(newcdout)
# so we're a few spots short. nothing we can't address with an all.x=T in the merge
tmpcd <- as.data.table(colData(lc3))

tmpcd <- merge.data.table(tmpcd,newcdout,by="rn",all.x=T)
stopifnot(all(colnames(lc3) %in% tmpcd$rn))

rm(newcdout,newcds)
## make the NMseg subclusters for LC and nonLC into one joint column
tmpcd[!is.na(NMseg_subclus),NMseg_subclus2:=NMseg_subclus]
tmpcd[!is.na(nonLC_NMseg_subclus),NMseg_subclus2:=nonLC_NMseg_subclus]

tmpcd[,NMseg_subclus:=NULL]
tmpcd[,nonLC_NMseg_subclus:=NULL]
setnames(tmpcd,"NMseg_subclus2","NMseg_subclus")

### 

stopifnot(length(unique(tmpcd$rn))==ncol(lc3))
stopifnot(all(tmpcd$rn %in% colnames(lc3)))
```

## save a spot id / main cluster / NM seg / LCadjacency labels table
```{r}
# how many spots are in each LCadjacent group and each LCNMadjacent group for the astrocytic clusters and vascular cluster?

cdout <- tmpcd[,.(rn,label,LCadjacent,NMseg_subclus)]
# cdout[,.N,by="LCadjacent"]
# ACHE_SERT_Npep_LCadj  2161
# ACHE_SERT_LC_WM_LCadj  2092
# Astro_Vasc_Mural_LCadj 2938  
# Astro_Oligo_LCadj  900
# Oligo_1_LCadj   594
# Oligo_2_LCadj   372
# Oligo_Astro_LCadj   751
# Oligo_Neuron_LCadj  724
# Vascular_LCadj   177

fwrite(cdout,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              SpatialExperiment_1.20.0   
 [5] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0               IRanges_2.44.0             
[11] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0           gridExtra_2.3              
[17] ggtext_0.1.2                ggplot2_4.0.2              
[19] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] SparseArray_1.10.8  xml2_1.5.2          lattice_0.22-7      digest_0.6.39      
 [5] magrittr_2.0.4      evaluate_1.0.5      grid_4.5.2          RColorBrewer_1.1-3 
 [9] fastmap_1.2.0       Matrix_1.7-4        rprojroot_2.1.1     scales_1.4.0       
[13] codetools_0.2-20    abind_1.4-8         cli_3.6.5           rlang_1.1.7        
[17] XVector_0.50.0      BiocNeighbors_2.4.0 DelayedArray_0.36.0 withr_3.0.2        
[21] yaml_2.3.12         otel_0.2.0          S4Arrays_1.10.1     tools_4.5.2        
[25] dplyr_1.2.0         here_1.0.2          vctrs_0.7.1         R6_2.6.1           
[29] magick_2.9.0        lifecycle_1.0.5     pkgconfig_2.0.3     pillar_1.11.1      
[33] gtable_0.3.6        rsconnect_1.7.0     glue_1.8.0          Rcpp_1.1.1         
[37] xfun_0.56           tibble_3.3.1        tidyselect_1.2.1    rstudioapi_0.18.0  
[41] knitr_1.51          dichromat_2.0-0.1   rjson_0.2.23        farver_2.1.2       
[45] htmltools_0.5.9     rmarkdown_2.30      compiler_4.5.2      S7_0.2.1           
[49] gridtext_0.1.5     
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-13
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
