---
title: "07_2-NMF decomp LC for GSEA weights"
author: "Bernie Mulvey"
date: "2026-02-14"
output: html_document
---

```{r}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(SpatialExperiment)
library(RcppML)
library(data.table)
library(Matrix)

```

setup
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

### annotate and subset to LC spots
clus <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(clus,2,"clusid")
anno <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")

clus <- merge.data.table(clus,anno,by="clusid",all.x=T)
clus <- DataFrame(clus,row.names=clus$rn)[colnames(lc3),]

lc3$label <- clus$anno

lc3 <- lc3[,lc3$label=="LC"]

lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

lclct <- logcounts(lc3)
lcrn <- rownames(lclct)
lccn <- colnames(lclct)

# get gene metadata to add symbol info later
rddt <- as.data.table(rowData(lc3))
```


## run NMF using Erik Nelson's final parameter set for some of these args (1k max iters, L1 penalty 0.1, diag = T) https://github.com/LieberInstitute/spatial_hpc/blob/main/snRNAseq_hpc/code/nmf/nmf.R

but let's set tol lower for more precision, which we can afford since we only have a few thousands spots here, and contempl8 a granular k
```{r}
setDTthreads(1,restore_after_fork = FALSE)
rm(lc3,anno,clus)
gc(full=T)
# let's suppose 10 factors per group that are distinct in each group. 2 sex * 4 apoe * 2 predominant ancestries = 16 groupings possible (12 represented) * 10 = 120 factors. stringency-ize tol from 10-6 to 10-9.
options(RcppML.threads = 11,RcppML.verbose = TRUE)

lc.nmf <-nmf(lclct,
    k=120,
    tol = 1e-09, 
    maxit = 10000,
    verbose = T,
    L1 = 0.1,
    seed = 42,
    mask_zeros = FALSE,
    diag = TRUE,
    nonneg = TRUE
)

w <- as.data.frame(lc.nmf$w)
rownames(w) <- lcrn
colnames(w) <- paste0("LCnmf",c(1:120))

h <- as.data.frame(lc.nmf$h)
rownames(h) <- paste0("LCnmf",c(1:120))
colnames(h) <- colnames(lclct)

lc.nmf2 <- list(w=w,h=h)
saveRDS(lc.nmf2,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/07_2-lcnmf120_tol1e-9.RDS")

rm(lc.nmf,lc.nmf2,w,h)
gc(full=T)

```
 convergence not reached in 5000 iterations (with k=160)
  (actual tol = 4.30e-07, target tol = 1.00e-09)
  with k=120: 5643 | 9.97e-10
  
  
### second run: same parameters, but dropping low/no detection genes. set back up:
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

### annotate and subset to LC spots
clus <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(clus,2,"clusid")
anno <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")

clus <- merge.data.table(clus,anno,by="clusid",all.x=T)
clus <- DataFrame(clus,row.names=clus$rn)[colnames(lc3),]

lc3$label <- clus$anno

lc3 <- lc3[,lc3$label=="LC"]

lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

# get gene metadata to add symbol info later
rddt <- as.data.table(rowData(lc3)) 
```

chuck low expr genes:
```{r}
check <- hist(log10(rowSums(logcounts(lc3))),breaks=seq(-2,8,0.25))
plot(check)
## looks like the minimum we want is somewhere around log10 1 to log10 2
View(cbind(check$breaks[2:length(check$breaks)],check$counts))
# 0.75-1: 1392; 1-1.25: 1254 
# <10^1.25 is the last major drop-off point, so let's set total gene logcounts 10^1.25 as the cutoff
dev.off()

## drop genes w less than 56 total counts
keepg <- rownames(lc3)[rowSums(logcounts(lc3))>=10^1.25]
length(keepg) ## 15170
lc3 <- lc3[keepg,]

```

## get new norms, logcount matrix, rownames colnames and gene info
```{r}
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)
lclct2 <- logcounts(lc3)
lcrn <- rownames(lclct2)
lccn <- colnames(lclct2)

# get gene metadata to add symbol info later
rddt <- as.data.table(rowData(lc3)) 

## clean up for second run
rm(lc3,keepg,check,lclct,anno,clus)
gc(full=T)
```

run again
# let's suppose 10 factors per group that are distinct in each group. 2 sex * 4 apoe * 2 predominant ancestries = 16 groupings (12 represented) * 10 = 120 factors. stringency-ize tol from 10-6 to 10-9.
```{r}
setDTthreads(1,restore_after_fork = FALSE)
options(RcppML.threads = 11)

lc.nmf.filt <-nmf(lclct2,
    k=120,
    tol = 1e-09, 
    maxit = 10000,
    verbose = T,
    L1 = 0.1,
    seed = 42,
    mask_zeros = FALSE,
    diag = TRUE,
    nonneg = TRUE
)
```
### didn't hit 1e-9 here with 10k iter--ended at 10000|9.78e-08

```{r}
w <- as.data.frame(lc.nmf.filt$w)
rownames(w) <- lcrn
colnames(w) <- paste0("LCnmf_filt",c(1:120))

h <- as.data.frame(lc.nmf.filt$h)
rownames(h) <- paste0("LCnmf_filt",c(1:120))
colnames(h) <- colnames(lclct2)

lc.nmf.filt2 <- list(w=w,h=h)
saveRDS(lc.nmf.filt2,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/07b-lcnmf120_tol1e-9_rowSumLogctFilt18.RDS")
```


repinfo
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] Matrix_1.7-4                data.table_1.18.2.1        
 [3] RcppML_0.5.6                SpatialExperiment_1.20.0   
 [5] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0               IRanges_2.44.0             
[11] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0          

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        beeswarm_0.4.0      rjson_0.2.23        xfun_0.56          
 [5] ggplot2_4.0.2       ggrepel_0.9.6       lattice_0.22-7      vctrs_0.7.1        
 [9] tools_4.5.2         parallel_4.5.2      tibble_3.3.1        pkgconfig_2.0.3    
[13] BiocNeighbors_2.4.0 RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5    
[17] compiler_4.5.2      farver_2.1.2        codetools_0.2-20    vipor_0.4.7        
[21] htmltools_0.5.9     yaml_2.3.12         pillar_1.11.1       rsconnect_1.7.0    
[25] BiocParallel_1.44.0 DelayedArray_0.36.0 magick_2.9.0        viridis_0.6.5      
[29] abind_1.4-8         tidyselect_1.2.1    rsvd_1.0.5          digest_0.6.39      
[33] BiocSingular_1.26.0 dplyr_1.2.0         rprojroot_2.1.1     fastmap_1.2.0      
[37] grid_4.5.2          here_1.0.2          cli_3.6.5           SparseArray_1.10.8 
[41] scater_1.38.0       magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1  
[45] scales_1.4.0        ggbeeswarm_0.7.3    rmarkdown_2.30      XVector_0.50.0     
[49] otel_0.2.0          gridExtra_2.3       ScaledMatrix_1.18.0 beachmat_2.26.0    
[53] evaluate_1.0.5      knitr_1.51          viridisLite_0.4.3   irlba_2.3.7        
[57] rlang_1.1.7         Rcpp_1.1.1          glue_1.8.0          scuttle_1.20.0     
[61] pkgload_1.5.0       rstudioapi_0.18.0   R6_2.6.1           
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-14
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beachmat               2.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel           1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2                4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra              2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7    2026-01-30 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix               * 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 pkgload                1.5.0    2026-02-03 [1] CRAN (R 4.5.2)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 RcppML               * 0.5.6    2026-02-07 [1] Github (zdebruine/RcppML@5449a5b)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5    2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3    2026-02-04 [1] CRAN (R 4.5.2)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
