---
title: "09_2-Enrichr_LCnmfFactorSpecificGenes"
author: "Bernie Mulvey"
date: "2026-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(enrichR)
library(SpatialExperiment)
library(biomaRt)
library(stringr)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

Regardless, we'll run all of the factors with specific genes from both unfilterd and pre-filtered NMF runs and see what there is to see. 

```{r}
unfilt <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens4_DEanalyses_removedsampsAndFinalNMseg/06_2-LCNMF_factors_and_factorSpecGenes_min25specGenes.txt")

filt <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens4_DEanalyses_removedsampsAndFinalNMseg/06b_2_LCnmfLogCt18Filt_factors_and_factorSpecGenes_min25specGenes.txt")

### we actually want the full NMF matrix and we need the SPE rowData to convert to entrez symbols
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")
rddt <- as.data.table(rowData(lc3))
rm(lc3)

### biomart query
biomart <- useMart("ensembl",host="https://useast.ensembl.org/",dataset="hsapiens_gene_ensembl")
lut <- biomaRt::getBM(attributes=c("ensembl_gene_id",
                                   "entrezgene_id",
                                   "entrezgene_accession",
                                   "entrezgene_description"),
                      mart=biomart,
                      filter="ensembl_gene_id",
                      values=unique(rddt$gene_id)) # uh ok "accession" = symbol

lut <- as.data.table(lut)
### this appears to have added some rows historic accessions so let's fix that this way
lut2 <- merge.data.table(lut,rddt,by.x="ensembl_gene_id",by.y="gene_id")
multig <- lut2[,.N,by=c("ensembl_gene_id")][N>1,ensembl_gene_id]

## now get the subset with multiple accessions per ensembl ID, pare down to those that match the gene names in the SPE.
lut3 <- lut2[ensembl_gene_id %in% multig&entrezgene_accession==gene_name]
## double check this ^ for still duped ensgs
lut3[,.N,by="ensembl_gene_id"][N>1] # none

## drop the duplicated gene ids altogether from the LUT
lut2 <- lut2[!(ensembl_gene_id %in% multig)]
# and now append back those where we had a matched gene symbol/accession.
lut2 <- as.data.table(rbind(lut2,lut3))

rddt <- copy(lut2)
rm(biomart,lut,lut2,lut3,multig)


## append these to the factor spec genes
unfilt <- merge.data.table(unfilt,rddt[,.(ensembl_gene_id,entrezgene_accession)],by.y="ensembl_gene_id",by.x="gene_id")

filt <- merge.data.table(filt,rddt[,.(ensembl_gene_id,entrezgene_accession)],by.y="ensembl_gene_id",by.x="gene_id")

## temporarily load the NMF matrices to get the background genes for filt and unfilt
nmf <- readRDS("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/07_2-lcnmf120_tol1e-9.RDS")
nmfbg <- rddt[ensembl_gene_id %in% rownames(nmf$w)&entrezgene_accession!="",unique(entrezgene_accession)]
rm(nmf)

nmffilt <- readRDS("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/07b-lcnmf120_tol1e-9_rowSumLogctFilt18.RDS")
filtbg <- rddt[ensembl_gene_id %in% rownames(nmffilt$w)&entrezgene_accession!="",unique(entrezgene_accession)]
rm(nmffilt)

### clean up
rm(lut,dupg,rddt)
```

### enrichr setup
## rummagenes don't work here unfortunately.
```{r}
checkdbs <- enrichR::listEnrichrDbs()
enrichr_dbs <- grep(checkdbs$libraryName,pattern="Aging|Allen|ARCHS4|Azimuth|BioCarta_2016|BioPlex|CellMarker|ChEA_2022|CORUM|DepMap_CRISPR_GeneDependency_CellLines_2023|Descartes|DGIdb_Drug_Targets_2024|Disease_Perturbations|Disease_Signatures|Drug_Perturbations|DrugMatrix|Elsevier_Pathway|ENCODE_Histone_Modifications_2015|ENCODE_TF_ChIP-seq_2015|Roadmap|FANTOM6|Gene_Perturbations|GO_Biological_Process_2025|GO_Cellular_Component_2025|GO_Molecular_Function_2025|HMS_LINCS|HuBMAP|huMAP|IDG_Drug_Targs_2022|Jensen_DISEASES_Curated_2025|KEGG_2021_Human|Kinase_Perturbations|KOMP2|L1000|Ligand_Perturbations|lncHUB|MAGNET|MGI_Mammalian_Phenotype_Level_4_2024|NIBR_DRUGseq_2025|PFOCR_Pathways_2023|PPI_Hub_Proteins|Sciplex_Drug_Perturbation_Signatures_2025|Proteomics|Reactome_Pathways_2024|SILAC|SubCell|SynGO_2024|Tabula_Muris|Tabula_Sapiens|TF_Perturbations|TF-LOF_Expression|TG_GATES_2020|The_Kinase_Library_2024|Tissue_Protein_Expression|Transcription_Factor_PPIs|TRANSFAC_and_JASPAR|TRRUST|WikiPathways_2024_Human|^DSigDB",value=T)

## filter out the ARCHS4 predictions re: rare disease, NIH PIs--interesting but not useful here
enrichr_dbs <- grep(enrichr_dbs,pattern="AutoRIF|GeneRIF",value=T,invert=T)

rm(checkdbs)
```

Interestingly, enrichr allows for the gene sets to have a "membership value of 0-1" in the second column. we COULD test this later on by adjusting the w matrix of the factor specific-genes to range 0-1 by subsetting to a table of those genes and that factor's values, then dt[,LCnmfX_scaled:=(LCnmfX-min(LCnmfX))/(max(LCnmfX)-min(LCnmfX))]

but for now, let's just feed the genes in as a straight-up vector (against the whole set of genes fed into the NMF run.)
```{r}
## need to run serially, otherwise server will start blocking requests and/or gene sets requested will get mixed up. plus other unexpected errors
unfilt.res <- list()
i <- 1
for (i in c(15:length(unique(unfilt$factor)))){
    f <- unique(unfilt$factor)[i]
    tmpunfilt <- copy(unfilt[factor==f])
    
    ## drop any dbs that throw weird factor-geneset-specific errors
    if(f=="LCnmf24"){
        cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="Elsevier_Pathway_Collection"]
    } else {
        cur_enrichr_dbs <- enrichr_dbs
    }

    if(f=="LCnmf96"|f=="LCnmf2"|f=="LCnmf22"){
        cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="HMS_LINCS_KinomeScan"]
    }
    
    if(f=="LCnmf4"){
        cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="Tabula_Sapiens"]
    }
    
    factlist <- enrichr(genes=tmpunfilt[factor==f,entrezgene_accession],
                        background = nmfbg,
                        databases=cur_enrichr_dbs)
    
  
    
    factlist.dt <- rbindlist(factlist,idcol="Enrichr_database")
    factlist.dt[,LCnmf_unfilt_factor:=f]
    
    unfilt.res[[i]] <- factlist.dt
    names(unfilt.res)[i] <- f
    
    rm(f,factlist.dt,factlist,tmpunfilt,cur_enrichr_dbs)
    gc(full=T)
}
rm(i)

#concat and save
unfilt.res2 <- rbindlist(unfilt.res)
# tabulate how many genes were hit in the set
unfilt.res2[,N_genes:=apply(.SD,MARGIN=1,simplify=FALSE,FUN=function(x){
  max(c(1,length(str_split_fixed(x,";",n=Inf)))) # if no separator, 1 gene matched
}),.SDcols="Genes"]

# get rid of deprecated columns
unfilt.res2[,Old.P.value:=NULL]
unfilt.res2[,Old.Adjusted.P.value:=NULL]

saveRDS(unfilt.res2,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/09_2-LCnmfUNfilt_factorSpecGenes_enrichr_results.RDS")
###
```

cleanup
```{r}
rm(unfilt,unfilt.res,unfilt.res2,nmfbg)
gc(full=T)
```

### pre-filtered-genes NMF factors
```{r}
filt.res <- list()
i<-1
for (i in c(17:length(unique(filt$factor)))){
    f <- unique(filt$factor)[i]
    tmpfilt <- copy(filt[factor==f])
    
    # preempt weird one-off factor-db pairs that throw errors
    if(f=="LCnmf_filt26"){
        cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="Elsevier_Pathway_Collection"]
    } else {
        cur_enrichr_dbs <- enrichr_dbs
    }
    # 
    if(f=="LCnmf_filt26"|f=="LCnmf_filt92"|f=="LCnmf_filt21"){
        cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="CellMarker_2024"]
    }
    # 
    if(f=="LCnmf_filt92"|f=="LCnmf_filt2"|f=="LCnmf_filt9"|f=="LCnmf_filt21"|f=="LCnmf_filt4"){
        cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="HMS_LINCS_KinomeScan"]
    }
    
    if(f=="LCnmf_filt3"|f=="LCnmf_filt9"){
        cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="Tabula_Sapiens"]
    }
    
    
    factlist <- enrichr(genes=tmpfilt[factor==f,entrezgene_accession],
                        background = filtbg,
                        databases=cur_enrichr_dbs)
    
    factlist.dt <- rbindlist(factlist,idcol="Enrichr_database")
    factlist.dt[,LCnmf_filt_factor:=f]
    
    filt.res[[i]] <- factlist.dt
    names(filt.res)[i] <- f
    
    rm(f,factlist.dt,factlist,tmpfilt)
}
rm(i)

# concat and save
filt.res2 <- rbindlist(filt.res)
# tabulate how many genes were hit in the set
filt.res2[,N_genes:=apply(.SD,MARGIN=1,simplify=FALSE,FUN=function(x){
  max(c(1,length(str_split_fixed(x,";",n=Inf)))) # if no separator, 1 gene matched
}),.SDcols="Genes"]

# get rid of deprecated columns
filt.res2[,Old.P.value:=NULL]
filt.res2[,Old.Adjusted.P.value:=NULL]
saveRDS(filt.res2,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/09_2-LCnmfFilt_factorSpecGenes_enrichr_results.RDS")
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              stringr_1.6.0              
 [5] biomaRt_2.66.0              SpatialExperiment_1.20.0   
 [7] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [9] Biobase_2.70.0              GenomicRanges_1.62.1       
[11] Seqinfo_1.0.0               IRanges_2.44.0             
[13] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[15] generics_0.1.4              MatrixGenerics_1.22.0      
[17] matrixStats_1.5.0           enrichR_3.4                
[19] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1     dplyr_1.2.0          farver_2.1.2        
 [4] blob_1.3.0           filelock_1.0.3       Biostrings_2.78.0   
 [7] S7_0.2.1             fastmap_1.2.0        BiocFileCache_3.0.0 
[10] digest_0.6.39        lifecycle_1.0.5      KEGGREST_1.50.0     
[13] RSQLite_2.4.5        magrittr_2.0.4       compiler_4.5.2      
[16] rlang_1.1.7          progress_1.2.3       tools_4.5.2         
[19] yaml_2.3.12          knitr_1.51           prettyunits_1.2.0   
[22] S4Arrays_1.10.1      bit_4.6.0            curl_7.0.0          
[25] here_1.0.2           DelayedArray_0.36.0  xml2_1.5.2          
[28] RColorBrewer_1.1-3   abind_1.4-8          rsconnect_1.7.0     
[31] withr_3.0.2          purrr_1.2.1          grid_4.5.2          
[34] ggplot2_4.0.2        scales_1.4.0         dichromat_2.0-0.1   
[37] cli_3.6.5            rmarkdown_2.30       crayon_1.5.3        
[40] otel_0.2.0           rstudioapi_0.18.0    httr_1.4.7          
[43] rjson_0.2.23         sessioninfo_1.2.3    DBI_1.2.3           
[46] cachem_1.1.0         parallel_4.5.2       AnnotationDbi_1.72.0
[49] XVector_0.50.0       WriteXLS_6.8.0       vctrs_0.7.1         
[52] Matrix_1.7-4         hms_1.1.4            bit64_4.6.0-1       
[55] magick_2.9.0         glue_1.8.0           codetools_0.2-20    
[58] stringi_1.8.7        gtable_0.3.6         tibble_3.3.1        
[61] pillar_1.11.1        rappdirs_0.3.4       htmltools_0.5.9     
[64] R6_2.6.1             dbplyr_2.5.1         httr2_1.2.2         
[67] rprojroot_2.1.1      evaluate_1.0.5       lattice_0.22-7      
[70] png_0.1-8            memoise_2.0.1        Rcpp_1.1.1          
[73] SparseArray_1.10.8   xfun_0.56            pkgconfig_2.0.3     
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-14
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache          3.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 biomaRt              * 2.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0    2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1  2025-01-16 [1] CRAN (R 4.5.0)
 blob                   1.3.0    2026-01-14 [1] CRAN (R 4.5.2)
 cachem                 1.1.0    2024-05-16 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0    2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3    2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr                 2.5.1    2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 enrichR              * 3.4      2025-02-02 [1] CRAN (R 4.5.0)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3    2023-12-11 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggplot2                4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4    2025-10-17 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 httr                   1.4.7    2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2    2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1    2021-11-26 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 png                    0.1-8    2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0    2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3    2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1    2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4    2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5    2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 stringi                1.8.7    2025-03-27 [1] CRAN (R 4.5.0)
 stringr              * 1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 WriteXLS               6.8.0    2025-05-22 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
