---
title: "01_2-LC samples only-hiDevGene_HVG_andSaveFeatSets_nobatch"
author: "Bernie Mulvey"
date: "2024-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(scater)
library(scran)
library(scry) # get null residuals
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

# load sectionwise, filtered, lognormed SPE
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")
```

# Right now, we do NOT want to incorporate any batch effects in any of these analyses -- we won't be able to determine what to correct for until we get thru dimensionality reduction.

###

# Feature type 1: binomial gene deviance residuals (an HVG-alike method that uses non-log normalized data. package scry). We already calculated high-deviance genes without covariates, covarying by slide, or covarying by donor in code/04-QC script #5. Since the spots themselves (and for that matter, their pairings to covariates of donor or slide) remain unchanged by splitting the capture areas in half, so we can (thankfully, as these calculations required on the order of 600 GB RAM) use those for identifying HDGs.

In other words, as Jacqui (power user of HDGs) says affirming this approach: "because you didn't specify any covariates/batch for the devianceFeatureSelection it doesn't matter if you did or did not split the capture area since the model was agnostic to the origin of any and all spots"

```{r}
devgen <- fread("processed-data/04_QC/05_2-LCsamplesonly_binomial_gene_deviances_panSPE_nobatch.txt")
## specific MTRNRLs are clearly a function of donor, per Jacqui and previously examined in this data as well: MTRNR2L1, MTRNR2L12, MTRNR2L8, so drop those and re-rank
devgen <- devgen[!gene_name %in% c("MTRNR2L1","MTRNR2L8","MTRNR2L12")]
devgen[,rank_nobatch:=rank(dev_nobatch)]
devgen[,rank_nobatch:=1+nrow(devgen)-rank_nobatch]

## for now we only want the deviances/ranks under a no batch condition
devgen <- devgen[,.(gene_id,gene_name,dev_nobatch,rank_nobatch)]
```

# Feature type 2: HVGs (typical approach, uses logcounts). Again, no batching variable to begin with. (note: if we specify sample_id as the batch, 920 of the top 1000 are the same as no batch specified)
```{r}
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- logNormCounts(lc3)

lv <- modelGeneVar(lc3)
## get all variable genes for now
hvg <- scran::getTopHVGs(lv,prop=1.0)

## specific MTRNRLs are clearly a function of donor, per Jacqui and previously examined in this data as well: MTRNR2L1, MTRNR2L12, MTRNR2L8, so drop those
# rowData(lc3)[rowData(lc3)$gene_name %in% c("MTRNR2L1","MTRNR2L8","MTRNR2L12"),"gene_id"] --> "ENSG00000269028", "ENSG00000255823","ENSG00000256618"

hvgsub <- hvg[hvg!=c("ENSG00000269028", "ENSG00000255823","ENSG00000256618")]
hvgtab <- cbind(hvgsub,seq(1,length(hvgsub),1))
colnames(hvgtab) <- c("gene_id","HVG_rank")
hvgtab <- as.data.table(hvgtab)

hvgtab <- merge.data.table(as.data.table(hvgtab),as.data.table(rowData(lc3))[,.(gene_id,gene_name)],by="gene_id")
hvgtab[,HVG_rank:=as.numeric(HVG_rank)]


fwrite(hvgtab,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/01a_2_LCsampsonly_TopHVGs_prop1.0_nobatch.txt",sep='\t',quote=F)

rm(lv,hvg,hvgsub)
```

# SVGs are calculated on a samplewise basis, so we just need to do some recalculating from the existing nnSVG results.
```{r}
nnsvgranks <- fread("processed-data/07_featureSelection_dimred_harmony_clustering/01c-nnsvg_s2n2_nobatch_meangeneranks.txt")

## drop MTRNRs from above
nnsvgranks <- nnsvgranks[!(gene_name %in% c("MTRNR2L1","MTRNR2L8","MTRNR2L12"))]

## get mean rank, number of samples signif, and proportion of samples from donor signif
keepname <- c("gene_id","gene_name",unique(lc3$sample_id)[unique(lc3$sample_id) %in% names(nnsvgranks)])
nnsvgranks2 <- nnsvgranks[,..keepname]
nnsvgranks2[,meanrank:=rowMeans(.SD,na.rm=T),.SDcols=c(3:ncol(nnsvgranks2))]
nnsvgranks2[,samples_signif:=rowSums(!is.na(.SD)),.SDcols=c(3:(ncol(nnsvgranks2)-1))]

fwrite(nnsvgranks2,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/01c_2-LCsampsonly_nnsvg_s2n2_nobatch_meangeneranks.txt",sep='\t',quote=F)

rm(nnsvgranks,keepname)
```


In the hypothalamus dataset, using  N = [the number of genes returned by scran::getTopHVGs with prop=0.1] for feature selection worked better than prop=0.2 for HVGs but also with SVGs. So let's determine that N and then make feature sets accordingly.
```{r}
nfeat <- round(0.1*nrow(hvgtab))
# 1409
```

•Feature sets minimizing effects of lognorm/zero inflation: 
Feature selection and GLMPCA dimensionality reduction using N features:
	1. top N highly deviant genes (HDGs) (i.e. also ascertained in a lognorm-free manner)
	2. top N SVGs (nnsvg uses lognorm) -
	3. Top N/2 HDGs and top N/2 SVGs
	4. Top N/4 HDGs and top 3N/4 SVGs
	5. Top 3N/4 HDGs and top N/4 SVGs
	
SVGs should be selected using their mean rank when only considering genes with nominally signif spatial variability in ≥50% (>38) of the 77 samples.
```{r}
nnsvg_tmp <- nnsvgranks2[samples_signif>38]
###

featlist <- list()

setorderv(devgen,"rank_nobatch")
featlist[[1]] <- devgen[1:nfeat,gene_id]

## check how many SVGs meet the 50% of samples cutoff
nrow(nnsvg_tmp) # 835 -- less than our 1409 minimum. so just take the 835 here (otherwise data table gives back empty rows to fill the remainder)
setorderv(nnsvg_tmp,"meanrank")
featlist[[2]] <- nnsvg_tmp[samples_signif>38][c(1:min(nfeat,nrow(nnsvg_tmp))),gene_id]

# nnsvg_tmp[!(gene_id %in% featlist[[1]]),gene_name]
## the HDG-SVG combo is almost pointless here. all but 11 of the SVGs are also in the top nfeat HDGs -- so at this point we're going to end up with a feature set that's still (nfeat-11)/nfeat = 99.2% of the HDGs.

## ## ## however, the SVG-only genes ARE interesting: GABRE, ADRA2A, SEZ6 , SV2B, SLC6A15 (!!), JPT1, CTXN1, APMAP, CDKN2D, TOLLIP

## so we'll make this set anyhow, though it probably will yield almost identical clustering. probably.
featlist[[3]] <- unique(c(featlist[[1]][c(1:(nfeat-11))],
                  nnsvg_tmp[!(gene_id %in% featlist[[1]][1:(nfeat-11)]),gene_id]))

### anywho, the other two sets we'll just make as described. these will be smaller sets than the first three.
featlist[[4]] <- unique(c(featlist[[1]][1:(nfeat/4)],featlist[[2]][1:min(3*nfeat/4,nrow(nnsvg_tmp))]))
## ^ 837 genes

featlist[[5]] <- unique(c(featlist[[1]][1:(3*nfeat/4)],featlist[[2]][1:min(nfeat/4,nrow(nnsvg_tmp))]))
## ^ 1063 genes

names(featlist) <- c("HDG_1409","nnSVG_1409","HDG1398_SVGexclusive11","HDG_SVG_2575","HDG_SVG_7525")
```

## Then another 4 sets using HVGs instead of HDGs:
```{r}
setorderv(hvgtab,"HVG_rank")
featlist[[6]] <- hvgtab[1:nfeat,gene_id]

# nnsvgranks2[!(gene_id %in% featlist[[6]])&samples_signif>38,gene_id]
# 585 of the 835 possible SVGs are not in the top 1409 HVGs so these are majority exclusive sets.
featlist[[7]] <- unique(c(featlist[[6]][1:(nfeat/2)],featlist[[2]][1:min(nfeat/2,nrow(nnsvg_tmp))]))
## 1224 genes ^

featlist[[8]] <- unique(c(featlist[[6]][1:(nfeat/4)],featlist[[2]][1:min(3*nfeat/4,nrow(nnsvg_tmp))]))
## 1028 ^

featlist[[9]] <- unique(c(featlist[[6]][1:(3*nfeat/4)],featlist[[2]][1:min(nfeat/4,nrow(nnsvg_tmp))]))
# 1272 ^

names(featlist)[6:9] <- c("HVG_1397","HVG_SVG_5050","HVG_SVG_2575","HVG_SVG_7525")

## make sure there's no NAs in the gene lists
stopifnot(sum(unlist(lapply(featlist,function(x){sum(is.na(x))})))==0)
# save
saveRDS(featlist,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/01d_2_LCsampsonly_HiDevG_HVG_SVG_andmixed_featsets.RDS")
```

### reprod info
# session inf #
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] scran_1.38.0                scater_1.38.0              
 [3] scuttle_1.20.0              ggtext_0.1.2               
 [5] scry_1.22.0                 BiocParallel_1.44.0        
 [7] parallelly_1.46.1           colorout_1.3-3             
 [9] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
[11] SummarizedExperiment_1.40.0 Biobase_2.70.0             
[13] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[15] IRanges_2.44.0              S4Vectors_0.48.0           
[17] BiocGenerics_0.56.0         generics_0.1.4             
[19] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[21] gridExtra_2.3               ggplot2_4.0.2              
[23] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1    viridisLite_0.4.2   dplyr_1.2.0         vipor_0.4.7        
 [5] farver_2.1.2        viridis_0.6.5       Biostrings_2.78.0   S7_0.2.1           
 [9] fastmap_1.2.0       bluster_1.20.0      digest_0.6.39       rsvd_1.0.5         
[13] lifecycle_1.0.5     cluster_2.1.8.1     statmod_1.5.1       magrittr_2.0.4     
[17] compiler_4.5.2      rlang_1.1.7         tools_4.5.2         igraph_2.2.1       
[21] yaml_2.3.12         knitr_1.51          dqrng_0.4.1         S4Arrays_1.10.1    
[25] here_1.0.2          DelayedArray_0.36.0 xml2_1.5.2          RColorBrewer_1.1-3 
[29] abind_1.4-8         rsconnect_1.7.0     withr_3.0.2         grid_4.5.2         
[33] beachmat_2.26.0     edgeR_4.8.2         scales_1.4.0        dichromat_2.0-0.1  
[37] cli_3.6.5           rmarkdown_2.30      crayon_1.5.3        metapod_1.18.0     
[41] otel_0.2.0          rstudioapi_0.18.0   rjson_0.2.23        sessioninfo_1.2.3  
[45] ggbeeswarm_0.7.3    XVector_0.50.0      vctrs_0.7.1         Matrix_1.7-4       
[49] BiocSingular_1.26.0 BiocNeighbors_2.4.0 ggrepel_0.9.6       irlba_2.3.7        
[53] beeswarm_0.4.0      magick_2.9.0        locfit_1.5-9.12     limma_3.66.0       
[57] glue_1.8.0          codetools_0.2-20    gtable_0.3.6        ScaledMatrix_1.18.0
[61] tibble_3.3.1        pillar_1.11.1       htmltools_0.5.9     R6_2.6.1           
[65] rprojroot_2.1.1     evaluate_1.0.5      lattice_0.22-7      gridtext_0.1.5     
[69] Rcpp_1.1.1          SparseArray_1.10.8  xfun_0.56           pkgconfig_2.0.3    
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-12
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beachmat               2.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bluster                1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1  2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 dqrng                  0.4.1    2024-05-28 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 igraph                 2.2.1    2025-10-27 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7    2026-01-30 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 metapod                1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scater               * 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 scran                * 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scry                 * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scuttle              * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 statmod                1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5    2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
