---
title: "07-Manual_expression_based_LC_definition_for_comparison_to_louvain"
author: "Bernie Mulvey"
date: "2025-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
```

```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

### genes we want to use for determining cluster labels: DBH, TH, SLC6A2, PHOX2A, PHOX2B, DDC, SLC18A2

tmp <- lc3
rownames(tmp) <- rowData(tmp)$gene_name

tmp <- tmp[c("DBH","TH","SLC6A2","PHOX2A","PHOX2B","DDC","SLC18A2"),]

###
### hi stringency definition: ≥5 counts across DBH TH and 6A2 (all 3 required), and two counts among any others

histr <- apply(counts(tmp),MARGIN=2,simplify = FALSE,FUN = function(x){
  if (x["DBH"]>=1 & (x["TH"]>=1 | x["SLC6A2"]>=1) & sum(x["DBH"],x["TH"],x["SLC6A2"])>=5 & sum(x["PHOX2A"], x["PHOX2B"], x["DDC"], x["SLC18A2"])>=2){return("histr_LC")} else{
      return("nonLC")
  }
})
# sum(histr=="histr_LC")
# --> 6111 spots


###
### low stringency definition: ≥1 count DBH, 1 count of either TH or 6A2, and one count among any others

lostr <- apply(counts(tmp),MARGIN=2,simplify = FALSE,FUN = function(x){
  if (x["DBH"]>=1 & (x["TH"]>=1 | x["SLC6A2"]>=1) & (x["PHOX2A"]>=1 | x["PHOX2B"]>=1 | x["DDC"]>=1 | x["SLC18A2"]>=1)){return("lowstr_LC")} else{
      return("nonLC")
  }
})

### make into a table
histr <- as.data.table(histr)
histr <- t(histr)
histr <- as.data.table(histr,keep.rownames=T)

lostr <- as.data.table(lostr)
lostr <- t(lostr)
lostr <- as.data.table(lostr,keep.rownames=T)

setnames(histr,2,"histringency_LC")
setnames(lostr,2,"lostringency_LC")

manlcs <- merge.data.table(lostr,histr,by="rn")

#
fwrite(manlcs,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/07_2_LCsampsonly_manual_expression_defined_LC.txt",sep='\t',quote=F)

```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] Matrix_1.7-4                bluster_1.20.0             
 [3] BiocParallel_1.44.0         parallelly_1.46.1          
 [5] colorout_1.3-3              harmony_1.2.4              
 [7] Rcpp_1.1.1                  scater_1.38.0              
 [9] scran_1.38.0                scuttle_1.20.0             
[11] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
[13] SummarizedExperiment_1.40.0 Biobase_2.70.0             
[15] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[17] IRanges_2.44.0              S4Vectors_0.48.0           
[19] BiocGenerics_0.56.0         generics_0.1.4             
[21] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[23] gridExtra_2.3               ggtext_0.1.2               
[25] ggplot2_4.0.2               data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] rlang_1.1.7         magrittr_2.0.4      RcppAnnoy_0.0.23    otel_0.2.0         
 [5] compiler_4.5.2      vctrs_0.7.1         pkgconfig_2.0.3     fastmap_1.2.0      
 [9] magick_2.9.0        XVector_0.50.0      rmarkdown_2.30      sessioninfo_1.2.3  
[13] ggbeeswarm_0.7.3    xfun_0.56           beachmat_2.26.0     DelayedArray_0.36.0
[17] irlba_2.3.7         cluster_2.1.8.1     R6_2.6.1            RColorBrewer_1.1-3 
[21] limma_3.66.0        knitr_1.51          igraph_2.2.1        tidyselect_1.2.1   
[25] rstudioapi_0.18.0   dichromat_2.0-0.1   abind_1.4-8         yaml_2.3.12        
[29] viridis_0.6.5       codetools_0.2-20    lattice_0.22-7      tibble_3.3.1       
[33] withr_3.0.2         S7_0.2.1            evaluate_1.0.5      xml2_1.5.2         
[37] pillar_1.11.1       rsconnect_1.7.0     rprojroot_2.1.1     scales_1.4.0       
[41] RhpcBLASctl_0.23-42 glue_1.8.0          metapod_1.18.0      tools_4.5.2        
[45] BiocNeighbors_2.4.0 ScaledMatrix_1.18.0 RSpectra_0.16-2     locfit_1.5-9.12    
[49] cowplot_1.2.0       grid_4.5.2          edgeR_4.8.2         beeswarm_0.4.0     
[53] BiocSingular_1.26.0 vipor_0.4.7         cli_3.6.5           rsvd_1.0.5         
[57] S4Arrays_1.10.1     viridisLite_0.4.2   dplyr_1.2.0         uwot_0.2.4         
[61] gtable_0.3.6        digest_0.6.39       SparseArray_1.10.8  ggrepel_0.9.6      
[65] dqrng_0.4.1         rjson_0.2.23        farver_2.1.2        htmltools_0.5.9    
[69] lifecycle_1.0.5     here_1.0.2          statmod_1.5.1       gridtext_0.1.5     
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-12
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beachmat               2.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bluster              * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1  2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 cowplot                1.2.0    2025-07-07 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 dqrng                  0.4.1    2024-05-28 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 harmony              * 1.2.4    2025-10-10 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 igraph                 2.2.1    2025-10-27 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7    2026-01-30 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix               * 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 metapod                1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                 * 1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 RcppAnnoy              0.0.23   2026-01-12 [1] CRAN (R 4.5.2)
 RhpcBLASctl            0.23-42  2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 RSpectra               0.16-2   2024-07-18 [1] CRAN (R 4.5.0)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scater               * 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 scran                * 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scuttle              * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 statmod                1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 uwot                   0.2.4    2025-11-10 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5    2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
