---
title: "01-Clust_of_interest_markers_for_plotting"
author: "Bernie Mulvey"
date: "2024-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


We already determined markers and made putative annotations for the clusters from glmpca-SNN k 10-louvain resolution 0.5. So load these in, filter, and save a subset of genes for which we'll plot (with a column added to indicate which cluster we think they mark).

```{r}
snnlouvmks <- readRDS("processed-data/07_Clustering/03-allgeneGLMPCA_snnk10-15-20_louvain_FDR0.1markers.RDS")

mktab <- snnlouvmks[["snn10_louv_res0.5"]]

## append cluster annots
annot <- fread("processed-data/07_Clustering/03-glmpca_snn10_louv0.5_tentativeannotsbytop10lfcmks.txt")

mktab <- merge.data.table(mktab,annot,by="clus")

## filter to only the clusters of interest: "astrocyte_1", "raphe", "LC_1", "Interneurons_1_SERT", "Interneurons_2_top10_2lnc_4maybegenes"
mktab <- mktab[annot %in% c("astrocyte_1", "raphe", "LC_1", "Interneurons_1_SERT", "Interneurons_2_top10_2lnc_4maybegenes")]

## get top 15 by LFC and top 15 by FDR (table is prefiltered to fdr 0.1 so may yield fewer than 30 genes) for each
toplfcs <- mktab[order(-logFC),head(.SD,15),by="clus"][,.(clus,gene,annot)]
topfdrs <- mktab[order(fdr),head(.SD,15),by="clus"][,.(clus,gene,annot)]
top30s <- unique(rbind(toplfcs,topfdrs))
top30s[,plottitle:=paste0(gene, " ", annot)]

fwrite(top30s,"processed-data/09_XenDesign_Analyses/01a-louv.5_top10markers_for_lc_dr_iN_astro.txt",sep='\t',quote=F)
```


We also already determined markers and made putative annotations for the clusters from glmpca-SNN k 10-louvain resolution 1. So same thing

```{r}
snnlouvmks <- readRDS("processed-data/07_Clustering/03-allgeneGLMPCA_snnk10-15-20_louvain_FDR0.1markers.RDS")

mktab <- snnlouvmks[["snn10_louv_res1"]]

## append cluster annots
annot <- fread("processed-data/07_Clustering/03-glmpca_snn10_louv1_tentativeannotsbytop10lfcmks.txt")

mktab <- merge.data.table(mktab,unique(annot[,.(clus,label)]),by="clus")

## filter to only the clusters of interest: "Exc.17A6.RBP4.TAC1.HTR2C"", "Astrocyte.1", "Microglia.MostlyPutativeXscrips", "Inh.SST.CCK.HTR2C", "5HT.1", "LC.1"
mktab <- mktab[label %in% c("Exc.17A6.RBP4.TAC1.HTR2C","Astrocyte.1","Microglia.MostlyPutativeXscrips","Inh.SST.CCK.HTR2C","5HT.1","LC.1")]

## get top 15 by LFC and top 15 by FDR (table is prefiltered to fdr 0.1 so may yield fewer than 30 genes) for each
toplfcs <- mktab[order(-logFC),head(.SD,15),by="clus"][,.(clus,gene,label)]
topfdrs <- mktab[order(fdr),head(.SD,15),by="clus"][,.(clus,gene,label)]
top30s <- unique(rbind(toplfcs,topfdrs))
top30s[,plottitle:=paste0(gene, " ", label)]

fwrite(top30s,"processed-data/09_XenDesign_Analyses/01b-louv1_top10markers_for_lc_dr_iN_TAC1ex_astro_mglia.txt",sep='\t',quote=F)
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] colorout_1.3-0.2            SpatialExperiment_1.14.0   
 [3] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
 [5] Biobase_2.64.0              GenomicRanges_1.56.1       
 [7] GenomeInfoDb_1.40.1         IRanges_2.38.1             
 [9] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[11] MatrixGenerics_1.16.0       matrixStats_1.3.0          
[13] data.table_1.15.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] Matrix_1.7-0            jsonlite_1.8.8          rjson_0.2.21           
 [4] compiler_4.4.1          crayon_1.5.3            Rcpp_1.0.13            
 [7] magick_2.8.3            yaml_2.3.9              fastmap_1.2.0          
[10] lattice_0.22-6          here_1.0.1              R6_2.5.1               
[13] XVector_0.44.0          S4Arrays_1.4.1          knitr_1.48             
[16] DelayedArray_0.30.1     rprojroot_2.0.4         GenomeInfoDbData_1.2.12
[19] xfun_0.45               SparseArray_1.4.8       cli_3.6.3              
[22] magrittr_2.0.3          zlibbioc_1.50.0         digest_0.6.36          
[25] grid_4.4.1              rstudioapi_0.16.0       evaluate_0.24.0        
[28] sessioninfo_1.2.2       abind_1.4-5             rmarkdown_2.27         
[31] httr_1.4.7              tools_4.4.1             htmltools_0.5.8.1      
[34] UCSC.utils_1.0.0       
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-12-11
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version date (UTC) lib source
   abind                  1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics         * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   DelayedArray           0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   evaluate               0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12  2024-05-01 [1] Bioconductor
   GenomicRanges        * 1.56.1  2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   jsonlite               1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr                  1.48    2024-07-07 [1] CRAN (R 4.4.1)
   lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   magick                 2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   rjson                  0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Arrays               1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SparseArray            1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SummarizedExperiment * 1.34.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   UCSC.utils             1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   xfun                   0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector                0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
