---
title: "04-allgeneGLMPCA_Harmony_UMAP"
output: html_document
date: "2024-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(harmony)
library(scater)
library(HDF5Array)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### function to load spe without image data for memory efficiency
```{r}
spesessionloader <- function(){
  lc <- readRDS("processed-data/05_QC/04-lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed.RDS")
  
  # make sure colnames of the spe are unique (i.e., colData(hyp2)$key)
  colnames(lc) <- lc$key
  
  ## for efficiency during these steps, drop the imgData slot
  lc@int_metadata$imgData <- DataFrame()
  gc(full=T)
  
  return(lc)
}
```
### 

## load GLMPCA and append to reduced dim of LC SPE
```{r}
lcglmpca <- HDF5Array("processed-data/06_FeatSelection_Dimred_Harmony/03b_lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed_allgeneGLMPCA.h5",name="GLM-PCA")

lc <- spesessionloader()

reducedDimNames(lc)
# still has 10x crap

# realize the H5 object which is (intentionally) not fully loaded to start by HDF5Array into the reducedDims slot
reducedDims(lc)$GLMPCA <- as.matrix(lcglmpca)
## get rid of 10x garbage
reducedDims(lc)[1:3] <- NULL
```

## run pre-HARMONY UMAP, HARMONY, and post-H umap
```{r}
## run pre-HARMONY UMAP
set.seed(42)
lc <- runUMAP(lc, dimred="GLMPCA",name = "preHarmony_UMAP")

## save pre-HARMONY UMAP in a separate file
saveRDS(reducedDim(lc,"preHarmony_UMAP"),"processed-data/06_FeatSelection_Dimred_Harmony/04a_allgeneGLMPCA_PREharmony_UMAP.RDS")


### Harmony (lambda = NA)--use the default Method (passing in separate matrices and metadata) to be absolutely certain its using the reduced dims we want it to
set.seed(42)
hmo_glmpca <- RunHarmony(data_mat=reducedDim(lc,"GLMPCA"),meta_data=colData(lc),lambda=NULL,vars_use="sample_id",plot_convergence=T,kmeans_init_iter_max=100)
dev.off()

saveRDS(hmo_glmpca,"processed-data/06_FeatSelection_Dimred_Harmony/04b_allgeneGLMPCA_Harmony.RDS")

stopifnot(identical(rownames(hmo_glmpca),colnames(lc)))
reducedDim(lc,"Harmony_allgeneGLMPCA") <- hmo_glmpca

## run post-HARMONY UMAP
set.seed(42)
lc <- runUMAP(lc, dimred="Harmony_allgeneGLMPCA",name = "postHarmony_UMAP")

# save
saveRDS(reducedDim(lc,"postHarmony_UMAP"),"processed-data/06_FeatSelection_Dimred_Harmony/04c_allgeneGLMPCA_POSTharmony_UMAP.RDS")
```

repro inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.38.0         parallelly_1.38.0          
 [3] colorout_1.3-0.2            HDF5Array_1.32.0           
 [5] rhdf5_2.48.0                DelayedArray_0.30.1        
 [7] SparseArray_1.4.8           S4Arrays_1.4.1             
 [9] abind_1.4-5                 Matrix_1.7-0               
[11] scater_1.32.0               scuttle_1.14.0             
[13] harmony_1.2.0               Rcpp_1.0.13                
[15] SpatialExperiment_1.14.0    SingleCellExperiment_1.26.0
[17] SummarizedExperiment_1.34.0 Biobase_2.64.0             
[19] GenomicRanges_1.56.1        GenomeInfoDb_1.40.1        
[21] IRanges_2.38.1              S4Vectors_0.42.1           
[23] BiocGenerics_0.50.0         MatrixGenerics_1.16.0      
[25] matrixStats_1.3.0           gridExtra_2.3              
[27] ggplot2_3.5.1               data.table_1.15.4          
[29] rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1          viridisLite_0.4.2         farver_2.1.2             
 [4] dplyr_1.1.4               vipor_0.4.7               viridis_0.6.5            
 [7] fastmap_1.2.0             digest_0.6.36             rsvd_1.0.5               
[10] lifecycle_1.0.4           magrittr_2.0.3            compiler_4.4.1           
[13] tools_4.4.1               utf8_1.2.4                yaml_2.3.9               
[16] knitr_1.48                labeling_0.4.3            here_1.0.1               
[19] withr_3.0.0               grid_4.4.1                fansi_1.0.6              
[22] beachmat_2.20.0           colorspace_2.1-0          Rhdf5lib_1.26.0          
[25] scales_1.3.0              cli_3.6.3                 rmarkdown_2.27           
[28] crayon_1.5.3              generics_0.1.3            rstudioapi_0.16.0        
[31] httr_1.4.7                rjson_0.2.21              sessioninfo_1.2.2        
[34] DelayedMatrixStats_1.26.0 ggbeeswarm_0.7.2          zlibbioc_1.50.0          
[37] parallel_4.4.1            ggrastr_1.0.2             XVector_0.44.0           
[40] vctrs_0.6.5               jsonlite_1.8.8            BiocSingular_1.20.0      
[43] BiocNeighbors_1.22.0      ggrepel_0.9.5             irlba_2.3.5.1            
[46] beeswarm_0.4.0            magick_2.8.3              glue_1.7.0               
[49] codetools_0.2-20          uwot_0.2.2                cowplot_1.1.3            
[52] RcppAnnoy_0.0.22          gtable_0.3.5              UCSC.utils_1.0.0         
[55] ScaledMatrix_1.12.0       munsell_0.5.1             tibble_3.2.1             
[58] pillar_1.9.0              htmltools_0.5.8.1         rhdf5filters_1.16.0      
[61] GenomeInfoDbData_1.2.12   R6_2.5.1                  sparseMatrixStats_1.16.0 
[64] rprojroot_2.0.4           evaluate_0.24.0           lattice_0.22-6           
[67] png_0.1-8                 RhpcBLASctl_0.23-42       xfun_0.45                
[70] pkgconfig_2.0.3          
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-18
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package              * version date (UTC) lib source
   abind                * 1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   beachmat               2.20.0  2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics         * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors          1.22.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular           1.20.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   cowplot                1.1.3   2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   DelayedArray         * 0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate               0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi                  1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12  2024-05-01 [1] Bioconductor
   GenomicRanges        * 1.56.1  2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5   2024-01-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   harmony              * 1.2.0   2023-11-29 [1] CRAN (R 4.4.0)
   HDF5Array            * 1.32.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr                  1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling               0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magick                 2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix               * 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   png                    0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
   R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                 * 1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   RcppAnnoy              0.0.22  2024-01-23 [1] CRAN (R 4.4.0)
   rhdf5                * 2.48.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters           1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib               1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RhpcBLASctl            0.23-42 2023-02-11 [1] CRAN (R 4.4.0)
   rjson                  0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
   S4Arrays             * 1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   ScaledMatrix           1.12.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.32.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle              * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SparseArray          * 1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SummarizedExperiment * 1.34.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   uwot                   0.2.2   2024-04-21 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector                0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
