---
title: "SF-Donorwise_louvLC_vs_ManualLC_labels"
author: "Bernie Mulvey"
date: "2025-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

already tabulated these into plottable form:
```{r}
louvman <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/07-manualLCdefs_vs_louvains.RDS")
cd <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")
cd <- unique(cd[,.(sample_id,brnum)])

louvman <- louvman$snnHARMONYlmbna_HDG_SVG_2575_louv_res1

pltl <- melt.data.table(louvman,id.vars="rn")
pltl[,variable:=gsub(variable,pattern="^.*__(.*)$",replacement="\\1")]

pltl <- merge.data.table(pltl,cd,by.x="rn",by.y="sample_id")
pltl[,value:=as.numeric(value)]

pltl[variable=="lostringency_LC",meancor_lo:=mean(value,na.rm=T),by="brnum"]
pltl[variable=="histringency_LC",meancor_hi:=mean(value,na.rm=T),by="brnum"]
pltl[is.na(meancor_hi),value:=meancor_lo]
pltl[is.na(meancor_lo),value:=meancor_hi]
pltl <- unique(pltl[,.(brnum,variable,value)])
```

make presentable and plot
```{r}
pltl[variable=="histringency_LC",variable:="High-Stringency"]
pltl[variable=="lostringency_LC",variable:="Low-Stringency"]

setnames(pltl,"variable","Expression-Based\nLC Definition")
setnames(pltl,"brnum","Donor ID")

pltl[`Donor ID`=="Br6119(re-dis)",`Donor ID`:="Br6119"]

pdf("plots/manuscript/SuppFigs/SF-LouvainLC_vs_ManualDefs_Pbulkcorrel_by_donor.pdf",height=5.5,width=4)
ggplot(pltl,aes(x=`Expression-Based\nLC Definition`,y=`Donor ID`,fill=value))+
    geom_tile()+
    scale_fill_continuous(breaks = seq(0.4,1,0.1),type="viridis")+
    scale_x_discrete(expand=c(0,0))+
    scale_y_discrete(expand=c(0,0))+
    labs(fill="Mean Samplewise\nSpearman Cor. of\nPseudobulk Expression\n(vs. Louvain LC)")+
    theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))
dev.off()
```


```{r}
sessionInfo()
sessioninfo::session_info()
```
─ Session info ───────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-09-09
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────────────
 !  package      * version date (UTC) lib source
    BiocParallel * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli            3.6.4   2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.3)
    colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    commonmark     1.9.5   2025-03-17 [1] CRAN (R 4.4.1)
    data.table   * 1.17.4  2025-05-26 [1] CRAN (R 4.4.3)
    dichromat      2.0-0.1 2022-05-02 [1] CRAN (R 4.4.0)
    digest         0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate       1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver         2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics       0.1.4   2025-05-09 [1] CRAN (R 4.4.1)
    ggplot2      * 3.5.2   2025-04-09 [1] CRAN (R 4.4.1)
    ggtext       * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue           1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext       0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable         0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    knitr          1.50    2025-03-16 [1] CRAN (R 4.4.2)
    labeling       0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    litedown       0.6     2025-02-27 [1] CRAN (R 4.4.3)
    magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    markdown       2.0     2025-03-23 [1] CRAN (R 4.4.3)
    parallelly   * 1.43.0  2025-03-24 [1] CRAN (R 4.4.1)
    pillar         1.10.2  2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6             2.6.1   2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer   1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp           1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.5   2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown      2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi     0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    scales         1.4.0   2025-04-24 [1] CRAN (R 4.4.1)
    sessioninfo    1.2.3   2025-02-05 [1] CRAN (R 4.4.1)
    stringi        1.8.7   2025-03-27 [1] CRAN (R 4.4.1)
    stringr        1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
    tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    withr          3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun           0.52    2025-04-02 [1] CRAN (R 4.4.1)
    xml2           1.3.8   2025-03-14 [1] CRAN (R 4.4.1)
    yaml           2.3.10  2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────────
