---
title: "SF-Additional_SpotNM_LMMs"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(ggtext)
library(gridExtra)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
# require(colorout)
# ColorOut()
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

### using an E4 carrier yes/no (i.e., E4 or E2) model
```{r}
nmlm <- readRDS("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/11-LCNMspots_NMmetric_LMMed_by_geneLogcounts_3apoeFactorizations.RDS")

nmi <- nmlm$E4carrierYN$Intensity_NM

nmi[variable==gene_name|variable=="APOE.x",bonfP:=p*(length(unique(nmi$gene_name)))]
nmi[bonfP>1,bonfP:=1]
```

genes mentioned in text but not in main fig: "MAP1LC3B", "PGRMC1", "SARAF", "MT2A", "MT1E", "MT1X", "MT3"

# we'll need the SPE with LC NM+ cells, log counts, and metadata with tissue NM scores
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

# calc spot log counts
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

# load clusters + annots
louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs[["HARMONYlmbna_HDG_SVG_2575"]][,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

finalann <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")

louvs <- merge.data.table(louvs,finalann,by="clusid")
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$anno
lc3 <- lc3[,!(lc3$brnum %in% paste0("Br",c(5517,5276,5712)))]

## read in nm subclusterings
lcnm <- fread("processed-data/10_subclustering/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnm <- lcnm[,.(rn,NMseg_subclus)]
lcnm <- DataFrame(lcnm,row.names=lcnm$rn)[colnames(lc3),]

## assign
lc3$lcnm <- lcnm$NMseg_subclus

## get LC NM+ spot spe and its coldata w/ nm metrics; calculate mean and median NM intensity per sample
lc3 <- lc3[,lc3$lcnm=="LC_NMpos"]
lcnmcd <- as.data.table(colData(lc3[,lc3$lcnm=="LC_NMpos"]))
lcnmcd[,med_NMI:=median(Intensity_NM,na.rm=T),by="sample_id"]
lcnmcd[,mn_NMI:=mean(Intensity_NM,na.rm=T),by="sample_id"]

# rownames to gene symbs
rownames(lc3) <- rowData(lc3)$gene_name
```

## facet wrap it
```{r}
pltg <- c("MAP1LC3B", "PGRMC1", "SARAF", "MT2A", "MT1E", "MT1X", "MT3")

pltdat <- as.data.frame(t(as.matrix(logcounts(lc3[pltg,]))))
pltdat <- as.data.table(pltdat,keep.rownames=T)
pltdat <- merge.data.table(lcnmcd[,.(rowname,sample_id,Intensity_NM)],pltdat,by.x="rowname",by.y="rn")
pltdat[,nspot:=.N,by="sample_id"] # determine the number of spots from each sample INCLUDING those with logcounts=0 (though we won't plot those)

pltdat<-melt.data.table(pltdat,id.vars = c("rowname","sample_id","Intensity_NM","nspot"))
pltdat[,variable:=as.character(variable)]

pltdat <- pltdat[value!=0]

i<-1
for (i in c(1:length(unique(pltdat$variable)))){
    curg <- unique(pltdat$variable)[i]
    tmpdat <- copy(pltdat)[variable==curg]
    qbreaks <- quantile(tmpdat$value, probs = seq(0, 1, by = 0.25), na.rm = TRUE)
    tmpdat[, x_quartile := cut(value, 
                 breaks = qbreaks, 
                 labels = 1:4,         # Label bins accoring to nth decile
                 include.lowest = TRUE)]
    if(i==1){
        pltdat2 <- as.data.table(tmpdat)
    } else {
        pltdat2 <- as.data.table(rbind(pltdat2,tmpdat))
    }
    rm(tmpdat,curg)
}

pltdat2[,calcfill:=mean(Intensity_NM),by=c("variable","x_quartile")]
pltdat2[,variable:=paste0("*",variable,"*")]

# encode bonferroni significance to mark in illustrator
pltdat2[variable %in% c("*SARAF*","*MT2A*","*MT1E*","*MAP1LC3B*","*PGRMC1*"),variable:=paste0(variable," (B)")]
# ibid for FDR
pltdat2[!(variable %in% c("*SARAF* (B)","*MT2A* (B)","*MT1E* (B)","*MAP1LC3B* (B)","*PGRMC1* (B)")),variable:=paste0(variable," (F)")]
pltdat2[,variable:=factor(variable,levels=c("*MT1E* (B)","*MT1X* (F)","*MT2A* (B)","*MT3* (F)","*SARAF* (B)","*MAP1LC3B* (B)","*PGRMC1* (B)"))]

p <- ggplot(unique(pltdat2[nspot>19,.(x_quartile,variable,Intensity_NM,calcfill)]),aes(x=x_quartile,group=x_quartile,y=Intensity_NM,fill=calcfill))+
    scale_fill_gradient(low="brown",high="beige")+
    guides(fill="none")+
    geom_boxplot(outlier.size = 0.4,outlier.alpha = 0.5)+
    facet_wrap(.~variable,ncol=2)+
    labs(y="Mean Spot NM+ Pixel Intensity",x=paste0("Quartile spot logcounts"))+
    theme(axis.title.x=element_text(size=9),
          axis.title.y=element_text(size=9),
          axis.text.x=element_text(size=8),
          axis.text.y=element_text(size=8),
          strip.text=element_markdown(size=9),
          strip.background = element_blank())


pdf("plots/manuscript/SuppFigs/SF-Additional_NMLMMhits.pdf",height=9.5,width=7)
p
dev.off()
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
 [4] Biobase_2.66.0              GenomicRanges_1.58.0        GenomeInfoDb_1.42.3        
 [7] IRanges_2.40.1              S4Vectors_0.44.0            BiocGenerics_0.52.0        
[10] MatrixGenerics_1.18.1       matrixStats_1.5.0           BiocParallel_1.40.0        
[13] parallelly_1.43.0           gridExtra_2.3               ggtext_0.1.2               
[16] ggplot2_3.5.2               data.table_1.17.4           rlang_1.1.5                

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1        viridisLite_0.4.2       dplyr_1.1.4            
 [4] vipor_0.4.7             farver_2.1.2            viridis_0.6.5          
 [7] fastmap_1.2.0           digest_0.6.37           rsvd_1.0.5             
[10] lifecycle_1.0.4         magrittr_2.0.3          compiler_4.4.3         
[13] tools_4.4.3             yaml_2.3.10             knitr_1.50             
[16] labeling_0.4.3          S4Arrays_1.6.0          here_1.0.1             
[19] DelayedArray_0.32.0     xml2_1.3.8              RColorBrewer_1.1-3     
[22] abind_1.4-8             withr_3.0.2             grid_4.4.3             
[25] beachmat_2.22.0         scales_1.4.0            dichromat_2.0-0.1      
[28] cli_3.6.4               rmarkdown_2.29          crayon_1.5.3           
[31] generics_0.1.4          rstudioapi_0.17.1       httr_1.4.7             
[34] rjson_0.2.23            commonmark_1.9.5        scuttle_1.16.0         
[37] ggbeeswarm_0.7.2        stringr_1.5.1           zlibbioc_1.52.0        
[40] datasets_4.4.3          parallel_4.4.3          XVector_0.46.0         
[43] vctrs_0.6.5             Matrix_1.7-3            jsonlite_2.0.0         
[46] litedown_0.6            BiocSingular_1.22.0     BiocNeighbors_2.0.1    
[49] ggrepel_0.9.6           irlba_2.3.5.1           beeswarm_0.4.0         
[52] scater_1.34.1           magick_2.8.6            glue_1.8.0             
[55] codetools_0.2-20        stringi_1.8.7           gtable_0.3.6           
[58] UCSC.utils_1.2.0        ScaledMatrix_1.14.0     tibble_3.2.1           
[61] pillar_1.10.2           htmltools_0.5.8.1       GenomeInfoDbData_1.2.13
[64] R6_2.6.1                rprojroot_2.0.4         evaluate_1.0.3         
[67] lattice_0.22-6          markdown_2.0            gridtext_0.1.5         
[70] Rcpp_1.0.14             SparseArray_1.6.2       xfun_0.52              
[73] pkgconfig_2.0.3        
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-09-17
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    beachmat               2.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocNeighbors          2.0.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel         * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 VP cli                    3.6.4   2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.3)
    commonmark             1.9.5   2025-03-17 [1] CRAN (R 4.4.1)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.17.4  2025-05-26 [1] CRAN (R 4.4.3)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.4   2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3  2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.2   2025-04-09 [1] CRAN (R 4.4.1)
    ggrepel                0.9.6   2024-09-07 [1] CRAN (R 4.4.1)
    ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
    jsonlite               2.0.0   2025-03-27 [1] CRAN (R 4.4.1)
    knitr                  1.50    2025-03-16 [1] CRAN (R 4.4.2)
    labeling               0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.3)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    litedown               0.6     2025-02-27 [1] CRAN (R 4.4.3)
    magick                 2.8.6   2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    markdown               2.0     2025-03-23 [1] CRAN (R 4.4.3)
    Matrix                 1.7-3   2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1  2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    parallelly           * 1.43.0  2025-03-24 [1] CRAN (R 4.4.1)
    pillar                 1.10.2  2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.6.1   2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5   2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ScaledMatrix           1.14.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0   2025-04-24 [1] CRAN (R 4.4.1)
    scater                 1.34.1  2025-03-03 [1] Bioconductor 3.20 (R 4.4.2)
    scuttle                1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.2   2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    stringi                1.8.7   2025-03-27 [1] CRAN (R 4.4.1)
    stringr                1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52    2025-04-02 [1] CRAN (R 4.4.1)
    xml2                   1.3.8   2025-03-14 [1] CRAN (R 4.4.1)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────────
