---
title: "SF-VisQC_cumprobs"
author: "Bernie Mulvey"
date: "2025-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(dplyr)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
rawcd <- fread("processed-data/02_build_spe/rawSPE_coldata.txt")
qccd <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")

rawcd[,original_barcode:=paste0(gsub(original_barcode,pattern="^(.*-1).*$",replacement="\\1"),"_",sample_id)]

rawcd[,APOE:=gsub(APOE,pattern="/ ",replacement="/")]
rawcd[,apo2:=ifelse(APOE %in% grep(APOE,pattern="E4",value=T),"E4","E2")]

rawcd[,group:=paste0(Sex,"/",Ancestry,"/",apo2)]

stopifnot(nrow(rawcd[original_barcode %in% qccd$original_spot_id])==nrow(qccd))

rawcd[,retained:=ifelse(original_barcode %in% qccd$original_spot_id,TRUE,FALSE)]
rawcd[,dropped:=ifelse(retained,FALSE,TRUE)]
## only consider in tissue spots
rawcd <- rawcd[in_tissue==TRUE]
```

```{r}
sample_summary <- copy(rawcd)
sample_summary <- sample_summary[,.(APOE,Ancestry,Sex,group,retained,dropped,.N),by="sample_id"]
setnames(sample_summary,"N","ntot")
sample_summary[,ndrop:=sum(dropped),by="sample_id"]
sample_summary[,propdrop:=ndrop/ntot]
sample_summary <- unique(sample_summary)
sample_summary[,retained:=NULL]
sample_summary[,dropped:=NULL]
sample_summary <- unique(sample_summary)

pdf("plots/manuscript/SuppFigs/SF_QCsummaries/proportion_dropped_by_SexAncApo_group.pdf",height=3,width=4)
ggplot(sample_summary,aes(x=group,y=propdrop,fill=group))+
    geom_boxplot(outliers=FALSE)+
    geom_jitter(size = 2,shape =16,alpha = 0.7,width = 0.2)+
    guides(fill="none")+
    xlab("Sex/Ancestry/*APOE* Genotype Group")+
    ylab("Proportion Spots Dropped")+
    theme(axis.title.x=element_markdown(size=9,colour = "#000000"),axis.text.x=element_text(angle=45,hjust=1,vjust=1,color="#000000",size=8))
dev.off()
```


# step plots for cumulative probs of inclusion, pretty much verbatim from Boyi:
https://github.com/LieberInstitute/spatialDLPFC_SCZ/blob/abb0fe73f4714a9797dd4dfc28358f0dbc1695d2/code/analysis/02_visium_qc/archived/plot_distribution_discard_spots.R#L123
```{r}
pdf("plots/manuscript/SuppFigs/SF_QCsummaries/InTissCumulDist_sumUMI.pdf",height=3,width=2.5)
rawcd[dropped==T] |>
  data.frame() |>
  ggplot() +
  stat_ecdf(
    aes(x = sum_umi, group = sample_id, color = group),
    alpha = 0.5,
    geom = "step"
  ) +
  coord_cartesian(xlim = c(1, 3000)) +
  scale_x_log10() +
  theme_minimal() +
  labs(
    title = "Cumulative Distribution of UMI Counts",
    x = "Log10(UMI Counts)",
    y = "Cumulative Probability",
    color = "Sex/Anc/*APOE* Group"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    legend.title=element_markdown(size=8,color="#000000")
  )
dev.off()

pdf("plots/manuscript/SuppFigs/SF_QCsummaries/InTissCumulDist_sumgene.pdf",height=3,width=2.5)
rawcd[dropped==T] |>
  data.frame() |>
  ggplot() +
  stat_ecdf(
    aes(x = sum_gene, group = sample_id, color = group),
    alpha = 0.5,
    geom = "step"
  ) +
  # scale_x_log10() +
  coord_cartesian(xlim = c(1, 4000)) +
  theme_minimal() +
  labs(
    title = "Cumulative Distribution of Unique Genes",
    x = "Log10(Unique Genes)",
    y = "Cumulative Probability",
    color = "Sex/Ances./*APOE* Group"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    legend.title=element_markdown(size=8,color="#000000")
  )
dev.off()

# 
pdf("plots/manuscript/SuppFigs/SF_QCsummaries/InTissCumulDist_mitoprop.pdf",height=3,width=2.5)
rawcd[dropped==TRUE] |>
  data.frame() |>
  ggplot() +
  stat_ecdf(
    aes(x = expr_chrM_ratio, group = sample_id, color = group),
    alpha = 0.5,
    geom = "step"
  ) +
  theme_minimal() +
  labs(
    title = "Cumulative Distribution of Mitochondrial Ratio",
    x = "Mitochondrial Ratio",
    y = "Cumulative Probability",
    color = "Sex/Ances./*APOE* Group"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    legend.title=element_markdown(size=8,color="#000000")
  )
dev.off()
```

seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] dplyr_1.1.4                 BiocParallel_1.40.0        
 [3] parallelly_1.43.0           colorout_1.3-0.2           
 [5] SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1
 [7] SummarizedExperiment_1.36.0 Biobase_2.66.0             
 [9] GenomicRanges_1.58.0        GenomeInfoDb_1.42.3        
[11] IRanges_2.40.1              S4Vectors_0.44.0           
[13] BiocGenerics_0.52.0         MatrixGenerics_1.18.1      
[15] matrixStats_1.5.0           gridExtra_2.3              
[17] ggtext_0.1.2                ggplot2_3.5.2              
[19] data.table_1.17.4           rlang_1.1.5                

loaded via a namespace (and not attached):
 [1] gtable_0.3.6            rjson_0.2.23            xfun_0.52              
 [4] lattice_0.22-6          vctrs_0.6.5             tools_4.4.3            
 [7] generics_0.1.4          parallel_4.4.3          tibble_3.2.1           
[10] pkgconfig_2.0.3         Matrix_1.7-3            RColorBrewer_1.1-3     
[13] lifecycle_1.0.4         GenomeInfoDbData_1.2.13 stringr_1.5.1          
[16] compiler_4.4.3          farver_2.1.2            codetools_0.2-20       
[19] litedown_0.6            htmltools_0.5.8.1       yaml_2.3.10            
[22] pillar_1.10.2           crayon_1.5.3            DelayedArray_0.32.0    
[25] sessioninfo_1.2.3       magick_2.8.6            abind_1.4-8            
[28] commonmark_1.9.5        tidyselect_1.2.1        digest_0.6.37          
[31] stringi_1.8.7           labeling_0.4.3          rprojroot_2.0.4        
[34] fastmap_1.2.0           grid_4.4.3              here_1.0.1             
[37] cli_3.6.4               SparseArray_1.6.2       magrittr_2.0.3         
[40] S4Arrays_1.6.0          dichromat_2.0-0.1       withr_3.0.2            
[43] scales_1.4.0            UCSC.utils_1.2.0        rmarkdown_2.29         
[46] XVector_0.46.0          httr_1.4.7              evaluate_1.0.3         
[49] knitr_1.50              markdown_2.0            gridtext_0.1.5         
[52] Rcpp_1.0.14             glue_1.8.0              xml2_1.3.8             
[55] rstudioapi_0.17.1       jsonlite_2.0.0          R6_2.6.1               
[58] zlibbioc_1.52.0        
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-10-12
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocParallel         * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli                    3.6.4   2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.3)
    colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    commonmark             1.9.5   2025-03-17 [1] CRAN (R 4.4.1)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.17.4  2025-05-26 [1] CRAN (R 4.4.3)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                * 1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.4   2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3  2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggplot2              * 3.5.2   2025-04-09 [1] CRAN (R 4.4.1)
    ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    jsonlite               2.0.0   2025-03-27 [1] CRAN (R 4.4.1)
    knitr                  1.50    2025-03-16 [1] CRAN (R 4.4.2)
    labeling               0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.3)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    litedown               0.6     2025-02-27 [1] CRAN (R 4.4.3)
    magick                 2.8.6   2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    markdown               2.0     2025-03-23 [1] CRAN (R 4.4.3)
    Matrix                 1.7-3   2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1  2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    parallelly           * 1.43.0  2025-03-24 [1] CRAN (R 4.4.1)
    pillar                 1.10.2  2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.6.1   2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5   2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0   2025-04-24 [1] CRAN (R 4.4.1)
    sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.2   2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    stringi                1.8.7   2025-03-27 [1] CRAN (R 4.4.1)
    stringr                1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    withr                  3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52    2025-04-02 [1] CRAN (R 4.4.1)
    xml2                   1.3.8   2025-03-14 [1] CRAN (R 4.4.1)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
