---
title: "SupFig-Spotplot_clusters"
author: "Bernie Mulvey"
date: "2025-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(ggrastr)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


read in the per-section SPE, already rotated into uniform orientation
```{r}
spe <- readRDS("processed-data/14_Samui_prep/01-Samui_TissSect_SPE_RotsMirrors.RDS")

# load clustering and the split-sections coldata, which has a column with pre-split spot ids we can use to match everything back to capture level
louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno,by="clusid")

## also attach LCnm subsets because we will use them in this figure
lcnm <- fread("processed-data/10_subclustering/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnm <- lcnm[,.(rn,NMseg_subclus)]

louvs <- merge.data.table(louvs,lcnm,by="rn",all.x=T)
louvs[anno=="LC"&NMseg_subclus=="LC_NMpos",plot_anno:="LC (NM+)"]
louvs[anno=="LC"&NMseg_subclus=="LC_NMneg",plot_anno:="LC (NM-)"]

louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(spe),]
stopifnot(nrow(louvs)==ncol(spe))

colLabels(spe) <- louvs$plot_anno

rm(lcnm,louvs,anno)
gc(full=T)
```

## define helper function to extract a single legend - from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

make plots
```{r}
# make a palette for the levels of QC and apply manually outside of spatiallibd
# unique(raw$QC)

pal <- readRDS("plots/finalpal_somuchfornorainbows.RDS")
# set seed and add a 10th color to jointly display LCNM+ and -
set.seed(42)
pal2 <- Polychrome::createPalette(10,seedcolors=pal)
names(pal2)[1:9] <- names(pal)
names(pal2)[10] <- "LC (NM-)"
names(pal2)[1] <- "LC (NM+)"
stopifnot(all(names(pal2) %in% unique(spe$label)))
rm(pal)
## make a plot for a sample containing all 10 clusters values to extract the legend from and append globally later: V13M06-332_B1
tmpdat <- as.data.table(colData(spe))[sample_id=="V13M06-332_B1s1"]
leg.tmp <- ggplot(tmpdat,aes(x=array_col,y=array_row,color=label))+
  geom_point()+
  scale_color_manual(values=pal2)+
  labs(color="Domain")+
  theme(legend.text=element_markdown(size=7),
        legend.title=element_text(size=8),
        legend.background = element_blank(),
        legend.box = element_blank(),
        legend.key=element_blank(),
        legend.key.height=unit(0.1,"in"))

leg.tmp <- extract_legend(leg.tmp)
dev.off()

# defactor
spe$label <- as.character(spe$label)

# get demo info to include with plots
tmpcd <- unique(as.data.table(colData(spe))[,.(sample_id,Sex,APOE,YRI,Age)])
tmpcd[,anc:=ifelse(YRI>0.5,"AA","EA")]

plt <- lapply(unique(spe$sample_id),function(s){
  p <- vis_clus(spe[,spe$sample_id==s],clustervar="label",auto_crop = T,point_size = 1.18,return_plots=T,height=1.875,width=1.875,spatial = T,sort_clust = FALSE,image_id = "lowres",alpha=0.5)
    sampsex <- tmpcd[sample_id==s,Sex]
    sampapo <- gsub(tmpcd[sample_id==s,APOE],pattern=" ",replacement="")
    sampanc <- tmpcd[sample_id==s,anc]
    sampage <- tmpcd[sample_id==s,Age]
    
    p <- p+
      ggtitle(paste0(s,"\n(",sampsex,", ",sampapo,", ",sampanc,", ",round(sampage),"yr)"))+
      guides(color="none",fill="none")+
      scale_color_manual(values=pal2)+
      scale_fill_manual(values=pal2)+
      theme(legend.position="none")+
      theme(axis.text.x=element_blank(),
            axis.text.y=element_blank(),
            plot.title=element_text(size=8,hjust=0.5),
            title=element_text(size=8,hjust=0.5))
        
    q <- rasterize(p,dpi=300,layers="Points",dev="cairo_png")
    return(q)
})

### save--3 plots per row x 2 rows (legend to go in the center of ea page)

pdf("plots/manuscript/SuppFigs/SupFig-Cluster_spotplots_allsamps.pdf",width=7.5,height=10,onefile=TRUE)
  i<-1
  for (i in seq(1,length(plt),6)){
    j <- i+5
    if(j>length(plt)){j <- length(plt)}
      do.call("grid.arrange",c(plt[i:j],ncol=3,nrow=2))
}

## save legend, mmodify in AI to place as desired withut messing around with other masks etc etc.  
p <- plt[[1]]
pdf("plots/manuscript/SuppFigs/Sfig-allspotplots_legendholder.pdf",height=5,width=5)
do.call("grid.arrange",list(arrangeGrob(p,leg.tmp),ncol=2,widths=c(3,2)))
dev.off()
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] colorout_1.3-0.2            spatialLIBD_1.18.0          SpatialExperiment_1.16.0   
 [4] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0 Biobase_2.66.0             
 [7] GenomicRanges_1.58.0        GenomeInfoDb_1.42.3         IRanges_2.40.1             
[10] S4Vectors_0.44.0            BiocGenerics_0.52.0         MatrixGenerics_1.18.1      
[13] matrixStats_1.5.0           gridExtra_2.3               ggplot2_3.5.2              
[16] data.table_1.17.4           rlang_1.1.5                

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3       rstudioapi_0.17.1        jsonlite_2.0.0          
  [4] magrittr_2.0.3           ggbeeswarm_0.7.2         magick_2.8.6            
  [7] farver_2.1.2             rmarkdown_2.29           BiocIO_1.16.0           
 [10] fields_16.3.1            zlibbioc_1.52.0          vctrs_0.6.5             
 [13] memoise_2.0.1            config_0.3.2             Rsamtools_2.22.0        
 [16] paletteer_1.6.0          RCurl_1.98-1.17          benchmarkme_1.0.8       
 [19] htmltools_0.5.8.1        S4Arrays_1.6.0           AnnotationHub_3.14.0    
 [22] curl_6.2.2               BiocNeighbors_2.0.1      SparseArray_1.6.2       
 [25] sass_0.4.10              bslib_0.9.0              htmlwidgets_1.6.4       
 [28] plotly_4.10.4            cachem_1.1.0             GenomicAlignments_1.42.0
 [31] mime_0.13                lifecycle_1.0.4          iterators_1.0.14        
 [34] pkgconfig_2.0.3          rsvd_1.0.5               Matrix_1.7-3            
 [37] R6_2.6.1                 fastmap_1.2.0            GenomeInfoDbData_1.2.13 
 [40] shiny_1.10.0             digest_0.6.37            rematch2_2.1.2          
 [43] AnnotationDbi_1.68.0     scater_1.34.1            rprojroot_2.0.4         
 [46] irlba_2.3.5.1            ExperimentHub_2.14.0     RSQLite_2.3.9           
 [49] beachmat_2.22.0          labeling_0.4.3           filelock_1.0.3          
 [52] httr_1.4.7               abind_1.4-8              compiler_4.4.3          
 [55] here_1.0.1               bit64_4.6.0-1            withr_3.0.2             
 [58] doParallel_1.0.17        attempt_0.3.1            BiocParallel_1.40.0     
 [61] viridis_0.6.5            DBI_1.2.3                maps_3.4.3              
 [64] sessioninfo_1.2.3        rappdirs_0.3.3           DelayedArray_0.32.0     
 [67] rjson_0.2.23             tools_4.4.3              vipor_0.4.7             
 [70] beeswarm_0.4.0           httpuv_1.6.16            glue_1.8.0              
 [73] restfulr_0.0.15          promises_1.3.2           grid_4.4.3              
 [76] generics_0.1.4           gtable_0.3.6             tidyr_1.3.1             
 [79] ScaledMatrix_1.14.0      BiocSingular_1.22.0      XVector_0.46.0          
 [82] ggrepel_0.9.6            BiocVersion_3.20.0       foreach_1.5.2           
 [85] pillar_1.10.2            spam_2.11-1              limma_3.62.2            
 [88] later_1.4.2              benchmarkmeData_1.0.4    dplyr_1.1.4             
 [91] BiocFileCache_2.14.0     lattice_0.22-6           rtracklayer_1.66.0      
 [94] bit_4.6.0                tidyselect_1.2.1         locfit_1.5-9.12         
 [97] scuttle_1.16.0           Biostrings_2.74.1        knitr_1.50              
[100] edgeR_4.4.2              xfun_0.52                statmod_1.5.0           
[103] DT_0.33                  UCSC.utils_1.2.0         lazyeval_0.2.2          
[106] yaml_2.3.10              shinyWidgets_0.9.0       evaluate_1.0.3          
[109] codetools_0.2-20         tibble_3.2.1             BiocManager_1.30.25     
[112] cli_3.6.4                xtable_1.8-4             jquerylib_0.1.4         
[115] golem_0.5.1              dichromat_2.0-0.1        Rcpp_1.0.14             
[118] dbplyr_2.5.0             png_0.1-8                XML_3.99-0.18           
[121] parallel_4.4.3           blob_1.2.4               dotCall64_1.2           
[124] bitops_1.0-9             viridisLite_0.4.2        scales_1.4.0            
[127] purrr_1.0.4              crayon_1.5.3             cowplot_1.1.3           
[130] KEGGREST_1.46.0         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-07-15
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────────────
 !  package              * version   date (UTC) lib source
    abind                  1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
    AnnotationDbi          1.68.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    AnnotationHub          3.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
    beachmat               2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
    benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
    benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocFileCache          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocIO                 1.16.0    2024-12-12 [1] Github (Bioconductor/BiocIO@8e80624)
    BiocManager            1.30.25   2024-08-28 [1] CRAN (R 4.4.1)
    BiocNeighbors          2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel           1.40.0    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocVersion            3.20.0    2024-05-04 [1] Bioconductor 3.20 (R 4.4.0)
    Biostrings             2.74.1    2024-12-16 [1] Bioconductor 3.20 (R 4.4.2)
    bit                    4.6.0     2025-03-06 [1] CRAN (R 4.4.1)
    bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.4.1)
    bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
    blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
    bslib                  0.9.0     2025-01-30 [1] CRAN (R 4.4.1)
    cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 VP cli                    3.6.4     2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.3)
    colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
    cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
    crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
    curl                   6.2.2     2025-03-24 [1] CRAN (R 4.4.1)
    data.table           * 1.17.4    2025-05-26 [1] CRAN (R 4.4.3)
    DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
    dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
    DelayedArray           0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37    2024-08-19 [1] CRAN (R 4.4.1)
    doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
    dotCall64              1.2       2024-10-04 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
    DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
    edgeR                  4.4.2     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate               1.0.3     2025-01-10 [1] CRAN (R 4.4.1)
    ExperimentHub          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
    fields                 16.3.1    2025-03-08 [1] CRAN (R 4.4.1)
    filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
    foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
    generics               0.1.4     2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13    2024-12-12 [1] Bioconductor
    GenomicAlignments      1.42.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    GenomicRanges        * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.2     2025-04-09 [1] CRAN (R 4.4.1)
    ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
    golem                  0.5.1     2024-08-27 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
    gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
    htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
    httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.4.1)
    httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
    iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
    jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
    jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.4.1)
    KEGGREST               1.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    knitr                  1.50      2025-03-16 [1] CRAN (R 4.4.2)
    labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
    later                  1.4.2     2025-04-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.3)
    lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
    lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
    limma                  3.62.2    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.4.1)
    magick                 2.8.6     2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
    maps                   3.4.3     2025-05-26 [1] CRAN (R 4.4.1)
    Matrix                 1.7-3     2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.4.1)
    memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
    mime                   0.13      2025-03-17 [1] CRAN (R 4.4.1)
    paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
    pillar                 1.10.2    2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
    plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
    png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
    promises               1.3.2     2024-11-28 [1] CRAN (R 4.4.1)
    purrr                  1.0.4     2025-02-05 [1] CRAN (R 4.4.1)
    R6                     2.6.1     2025-02-15 [1] CRAN (R 4.4.1)
    rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
    RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14    2025-01-12 [1] CRAN (R 4.4.1)
    RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.4.1)
    rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
    restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
    rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5     2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown              2.29      2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
    Rsamtools              2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    RSQLite                2.3.9     2024-12-03 [1] CRAN (R 4.4.1)
    rstudioapi             0.17.1    2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
    rtracklayer            1.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Arrays               1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sass                   0.4.10    2025-04-11 [1] CRAN (R 4.4.1)
    ScaledMatrix           1.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0     2025-04-24 [1] CRAN (R 4.4.1)
    scater                 1.34.1    2025-03-03 [1] Bioconductor 3.20 (R 4.4.2)
    scuttle                1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.4.1)
    shiny                  1.10.0    2024-12-14 [1] CRAN (R 4.4.1)
    shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    spam                   2.11-1    2025-01-20 [1] CRAN (R 4.4.1)
    SparseArray            1.6.2     2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    spatialLIBD          * 1.18.0    2024-11-07 [1] Bioconductor 3.20 (R 4.4.3)
    statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52      2025-04-02 [1] CRAN (R 4.4.1)
    XML                    3.99-0.18 2025-01-01 [1] CRAN (R 4.4.1)
    xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
    XVector                0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────────────────
