---
title: "SupFig-Spotplot_clusters"
author: "Bernie Mulvey"
date: "2025-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(ggrastr)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


read in the full per CAPTURE SPE and plot capture areas 
```{r}
spe <- readRDS("processed-data/04_QC/04-lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed.RDS")

# clean brain id 6119 (has (re-dis) after it))
spe$brnum[spe$brnum=="Br6119(re-dis)"] <- "Br6119"

# load clustering and the split-sections coldata, which has a column with pre-split spot ids we can use to match everything back to capture level
louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

lccd <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")
louvs <- merge.data.table(louvs,lccd[,.(rowname,original_spot_id)],by.x="rn",by.y="rowname")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno,by="clusid")

## also attach LCnm subsets because we will use them in this figure
lcnm <- fread("processed-data/10_subclustering/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnm <- lcnm[,.(rn,NMseg_subclus)]

louvs <- merge.data.table(louvs,lcnm,by="rn",all.x=T)
louvs[anno=="LC"&NMseg_subclus=="LC_NMpos",plot_anno:="LC (NM+)"]
louvs[anno=="LC"&NMseg_subclus=="LC_NMneg",plot_anno:="LC (NM-)"]

louvs <- DataFrame(louvs,row.names=louvs$original_spot_id)[colnames(spe),]
stopifnot(nrow(louvs)==ncol(spe))

colLabels(spe) <- louvs$plot_anno

rm(lcnm,louvs,anno)
gc(full=T)
```

### append demos info (YRI prop); clean up APOE genotypes
```{r}
colData(spe)$APOE[colData(spe)$APOE=="E4/ E4"] <- "E4/E4"
colData(spe)$APOE[colData(spe)$APOE=="E2/ E2"] <- "E2/E2"
colData(spe)$APOE[colData(spe)$APOE=="E3/ E4"] <- "E3/E4"
colData(spe)$APOE[colData(spe)$APOE=="E2/ E3"] <- "E2/E3"

yri <- fread("raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")
stopifnot(all(unique(spe$brnum) %in% yri$ID))
tmpcd <- as.data.table(colData(spe),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,yri[ID %in% tmpcd$brnum],by.x="brnum",by.y="ID")
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(spe),]
tmpcd$rn <- NULL
colData(spe) <- tmpcd

colData(spe)$Ancestry <- ifelse(colData(spe)$YRI>0.5,yes="AA",no="EA")

rm(tmpcd,yri)
```

## define helper function to extract a single legend - from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

make plots
```{r}
# load palette, which has 9 colors, and add a 10th for LC (NM-)

pal <- readRDS("plots/finalpal_somuchfornorainbows.RDS")
# set seed and add a 10th color to jointly display LCNM+ and -
set.seed(42)
pal2 <- Polychrome::createPalette(10,seedcolors=pal)

names(pal2)[1:9] <- names(pal)
names(pal2)[10] <- "LC (NM-)"
names(pal2)[names(pal2)=="LC"] <- "LC (NM+)"

stopifnot(all(names(pal2) %in% unique(spe$label)))
rm(pal)

## make a plot for a sample containing all 10 clusters values to extract the legend from and append globally later: V13M06-332_B1
tmpdat <- as.data.table(colData(spe))[sample_id=="V13M06-332_B1"]
leg.tmp <- ggplot(tmpdat,aes(x=array_col,y=array_row,color=label))+
  geom_point()+
  scale_color_manual(values=pal2)+
  labs(color="Domain")+
  theme(legend.text=element_markdown(size=7),
        legend.title=element_text(size=8),
        legend.background = element_blank(),
        legend.box = element_blank(),
        legend.key=element_blank(),
        legend.key.height=unit(0.1,"in"))

leg.tmp <- extract_legend(leg.tmp)
dev.off()

# defactor label
spe$label <- as.character(spe$label)

# get coldata as d.t. for ease of use below:
tmpcd <- as.data.table(colData(spe),keep.rownames=T)

# generate plots
plt <- lapply(unique(spe$sample_id),function(s){
  p <- vis_clus(spe[,spe$sample_id==s],clustervar="label",auto_crop = T,point_size = 0.75,return_plots=T,height=2.25,width=2.25,spatial = T,sort_clust = FALSE,image_id = "lowres",alpha=0.5)
    sampsex <- tmpcd[sample_id==s,Sex]
    sampapo <- tmpcd[sample_id==s,APOE]
    sampanc <- tmpcd[sample_id==s,Ancestry]
    sampage <- tmpcd[sample_id==s,Age]
    
    p <- p+
      ggtitle(paste0(s,"\n(",sampsex,", ",sampapo,", ",sampanc,", ",round(sampage),"yr)"))+
      guides(color="none",fill="none")+
      scale_fill_manual(values=pal2)+
      theme(legend.position="none")+
      theme(axis.text.x=element_blank(),
            axis.text.y=element_blank(),
            plot.title=element_text(size=8,hjust=0.5),
            title=element_text(size=8,hjust=0.5))
        
    q <- rasterize(p,dpi=300,layers="Points",dev="cairo_png")
    return(q)
})

### save--3 plots per row x 3 rows (legend to go across bottom of ea page in a 1.5 inch bottom bar to add per page in illustrator)

pdf("plots/manuscript/SuppFigs/SupFig-Cluster_spotplots_allsamps.pdf",width=7.5,height=8.5,onefile=TRUE)
i<-1
for (i in seq(1,length(plt),9)){
  j <- i+8
  if(j>length(plt)){j <- length(plt)}
    do.call("grid.arrange",c(plt[i:j],ncol=3,nrow=3))
}
dev.off()

## save legend, mmodify in AI to place as desired withut messing around with other masks etc etc.  
p <- plt[[1]]
pdf("plots/manuscript/SuppFigs/Sfig-allspotplots_legendholder.pdf",height=5,width=5)
do.call("grid.arrange",list(arrangeGrob(p,leg.tmp),ncol=2,widths=c(3,2)))
dev.off()
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] colorout_1.3-0.2            ggrastr_1.0.2              
 [3] spatialLIBD_1.18.0          SpatialExperiment_1.16.0   
 [5] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
 [7] Biobase_2.66.0              GenomicRanges_1.58.0       
 [9] GenomeInfoDb_1.42.3         IRanges_2.40.1             
[11] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[13] MatrixGenerics_1.18.1       matrixStats_1.5.0          
[15] gridExtra_2.3               ggtext_0.1.2               
[17] ggplot2_3.5.2               data.table_1.17.4          
[19] rlang_1.1.5                

loaded via a namespace (and not attached):
  [1] later_1.4.2              BiocIO_1.16.0            bitops_1.0-9            
  [4] filelock_1.0.3           fields_16.3.1            tibble_3.2.1            
  [7] XML_3.99-0.18            lifecycle_1.0.4          edgeR_4.4.2             
 [10] doParallel_1.0.17        rprojroot_2.0.4          lattice_0.22-6          
 [13] magrittr_2.0.3           limma_3.62.2             plotly_4.10.4           
 [16] sass_0.4.10              rmarkdown_2.29           jquerylib_0.1.4         
 [19] yaml_2.3.10              httpuv_1.6.16            spam_2.11-1             
 [22] sessioninfo_1.2.3        cowplot_1.1.3            DBI_1.2.3               
 [25] RColorBrewer_1.1-3       golem_0.5.1              maps_3.4.3              
 [28] abind_1.4-8              zlibbioc_1.52.0          purrr_1.0.4             
 [31] RCurl_1.98-1.17          rappdirs_0.3.3           GenomeInfoDbData_1.2.13 
 [34] ggrepel_0.9.6            irlba_2.3.5.1            commonmark_1.9.5        
 [37] codetools_0.2-20         DelayedArray_0.32.0      DT_0.33                 
 [40] scuttle_1.16.0           xml2_1.3.8               tidyselect_1.2.1        
 [43] UCSC.utils_1.2.0         farver_2.1.2             ScaledMatrix_1.14.0     
 [46] viridis_0.6.5            shinyWidgets_0.9.0       BiocFileCache_2.14.0    
 [49] GenomicAlignments_1.42.0 jsonlite_2.0.0           BiocNeighbors_2.0.1     
 [52] scater_1.34.1            iterators_1.0.14         foreach_1.5.2           
 [55] tools_4.4.3              Rcpp_1.0.14              glue_1.8.0              
 [58] SparseArray_1.6.2        xfun_0.52                here_1.0.1              
 [61] dplyr_1.1.4              withr_3.0.2              BiocManager_1.30.25     
 [64] fastmap_1.2.0            litedown_0.6             digest_0.6.37           
 [67] rsvd_1.0.5               R6_2.6.1                 mime_0.13               
 [70] colorspace_2.1-1         dichromat_2.0-0.1        markdown_2.0            
 [73] RSQLite_2.3.9            config_0.3.2             tidyr_1.3.1             
 [76] generics_0.1.4           rtracklayer_1.66.0       httr_1.4.7              
 [79] htmlwidgets_1.6.4        S4Arrays_1.6.0           scatterplot3d_0.3-44    
 [82] pkgconfig_2.0.3          gtable_0.3.6             blob_1.2.4              
 [85] XVector_0.46.0           htmltools_0.5.8.1        dotCall64_1.2           
 [88] scales_1.4.0             png_0.1-8                attempt_0.3.1           
 [91] knitr_1.50               rstudioapi_0.17.1        rjson_0.2.23            
 [94] curl_6.2.2               cachem_1.1.0             stringr_1.5.1           
 [97] Polychrome_1.5.1         BiocVersion_3.20.0       parallel_4.4.3          
[100] vipor_0.4.7              AnnotationDbi_1.68.0     restfulr_0.0.15         
[103] pillar_1.10.2            grid_4.4.3               vctrs_0.6.5             
[106] promises_1.3.2           BiocSingular_1.22.0      dbplyr_2.5.0            
[109] beachmat_2.22.0          xtable_1.8-4             beeswarm_0.4.0          
[112] paletteer_1.6.0          evaluate_1.0.3           magick_2.8.6            
[115] cli_3.6.4                locfit_1.5-9.12          compiler_4.4.3          
[118] Rsamtools_2.22.0         crayon_1.5.3             labeling_0.4.3          
[121] rematch2_2.1.2           ggbeeswarm_0.7.2         stringi_1.8.7           
[124] viridisLite_0.4.2        BiocParallel_1.40.0      Biostrings_2.74.1       
[127] lazyeval_0.2.2           Matrix_1.7-3             ExperimentHub_2.14.0    
[130] benchmarkme_1.0.8        bit64_4.6.0-1            KEGGREST_1.46.0         
[133] statmod_1.5.0            shiny_1.10.0             AnnotationHub_3.14.0    
[136] gridtext_0.1.5           memoise_2.0.1            bslib_0.9.0             
[139] benchmarkmeData_1.0.4    bit_4.6.0               
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-10-07
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package              * version   date (UTC) lib source
    abind                  1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
    AnnotationDbi          1.68.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    AnnotationHub          3.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
    beachmat               2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
    benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
    benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocFileCache          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocIO                 1.16.0    2024-12-12 [1] Github (Bioconductor/BiocIO@8e80624)
    BiocManager            1.30.25   2024-08-28 [1] CRAN (R 4.4.1)
    BiocNeighbors          2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel           1.40.0    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocVersion            3.20.0    2024-05-04 [1] Bioconductor 3.20 (R 4.4.0)
    Biostrings             2.74.1    2024-12-16 [1] Bioconductor 3.20 (R 4.4.2)
    bit                    4.6.0     2025-03-06 [1] CRAN (R 4.4.1)
    bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.4.1)
    bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
    blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
    bslib                  0.9.0     2025-01-30 [1] CRAN (R 4.4.1)
    cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 VP cli                    3.6.4     2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.3)
    colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1     2024-07-26 [1] CRAN (R 4.4.0)
    commonmark             1.9.5     2025-03-17 [1] CRAN (R 4.4.1)
    config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
    cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
    crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
    curl                   6.2.2     2025-03-24 [1] CRAN (R 4.4.1)
    data.table           * 1.17.4    2025-05-26 [1] CRAN (R 4.4.3)
    DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
    dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
    DelayedArray           0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37    2024-08-19 [1] CRAN (R 4.4.1)
    doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
    dotCall64              1.2       2024-10-04 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
    DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
    edgeR                  4.4.2     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate               1.0.3     2025-01-10 [1] CRAN (R 4.4.1)
    ExperimentHub          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
    fields                 16.3.1    2025-03-08 [1] CRAN (R 4.4.1)
    filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
    foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
    generics               0.1.4     2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13    2024-12-12 [1] Bioconductor
    GenomicAlignments      1.42.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    GenomicRanges        * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.2     2025-04-09 [1] CRAN (R 4.4.1)
    ggrastr              * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.4.1)
    ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
    golem                  0.5.1     2024-08-27 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
    htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
    httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.4.1)
    httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
    iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
    jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
    jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.4.1)
    KEGGREST               1.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    knitr                  1.50      2025-03-16 [1] CRAN (R 4.4.2)
    labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
    later                  1.4.2     2025-04-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.3)
    lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
    lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
    limma                  3.62.2    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    litedown               0.6       2025-02-27 [1] CRAN (R 4.4.3)
    locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.4.1)
    magick                 2.8.6     2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
    maps                   3.4.3     2025-05-26 [1] CRAN (R 4.4.1)
    markdown               2.0       2025-03-23 [1] CRAN (R 4.4.3)
    Matrix                 1.7-3     2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.4.1)
    memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
    mime                   0.13      2025-03-17 [1] CRAN (R 4.4.1)
    paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
    pillar                 1.10.2    2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
    plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
    png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
    Polychrome             1.5.1     2022-05-03 [1] CRAN (R 4.4.0)
    promises               1.3.2     2024-11-28 [1] CRAN (R 4.4.1)
    purrr                  1.0.4     2025-02-05 [1] CRAN (R 4.4.1)
    R6                     2.6.1     2025-02-15 [1] CRAN (R 4.4.1)
    rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
    RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14    2025-01-12 [1] CRAN (R 4.4.1)
    RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.4.1)
    rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
    restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
    rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5     2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown              2.29      2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
    Rsamtools              2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    RSQLite                2.3.9     2024-12-03 [1] CRAN (R 4.4.1)
    rstudioapi             0.17.1    2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
    rtracklayer            1.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Arrays               1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sass                   0.4.10    2025-04-11 [1] CRAN (R 4.4.1)
    ScaledMatrix           1.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0     2025-04-24 [1] CRAN (R 4.4.1)
    scater                 1.34.1    2025-03-03 [1] Bioconductor 3.20 (R 4.4.2)
    scatterplot3d          0.3-44    2023-05-05 [1] CRAN (R 4.4.0)
    scuttle                1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.4.1)
    shiny                  1.10.0    2024-12-14 [1] CRAN (R 4.4.1)
    shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    spam                   2.11-1    2025-01-20 [1] CRAN (R 4.4.1)
    SparseArray            1.6.2     2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    spatialLIBD          * 1.18.0    2024-11-07 [1] Bioconductor 3.20 (R 4.4.3)
    statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
    stringi                1.8.7     2025-03-27 [1] CRAN (R 4.4.1)
    stringr                1.5.1     2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52      2025-04-02 [1] CRAN (R 4.4.1)
    XML                    3.99-0.18 2025-01-01 [1] CRAN (R 4.4.1)
    xml2                   1.3.8     2025-03-14 [1] CRAN (R 4.4.1)
    xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
    XVector                0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
