---
title: "01to03-Plot_initial_dimreds_by_covars"
author: "Bernie Mulvey"
date: "2025-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(Cairo)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =8), plot.title = element_text(size = 9,hjust=0.5), strip.text = element_text(size=8), legend.text = element_text(size=8), legend.title = element_text(size=8,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load dimred mats and metadata
```{r}
metad <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")

dimr <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/02-PCA_harmo_UMAP_matrices_HiDevG_HVG_SVG_andmixed_featsets.RDS")

# covars to plot by: 
# categorical sample_id, capture_id,  APOE, Ancestry, Sex, brnum
# continuous sum_gene, sum_umi, expr_chrM_ratio, Age. make separate ones for clusters

dimr <- lapply(dimr,as.data.table,keep.rownames=T)
dimr <- dimr[c("PCA_HDG_SVG_2575", "UMAP_HDG_SVG_2575", "HARMONYlmbna_HDG_SVG_2575", "HARMONYlmbna_HDG_SVG_2575_UMAP")]

names(dimr) <- c("PCA","UMAP","HARMONY","HARMONY-UMAP")

covs <- c("sample_id", "capture_id", "APOE", "Ancestry", "Sex", "brnum", "sum_gene", "sum_umi", "expr_chrM_ratio", "Age")
```

## dimred plots
```{r}
bpmapply(d=dimr,f=names(dimr),SIMPLIFY=FALSE,BPPARAM = MulticoreParam(4),FUN=function(d,f){
    plotdr <- names(d)[c(1:3)] ## spot id and first 2 PCs/Harmos or 2 UMAP dims
    tmp <- copy(d)[,..plotdr]
    
    mdcol <- c("rowname",covs)
    
    tmp <- merge.data.table(tmp,metad[,..mdcol],by.x="rn",by.y="rowname")
    plts <- lapply(covs,function(x){
        if(x %in% c("sum_gene", "sum_umi", "expr_chrM_ratio", "Age")){
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=as.numeric(.data[[x]])))+
                rasterize(geom_point(size=0.2,stroke=0),dpi = 450,dev="cairo_png")+
                labs(title=paste0(f, " by ",x))+
                guides(color="none")
            
        } else {
            xv <- names(tmp)[2]
            yv <- names(tmp)[3]
            p <- ggplot(tmp,aes(x=.data[[xv]],y=.data[[yv]],color=.data[[x]]))+
                rasterize(geom_point(size=0.2,stroke=0),dpi = 450,dev="cairo")+
                labs(title=paste0(f, " by ",x))+
                guides(color="none")
        }
        
        return(p)
    })
    
    pdf(paste0("plots/manuscript/SuppFigs/SupFig-DimredCovarPlots_",f,".pdf"),width=8,height=10)
    do.call("grid.arrange",c(plts,ncol=2))
    dev.off()
})

```

## plotting clusters:
## define helper function to extract a single legend - from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

```{r}
louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
anno[plot_anno=="Oligo and Other Types Mixed",plot_anno:="Oligo & Other Mixed"]
anno[plot_anno=="*ACHE*, *SERT*, Neuropeptide Mixed",plot_anno:="*ACHE*, *SERT*, & Npep"]
anno[plot_anno=="*ACHE*, *SERT*, LC, Oligo Mixed",plot_anno:="*ACHE*, *SERT*, LC, & WM"]
anno[plot_anno=="Astro-Oligo Mixed",plot_anno:="Astro-Oligo  Mixed"] # single space prints badly with ggtext
anno[plot_anno=="Oligo-Astro Mixed",plot_anno:="Oligo-Astro  Mixed"] # single space prints badly with ggtext



louvs <- merge.data.table(louvs,anno,by="clusid")

dimr <- lapply(dimr,function(x){
   y <- merge.data.table(x,louvs[,.(rn,plot_anno)],by="rn")
   return(y)
})

pal <- readRDS("plots/finalpal_somuchfornorainbows.RDS")
# names(pal)
names(pal)[2] <- "*ACHE*, *SERT*, LC, & WM"
names(pal)[3] <- "*ACHE*, *SERT*, & Npep"
names(pal)[9] <- "Oligo & Other Mixed"
names(pal)[5] <- "Astro-Oligo  Mixed"
names(pal)[8] <- "Oligo-Astro  Mixed"
stopifnot(all(names(pal) %in% unique(louvs$plot_anno)))

## get legend then make all plots without
d <- dimr[[1]]
f <- names(dimr)[1]

plotdr <- c(names(d)[c(1:3)], "plot_anno") ## spot id and first 2 PCs/Harmos or 2 UMAP dims and cluster
    
xv <- names(d)[2]
yv <- names(d)[3]
p <- ggplot(d,aes(x=.data[[xv]],y=.data[[yv]],color=plot_anno))+
   rasterize(geom_point(size=0.2,stroke=0),dpi = 450,dev="cairo")+
   labs(title=paste0(f, " by Domain"))+
   scale_color_manual(values=pal)+
   guides(color=guide_legend(override.aes = list(size=2,stroke=0)))+
   labs(color="Domain")+
   theme(legend.text = element_markdown(size=8),legend.background = element_blank(),legend.key = element_blank())

tmpleg <- extract_legend(p)
dev.off()

rm(d,f,plotdr,xv,yv,p)
```


```{r}
plts <- bpmapply(d=dimr,f=names(dimr),SIMPLIFY=FALSE,BPPARAM = MulticoreParam(4),FUN=function(d,f){
    plotdr <- c(names(d)[c(1:3)], "plot_anno") ## spot id and first 2 PCs/Harmos or 2 UMAP dims and cluster
    
   xv <- names(d)[2]
   yv <- names(d)[3]
   p <- ggplot(d,aes(x=.data[[xv]],y=.data[[yv]],color=plot_anno))+
      rasterize(geom_point(size=0.3,stroke=0),dpi = 600,dev="cairo")+
      labs(title=paste0(f, " by Domain"))+
      scale_color_manual(values=pal)+
      guides(color="none")
   
   return(p)
})

pdf(paste0("plots/manuscript/SuppFigs/SupFig-DimredDomainPlots.pdf"),width=8,height=10)
do.call("grid.arrange",list(arrangeGrob(grobs=plts,ncol=2),tmpleg,ncol=2,widths=c(6.25,1.75)))
dev.off()

```

reprod
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0         parallelly_1.43.0           Cairo_1.6-2                
 [4] ggtext_0.1.2                ggrastr_1.0.2               colorout_1.3-0.2           
 [7] spatialLIBD_1.18.0          SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1
[10] SummarizedExperiment_1.36.0 Biobase_2.66.0              GenomicRanges_1.58.0       
[13] GenomeInfoDb_1.42.3         IRanges_2.40.1              S4Vectors_0.44.0           
[16] BiocGenerics_0.52.0         MatrixGenerics_1.18.1       matrixStats_1.5.0          
[19] gridExtra_2.3               ggplot2_3.5.2               data.table_1.17.4          
[22] rlang_1.1.5                

loaded via a namespace (and not attached):
  [1] later_1.4.2              BiocIO_1.16.0            bitops_1.0-9            
  [4] filelock_1.0.3           fields_16.3.1            tibble_3.2.1            
  [7] XML_3.99-0.18            lifecycle_1.0.4          edgeR_4.4.2             
 [10] doParallel_1.0.17        rprojroot_2.0.4          lattice_0.22-6          
 [13] magrittr_2.0.3           limma_3.62.2             plotly_4.10.4           
 [16] sass_0.4.10              rmarkdown_2.29           jquerylib_0.1.4         
 [19] yaml_2.3.10              httpuv_1.6.16            spam_2.11-1             
 [22] sessioninfo_1.2.3        cowplot_1.1.3            DBI_1.2.3               
 [25] RColorBrewer_1.1-3       golem_0.5.1              maps_3.4.3              
 [28] abind_1.4-8              zlibbioc_1.52.0          purrr_1.0.4             
 [31] RCurl_1.98-1.17          rappdirs_0.3.3           GenomeInfoDbData_1.2.13 
 [34] ggrepel_0.9.6            irlba_2.3.5.1            commonmark_1.9.5        
 [37] codetools_0.2-20         DelayedArray_0.32.0      DT_0.33                 
 [40] scuttle_1.16.0           xml2_1.3.8               tidyselect_1.2.1        
 [43] UCSC.utils_1.2.0         farver_2.1.2             ScaledMatrix_1.14.0     
 [46] viridis_0.6.5            shinyWidgets_0.9.0       BiocFileCache_2.14.0    
 [49] GenomicAlignments_1.42.0 jsonlite_2.0.0           BiocNeighbors_2.0.1     
 [52] scater_1.34.1            iterators_1.0.14         foreach_1.5.2           
 [55] tools_4.4.3              Rcpp_1.0.14              glue_1.8.0              
 [58] SparseArray_1.6.2        xfun_0.52                here_1.0.1              
 [61] dplyr_1.1.4              withr_3.0.2              BiocManager_1.30.25     
 [64] fastmap_1.2.0            litedown_0.6             digest_0.6.37           
 [67] rsvd_1.0.5               R6_2.6.1                 mime_0.13               
 [70] dichromat_2.0-0.1        markdown_2.0             RSQLite_2.3.9           
 [73] config_0.3.2             tidyr_1.3.1              generics_0.1.4          
 [76] rtracklayer_1.66.0       httr_1.4.7               htmlwidgets_1.6.4       
 [79] S4Arrays_1.6.0           pkgconfig_2.0.3          gtable_0.3.6            
 [82] blob_1.2.4               XVector_0.46.0           htmltools_0.5.8.1       
 [85] dotCall64_1.2            scales_1.4.0             png_0.1-8               
 [88] attempt_0.3.1            knitr_1.50               rstudioapi_0.17.1       
 [91] rjson_0.2.23             curl_6.2.2               cachem_1.1.0            
 [94] stringr_1.5.1            BiocVersion_3.20.0       parallel_4.4.3          
 [97] vipor_0.4.7              AnnotationDbi_1.68.0     restfulr_0.0.15         
[100] pillar_1.10.2            grid_4.4.3               vctrs_0.6.5             
[103] promises_1.3.2           BiocSingular_1.22.0      dbplyr_2.5.0            
[106] beachmat_2.22.0          xtable_1.8-4             beeswarm_0.4.0          
[109] paletteer_1.6.0          evaluate_1.0.3           magick_2.8.6            
[112] cli_3.6.4                locfit_1.5-9.12          compiler_4.4.3          
[115] Rsamtools_2.22.0         crayon_1.5.3             labeling_0.4.3          
[118] rematch2_2.1.2           ggbeeswarm_0.7.2         stringi_1.8.7           
[121] viridisLite_0.4.2        Biostrings_2.74.1        lazyeval_0.2.2          
[124] Matrix_1.7-3             ExperimentHub_2.14.0     benchmarkme_1.0.8       
[127] bit64_4.6.0-1            KEGGREST_1.46.0          statmod_1.5.0           
[130] shiny_1.10.0             AnnotationHub_3.14.0     gridtext_0.1.5          
[133] memoise_2.0.1            bslib_0.9.0              benchmarkmeData_1.0.4   
[136] bit_4.6.0               
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-07-18
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────────────
 !  package              * version   date (UTC) lib source
    abind                  1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
    AnnotationDbi          1.68.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    AnnotationHub          3.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
    beachmat               2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
    benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
    benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocFileCache          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocIO                 1.16.0    2024-12-12 [1] Github (Bioconductor/BiocIO@8e80624)
    BiocManager            1.30.25   2024-08-28 [1] CRAN (R 4.4.1)
    BiocNeighbors          2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel         * 1.40.0    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocVersion            3.20.0    2024-05-04 [1] Bioconductor 3.20 (R 4.4.0)
    Biostrings             2.74.1    2024-12-16 [1] Bioconductor 3.20 (R 4.4.2)
    bit                    4.6.0     2025-03-06 [1] CRAN (R 4.4.1)
    bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.4.1)
    bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
    blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
    bslib                  0.9.0     2025-01-30 [1] CRAN (R 4.4.1)
    cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
    Cairo                * 1.6-2     2023-11-28 [1] CRAN (R 4.4.0)
 VP cli                    3.6.4     2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.3)
    colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    commonmark             1.9.5     2025-03-17 [1] CRAN (R 4.4.1)
    config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
    cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
    crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
    curl                   6.2.2     2025-03-24 [1] CRAN (R 4.4.1)
    data.table           * 1.17.4    2025-05-26 [1] CRAN (R 4.4.3)
    DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
    dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
    DelayedArray           0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37    2024-08-19 [1] CRAN (R 4.4.1)
    doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
    dotCall64              1.2       2024-10-04 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
    DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
    edgeR                  4.4.2     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate               1.0.3     2025-01-10 [1] CRAN (R 4.4.1)
    ExperimentHub          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
    fields                 16.3.1    2025-03-08 [1] CRAN (R 4.4.1)
    filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
    foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
    generics               0.1.4     2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13    2024-12-12 [1] Bioconductor
    GenomicAlignments      1.42.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    GenomicRanges        * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.2     2025-04-09 [1] CRAN (R 4.4.1)
    ggrastr              * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.4.1)
    ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
    golem                  0.5.1     2024-08-27 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
    htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
    httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.4.1)
    httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
    iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
    jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
    jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.4.1)
    KEGGREST               1.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    knitr                  1.50      2025-03-16 [1] CRAN (R 4.4.2)
    labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
    later                  1.4.2     2025-04-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.3)
    lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
    lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
    limma                  3.62.2    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    litedown               0.6       2025-02-27 [1] CRAN (R 4.4.3)
    locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.4.1)
    magick                 2.8.6     2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
    maps                   3.4.3     2025-05-26 [1] CRAN (R 4.4.1)
    markdown               2.0       2025-03-23 [1] CRAN (R 4.4.3)
    Matrix                 1.7-3     2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.4.1)
    memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
    mime                   0.13      2025-03-17 [1] CRAN (R 4.4.1)
    paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
    parallelly           * 1.43.0    2025-03-24 [1] CRAN (R 4.4.1)
    pillar                 1.10.2    2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
    plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
    png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
    promises               1.3.2     2024-11-28 [1] CRAN (R 4.4.1)
    purrr                  1.0.4     2025-02-05 [1] CRAN (R 4.4.1)
    R6                     2.6.1     2025-02-15 [1] CRAN (R 4.4.1)
    rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
    RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14    2025-01-12 [1] CRAN (R 4.4.1)
    RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.4.1)
    rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
    restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
    rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5     2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown              2.29      2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
    Rsamtools              2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    RSQLite                2.3.9     2024-12-03 [1] CRAN (R 4.4.1)
    rstudioapi             0.17.1    2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
    rtracklayer            1.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Arrays               1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sass                   0.4.10    2025-04-11 [1] CRAN (R 4.4.1)
    ScaledMatrix           1.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0     2025-04-24 [1] CRAN (R 4.4.1)
    scater                 1.34.1    2025-03-03 [1] Bioconductor 3.20 (R 4.4.2)
    scuttle                1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.4.1)
    shiny                  1.10.0    2024-12-14 [1] CRAN (R 4.4.1)
    shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    spam                   2.11-1    2025-01-20 [1] CRAN (R 4.4.1)
    SparseArray            1.6.2     2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    spatialLIBD          * 1.18.0    2024-11-07 [1] Bioconductor 3.20 (R 4.4.3)
    statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
    stringi                1.8.7     2025-03-27 [1] CRAN (R 4.4.1)
    stringr                1.5.1     2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52      2025-04-02 [1] CRAN (R 4.4.1)
    XML                    3.99-0.18 2025-01-01 [1] CRAN (R 4.4.1)
    xml2                   1.3.8     2025-03-14 [1] CRAN (R 4.4.1)
    xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
    XVector                0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────────────────
