---
title: "SupFigPanel-RNAscope_QC_and_GeneXpr"
author: "Bernie Mulvey"
date: "2025-07-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = "#000000"), axis.title.x = element_text(size = 9,color = "#000000"), axis.text.y = element_text(size = 8,color = "#000000"), axis.title.y = element_text(size =9,color = "#000000"), plot.title = element_markdown(size = 10,hjust=0.5,color = "#000000"), strip.text = element_text(size=11), legend.text = element_text(size=7,color = "#000000"), legend.title = element_text(size=8,hjust=0.5,color = "#000000"),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## per capture area for 4 donors: 
Br1556, Br6423, Br1039 and Br5460
```{r}
spe <- readRDS("processed-data/04_QC/04-lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed.RDS")

spe <- spe[,spe$brnum %in% paste0("Br",c(1556,6423,1039,5460))]

spe <- scater::computeLibraryFactors(spe)
spe <- scater::logNormCounts(spe)
```


```{r}
samps <- lapply(unique(spe$sample_id),function(s){
    spe[,spe$sample_id==s] 
})
names(samps) <- unique(spe$sample_id)
rm(spe)

samps <- lapply(samps,function(x){
    rownames(x) <- rowData(x)$gene_name
return(x)})


plt <- mapply(x=samps,s=names(samps),SIMPLIFY=FALSE,function(x,s){
    
    maxth <- max(as.matrix(logcounts(x["TH",])))
    minth <- as.matrix(logcounts(x["TH",]))
    minth <- minth[minth>0]
    minth <- min(minth)
    
    p <- vis_gene(x, sampleid = s,geneid = "TH",spatial = T,image_id = "hires",alpha = 0.6,point_size = 1.5,return_plots=T,auto_crop = FALSE)+
        ## make the color range 0 to whatever the maxth value is
        scale_color_continuous(limits=c(minth,maxth),type="viridis",na.value=NA)+
        scale_fill_continuous(limits=c(minth,maxth),type="viridis",na.value=NA)+
        ggtitle(s)+
        labs(fill="logcounts<br>*TH*",color="none")+
        theme(axis.text.x=element_blank(),
              axis.text.y=element_blank(),
              legend.title = element_markdown(size=9,hjust=0.5),
              title=element_text(size=10,hjust=0.5),
              plot.title=element_text(size=10,hjust=0.5))
    return(p)
})

i<-1
for (i in c(1:length(plt))){
    s <- names(plt)[i]
    pdf(paste0("plots/manuscript/SuppFigs/SF_tissueQC_THexpr/",s,".pdf"),height=5,width=6)
    print(plt[[i]])
    dev.off()
}
```

repinf      
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.0 Patched (2025-05-21 r88220)- 06/03/2026
Platform: x86_64-conda-linux-gnu
Running under: Rocky Linux 9.4 (Blue Onyx)

Matrix products: default
BLAS:   /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/lib/libRblas.so 
LAPACK: /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/lib/libRlapack.so;  LAPACK version 3.12.1

locale:
 [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
 [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
 [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
 [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
 [9] LC_ADDRESS=C               LC_TELEPHONE=C            
[11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       

time zone: US/Eastern
tzcode source: system (glibc)

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.42.0         parallelly_1.44.0          
 [3] spatialLIBD_1.20.1          SpatialExperiment_1.18.1   
 [5] SingleCellExperiment_1.30.1 SummarizedExperiment_1.38.1
 [7] Biobase_2.68.0              GenomicRanges_1.60.0       
 [9] GenomeInfoDb_1.44.0         IRanges_2.42.0             
[11] S4Vectors_0.46.0            BiocGenerics_0.54.0        
[13] generics_0.1.4              MatrixGenerics_1.20.0      
[15] matrixStats_1.5.0           gridExtra_2.3              
[17] ggtext_0.1.2                ggplot2_3.5.2              
[19] data.table_1.17.2           testthat_3.2.3             
[21] devtools_2.4.5              usethis_3.1.0              
[23] colorout_1.3-3             

loaded via a namespace (and not attached):
  [1] later_1.4.2              BiocIO_1.18.0            bitops_1.0-9            
  [4] filelock_1.0.3           tibble_3.2.1             XML_3.99-0.18           
  [7] lifecycle_1.0.4          edgeR_4.6.2              doParallel_1.0.17       
 [10] lattice_0.22-7           magrittr_2.0.3           limma_3.64.0            
 [13] plotly_4.10.4            sass_0.4.10              jquerylib_0.1.4         
 [16] yaml_2.3.10              remotes_2.5.0            httpuv_1.6.16           
 [19] sessioninfo_1.2.3        pkgbuild_1.4.7           cowplot_1.1.3           
 [22] DBI_1.2.3                RColorBrewer_1.1-3       golem_0.5.1             
 [25] abind_1.4-8              pkgload_1.4.0            purrr_1.0.4             
 [28] RCurl_1.98-1.17          rappdirs_0.3.3           circlize_0.4.16         
 [31] GenomeInfoDbData_1.2.14  ggrepel_0.9.6            irlba_2.3.5.1           
 [34] commonmark_1.9.5         codetools_0.2-20         DelayedArray_0.34.1     
 [37] DT_0.33                  scuttle_1.18.0           xml2_1.3.8              
 [40] tidyselect_1.2.1         shape_1.4.6.1            UCSC.utils_1.4.0        
 [43] farver_2.1.2             ScaledMatrix_1.16.0      viridis_0.6.5           
 [46] shinyWidgets_0.9.0       BiocFileCache_2.16.0     GenomicAlignments_1.44.0
 [49] jsonlite_2.0.0           GetoptLong_1.0.5         BiocNeighbors_2.2.0     
 [52] ellipsis_0.3.2           scater_1.36.0            iterators_1.0.14        
 [55] foreach_1.5.2            tools_4.5.0              Rcpp_1.0.14             
 [58] glue_1.8.0               SparseArray_1.8.0        xfun_0.52               
 [61] dplyr_1.1.4              withr_3.0.2              BiocManager_1.30.25     
 [64] fastmap_1.2.0            litedown_0.7             digest_0.6.37           
 [67] rsvd_1.0.5               R6_2.6.1                 mime_0.13               
 [70] colorspace_2.1-1         markdown_2.0             dichromat_2.0-0.1       
 [73] RSQLite_2.3.11           config_0.3.2             tidyr_1.3.1             
 [76] rtracklayer_1.68.0       httr_1.4.7               htmlwidgets_1.6.4       
 [79] S4Arrays_1.8.0           pkgconfig_2.0.3          gtable_0.3.6            
 [82] blob_1.2.4               ComplexHeatmap_2.24.0    XVector_0.48.0          
 [85] brio_1.1.5               htmltools_0.5.8.1        profvis_0.4.0           
 [88] clue_0.3-66              scales_1.4.0             png_0.1-8               
 [91] attempt_0.3.1            rjson_0.2.23             curl_6.2.2              
 [94] cachem_1.1.0             GlobalOptions_0.1.2      stringr_1.5.1           
 [97] BiocVersion_3.21.1       parallel_4.5.0           miniUI_0.1.2            
[100] vipor_0.4.7              AnnotationDbi_1.70.0     restfulr_0.0.15         
[103] pillar_1.10.2            grid_4.5.0               vctrs_0.6.5             
[106] urlchecker_1.0.1         promises_1.3.2           BiocSingular_1.24.0     
[109] dbplyr_2.5.0             beachmat_2.24.0          xtable_1.8-4            
[112] cluster_2.1.8.1          beeswarm_0.4.0           paletteer_1.6.0         
[115] magick_2.8.6             cli_3.6.5                locfit_1.5-9.12         
[118] compiler_4.5.0           Rsamtools_2.24.0         rlang_1.1.6             
[121] crayon_1.5.3             labeling_0.4.3           rematch2_2.1.2          
[124] fs_1.6.6                 ggbeeswarm_0.7.2         stringi_1.8.7           
[127] viridisLite_0.4.2        Biostrings_2.76.0        lazyeval_0.2.2          
[130] Matrix_1.7-3             ExperimentHub_2.16.0     benchmarkme_1.0.8       
[133] bit64_4.6.0-1            KEGGREST_1.48.0          statmod_1.5.0           
[136] shiny_1.10.0             AnnotationHub_3.16.0     gridtext_0.1.6          
[139] memoise_2.0.1            bslib_0.9.0              benchmarkmeData_1.0.4   
[142] bit_4.6.0               
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.0 Patched (2025-05-21 r88220)
 os       Rocky Linux 9.4 (Blue Onyx)
 system   x86_64, linux-gnu
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       US/Eastern
 date     2026-06-03
 pandoc   3.7.0.1 @ /jhpce/shared/community/core/conda_R/4.5/bin/pandoc
 quarto   NA

─ Packages ──────────────────────────────────────────────────────────────────────────────────────────────────────
 package              * version   date (UTC) lib source
 abind                  1.4-8     2024-09-12 [2] CRAN (R 4.5.0)
 AnnotationDbi          1.70.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 AnnotationHub          3.16.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 attempt                0.3.1     2020-05-03 [2] CRAN (R 4.5.0)
 beachmat               2.24.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 beeswarm               0.4.0     2021-06-01 [2] CRAN (R 4.5.0)
 benchmarkme            1.0.8     2022-06-12 [2] CRAN (R 4.5.0)
 benchmarkmeData        1.0.4     2020-04-23 [2] CRAN (R 4.5.0)
 Biobase              * 2.68.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 BiocFileCache          2.16.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 BiocGenerics         * 0.54.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 BiocIO                 1.18.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 BiocManager            1.30.25   2024-08-28 [2] CRAN (R 4.5.0)
 BiocNeighbors          2.2.0     2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 BiocParallel         * 1.42.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 BiocSingular           1.24.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 BiocVersion            3.21.1    2024-10-29 [2] Bioconductor 3.21 (R 4.5.0)
 Biostrings             2.76.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 bit                    4.6.0     2025-03-06 [2] CRAN (R 4.5.0)
 bit64                  4.6.0-1   2025-01-16 [2] CRAN (R 4.5.0)
 bitops                 1.0-9     2024-10-03 [2] CRAN (R 4.5.0)
 blob                   1.2.4     2023-03-17 [2] CRAN (R 4.5.0)
 brio                   1.1.5     2024-04-24 [2] CRAN (R 4.5.0)
 bslib                  0.9.0     2025-01-30 [2] CRAN (R 4.5.0)
 cachem                 1.1.0     2024-05-16 [2] CRAN (R 4.5.0)
 circlize               0.4.16    2024-02-20 [2] CRAN (R 4.5.0)
 cli                    3.6.5     2025-04-23 [2] CRAN (R 4.5.0)
 clue                   0.3-66    2024-11-13 [2] CRAN (R 4.5.0)
 cluster                2.1.8.1   2025-03-12 [3] CRAN (R 4.5.0)
 codetools              0.2-20    2024-03-31 [3] CRAN (R 4.5.0)
 colorout             * 1.3-3     2026-05-09 [1] Github (jalvesaq/colorout@64863bb)
 colorspace             2.1-1     2024-07-26 [2] CRAN (R 4.5.0)
 commonmark             1.9.5     2025-03-17 [2] CRAN (R 4.5.0)
 ComplexHeatmap         2.24.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 config                 0.3.2     2023-08-30 [2] CRAN (R 4.5.0)
 cowplot                1.1.3     2024-01-22 [2] CRAN (R 4.5.0)
 crayon                 1.5.3     2024-06-20 [2] CRAN (R 4.5.0)
 curl                   6.2.2     2025-03-24 [2] CRAN (R 4.5.0)
 data.table           * 1.17.2    2025-05-12 [2] CRAN (R 4.5.0)
 DBI                    1.2.3     2024-06-02 [2] CRAN (R 4.5.0)
 dbplyr                 2.5.0     2024-03-19 [2] CRAN (R 4.5.0)
 DelayedArray           0.34.1    2025-04-17 [2] Bioconductor 3.21 (R 4.5.0)
 devtools             * 2.4.5     2022-10-11 [2] CRAN (R 4.5.0)
 dichromat              2.0-0.1   2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.37    2024-08-19 [2] CRAN (R 4.5.0)
 doParallel             1.0.17    2022-02-07 [2] CRAN (R 4.5.0)
 dplyr                  1.1.4     2023-11-17 [2] CRAN (R 4.5.0)
 DT                     0.33      2024-04-04 [2] CRAN (R 4.5.0)
 edgeR                  4.6.2     2025-05-07 [2] Bioconductor 3.21 (R 4.5.0)
 ellipsis               0.3.2     2021-04-29 [2] CRAN (R 4.5.0)
 ExperimentHub          2.16.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 farver                 2.1.2     2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0     2024-05-15 [2] CRAN (R 4.5.0)
 filelock               1.0.3     2023-12-11 [2] CRAN (R 4.5.0)
 foreach                1.5.2     2022-02-02 [2] CRAN (R 4.5.0)
 fs                     1.6.6     2025-04-12 [2] CRAN (R 4.5.0)
 generics             * 0.1.4     2025-05-09 [2] CRAN (R 4.5.0)
 GenomeInfoDb         * 1.44.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 GenomeInfoDbData       1.2.14    2025-05-21 [2] Bioconductor
 GenomicAlignments      1.44.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 GenomicRanges        * 1.60.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 GetoptLong             1.0.5     2020-12-15 [2] CRAN (R 4.5.0)
 ggbeeswarm             0.7.2     2023-04-29 [2] CRAN (R 4.5.0)
 ggplot2              * 3.5.2     2025-04-09 [2] CRAN (R 4.5.0)
 ggrepel                0.9.6     2024-09-07 [2] CRAN (R 4.5.0)
 ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.5.0)
 GlobalOptions          0.1.2     2020-06-10 [2] CRAN (R 4.5.0)
 glue                   1.8.0     2024-09-30 [2] CRAN (R 4.5.0)
 golem                  0.5.1     2024-08-27 [2] CRAN (R 4.5.0)
 gridExtra            * 2.3       2017-09-09 [2] CRAN (R 4.5.0)
 gridtext               0.1.6     2026-02-19 [1] CRAN (R 4.5.0)
 gtable                 0.3.6     2024-10-25 [2] CRAN (R 4.5.0)
 htmltools              0.5.8.1   2024-04-04 [2] CRAN (R 4.5.0)
 htmlwidgets            1.6.4     2023-12-06 [2] CRAN (R 4.5.0)
 httpuv                 1.6.16    2025-04-16 [2] CRAN (R 4.5.0)
 httr                   1.4.7     2023-08-15 [2] CRAN (R 4.5.0)
 IRanges              * 2.42.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 irlba                  2.3.5.1   2022-10-03 [2] CRAN (R 4.5.0)
 iterators              1.0.14    2022-02-05 [2] CRAN (R 4.5.0)
 jquerylib              0.1.4     2021-04-26 [2] CRAN (R 4.5.0)
 jsonlite               2.0.0     2025-03-27 [2] CRAN (R 4.5.0)
 KEGGREST               1.48.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 labeling               0.4.3     2023-08-29 [2] CRAN (R 4.5.0)
 later                  1.4.2     2025-04-08 [2] CRAN (R 4.5.0)
 lattice                0.22-7    2025-04-02 [3] CRAN (R 4.5.0)
 lazyeval               0.2.2     2019-03-15 [2] CRAN (R 4.5.0)
 lifecycle              1.0.4     2023-11-07 [2] CRAN (R 4.5.0)
 limma                  3.64.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 litedown               0.7       2025-04-08 [2] CRAN (R 4.5.0)
 locfit                 1.5-9.12  2025-03-05 [2] CRAN (R 4.5.0)
 magick                 2.8.6     2025-03-23 [2] CRAN (R 4.5.0)
 magrittr               2.0.3     2022-03-30 [2] CRAN (R 4.5.0)
 markdown               2.0       2025-03-23 [2] CRAN (R 4.5.0)
 Matrix                 1.7-3     2025-03-11 [3] CRAN (R 4.5.0)
 MatrixGenerics       * 1.20.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 matrixStats          * 1.5.0     2025-01-07 [2] CRAN (R 4.5.0)
 memoise                2.0.1     2021-11-26 [2] CRAN (R 4.5.0)
 mime                   0.13      2025-03-17 [2] CRAN (R 4.5.0)
 miniUI                 0.1.2     2025-04-17 [2] CRAN (R 4.5.0)
 paletteer              1.6.0     2024-01-21 [2] CRAN (R 4.5.0)
 parallelly           * 1.44.0    2025-05-07 [2] CRAN (R 4.5.0)
 pillar                 1.10.2    2025-04-05 [2] CRAN (R 4.5.0)
 pkgbuild               1.4.7     2025-03-24 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3     2019-09-22 [2] CRAN (R 4.5.0)
 pkgload                1.4.0     2024-06-28 [2] CRAN (R 4.5.0)
 plotly                 4.10.4    2024-01-13 [2] CRAN (R 4.5.0)
 png                    0.1-8     2022-11-29 [2] CRAN (R 4.5.0)
 profvis                0.4.0     2024-09-20 [2] CRAN (R 4.5.0)
 promises               1.3.2     2024-11-28 [2] CRAN (R 4.5.0)
 purrr                  1.0.4     2025-02-05 [2] CRAN (R 4.5.0)
 R6                     2.6.1     2025-02-15 [2] CRAN (R 4.5.0)
 rappdirs               0.3.3     2021-01-31 [2] CRAN (R 4.5.0)
 RColorBrewer           1.1-3     2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.0.14    2025-01-12 [2] CRAN (R 4.5.0)
 RCurl                  1.98-1.17 2025-03-22 [2] CRAN (R 4.5.0)
 rematch2               2.1.2     2020-05-01 [2] CRAN (R 4.5.0)
 remotes                2.5.0     2024-03-17 [2] CRAN (R 4.5.0)
 restfulr               0.0.15    2022-06-16 [2] CRAN (R 4.5.0)
 rjson                  0.2.23    2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.6     2025-04-11 [2] CRAN (R 4.5.0)
 Rsamtools              2.24.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 RSQLite                2.3.11    2025-05-04 [2] CRAN (R 4.5.0)
 rsvd                   1.0.5     2021-04-16 [2] CRAN (R 4.5.0)
 rtracklayer            1.68.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 S4Arrays               1.8.0     2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 S4Vectors            * 0.46.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 sass                   0.4.10    2025-04-11 [2] CRAN (R 4.5.0)
 ScaledMatrix           1.16.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 scales                 1.4.0     2025-04-24 [2] CRAN (R 4.5.0)
 scater                 1.36.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 scuttle                1.18.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 sessioninfo            1.2.3     2025-02-05 [2] CRAN (R 4.5.0)
 shape                  1.4.6.1   2024-02-23 [2] CRAN (R 4.5.0)
 shiny                  1.10.0    2024-12-14 [2] CRAN (R 4.5.0)
 shinyWidgets           0.9.0     2025-02-21 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.30.1    2025-05-07 [2] Bioconductor 3.21 (R 4.5.0)
 SparseArray            1.8.0     2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 SpatialExperiment    * 1.18.1    2025-05-11 [2] Bioconductor 3.21 (R 4.5.0)
 spatialLIBD          * 1.20.1    2025-05-01 [2] Bioconductor 3.21 (R 4.5.0)
 statmod                1.5.0     2023-01-06 [2] CRAN (R 4.5.0)
 stringi                1.8.7     2025-03-27 [2] CRAN (R 4.5.0)
 stringr                1.5.1     2023-11-14 [2] CRAN (R 4.5.0)
 SummarizedExperiment * 1.38.1    2025-04-30 [2] Bioconductor 3.21 (R 4.5.0)
 testthat             * 3.2.3     2025-01-13 [2] CRAN (R 4.5.0)
 tibble                 3.2.1     2023-03-20 [2] CRAN (R 4.5.0)
 tidyr                  1.3.1     2024-01-24 [2] CRAN (R 4.5.0)
 tidyselect             1.2.1     2024-03-11 [2] CRAN (R 4.5.0)
 UCSC.utils             1.4.0     2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 urlchecker             1.0.1     2021-11-30 [2] CRAN (R 4.5.0)
 usethis              * 3.1.0     2024-11-26 [2] CRAN (R 4.5.0)
 vctrs                  0.6.5     2023-12-01 [2] CRAN (R 4.5.0)
 vipor                  0.4.7     2023-12-18 [2] CRAN (R 4.5.0)
 viridis                0.6.5     2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite            0.4.2     2023-05-02 [2] CRAN (R 4.5.0)
 withr                  3.0.2     2024-10-28 [2] CRAN (R 4.5.0)
 xfun                   0.52      2025-04-02 [2] CRAN (R 4.5.0)
 XML                    3.99-0.18 2025-01-01 [2] CRAN (R 4.5.0)
 xml2                   1.3.8     2025-03-14 [2] CRAN (R 4.5.0)
 xtable                 1.8-4     2019-04-21 [2] CRAN (R 4.5.0)
 XVector                0.48.0    2025-04-15 [2] Bioconductor 3.21 (R 4.5.0)
 yaml                   2.3.10    2024-07-26 [2] CRAN (R 4.5.0)

 [1] /users/hdivecha/R/4.5
 [2] /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/site-library
 [3] /jhpce/shared/community/core/conda_R/4.5/R/lib64/R/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────────────────────────────────
