---
title: "SuppFig-CARTPT_LClocalSVG"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(escheR)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
options('styler.addins_style_transformer' = 'biocthis::bioc_style()')
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn='')
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost='localhost')


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8,color = '#000000'), axis.title.x = element_text(size = 9,color = '#000000'), axis.text.y = element_text(size = 8,color = '#000000'), axis.title.y = element_text(size =9,color = '#000000'), plot.title = element_markdown(size = 10,hjust=0.5,color = '#000000'), strip.text = element_text(size=11), legend.text = element_text(size=7,color = '#000000'), legend.title = element_text(size=8,hjust=0.5,color = '#000000'),axis.ticks = element_line(linewidth=0.5),panel.grid.major = element_line(linewidth=0.5),panel.grid.minor=element_line(linewidth=0.25),plot.title.position='plot'))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace('datasets')
```

```{r}
lc3 <- readRDS("processed-data/14_Samui_prep/01-Samui_TissSect_SPE_RotsMirrors.RDS")
lc3@int_metadata$imgData <- DataFrame()
gc(full=T)
## recreate logcounts assay which is absent from this lite RDS
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)
## and restore rownames to ensembl identifiers
rownames(lc3) <- rowData(lc3)$gene_id


louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno[,.(clusid,anno)],by="clusid",all.x=T)
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$anno


# load SVG table of nomsignif by sample (yes=rank listed, no=NA)

svgsum <- fread("processed-data/09_local_svgs/01d-LC prelim nnSVG 5addnl recovered samples nomsig genes and ranks per sample.txt")

## subset to genes with ≥ 7 tissue sections significant of 70 analyzed and previously marked interesting
keepg <- "CARTPT"



plotg <- svgsum[gene_name %in% keepg,gene_id]
## melt into a table of gene-nonNA samples
plotg <- melt.data.table(svgsum[gene_id %in% plotg],id.vars=c("gene_id","gene_name","meanrank"))[!is.na(value)]


# for memory conservation during parallelization, subset the SPE to the gene

lc3 <- lc3[unique(plotg$gene_id),]

rm(svgsum,louvs,anno)
gc(full=T)

## set labels to LC or other; make a palette of neon blue, NA
colData(lc3)$label <- ifelse(colData(lc3)$label=="LC","LC","other")
lc3$label <- factor(colData(lc3)$label,levels=c("LC","other"))
pal <- c("LC"="darkred","other"=NA)

## plot the SVGs (20 samples)
## plot the SVGs
library(parallel)
setDTthreads(1,restore_after_fork = FALSE)
lapply(rownames(lc3),FUN=function(g){
    ## append significance info to sample ids, drop samples that weren't analyzed
    tmpspe <- lc3[g,lc3$sample_id %in% plotg[gene_id==g,variable]]

    samps <- lapply(unique(tmpspe$sample_id),FUN=function(s){
        tmp <- tmpspe[,tmpspe$sample_id==s]
        return(tmp)
    })
    names(samps) <- unique(tmpspe$sample_id)
    rm(tmpspe)
    gc(full=T)

    plts <- bpmapply(s=samps,n=names(samps),SIMPLIFY=FALSE,BPPARAM=MulticoreParam(8),FUN=function(s,n){
        s$`log counts` <- as.numeric(logcounts(s)[g,])
        s$pltalph <- ifelse(s$label=="LC",yes=1,no=0.5)

        p <- make_escheR(s)
        p <- p |> add_ground("label",stroke=0.8,point_size=1.7)
        p <- p |> add_fill("log counts",size=1.7,point_size = 1.7)
        p <- p+aes(alpha=pltalph)


        p <- p+
            scale_alpha_continuous(range = c(0.5,1),guide="none")+
            scale_fill_continuous(type = "viridis")+
            scale_color_manual(values=pal,na.value = NA)+
            guides(color="none",fill=guide_colorbar(override.aes=list(size=1.5,alpha=1)))+
            ggtitle(label=paste0(n," ",unique(plotg[gene_id==g,gene_name])))+
            theme(plot.title.position = "plot",plot.title = element_text(size=9,hjust=0.5),legend.key.size = ggplot2::unit(0.075,"in"),legend.title=element_text(size=7,hjust=0.5),legend.text=element_text(size=6))


    })

    pdf(paste0("plots/manuscript/SuppFigs/",unique(plotg[gene_id==g,gene_name]),"_LClocalSVG.pdf"),height=10,width=7.5,onefile=TRUE)
    i<-1
    for(i in seq(1,length(plts),8)){
        if(length(plts)<=i+7){
            do.call(grid.arrange,c(plts[c(i:length(plts))],ncol=2))
        } else {
            do.call(grid.arrange,c(plts[c(i:(i+7))],ncol=2))
        }
    }

    dev.off()
})
```


seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics 
[5] grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0        
 [2] parallelly_1.43.0          
 [3] escheR_1.6.0               
 [4] SpatialExperiment_1.16.0   
 [5] SingleCellExperiment_1.28.1
 [6] SummarizedExperiment_1.36.0
 [7] Biobase_2.66.0             
 [8] GenomicRanges_1.58.0       
 [9] GenomeInfoDb_1.42.3        
[10] IRanges_2.40.1             
[11] S4Vectors_0.44.0           
[12] BiocGenerics_0.52.0        
[13] MatrixGenerics_1.18.1      
[14] matrixStats_1.5.0          
[15] gridExtra_2.3              
[16] ggtext_0.1.2               
[17] ggplot2_3.5.2              
[18] data.table_1.17.4          
[19] rlang_1.1.5                

loaded via a namespace (and not attached):
 [1] beeswarm_0.4.0         
 [2] gtable_0.3.6           
 [3] rjson_0.2.23           
 [4] xfun_0.52              
 [5] ggrepel_0.9.6          
 [6] lattice_0.22-6         
 [7] vctrs_0.6.5            
 [8] tools_4.4.3            
 [9] generics_0.1.4         
[10] tibble_3.2.1           
[11] pkgconfig_2.0.3        
[12] BiocNeighbors_2.0.1    
[13] Matrix_1.7-3           
[14] RColorBrewer_1.1-3     
[15] lifecycle_1.0.4        
[16] GenomeInfoDbData_1.2.13
[17] compiler_4.4.3         
[18] farver_2.1.2           
[19] codetools_0.2-20       
[20] vipor_0.4.7            
[21] pillar_1.10.2          
[22] crayon_1.5.3           
[23] DelayedArray_0.32.0    
[24] viridis_0.6.5          
[25] magick_2.8.6           
[26] abind_1.4-8            
[27] rsvd_1.0.5             
[28] tidyselect_1.2.1       
[29] BiocSingular_1.22.0    
[30] dplyr_1.1.4            
[31] labeling_0.4.3         
[32] rprojroot_2.0.4        
[33] grid_4.4.3             
[34] here_1.0.1             
[35] cli_3.6.4              
[36] SparseArray_1.6.2      
[37] scater_1.34.1          
[38] magrittr_2.0.3         
[39] S4Arrays_1.6.0         
[40] dichromat_2.0-0.1      
[41] withr_3.0.2            
[42] scales_1.4.0           
[43] UCSC.utils_1.2.0       
[44] ggbeeswarm_0.7.2       
[45] XVector_0.46.0         
[46] httr_1.4.7             
[47] ScaledMatrix_1.14.0    
[48] beachmat_2.22.0        
[49] evaluate_1.0.3         
[50] knitr_1.50             
[51] viridisLite_0.4.2      
[52] irlba_2.3.5.1          
[53] gridtext_0.1.5         
[54] Rcpp_1.0.14            
[55] glue_1.8.0             
[56] scuttle_1.16.0         
[57] xml2_1.3.8             
[58] jsonlite_2.0.0         
[59] R6_2.6.1               
[60] zlibbioc_1.52.0        
> sessioninfo::session_info()
─ Session info ──────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-07-30
 pandoc   NA
 quarto   NA

─ Packages ──────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    beachmat               2.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocNeighbors          2.0.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel         * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 VP cli                    3.6.4   2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.3)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.17.4  2025-05-26 [1] CRAN (R 4.4.3)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1 2022-05-02 [1] CRAN (R 4.4.0)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    escheR               * 1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    generics               0.1.4   2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3  2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.2   2025-04-09 [1] CRAN (R 4.4.1)
    ggrepel                0.9.6   2024-09-07 [1] CRAN (R 4.4.1)
    ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
    jsonlite               2.0.0   2025-03-27 [1] CRAN (R 4.4.1)
    knitr                  1.50    2025-03-16 [1] CRAN (R 4.4.2)
    labeling               0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.3)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magick                 2.8.6   2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    Matrix                 1.7-3   2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1  2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    parallelly           * 1.43.0  2025-03-24 [1] CRAN (R 4.4.1)
    pillar                 1.10.2  2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.6.1   2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer           1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5   2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ScaledMatrix           1.14.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0   2025-04-24 [1] CRAN (R 4.4.1)
    scater                 1.34.1  2025-03-03 [1] Bioconductor 3.20 (R 4.4.2)
    scuttle                1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.3   2025-02-05 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.2   2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52    2025-04-02 [1] CRAN (R 4.4.1)
    xml2                   1.3.8   2025-03-14 [1] CRAN (R 4.4.1)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────