---
title: "02-Compare LFF LC-Other Markersto Mouse LC Perturbation TRAP-seq Experiments"
author: "Bernie Mulvey"
date: "2025-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### load registrations
```{r}
trapregs <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/02-LouvLC_within_epidemRiskGrps_reg_to_msLCtrap.txt")

trapregs <- melt.data.table(trapregs,id.vars=c("LFF_subgroup","topNtrapgene"))
trapregs[variable=="Sal",variable:="Saline (DSP4 expt)"]
trapregs[variable=="YFP",variable:="YFP (hTyr expt)"]

pdf("plots/08_validitycheck_25hdg75svg_louv1/02-louvLC-other_by_E4_registr_to_LCtrapseq.pdf",height=6,width=7.5)
ggplot(trapregs,aes(x=log10(topNtrapgene),y=value,color=LFF_subgroup))+
    geom_line()+
    geom_point()+
    facet_wrap(.~variable)
dev.off()
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics 
[5] grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0        
 [2] parallelly_1.40.1          
 [3] spatialLIBD_1.18.0         
 [4] SpatialExperiment_1.16.0   
 [5] SingleCellExperiment_1.28.1
 [6] SummarizedExperiment_1.36.0
 [7] Biobase_2.66.0             
 [8] GenomicRanges_1.58.0       
 [9] GenomeInfoDb_1.42.1        
[10] IRanges_2.40.1             
[11] S4Vectors_0.44.0           
[12] BiocGenerics_0.52.0        
[13] MatrixGenerics_1.18.0      
[14] matrixStats_1.4.1          
[15] gridExtra_2.3              
[16] ggtext_0.1.2               
[17] ggplot2_3.5.1              
[18] data.table_1.16.4          
[19] rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.4.1             
  [2] BiocIO_1.16.0           
  [3] bitops_1.0-9            
  [4] filelock_1.0.3          
  [5] fields_16.3             
  [6] tibble_3.2.1            
  [7] XML_3.99-0.17           
  [8] lifecycle_1.0.4         
  [9] edgeR_4.4.1             
 [10] doParallel_1.0.17       
 [11] rprojroot_2.0.4         
 [12] lattice_0.22-6          
 [13] magrittr_2.0.3          
 [14] limma_3.62.1            
 [15] plotly_4.10.4           
 [16] sass_0.4.9              
 [17] jquerylib_0.1.4         
 [18] yaml_2.3.10             
 [19] httpuv_1.6.15           
 [20] spam_2.11-0             
 [21] sessioninfo_1.2.2       
 [22] cowplot_1.1.3           
 [23] DBI_1.2.3               
 [24] RColorBrewer_1.1-3      
 [25] golem_0.5.1             
 [26] maps_3.4.2.1            
 [27] abind_1.4-8             
 [28] zlibbioc_1.52.0         
 [29] purrr_1.0.2             
 [30] RCurl_1.98-1.16         
 [31] rappdirs_0.3.3          
 [32] GenomeInfoDbData_1.2.13 
 [33] ggrepel_0.9.6           
 [34] irlba_2.3.5.1           
 [35] codetools_0.2-20        
 [36] DelayedArray_0.32.0     
 [37] DT_0.33                 
 [38] scuttle_1.16.0          
 [39] xml2_1.3.6              
 [40] tidyselect_1.2.1        
 [41] farver_2.1.2            
 [42] UCSC.utils_1.2.0        
 [43] ScaledMatrix_1.14.0     
 [44] viridis_0.6.5           
 [45] shinyWidgets_0.8.7      
 [46] BiocFileCache_2.14.0    
 [47] GenomicAlignments_1.42.0
 [48] jsonlite_1.8.9          
 [49] BiocNeighbors_2.0.1     
 [50] scater_1.34.0           
 [51] iterators_1.0.14        
 [52] foreach_1.5.2           
 [53] tools_4.4.1             
 [54] Rcpp_1.0.13-1           
 [55] glue_1.8.0              
 [56] SparseArray_1.6.0       
 [57] xfun_0.49               
 [58] here_1.0.1              
 [59] dplyr_1.1.4             
 [60] withr_3.0.2             
 [61] BiocManager_1.30.25     
 [62] fastmap_1.2.0           
 [63] fansi_1.0.6             
 [64] digest_0.6.37           
 [65] rsvd_1.0.5              
 [66] R6_2.5.1                
 [67] mime_0.12               
 [68] colorspace_2.1-1        
 [69] RSQLite_2.3.9           
 [70] config_0.3.2            
 [71] utf8_1.2.4              
 [72] tidyr_1.3.1             
 [73] generics_0.1.3          
 [74] rtracklayer_1.66.0      
 [75] httr_1.4.7              
 [76] htmlwidgets_1.6.4       
 [77] S4Arrays_1.6.0          
 [78] pkgconfig_2.0.3         
 [79] gtable_0.3.6            
 [80] blob_1.2.4              
 [81] XVector_0.46.0          
 [82] htmltools_0.5.8.1       
 [83] dotCall64_1.2           
 [84] scales_1.3.0            
 [85] png_0.1-8               
 [86] attempt_0.3.1           
 [87] knitr_1.49              
 [88] rjson_0.2.23            
 [89] curl_6.0.1              
 [90] cachem_1.1.0            
 [91] BiocVersion_3.20.0      
 [92] vipor_0.4.7             
 [93] AnnotationDbi_1.68.0    
 [94] restfulr_0.0.15         
 [95] pillar_1.9.0            
 [96] grid_4.4.1              
 [97] vctrs_0.6.5             
 [98] promises_1.3.2          
 [99] BiocSingular_1.22.0     
[100] dbplyr_2.5.0            
[101] beachmat_2.22.0         
[102] xtable_1.8-4            
[103] beeswarm_0.4.0          
[104] paletteer_1.6.0         
[105] evaluate_1.0.1          
[106] magick_2.8.5            
[107] cli_3.6.3               
[108] locfit_1.5-9.10         
[109] compiler_4.4.1          
[110] Rsamtools_2.22.0        
[111] crayon_1.5.3            
[112] labeling_0.4.3          
[113] rematch2_2.1.2          
[114] ggbeeswarm_0.7.2        
[115] viridisLite_0.4.2       
[116] assertthat_0.2.1        
[117] topconfects_1.22.0      
[118] munsell_0.5.1           
[119] Biostrings_2.74.0       
[120] lazyeval_0.2.2          
[121] Matrix_1.7-1            
[122] ExperimentHub_2.14.0    
[123] benchmarkme_1.0.8       
[124] bit64_4.5.2             
[125] KEGGREST_1.46.0         
[126] statmod_1.5.0           
[127] shiny_1.9.1             
[128] AnnotationHub_3.14.0    
[129] gridtext_0.1.5          
[130] memoise_2.0.1           
[131] bslib_0.8.0             
[132] benchmarkmeData_1.0.4   
[133] bit_4.5.0.1             
> sessioninfo::session_info()
─ Session info ───────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-01-26
 pandoc   NA

─ Packages ───────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
   AnnotationDbi          1.68.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   AnnotationHub          3.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   assertthat             0.2.1     2019-03-21 [1] CRAN (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocFileCache          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics         * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocIO                 1.16.0    2024-12-12 [1] Github (Bioconductor/BiocIO@8e80624)
   BiocManager            1.30.25   2024-08-28 [1] CRAN (R 4.4.1)
   BiocNeighbors          2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   BiocParallel         * 1.40.0    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   BiocSingular           1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocVersion            3.20.0    2024-05-04 [1] Bioconductor 3.20 (R 4.4.0)
   Biostrings             2.74.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   bit                    4.5.0.1   2024-12-03 [1] CRAN (R 4.4.1)
   bit64                  4.5.2     2024-09-22 [1] CRAN (R 4.4.1)
   bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bslib                  0.8.0     2024-07-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorspace             2.1-1     2024-07-26 [1] CRAN (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   6.0.1     2024-11-14 [1] CRAN (R 4.4.1)
   data.table           * 1.16.4    2024-12-06 [1] CRAN (R 4.4.1)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   digest                 0.6.37    2024-08-19 [1] CRAN (R 4.4.1)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64              1.2       2024-10-04 [1] CRAN (R 4.4.1)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.4.1     2024-12-02 [1] Bioconductor 3.20 (R 4.4.2)
   evaluate               1.0.1     2024-10-10 [1] CRAN (R 4.4.1)
   ExperimentHub          2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.3      2024-09-30 [1] CRAN (R 4.4.1)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.42.1    2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData       1.2.13    2024-12-12 [1] Bioconductor
   GenomicAlignments      1.42.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   GenomicRanges        * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.4.1)
   ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   glue                   1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
   golem                  0.5.1     2024-08-27 [1] CRAN (R 4.4.1)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.9     2024-09-20 [1] CRAN (R 4.4.1)
   KEGGREST               1.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   knitr                  1.49      2024-11-08 [1] CRAN (R 4.4.1)
   labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.4.1     2024-11-27 [1] CRAN (R 4.4.1)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.62.1    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.5     2024-09-20 [1] CRAN (R 4.4.1)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2.1   2024-11-10 [1] CRAN (R 4.4.1)
   Matrix                 1.7-1     2024-10-18 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.18.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   matrixStats          * 1.4.1     2024-09-08 [1] CRAN (R 4.4.1)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.40.1    2024-12-04 [1] CRAN (R 4.4.1)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises               1.3.2     2024-11-28 [1] CRAN (R 4.4.1)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13-1  2024-11-02 [1] CRAN (R 4.4.1)
   RCurl                  1.98-1.16 2024-07-11 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   RSQLite                2.3.9     2024-12-03 [1] CRAN (R 4.4.1)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Arrays               1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors            * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                 1.34.0    2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
   scuttle                1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   shiny                  1.9.1     2024-08-01 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.7     2024-09-23 [1] CRAN (R 4.4.1)
   SingleCellExperiment * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   spam                   2.11-0    2024-10-03 [1] CRAN (R 4.4.1)
   SparseArray            1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SpatialExperiment    * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   spatialLIBD          * 1.18.0    2024-11-07 [1] Bioconductor 3.20 (R 4.4.1)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   topconfects            1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   UCSC.utils             1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
   xfun                   0.49      2024-10-31 [1] CRAN (R 4.4.1)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xml2                   1.3.6     2023-12-04 [1] CRAN (R 4.4.0)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                   2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
   zlibbioc               1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.
