---
title: "04-integratedUMAP_oneClus_codes_by_covars"
author: "Bernie Mulvey"
date: "2025-01-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(ggrastr)
library(Cairo)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## load the dimensionality reduction matrices, pull out the UMAP for HDG-SVG 2575 after harmony; load cluster assignments at louvain res 1 and append
```{r}
dimr <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/02-PCA_harmo_UMAP_matrices_HiDevG_HVG_SVG_andmixed_featsets.RDS")

udat <- as.data.table(dimr$HARMONYlmbna_HDG_SVG_2575_UMAP,keep.rownames=T)

louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs[["HARMONYlmbna_HDG_SVG_2575"]][,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

udat <- merge.data.table(louvs,udat,by="rn")

rm(dimr)
```

## get metadata sheet for covars
covars to plot by: 
categorical sample_id, capture_id,  APOE, Ancestry, Sex, brnum
continuous sum_gene, sum_umi, expr_chrM_ratio, Age
```{r}
metad <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")
udat <- merge.data.table(udat,metad,by.x="rn",by.y="rowname")

covs <- c("sample_id", "capture_id", "APOE", "Ancestry", "Sex", "brnum", "sum_gene", "sum_umi", "expr_chrM_ratio", "Age")

bplapply(unique(udat$clusid),BPPARAM = MulticoreParam(8),FUN=function(d){
    tmp <- copy(udat)[clusid==d]
    
    mdcol <- c("rowname",covs)
    
    plts <- lapply(covs,function(x){
        if(x %in% c("sum_gene", "sum_umi", "expr_chrM_ratio", "Age")){
            p <- ggplot(tmp,aes(x=UMAP1,y=UMAP2,color=as.numeric(.data[[x]])))+
                 geom_jitter(size=0.75,stroke = 0,position=position_jitterdodge(jitter.height=0.05,jitter.width=0.05,dodge.width=0.1))+
                labs(title=paste0("HmnyUMAP ",d," by ",x),color="none")
        } else {
            p <- ggplot(tmp,aes(x=UMAP1,y=UMAP2,color=.data[[x]]))+
                geom_jitter(size=0.75,stroke = 0,position=position_jitterdodge(jitter.height=0.05,jitter.width=0.05,dodge.width=0.1))+
                labs(title=paste0("HmnyUMAP ",d," by ",x))+
                guides(color="none")+
                scale_x_continuous(limits=c(min(tmp$UMAP1),max(tmp$UMAP1)),expand=c(0.1,0.1))+
                scale_y_continuous(limits=c(min(tmp$UMAP2),max(tmp$UMAP2)),expand=c(0.1,0.1))
        }
        
        q <- rasterize(p,layers="Points",dev="cairo_png",dpi=450)
        return(q)
    })
    pdf(paste0("plots/08_validitycheck_25hdg75svg_louv1/04-postHarmony_UMAP_by_covars_clus_",d,"_only.pdf"),width=18,height=16,onefile = T)
    i<-1
    for (i in c(1:length(plts))){
        print(plts[[i]])
    }
    dev.off()
})
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices
[4] utils     methods   base     

other attached packages:
 [1] Cairo_1.6-2        
 [2] ggrastr_1.0.2      
 [3] BiocParallel_1.40.0
 [4] parallelly_1.40.1  
 [5] colorout_1.3-0.2   
 [6] gridExtra_2.3      
 [7] ggtext_0.1.2       
 [8] ggplot2_3.5.1      
 [9] data.table_1.16.4  
[10] rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6     
 [2] dplyr_1.1.4      
 [3] compiler_4.4.1   
 [4] tidyselect_1.2.1 
 [5] Rcpp_1.0.14      
 [6] ggbeeswarm_0.7.2 
 [7] xml2_1.3.6       
 [8] parallel_4.4.1   
 [9] scales_1.3.0     
[10] here_1.0.1       
[11] R6_2.5.1         
[12] generics_0.1.3   
[13] knitr_1.49       
[14] tibble_3.2.1     
[15] munsell_0.5.1    
[16] rprojroot_2.0.4  
[17] pillar_1.10.1    
[18] xfun_0.49        
[19] cli_3.6.3        
[20] withr_3.0.2      
[21] magrittr_2.0.3   
[22] grid_4.4.1       
[23] gridtext_0.1.5   
[24] beeswarm_0.4.0   
[25] lifecycle_1.0.4  
[26] vipor_0.4.7      
[27] vctrs_0.6.5      
[28] evaluate_1.0.1   
[29] glue_1.8.0       
[30] sessioninfo_1.2.2
[31] codetools_0.2-20 
[32] colorspace_2.1-1 
[33] tools_4.4.1      
[34] pkgconfig_2.0.3  
> sessioninfo::session_info()
─ Session info ────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-01-31
 pandoc   NA

─ Packages ────────────────────────
 !  package      * version date (UTC) lib source
    beeswarm       0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
    BiocParallel * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    Cairo        * 1.6-2   2023-11-28 [1] CRAN (R 4.4.0)
 P  cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
    codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
    colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace     2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    data.table   * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
    dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate       1.0.1   2024-10-10 [1] CRAN (R 4.4.1)
    generics       0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
    ggbeeswarm     0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2      * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
    ggrastr      * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
    ggtext       * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue           1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext       0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable         0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    knitr          1.49    2024-11-08 [1] CRAN (R 4.4.1)
    lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    munsell        0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
    parallelly   * 1.40.1  2024-12-04 [1] CRAN (R 4.4.1)
    pillar         1.10.1  2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6             2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
    Rcpp           1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.4   2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    scales         1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
    sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
    tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    vipor          0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
    withr          3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun           0.49    2024-10-31 [1] CRAN (R 4.4.1)
    xml2           1.3.6   2023-12-04 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

───────────────────────────────────