---
title: "08-NM_per_cluster"
author: "Bernie Mulvey"
date: "2025-02-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## load coldata (has the spot-level NM annots in it) and append cluster labels
```{r}
lc3cd <- fread("processed-data/06-QCed_SPE_splitToTissSect_colData.txt")

### append labels
clustab <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
clustab <- clustab$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(clustab,2,"clusid")

clusano <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/05b-init_clusterings_autoannoLCrapheAndGABAergic.RDS")
clusano <- clusano$snnHARMONYlmbna_HDG_SVG_2575_louv_res1
clusano[autoanno %in% c("5HT.1","5HT.2"),autoanno:=paste0("x",autoanno)]

clustab <- merge.data.table(clustab,clusano,by="clusid")

lc3cd <- merge.data.table(lc3cd,clustab[,.(rn,autoanno)],by.x="rowname",by.y="rn")
setnames(lc3cd,"autoanno","cluster")
```

## make plots of %area NM+ per cluster
```{r}
## calculate proportion of spots with any NM for a cluster
ans <- lc3cd[,.(sum(Prop_NM>0)/.N),by=cluster]
setnames(ans,"V1","propNMspots")
ans[,propNMspots:=round(propNMspots,digits=3)]
ans[,propNMspots:=100*propNMspots]

lc3cd <- merge.data.table(lc3cd,ans,by="cluster",all.x=T)
lc3cd[,clustersimple:=cluster]
lc3cd[,cluster:=paste0(cluster," (",propNMspots,"% spots\n>0 NM area)")]

### histogram
pdf("plots/08_validitycheck_25hdg75svg_louv1/09a-NMarea_per_cluster_hist.pdf",width=9,height=9)
ggplot(lc3cd[Prop_NM>0],aes(x=Prop_NM))+
    geom_histogram(bins=75)+
    facet_wrap(~cluster,scales="fixed",ncol=3)
dev.off()

### weirdly, there's a spike for every cluster around 0.002. so try above this instead
ans <- lc3cd[,.(sum(Prop_NM>0.003)/.N),by=clustersimple]
setnames(ans,"V1","propNMspots")
ans[,propNMspots:=round(propNMspots,digits=3)]
ans[,propNMspots:=100*propNMspots]

## remove the old columns being used for facet titles/propNMspots
lc3cd[,propNMspots:=NULL]
lc3cd[,cluster:=NULL]

lc3cd <- merge.data.table(lc3cd,ans,by="clustersimple",all.x=T)
lc3cd[,cluster:=paste0(clustersimple," (",propNMspots,"% spots\n>0.3% NM area)")]


pdf("plots/08_validitycheck_25hdg75svg_louv1/09b-NMarea_per_cluster_hist_above0.003.pdf",height=9,width=9)
ggplot(lc3cd[Prop_NM>0.003],aes(x=Prop_NM))+
    geom_histogram(bins=75)+
    facet_wrap(~cluster,scales="fixed",ncol=3)
dev.off()


### make one more where density plots show each LC and one other cluster
lc3cd.dens <- copy(lc3cd[Prop_NM>0.003,.(cluster,clustersimple,Prop_NM)])

lc3cd.dens[,cluster:=gsub(cluster,pattern="\n",replacement=" ")]
lc3cd.dens[,cluster:=gsub(cluster,pattern="^(.* )(>.*) NM area\\)$",replacement=paste0("\\1","NM area","\\2",")"))]
densplt <- lapply(unique(lc3cd.dens[clustersimple!="LC.1",clustersimple]),function(x){
    curclusstat <- unique(lc3cd.dens[clustersimple==x,cluster])
    pltdat <- lc3cd[clustersimple %in% c("LC.1",x),.(cluster,Prop_NM)]
    pltdat[,cluster:=gsub(cluster,pattern="\n",replacement="<br>")]
    p <- ggplot(pltdat,aes(x=Prop_NM,col=cluster))+
        geom_density(alpha=0.5)+
        theme(legend.position = "none")+
        ggtitle(paste0("LC.1 (red, 39.8% NM>0.3%) vs.<br>",curclusstat))
    return(p)
})



pdf("plots/08_validitycheck_25hdg75svg_louv1/09c-NMarea_over0.3pct_per_cluster_densityplot_LC1_vs_others.pdf",height=9,width=9)
do.call("grid.arrange",c(densplt,ncol=3))
dev.off()

```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.2 (2024-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] colorout_1.3-0.2            SpatialExperiment_1.16.0   
 [3] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
 [5] Biobase_2.66.0              GenomicRanges_1.58.0       
 [7] GenomeInfoDb_1.42.1         IRanges_2.40.1             
 [9] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[11] MatrixGenerics_1.18.0       matrixStats_1.5.0          
[13] gridExtra_2.3               ggtext_0.1.2               
[15] ggplot2_3.5.1               data.table_1.16.4          
[17] rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] gtable_0.3.6            rjson_0.2.23            xfun_0.50              
 [4] lattice_0.22-6          vctrs_0.6.5             tools_4.4.2            
 [7] generics_0.1.3          tibble_3.2.1            pkgconfig_2.0.3        
[10] Matrix_1.7-1            lifecycle_1.0.4         GenomeInfoDbData_1.2.13
[13] stringr_1.5.1           compiler_4.4.2          farver_2.1.2           
[16] munsell_0.5.1           htmltools_0.5.8.1       yaml_2.3.10            
[19] pillar_1.10.1           crayon_1.5.3            DelayedArray_0.32.0    
[22] sessioninfo_1.2.2       magick_2.8.5            abind_1.4-8            
[25] commonmark_1.9.2        tidyselect_1.2.1        digest_0.6.37          
[28] stringi_1.8.4           dplyr_1.1.4             labeling_0.4.3         
[31] rprojroot_2.0.4         fastmap_1.2.0           grid_4.4.2             
[34] here_1.0.1              colorspace_2.1-1        cli_3.6.3              
[37] SparseArray_1.6.0       magrittr_2.0.3          S4Arrays_1.6.0         
[40] withr_3.0.2             scales_1.3.0            UCSC.utils_1.2.0       
[43] rmarkdown_2.29          XVector_0.46.0          httr_1.4.7             
[46] evaluate_1.0.3          knitr_1.49              markdown_1.13          
[49] gridtext_0.1.5          Rcpp_1.0.14             glue_1.8.0             
[52] xml2_1.3.6              rstudioapi_0.17.1       jsonlite_1.8.9         
[55] R6_2.5.1                zlibbioc_1.52.0        
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.2 (2024-10-31)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-15
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 P  cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
    colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    commonmark             1.9.2   2024-10-04 [1] CRAN (R 4.4.1)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1  2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
    ggtext               * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    jsonlite               1.8.9   2024-09-20 [1] CRAN (R 4.4.1)
    knitr                  1.49    2024-11-08 [1] CRAN (R 4.4.1)
    labeling               0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.2)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magick                 2.8.5   2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    markdown               1.13    2024-06-04 [1] CRAN (R 4.4.0)
    Matrix                 1.7-1   2024-10-18 [1] CRAN (R 4.4.2)
    MatrixGenerics       * 1.18.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
    pillar                 1.10.1  2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4   2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
    sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    stringi                1.8.4   2024-05-06 [1] CRAN (R 4.4.0)
    stringr                1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    withr                  3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.50    2025-01-07 [1] CRAN (R 4.4.1)
    xml2                   1.3.6   2023-12-04 [1] CRAN (R 4.4.0)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
