---
title: "11_2-RCTD_perSPEtissSection_WebDivSnseqLabelMerged_and_granular_clusts"
author: "Bernie Mulvey"
date: "2026-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(SingleCellExperiment)
# devtools::install_github("dmcable/spacexr@744153c", build_vignettes = FALSE, force = TRUE)
library(spacexr) # Bioc version is a fork of the github version without some of the functionality yet incorporated and totally different commands, so using same github version as in original RCTD run, per above command commented out
library(SummarizedExperiment)
library(WeberDivechaLCdata)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

## load SPE (we need untransformed counts) and its annots
## load pre-annotated webdiv snseq 
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

louvs <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(louvs,2,"clusid")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")

louvs <- merge.data.table(louvs,anno[,.(clusid,anno)],by="clusid")
lc3$cell_type <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),"anno"]
lc3$cell_type <- as.factor(lc3$cell_type)
rm(louvs,anno)

#######
lcsn <- WeberDivechaLCdata_singleNucleus()

## determine what hierarchical clustering levels there are
tmpcd <- as.data.table(colData(lcsn),keep.rownames=T)
checklabs <- unique(tmpcd[,.(label,label_merged,label_inhibitory)])
## so the inhibitories are re-clusterings (multiple label_inhibitory per label)
View(checklabs[label_merged!="inhibitory"])
## the other 23 labels (of 30) have single label merged assignments.
## so we will make a cluster hierarchy as follows: cluster_merged > label > label or inhibitory subcluster

# to ensure there's enough LCNE cells for RCTD to work with, change label_merged and then label itself to NE for any cell classified as supervised NE OR unsup. NE
tmpcd[,label_merged:=as.character(label_merged)]
tmpcd[unsupervised_NE==TRUE,label_merged:="NE"]
tmpcd[supervised_NE==TRUE,label_merged:="NE"]

## then, make label a character column of merged_label (parent) and label (granular clust #)
tmpcd[,label:=as.character(label)]
tmpcd[,label:=paste0(label_merged,"_",label)]
## for NE neurons, disregard the granular cluster of origin since we're integrating several definition types now
tmpcd[label_merged=="NE",label:="NE"]
## finally, add info about inhibitory subcluster; ONLY for inhibitory merged clusters
tmpcd[,sublabel:=label]
tmpcd[label_merged=="inhibitory"&!is.na(label_inhibitory),sublabel:=paste0(sublabel,
                                                "_sub",label_inhibitory)]
## for sublabels, we will need to filter out sublabels with fewer than RCTD's suggested minimum of 25 cells, if any. performed in the main loop next block.

stopifnot(length(unique(tmpcd$label))>length(unique(tmpcd$label_merged)))
stopifnot(length(unique(tmpcd$sublabel))>length(unique(tmpcd$label)))

colData(lcsn) <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcsn),]

rm(tmpcd,checklabs)
```

going by the bioc RCTD vignette; iterate thru each cluster level per spatial sample
```{r}
cluslevs <- c("label_merged","label","sublabel")

## iterate over cluster types to prevent data loss if errors partway
rctdresults <- list()
i <- 1
for (i in c(1:length(cluslevs))){
  cluslev <- cluslevs[i]
  setDTthreads(1,restore_after_fork=FALSE) # prevent resource overcommitting while running RCTD steps
  
  ## check, remove any cells for which there are fewer than 25 cells for the current label set
  checkcts <- as.data.table(table(as.character(colData(lcsn)[,cluslev])))
  
  if(nrow(checkcts[N<25])>0){
    checkcts <- checkcts[N>=25]
    tmplcsn <- lcsn[,colData(lcsn)[,cluslev] %in% checkcts$V1]
    cts <- colData(tmplcsn)[,cluslev]
    names(cts) <- rownames(colData(tmplcsn))
    cts <- as.factor(cts)
  } else {
    tmplcsn <- lcsn
    cts <- colData(lcsn)[,cluslev]
    names(cts) <- rownames(colData(lcsn))
    cts <- as.factor(cts)
  }
  rm(checkcts)
  
  stopifnot(all(names(cts)==colnames(counts(tmplcsn))))
  
  # convert snseq ref to an RCTD Reference class
  # theres 12k of one merged label type i guess so raise the max limit to accomodate
  # expects the cell type annots to be named "cell_types"
  reflcsn <- spacexr::Reference(counts=counts(tmplcsn),cell_types=cts,n_max_cells=15000)  
  # create RCTD, run rctd per visium sample
  
  samps <- unique(lc3$sample_id)
  rctdsampres <- lapply(X=samps,FUN=function(s){
      setDTthreads(1,restore_after_fork=FALSE) # make certain of preventing resource overcommitting while running RCTD steps by also embedding this in apply steps
      ## make an RCTD spatialrna class object for input
      lccoord <- as.data.frame(spatialCoords(lc3[,lc3$sample_id==s]))
      colnames(lccoord) <- c("x","y")
      lc3sxr <- spacexr::SpatialRNA(coords=lccoord,counts = counts(lc3[,lc3$sample_id==s]))

      # include low count spots, because they are frequent; do not penalize for treating spots as multitype ("double_threshold") because they pretty obviously are; counts_min is default and shown for posterity
      rctd_data <- create.RCTD(spatialRNA = lc3sxr,
                           reference = reflcsn,
                           max_cores = 10,
                           UMI_max = Inf,
                           UMI_min=50,
                           MAX_MULTI_TYPES = 4,
                           DOUBLET_THRESHOLD = 0,
                           counts_MIN = 10,
                           keep_reference = TRUE)
  
      # run RCTD multi, allowing up to 4 types per spot
      lc.rctd <- run.RCTD(rctd_data,doublet_mode = "multi")
  
      # extract the spotwise RCTD results and their corresponding spot IDs to a table; save
      wat <- lc.rctd@results
      spotnames <- colnames(lc.rctd@spatialRNA@counts)
  
      # extract the spotwise RCTD results and their corresponding spot IDs to a table; save
      wat <- lc.rctd@results
      spotnames <- colnames(lc.rctd@spatialRNA@counts)
  
      
      setDTthreads(10,restore_after_fork=FALSE) # multithread for this part
      #tabulate spotwise results:
      rctdrestab <- mapply(x=wat,n=spotnames,SIMPLIFY=FALSE,function(x,n){
          setDTthreads(10,restore_after_fork=FALSE) # multithread for this part
          y <- as.data.table(t(x$all_weights))
          setnames(y,"5HT","x5HT",skip_absent=TRUE)
      
          # change sub_weights to match x5HT where applicable
          names(x$sub_weights)[names(x$sub_weights)=="5HT"] <- "x5HT"
      
          z <- copy(y)
          keepn <- names(z)[!(names(z) %in% names(x$sub_weights))]
          z <- z[,..keepn]
          set(z,j = c(1:ncol(z)),value=0)
          z <- as.data.table(cbind(z,t(x$sub_weights)))
          setnames(z,names(z),paste0(names(z),"_multi_weight"))
          
          z <- as.data.table(cbind(y,z))
          z[,rn:=n]
          return(z)
      })
      # convert the spotwise tables (1 row by many columns) to one table with a row per spot
      rctdrestab <- rbindlist(rctdrestab,fill=T)
      return(rctdrestab)
  })
  # make sure sample_id is passed to the corresponding sample's results (one sample = one list entry with a single data.table; we can rbindlist the samplewise results to make a single table to join into the SPE colData when we want to use the results.)
  names(rctdsampres) <- samps
  
  # return results to output-holding list
  
  rctdresults[[i]] <- rctdsampres
  names(rctdresults)[i] <- cluslevs[i]
  
  rm(cts,reflcsn,rctdsampres,tmplcsn)
  gc(full=T)
}

names(rctdresults) <- cluslevs
```

```{r}
### gah, didn't label the first set of columns in each output table as [label]_full_weights. append that to make clear what those columns are for later:
rctdresults <- lapply(rctdresults,function(x){
  z <- lapply(x,function(y){
    fixnames <- names(y)[!(names(y) %in% c("rn",grep(names(y),pattern="multi_weight",value=T)))]
    setnames(y,fixnames,paste0(fixnames,"_full_weights"))
    return(y)
  })
  return(z)
})

rctdconcat <- lapply(rctdresults,rbindlist,idcol="sample_id",fill=T)
rctdconcat <- lapply(rctdconcat,function(x){
  nameord <- c("rn",
               "sample_id",
               grep(names(x),pattern="full_we",value=T),
               grep(names(x),pattern="multi_we",value=T))
  # drop weird glitchy noncluster results that have all NAs--showed up only in some samples
  nameord <- grep(nameord,pattern="V1",value=T,invert =T)
  # reorder columns
  y <- x[,..nameord]
  return(y) 
})
# and save
# saveRDS(rctdresults,"processed-data/10_subclustering/11-RCTDmulti4_webdivLabel-LabMerge-LabWithInhibSubclusts_samplewise_results.RDS")
saveRDS(rctdconcat,"processed-data/10_2_subclustering_and_NMF/11_2-RCTDmulti4_webdivLabel-LabelMerged-LabelWithInhibSubclusts_concatresults.RDS")
```

repinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              WeberDivechaLCdata_1.12.0  
 [5] ExperimentHub_3.0.0         AnnotationHub_4.0.0        
 [7] BiocFileCache_3.0.0         dbplyr_2.5.1               
 [9] spacexr_2.2.1               SpatialExperiment_1.20.0   
[11] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[13] Biobase_2.70.0              GenomicRanges_1.62.1       
[15] Seqinfo_1.0.0               IRanges_2.44.0             
[17] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[19] generics_0.1.4              MatrixGenerics_1.22.0      
[21] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] KEGGREST_1.50.0      rjson_0.2.23         xfun_0.56           
 [4] httr2_1.2.2          lattice_0.22-7       quadprog_1.5-8      
 [7] vctrs_0.7.1          tools_4.5.2          curl_7.0.0          
[10] parallel_4.5.2       AnnotationDbi_1.72.0 tibble_3.3.1        
[13] RSQLite_2.4.5        blob_1.3.0           pkgconfig_2.0.3     
[16] Matrix_1.7-4         lifecycle_1.0.5      compiler_4.5.2      
[19] Biostrings_2.78.0    codetools_0.2-20     htmltools_0.5.9     
[22] yaml_2.3.12          crayon_1.5.3         pillar_1.11.1       
[25] rsconnect_1.7.0      DelayedArray_0.36.0  cachem_1.1.0        
[28] magick_2.9.0         iterators_1.0.14     abind_1.4-8         
[31] foreach_1.5.2        tidyselect_1.2.1     digest_0.6.39       
[34] purrr_1.2.1          dplyr_1.2.0          BiocVersion_3.22.0  
[37] rprojroot_2.1.1      fastmap_1.2.0        grid_4.5.2          
[40] here_1.0.2           cli_3.6.5            SparseArray_1.10.8  
[43] magrittr_2.0.4       S4Arrays_1.10.1      withr_3.0.2         
[46] filelock_1.0.3       rappdirs_0.3.4       bit64_4.6.0-1       
[49] httr_1.4.7           rmarkdown_2.30       XVector_0.50.0      
[52] bit_4.6.0            otel_0.2.0           png_0.1-8           
[55] memoise_2.0.1        evaluate_1.0.5       knitr_1.51          
[58] doParallel_1.0.17    rlang_1.1.7          Rcpp_1.1.1          
[61] glue_1.8.0           DBI_1.2.3            BiocManager_1.30.27 
[64] rstudioapi_0.18.0    R6_2.6.1            
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-16
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationHub        * 4.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocFileCache        * 3.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocManager            1.30.27  2025-11-14 [1] CRAN (R 4.5.2)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocVersion            3.22.0   2025-10-07 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0    2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1  2025-01-16 [1] CRAN (R 4.5.0)
 blob                   1.3.0    2026-01-14 [1] CRAN (R 4.5.2)
 cachem                 1.1.0    2024-05-16 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0    2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3    2024-06-02 [1] CRAN (R 4.5.0)
 dbplyr               * 2.5.1    2025-09-10 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 doParallel             1.0.17   2022-02-07 [1] CRAN (R 4.5.0)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 ExperimentHub        * 3.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 filelock               1.0.3    2023-12-11 [1] CRAN (R 4.5.0)
 foreach                1.5.2    2022-02-02 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 httr                   1.4.7    2023-08-15 [1] CRAN (R 4.5.0)
 httr2                  1.2.2    2025-12-08 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 iterators              1.0.14   2022-02-05 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1    2021-11-26 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 png                    0.1-8    2022-11-29 [1] CRAN (R 4.5.0)
 purrr                  1.2.1    2026-01-09 [1] CRAN (R 4.5.2)
 quadprog               1.5-8    2019-11-20 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4    2026-01-17 [1] CRAN (R 4.5.2)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5    2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 spacexr              * 2.2.1    2026-02-15 [1] Github (dmcable/spacexr@744153c)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 WeberDivechaLCdata   * 1.12.0   2025-11-04 [1] Bioconductor 3.22 (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
