---
title: "03bi-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEAmsigdb"
author: "Bernie Mulvey"
date: "2026-02-13"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(clusterProfiler) # otherwise fgsea parallelization doesnt seem to work
library(fgsea)
library(msigdbr)

require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallel)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallelly)
options(bphost="localhost")
library(BiocParallel)
```

### fgsea sets retrieval (2025.1 HS)
```{r}
allsigs <- msigdbr("HS",category="C1")

# tack on the others, make d.t.
msigs <- paste0("C",c(2,3,5,6,8))
for (msig in msigs){
    allsigs <- rbind(allsigs,msigdbr("HS",category=msig))
}
rm(msig,msigs)
allsigs <- as.data.table(allsigs)

# remove legacy sets
allsigs <- allsigs[gs_subcat %in% grep(gs_subcat,pattern="legacy|LEGACY|Legacy",value=T,invert=T)]

# make a list object of identifiers per  gene set
# we can map these back to the msigdb table later for reference purposes by their list name (gs_name)...
sets <- split(allsigs$ensembl_gene,allsigs$gs_name)

rm(allsigs)
gc(full=T)
```


# PART 1: get signed DE t stat vectors per cluster per test
```{r}
de2b <- readRDS("processed-data/12_2_DEanalyses_removedsampsAndFinalNMseg/02b_2-voomLmFit_objs_and_DE_tables_LabSex_LabE2vE4_LabSingleGenos_LabAnces_LabAncesE4E2_noMixClusts_YRIcontAndCat_Feb26.RDS")

## we want continuous ancestry analyses, except for the ancestry-strat analyses (predom ancestry, and within-predom-ancestry APOE)
de2b2 <- list(sexlabel=de2b$cont_ances$sexlabel$detabs,
              E2E4label=de2b$cont_ances$E2E4label$detabs,
              ApoAllelesLabel=de2b$cont_ances$ApoAllelesLabel$detabs,
              AncesByLabel=de2b$cat_ances$AncesByLabel$detabs,
              YRIAPOlabel=de2b$cat_ances$YRIAPOlabel$detabs)

de2b3 <- list()
i<-1
for (i in c(1:length(de2b2))){
    j<-1
    for (j in c(1:length(de2b2[[i]]))){
        de2b3[length(de2b3)+1] <- de2b2[[i]][j]
        names(de2b3)[length(de2b3)] <- names(de2b2[[i]])[j]
    }
    rm(j)
}
rm(i)

### drop sex chr genes from sex DE tests
de2b3[grep(names(de2b3),pattern="Sex",value=T)] <- lapply(de2b3[grep(names(de2b3),pattern="Sex",value=T)],FUN=function(x){
    x[!(chr %in% grep(chr,pattern="chrX|chrY",value=T))]})

### drop mito genes from all tests
de2b3 <- lapply(de2b3,FUN=function(x){
    x[!(chr %in% grep(chr,pattern="chrMT",value=T))]})

### get named, SIGNED vectors of DE t-stats from each test
### need to use gene symbols for compatibility with the TF-target sets
deres <- list()
i<-1
for (i in c(1:length(de2b3))){
    setnames(de2b3[[i]],"t","tstat")
    tmp <- copy(de2b3[[i]])[,.(gene_id,tstat)]
    
    dropg <- tmp[,.N,by="gene_id"][N>1,gene_id]
    tmp <- tmp[!(gene_id %in% dropg)]
    
    setorderv(tmp,"tstat",-1)
    ## get the stat column as a vector and give it the gene ids as names for fgsea
    ## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
    tmp <- as.data.frame(tmp)
    v <- tmp$tstat
    names(v) <- tmp$gene_id
    deres[[length(deres)+1]] <- v
    names(deres)[length(deres)] <- names(de2b3)[i]
    rm(v,tmp)
}
rm(i,de2b,de2b2,dropg)
gc(full=T)
```

### run fgsea-signed tests
### looping through each cluster-variable interaction (each run is very fast with 10 cpus in biocparallel, so just serially run 10cpu GSEA for each cluster)

```{r}
gseares <- mapply(X=deres,Y=names(deres),SIMPLIFY=FALSE,FUN=function(X,Y){
  register(MulticoreParam())
   
   # get cond1-up enriched results
   z1<-fgseaMultilevel(pathways=sets,stats = X,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")
   # coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
   z1$leadingEdge <- as.character(z1$leadingEdge)
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')
   
   # drop newlines from leadingEdge
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   # somehow, one run of this sometimes doesnt get them all so repeat for good meas
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   
   z1 <- as.data.table(z1)
   z1[,test:=Y]
   
   if(length(grep(Y,pattern="Sex",value=T))>0){
       z1[,dir:="male_up"]
   } else if(length(grep(Y,pattern="E4vE2",value=T))>0){
       z1[,dir:="E4carriers_up"]
   } else if(length(grep(Y,pattern="_YRI",value=T))>0){
       z1[,dir:="YRI_up"]
   } else {
       z1[,dir:=gsub(Y,pattern="^.*_(E....)_vs_(.*)",replacement=paste0("\\1","_up_vs_","\\2"))]
   }
   
   # get condition 2-up (e.g., female-up) enriched results
   register(MulticoreParam())
   z2<-fgseaMultilevel(pathways=sets,stats = X,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "neg")
   # coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
   z2$leadingEdge <- as.character(z2$leadingEdge)
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern='"',replacement='')
   
   # drop newlines from leadingEdge
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")
   # somehow, one run of this sometimes doesnt get them all so repeat for good meas
   z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")
   
   z2 <- as.data.table(z2)
   z2[,test:=Y]
   
    if(length(grep(Y,pattern="Sex",value=T))>0){
       z2[,dir:="female_up"]
   } else if(length(grep(Y,pattern="E4vE2",value=T))>0){
       z2[,dir:="E4carriers_down"]
   } else if(length(grep(Y,pattern="_YRI",value=T))>0){
       z2[,dir:="YRI_down"]
   } else {
       z2[,dir:=gsub(Y,pattern="^.*_(E....)_vs_(.*)",replacement=paste0("\\1","_down_vs_","\\2"))]
   }

   z <- as.data.table(rbind(z1,z2))
   return(z)
})
names(gseares) <- names(deres)

saveRDS(gseares,"processed-data/12_2_DEanalyses_removedsampsAndFinalNMseg/03bi_2-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEAmsigdb.RDS")

rm(deres,gseares)
```

### PART 2: get UNsigned DE t stat vectors per cluster per test
```{r}
### get named, UNsigned vectors of DE t-stats from each test
deres <- list()
i<-1
for (i in c(1:length(de2b3))){
    tmp <- copy(de2b3[[i]])[,.(gene_id,abs(tstat))]

    dropg <- tmp[,.N,by="gene_id"][N>1,gene_id]
    tmp <- tmp[!(gene_id %in% dropg)]
    
    setnames(tmp,2,"tstat")
    setorderv(tmp,"tstat",-1)
    ## get the stat column as a vector and give it the gene ids as names for fgsea
    ## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
    tmp <- as.data.frame(tmp)
    v <- tmp$tstat
    names(v) <- tmp$gene_id
    deres[[length(deres)+1]] <- v
    names(deres)[length(deres)] <- names(de2b3)[i]
    rm(v,tmp)
}
rm(i)
```

run UNSIGNED fgsea (genes that tend to be affected, regardless of direction)
```{r}
gseares.unsign <- mapply(X=deres,Y=names(deres),SIMPLIFY=FALSE,FUN=function(X,Y){
   register(MulticoreParam())
   
   # get DE-enriched results
   z1<-fgseaMultilevel(pathways=sets,stats = X,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")
   
   # coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
   z1$leadingEdge <- as.character(z1$leadingEdge)
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')
   # drop newlines from leadingEdge
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   # belt and suspenders
   z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
   
   z1 <- as.data.table(z1)
   z1[,test:=Y]
   z1[,dir:="either"]
   
   return(z1)
})

names(gseares.unsign) <- names(deres)

saveRDS(gseares.unsign,"processed-data/12_2_DEanalyses_removedsampsAndFinalNMseg/03bi_2-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEAmsigdb_UNsigned.RDS")
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] BiocParallel_1.44.0    parallelly_1.46.1      default_1.0.0         
[4] colorout_1.3-3         msigdbr_25.1.1         fgsea_1.36.2          
[7] clusterProfiler_4.18.4 data.table_1.18.2.1   

loaded via a namespace (and not attached):
  [1] DBI_1.2.3               gson_0.1.0              rlang_1.1.7            
  [4] magrittr_2.0.4          DOSE_4.4.0              otel_0.2.0             
  [7] compiler_4.5.2          RSQLite_2.4.5           png_0.1-8              
 [10] systemfonts_1.3.1       vctrs_0.7.1             reshape2_1.4.5         
 [13] stringr_1.6.0           pkgconfig_2.0.3         crayon_1.5.3           
 [16] fastmap_1.2.0           XVector_0.50.0          rmarkdown_2.30         
 [19] sessioninfo_1.2.3       enrichplot_1.30.4       purrr_1.2.1            
 [22] bit_4.6.0               xfun_0.56               cachem_1.1.0           
 [25] aplot_0.2.9             jsonlite_2.0.0          blob_1.3.0             
 [28] tidydr_0.0.6            tweenr_2.0.3            cluster_2.1.8.1        
 [31] R6_2.6.1                stringi_1.8.7           RColorBrewer_1.1-3     
 [34] GOSemSim_2.36.0         assertthat_0.2.1        Rcpp_1.1.1             
 [37] Seqinfo_1.0.0           knitr_1.51              ggtangle_0.1.1         
 [40] R.utils_2.13.0          IRanges_2.44.0          Matrix_1.7-4           
 [43] splines_4.5.2           igraph_2.2.1            tidyselect_1.2.1       
 [46] qvalue_2.42.0           rstudioapi_0.18.0       dichromat_2.0-0.1      
 [49] yaml_2.3.12             codetools_0.2-20        curl_7.0.0             
 [52] lattice_0.22-7          tibble_3.3.1            plyr_1.8.9             
 [55] withr_3.0.2             treeio_1.34.0           Biobase_2.70.0         
 [58] KEGGREST_1.50.0         S7_0.2.1                evaluate_1.0.5         
 [61] gridGraphics_0.5-1      polyclip_1.10-7         scatterpie_0.2.6       
 [64] Biostrings_2.78.0       pillar_1.11.1           ggtree_4.0.4           
 [67] rsconnect_1.7.0         stats4_4.5.2            ggfun_0.2.0            
 [70] generics_0.1.4          rprojroot_2.1.1         S4Vectors_0.48.0       
 [73] ggplot2_4.0.2           scales_1.4.0            tidytree_0.4.7         
 [76] glue_1.8.0              gdtools_0.4.4           lazyeval_0.2.2         
 [79] tools_4.5.2             ggnewscale_0.5.2        babelgene_22.9         
 [82] ggiraph_0.9.3           fs_1.6.6                fastmatch_1.1-8        
 [85] cowplot_1.2.0           grid_4.5.2              tidyr_1.3.2            
 [88] ape_5.8-1               AnnotationDbi_1.72.0    nlme_3.1-168           
 [91] patchwork_1.3.2         ggforce_0.5.0           cli_3.6.5              
 [94] rappdirs_0.3.4          fontBitstreamVera_0.1.1 dplyr_1.2.0            
 [97] gtable_0.3.6            R.methodsS3_1.8.2       yulab.utils_0.2.4      
[100] digest_0.6.39           fontquiver_0.2.1        BiocGenerics_0.56.0    
[103] ggrepel_0.9.6           ggplotify_0.1.3         htmlwidgets_1.6.4      
[106] farver_2.1.2            memoise_2.0.1           htmltools_0.5.9        
[109] R.oo_1.27.1             lifecycle_1.0.5         httr_1.4.7             
[112] here_1.0.2              GO.db_3.22.0            MASS_7.3-65            
[115] fontLiberation_0.1.0    bit64_4.6.0-1          
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-13
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package           * version  date (UTC) lib source
 AnnotationDbi       1.72.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 ape                 5.8-1    2024-12-16 [1] CRAN (R 4.5.0)
 aplot               0.2.9    2025-09-12 [1] CRAN (R 4.5.0)
 assertthat          0.2.1    2019-03-21 [1] CRAN (R 4.5.0)
 babelgene           22.9     2022-09-29 [1] CRAN (R 4.5.0)
 Biobase             2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics        0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel      * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings          2.78.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                 4.6.0    2025-03-06 [1] CRAN (R 4.5.0)
 bit64               4.6.0-1  2025-01-16 [1] CRAN (R 4.5.0)
 blob                1.3.0    2026-01-14 [1] CRAN (R 4.5.2)
 cachem              1.1.0    2024-05-16 [1] CRAN (R 4.5.0)
 cli                 3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 cluster             2.1.8.1  2025-03-12 [1] CRAN (R 4.5.0)
 clusterProfiler   * 4.18.4   2025-12-15 [1] Bioconductor 3.22 (R 4.5.2)
 codetools           0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout          * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 cowplot             1.2.0    2025-07-07 [1] CRAN (R 4.5.0)
 crayon              1.5.3    2024-06-20 [1] CRAN (R 4.5.0)
 curl                7.0.0    2025-08-19 [1] CRAN (R 4.5.0)
 data.table        * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DBI                 1.2.3    2024-06-02 [1] CRAN (R 4.5.0)
 default           * 1.0.0    2017-08-07 [1] CRAN (R 4.5.0)
 dichromat           2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest              0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 DOSE                4.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dplyr               1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 enrichplot          1.30.4   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 evaluate            1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver              2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap             1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 fastmatch           1.1-8    2026-01-17 [1] CRAN (R 4.5.2)
 fgsea             * 1.36.2   2026-01-05 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 fontBitstreamVera   0.1.1    2017-02-01 [1] CRAN (R 4.5.0)
 fontLiberation      0.1.0    2016-10-15 [1] CRAN (R 4.5.0)
 fontquiver          0.2.1    2017-02-01 [1] CRAN (R 4.5.0)
 fs                  1.6.6    2025-04-12 [1] CRAN (R 4.5.0)
 gdtools             0.4.4    2025-10-06 [1] CRAN (R 4.5.0)
 generics            0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 ggforce             0.5.0    2025-06-18 [1] CRAN (R 4.5.0)
 ggfun               0.2.0    2025-07-15 [1] CRAN (R 4.5.0)
 ggiraph             0.9.3    2026-01-19 [1] CRAN (R 4.5.2)
 ggnewscale          0.5.2    2025-06-20 [1] CRAN (R 4.5.0)
 ggplot2             4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggplotify           0.1.3    2025-09-20 [1] CRAN (R 4.5.0)
 ggrepel             0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 ggtangle            0.1.1    2026-01-16 [1] CRAN (R 4.5.2)
 ggtree              4.0.4    2026-01-05 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 glue                1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 GO.db               3.22.0   2026-01-31 [1] Bioconductor
 GOSemSim            2.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gridGraphics        0.5-1    2020-12-13 [1] CRAN (R 4.5.0)
 gson                0.1.0    2023-03-07 [1] CRAN (R 4.5.0)
 gtable              0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools           0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets         1.6.4    2023-12-06 [1] CRAN (R 4.5.0)
 httr                1.4.7    2023-08-15 [1] CRAN (R 4.5.0)
 igraph              2.2.1    2025-10-27 [1] CRAN (R 4.5.0)
 IRanges             2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 jsonlite            2.0.0    2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST            1.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr               1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice             0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval            0.2.2    2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle           1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magrittr            2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 MASS                7.3-65   2025-02-28 [1] CRAN (R 4.5.2)
 Matrix              1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 memoise             2.0.1    2021-11-26 [1] CRAN (R 4.5.0)
 msigdbr           * 25.1.1   2025-07-21 [1] CRAN (R 4.5.2)
 nlme                3.1-168  2025-03-31 [1] CRAN (R 4.5.2)
 otel                0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly        * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 patchwork           1.3.2    2025-08-25 [1] CRAN (R 4.5.0)
 pillar              1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig           2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 plyr                1.8.9    2023-10-02 [1] CRAN (R 4.5.0)
 png                 0.1-8    2022-11-29 [1] CRAN (R 4.5.0)
 polyclip            1.10-7   2024-07-23 [1] CRAN (R 4.5.0)
 purrr               1.2.1    2026-01-09 [1] CRAN (R 4.5.2)
 qvalue              2.42.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 R.methodsS3         1.8.2    2022-06-13 [1] CRAN (R 4.5.0)
 R.oo                1.27.1   2025-05-02 [1] CRAN (R 4.5.0)
 R.utils             2.13.0   2025-02-24 [1] CRAN (R 4.5.0)
 R6                  2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs            0.3.4    2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer        1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 reshape2            1.4.5    2025-11-12 [1] CRAN (R 4.5.0)
 rlang               1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown           2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot           2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect           1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite             2.4.5    2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi          0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Vectors           0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                  0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales              1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scatterpie          0.2.6    2025-09-12 [1] CRAN (R 4.5.0)
 Seqinfo             1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo         1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 stringi             1.8.7    2025-03-27 [1] CRAN (R 4.5.0)
 stringr             1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 systemfonts         1.3.1    2025-10-01 [1] CRAN (R 4.5.0)
 tibble              3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidydr              0.0.6    2025-07-25 [1] CRAN (R 4.5.0)
 tidyr               1.3.2    2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect          1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 tidytree            0.4.7    2026-01-08 [1] CRAN (R 4.5.2)
 treeio              1.34.0   2025-10-30 [1] Bioconductor 3.22 (R 4.5.1)
 tweenr              2.0.3    2024-02-26 [1] CRAN (R 4.5.0)
 vctrs               0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr               3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector             0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                2.3.12   2025-12-10 [1] CRAN (R 4.5.2)
 yulab.utils         0.2.4    2026-02-02 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
