---
title: "05-LCNMpos-vs-neg_GSEA_msigdb_and_TFtargs"
author: "Bernie Mulvey"
date: "2025-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(clusterProfiler) # otherwise fgsea parallelization doesnt seem to work
library(fgsea)
library(msigdbr)

require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallel)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallelly)
options(bphost="localhost")
library(BiocParallel)
```

### msigdb sets retrieval
```{r}
allsigs <- msigdbr("hs",category="C1")

# tack on the others, make d.t.
msigs <- paste0("C",c(2,3,5,6,8))
for (msig in msigs){
    allsigs <- rbind(allsigs,msigdbr("hs",category=msig))
}
rm(msig,msigs)
allsigs <- as.data.table(allsigs)

# remove legacy sets
allsigs <- allsigs[gs_subcat %in% grep(gs_subcat,pattern="legacy|LEGACY|Legacy",value=T,invert=T)]

# make a list object of identifiers per  gene set
# we can map these back to the msigdb table later for reference purposes by their list name (gs_name)...
sets <- split(allsigs$ensembl_gene,allsigs$gs_name)

rm(allsigs)
gc(full=T)
```


# PART 1: get signed DE t stat vectors per cluster per test. 
# note that we need SYMBOLS ("gene" column), not ens IDs for the TF-targ tests. the TF-target lists are only available as symbols
```{r}
lcnmde <- fread("processed-data/12_2_DEanalyses_removedsampsAndFinalNMseg/04_2-LCNMpos_vs_LCNMneg_DE.txt")

### drop mito genes
lcnmde <- lcnmde[!(chr %in% grep(chr,pattern="chrMT",value=T))]

### get named, SIGNED vectors of DE t-stats from each test
setnames(lcnmde,"t","tstat")
tmp <- copy(lcnmde)[,.(gene_id,gene_name,tstat)]
setorderv(tmp,"tstat",-1)
## get the stat column as a vector and give it the gene ids as names for fgsea
## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
tmp <- as.data.frame(tmp)
v.id <- tmp$tstat
v.name <- tmp$tstat

names(v.id) <- tmp$gene_id
names(v.name) <- tmp$gene_name
deres.id <- v.id
deres.name <- v.name

rm(v.id,v.name,tmp)
```

### run fgsea-signed tests
### looping through each cluster-variable interaction (each run is very fast with 10 cpus in biocparallel, so just serially run 10cpu GSEA for each cluster)

```{r}
register(MulticoreParam())
# get LCNM-up enriched results
z1<-fgseaMultilevel(pathways=sets,stats = deres.id,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")
# coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
z1$leadingEdge <- as.character(z1$leadingEdge)
z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')

# drop newlines from leadingEdge
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
# somehow, one run of this sometimes doesnt get them all so repeat for good meas
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")

z1 <- as.data.table(z1)
z1[,dir:="LCNM_pos_up"]

   
# get LCNM_neg enriched results
register(MulticoreParam())
z2<-fgseaMultilevel(pathways=sets,stats = deres.id,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "neg")
# coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
z2$leadingEdge <- as.character(z2$leadingEdge)
z2$leadingEdge <- gsub(z2$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
z2$leadingEdge <- gsub(z2$leadingEdge,pattern='"',replacement='')

# drop newlines from leadingEdge
z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")
# somehow, one run of this sometimes doesnt get them all so repeat for good meas
z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")

z2 <- as.data.table(z2)
z2[,dir:="LCNM_neg_up"]

nmgseares.msigdb.signed <- as.data.table(rbind(z1,z2))

rm(z1,z2)
```


### PART 2: get SIGNED tf-target enrichment results DE t stat vectors per cluster per test for msigdb
```{r}
tfsets <- readRDS("raw-data/ref_data/00a-enrichrTFsets_forfgseainput_121823.RDS")

# get LCNM-up enriched results
z1<-fgseaMultilevel(pathways=tfsets,stats = deres.name,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")
# coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
z1$leadingEdge <- as.character(z1$leadingEdge)
z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')

# drop newlines from leadingEdge
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
# somehow, one run of this sometimes doesnt get them all so repeat for good meas
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")

z1 <- as.data.table(z1)
z1[,dir:="LCNM_pos_up"]

   
# get LCNM_neg enriched results
register(MulticoreParam())
z2<-fgseaMultilevel(pathways=tfsets,stats = deres.name,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "neg")
# coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
z2$leadingEdge <- as.character(z2$leadingEdge)
z2$leadingEdge <- gsub(z2$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
z2$leadingEdge <- gsub(z2$leadingEdge,pattern='"',replacement='')

# drop newlines from leadingEdge
z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")
# somehow, one run of this sometimes doesnt get them all so repeat for good meas
z2$leadingEdge <- gsub(z2$leadingEdge,pattern=", \n",replacement=", ")

z2 <- as.data.table(z2)
z2[,dir:="LCNM_neg_up"]

nmgseares.tftarg.signed <- as.data.table(rbind(z1,z2))

rm(z1,z2,deres.name,deres.id)
```

### part 3: run UNSIGNED fgsea (genes that tend to be affected, regardless of direction)-msigdb
```{r}
lcnmde <- fread("processed-data/12_DEanalyses_removedsampsAndFinalNMseg/04-LCNMpos_vs_LCNMneg_DE.txt")

### drop mito genes
lcnmde <- lcnmde[!(chr %in% grep(chr,pattern="chrMT",value=T))]

### get named, SIGNED vectors of DE t-stats from each test
setnames(lcnmde,"t","tstat")
tmp <- copy(lcnmde)[,.(gene_id,gene_name,abs(tstat))]
setnames(tmp,3,"tstat")
setorderv(tmp,"tstat",-1)
## get the stat column as a vector and give it the gene ids as names for fgsea
## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
setorderv(tmp,"tstat",-1)
## get the stat column as a vector and give it the gene ids as names for fgsea
## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
tmp <- as.data.frame(tmp)
v.id <- tmp$tstat
v.name <- tmp$tstat

names(v.id) <- tmp$gene_id
names(v.name) <- tmp$gene_name
deres.id.us <- v.id
deres.name.us <- v.name

rm(v.id,v.name,tmp)
```

```{r}
register(MulticoreParam())

# get DE-enriched results
z1<-fgseaMultilevel(pathways=sets,stats = deres.id.us,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")

# coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
z1$leadingEdge <- as.character(z1$leadingEdge)
z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')
# drop newlines from leadingEdge
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
# belt and suspenders
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")

z1 <- as.data.table(z1)
z1[,dir:="either"]
nmgseares.msigdb.unsign <- copy(z1)

rm(z1,sets)
```

## part 4: UNsigned TFtarg enrichment
```{r}
register(MulticoreParam())

# get DE-enriched results
z1<-fgseaMultilevel(pathways=tfsets,stats = deres.name.us,minSize = 15,maxSize = 500,eps=0.0,nPermSimple = 50000,nproc=10,scoreType = "pos")

# coonvert leadingedge to one long string (it's returned as a list which is not compatible with data.table)
z1$leadingEdge <- as.character(z1$leadingEdge)
z1$leadingEdge <- gsub(z1$leadingEdge,pattern="^c\\((.*)\\)$",replacement="\\1")
z1$leadingEdge <- gsub(z1$leadingEdge,pattern='"',replacement='')
# drop newlines from leadingEdge
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")
# belt and suspenders
z1$leadingEdge <- gsub(z1$leadingEdge,pattern=", \n",replacement=", ")

z1 <- as.data.table(z1)
z1[,dir:="either"]
nmgseares.tftarg.unsign <- copy(z1)

rm(z1,sets)
```

## save
```{r}
nmgseas <- list(
    msigdb_signed = nmgseares.msigdb.signed,
    tftarg_signed = nmgseares.tftarg.signed,
    msigdb_unsigned = nmgseares.msigdb.unsign,
    tftarg_unsigned = nmgseares.tftarg.unsign
)

saveRDS(nmgseas,"processed-data/12_2_DEanalyses_removedsampsAndFinalNMseg/05_2-LCNMposVsNeg_msigdb_and_tftarg_GSEAs.RDS")
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] msigdbr_25.1.1              fgsea_1.36.2               
 [3] clusterProfiler_4.18.4      BiocParallel_1.44.0        
 [5] parallelly_1.46.1           default_1.0.0              
 [7] colorout_1.3-3              SpatialExperiment_1.20.0   
 [9] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[11] Biobase_2.70.0              GenomicRanges_1.62.1       
[13] Seqinfo_1.0.0               IRanges_2.44.0             
[15] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[17] generics_0.1.4              MatrixGenerics_1.22.0      
[19] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3      rstudioapi_0.18.0       jsonlite_2.0.0         
  [4] tidydr_0.0.6            magrittr_2.0.4          ggtangle_0.1.1         
  [7] magick_2.9.0            farver_2.1.2            rmarkdown_2.30         
 [10] fs_1.6.6                vctrs_0.7.1             memoise_2.0.1          
 [13] ggtree_4.0.4            htmltools_0.5.9         S4Arrays_1.10.1        
 [16] curl_7.0.0              SparseArray_1.10.8      gridGraphics_0.5-1     
 [19] htmlwidgets_1.6.4       plyr_1.8.9              cachem_1.1.0           
 [22] igraph_2.2.1            lifecycle_1.0.5         pkgconfig_2.0.3        
 [25] Matrix_1.7-4            R6_2.6.1                fastmap_1.2.0          
 [28] gson_0.1.0              digest_0.6.39           aplot_0.2.9            
 [31] enrichplot_1.30.4       ggnewscale_0.5.2        patchwork_1.3.2        
 [34] AnnotationDbi_1.72.0    rprojroot_2.1.1         RSQLite_2.4.5          
 [37] httr_1.4.7              polyclip_1.10-7         abind_1.4-8            
 [40] compiler_4.5.2          here_1.0.2              bit64_4.6.0-1          
 [43] fontquiver_0.2.1        withr_3.0.2             S7_0.2.1               
 [46] DBI_1.2.3               ggforce_0.5.0           R.utils_2.13.0         
 [49] MASS_7.3-65             rappdirs_0.3.4          DelayedArray_0.36.0    
 [52] rjson_0.2.23            tools_4.5.2             otel_0.2.0             
 [55] ape_5.8-1               scatterpie_0.2.6        R.oo_1.27.1            
 [58] glue_1.8.0              nlme_3.1-168            GOSemSim_2.36.0        
 [61] grid_4.5.2              rsconnect_1.7.0         cluster_2.1.8.1        
 [64] reshape2_1.4.5          gtable_0.3.6            R.methodsS3_1.8.2      
 [67] tidyr_1.3.2             XVector_0.50.0          ggrepel_0.9.6          
 [70] pillar_1.11.1           stringr_1.6.0           babelgene_22.9         
 [73] yulab.utils_0.2.4       splines_4.5.2           dplyr_1.2.0            
 [76] tweenr_2.0.3            treeio_1.34.0           lattice_0.22-7         
 [79] bit_4.6.0               tidyselect_1.2.1        fontLiberation_0.1.0   
 [82] GO.db_3.22.0            Biostrings_2.78.0       knitr_1.51             
 [85] fontBitstreamVera_0.1.1 xfun_0.56               stringi_1.8.7          
 [88] lazyeval_0.2.2          ggfun_0.2.0             yaml_2.3.12            
 [91] evaluate_1.0.5          codetools_0.2-20        gdtools_0.4.4          
 [94] tibble_3.3.1            qvalue_2.42.0           ggplotify_0.1.3        
 [97] cli_3.6.5               systemfonts_1.3.1       dichromat_2.0-0.1      
[100] Rcpp_1.1.1              png_0.1-8               assertthat_0.2.1       
[103] ggplot2_4.0.2           blob_1.3.0              DOSE_4.4.0             
[106] tidytree_0.4.7          ggiraph_0.9.3           scales_1.4.0           
[109] purrr_1.2.1             crayon_1.5.3            rlang_1.1.7            
[112] cowplot_1.2.0           fastmatch_1.1-8         KEGGREST_1.50.0        
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-22
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 ape                    5.8-1    2024-12-16 [1] CRAN (R 4.5.0)
 aplot                  0.2.9    2025-09-12 [1] CRAN (R 4.5.0)
 assertthat             0.2.1    2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9     2022-09-29 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0    2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1  2025-01-16 [1] CRAN (R 4.5.0)
 blob                   1.3.0    2026-01-14 [1] CRAN (R 4.5.2)
 cachem                 1.1.0    2024-05-16 [1] CRAN (R 4.5.0)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1  2025-03-12 [1] CRAN (R 4.5.0)
 clusterProfiler      * 4.18.4   2025-12-15 [1] Bioconductor 3.22 (R 4.5.2)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 cowplot                1.2.0    2025-07-07 [1] CRAN (R 4.5.0)
 crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0    2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3    2024-06-02 [1] CRAN (R 4.5.0)
 default              * 1.0.0    2017-08-07 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 DOSE                   4.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 enrichplot             1.30.4   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 fastmatch              1.1-8    2026-01-17 [1] CRAN (R 4.5.2)
 fgsea                * 1.36.2   2026-01-05 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 fontBitstreamVera      0.1.1    2017-02-01 [1] CRAN (R 4.5.0)
 fontLiberation         0.1.0    2016-10-15 [1] CRAN (R 4.5.0)
 fontquiver             0.2.1    2017-02-01 [1] CRAN (R 4.5.0)
 fs                     1.6.6    2025-04-12 [1] CRAN (R 4.5.0)
 gdtools                0.4.4    2025-10-06 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggforce                0.5.0    2025-06-18 [1] CRAN (R 4.5.0)
 ggfun                  0.2.0    2025-07-15 [1] CRAN (R 4.5.0)
 ggiraph                0.9.3    2026-01-19 [1] CRAN (R 4.5.2)
 ggnewscale             0.5.2    2025-06-20 [1] CRAN (R 4.5.0)
 ggplot2                4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggplotify              0.1.3    2025-09-20 [1] CRAN (R 4.5.0)
 ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 ggtangle               0.1.1    2026-01-16 [1] CRAN (R 4.5.2)
 ggtree                 4.0.4    2026-01-05 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 GO.db                  3.22.0   2026-01-31 [1] Bioconductor
 GOSemSim               2.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gridGraphics           0.5-1    2020-12-13 [1] CRAN (R 4.5.0)
 gson                   0.1.0    2023-03-07 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 htmlwidgets            1.6.4    2023-12-06 [1] CRAN (R 4.5.0)
 httr                   1.4.7    2023-08-15 [1] CRAN (R 4.5.0)
 igraph                 2.2.1    2025-10-27 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 jsonlite               2.0.0    2025-03-27 [1] CRAN (R 4.5.0)
 KEGGREST               1.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lazyeval               0.2.2    2019-03-15 [1] CRAN (R 4.5.0)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 MASS                   7.3-65   2025-02-28 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1    2021-11-26 [1] CRAN (R 4.5.0)
 msigdbr              * 25.1.1   2025-07-21 [1] CRAN (R 4.5.2)
 nlme                   3.1-168  2025-03-31 [1] CRAN (R 4.5.2)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 patchwork              1.3.2    2025-08-25 [1] CRAN (R 4.5.0)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9    2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8    2022-11-29 [1] CRAN (R 4.5.0)
 polyclip               1.10-7   2024-07-23 [1] CRAN (R 4.5.0)
 purrr                  1.2.1    2026-01-09 [1] CRAN (R 4.5.2)
 qvalue                 2.42.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 R.methodsS3            1.8.2    2022-06-13 [1] CRAN (R 4.5.0)
 R.oo                   1.27.1   2025-05-02 [1] CRAN (R 4.5.0)
 R.utils                2.13.0   2025-02-24 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 rappdirs               0.3.4    2026-01-17 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 reshape2               1.4.5    2025-11-12 [1] CRAN (R 4.5.0)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5    2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scatterpie             0.2.6    2025-09-12 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 stringi                1.8.7    2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0    2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1    2025-10-01 [1] CRAN (R 4.5.0)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidydr                 0.0.6    2025-07-25 [1] CRAN (R 4.5.0)
 tidyr                  1.3.2    2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 tidytree               0.4.7    2026-01-08 [1] CRAN (R 4.5.2)
 treeio                 1.34.0   2025-10-30 [1] Bioconductor 3.22 (R 4.5.1)
 tweenr                 2.0.3    2024-02-26 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)
 yulab.utils            0.2.4    2026-02-02 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
