---
title: "06b_2-LC_NMFLogCt18Filt_factorSpecGenes"
author: "Bernie Mulvey"
date: "2026-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallel)
library(BiocParallel)
options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## load the PRE-GENE-FILTERED NMF results from LC spots; identify factors with relatively factor-specific gene loadings for many genes
```{r}
# get gene metadata
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")
rddt <- as.data.table(rowData(lc3)) 
rm(lc3)
gc(full=T)


### now we can load and tackle the GENE FILTERED nmf results
lc.nmf2 <- readRDS("processed-data/10_2_subclustering_and_NMF/07b-lcnmf120_tol1e-9_rowSumLogctFilt18.RDS")
w2 <- as.data.table(lc.nmf2$w,keep.rownames=T)
w2 <- merge.data.table(w2,rddt[,.(gene_id,gene_name)],by.x="rn",by.y="gene_id")
nameord <- c("rn","gene_name",paste0("LCnmf_filt",c(1:120)))
w2 <- w2[,..nameord]
setnames(w2,"rn","gene_id")

## find factors with a "factor-specific"ish gene based on the loading for that factor and median abs deviation of that genes loadings across all factors.

w2m <- melt.data.table(w2,id.vars=c("gene_id","gene_name"),variable.name = "factor")
w2m[,medianloading:=median(value),by="gene_id"]
w2m[,factorGene:=ifelse(value>1.5*mad(value,constant=1,na.rm = T),"y","n")]

w2m.sub <- w2m[factorGene=="y"]

## get genes that are "factor specific" in 3 or fewer factors
fewfactorgenes <- w2m.sub[,.N,by="gene_id"][N<=3,gene_id]
# 1367 genes total there

# now subset to these genes and identify factors with at least 25 such  factor specific-ish genes -- by only considering genes that meet the factor-specificish condition
w2m.select <- w2m.sub[gene_id %in% fewfactorgenes]
## how many factors have more than, say, 25 genes?
keepfacts <- w2m.select[,.N,by="factor"][N>=25,as.character(factor)]
## 17 factors, nice
w2m.select <- w2m.select[factor %in% keepfacts] # these are our factors of interest AND their specific-ish genes

## save our table of factors with 25+ specific genes and which genes go with each
fwrite(w2m.select,"processed-data/12_2_DEanalyses_removedsampsAndFinalNMseg/06b_2_LCnmfLogCt18Filt_factors_and_factorSpecGenes_min25specGenes.txt",sep='\t',quote=F,row.names=F,col.names=T)
```


repinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              SpatialExperiment_1.20.0   
 [5] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0               IRanges_2.44.0             
[11] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0           data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] Matrix_1.7-4        compiler_4.5.2      rjson_0.2.23        Rcpp_1.1.1         
 [5] magick_2.9.0        yaml_2.3.12         fastmap_1.2.0       lattice_0.22-7     
 [9] here_1.0.2          XVector_0.50.0      S4Arrays_1.10.1     knitr_1.51         
[13] DelayedArray_0.36.0 rprojroot_2.1.1     rlang_1.1.7         xfun_0.56          
[17] otel_0.2.0          SparseArray_1.10.8  cli_3.6.5           magrittr_2.0.4     
[21] digest_0.6.39       grid_4.5.2          rstudioapi_0.18.0   lifecycle_1.0.5    
[25] evaluate_1.0.5      codetools_0.2-20    abind_1.4-8         rsconnect_1.7.0    
[29] rmarkdown_2.30      tools_4.5.2         htmltools_0.5.9    
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-14
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
