---
title: "05-GlobalAncestry_DEanalysis"
author: "Bernie Mulvey"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(edgeR)
library(limma)

# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

This is almost the same as script 11/02, but with E4 as 0, 1, 2 (number of E4 alleles) and with FLARE CEU global ancestry proportion estimates as the variable of interest


## load DGElist objects
```{r}
dgels <- readRDS("processed-data/11_DEanalyses/01-DGElist_per_25HDG75SVGlouv1_cluster_normFilt.RDS")

## append DGElist for LC NM+ and NM- fractions

nmdgels <- readRDS("processed-data/11_DEanalyses/01b-DGElist_per_LCNMadj_subcluster_normFilt.RDS")

dgels <- c(dgels,list(nmdgels$LC.1_NMpos,nmdgels$LC.1_NMneg))
names(dgels)[10:11] <- c("LC.1_NMpos","LC.1_NMneg")

rm(nmdgels)
### to avoid any issues, replace "Br6119(re-dis)" with "Br6119" in the samples tables
dgels <- lapply(dgels,FUN=function(d){
    d$samples$brnum[d$samples$brnum=="Br6119(re-dis)"] <- "Br6119"

    ## convert APOE genotype to E4 copy number
    
    d$samples$e4copies <- 0
    d$samples$e4copies[d$samples$APOE=="E4/ E4"] <- 2
    d$samples$e4copies[d$samples$APOE=="E3/ E4"] <- 1 # all hets are E4/E3
    
    ## get rid of the other apoe columns to force an error to return if we accidentally are calling for them later
    d$samples$e4 <- NULL
    d$samples$APOE <- NULL
    
    return(d)
})
```

## append global genetic ancestry proportion estimates (FLARE, 3-component CEU ancestry--the other two being YRI and CHB). this source file is intentionally not included in in github but is on JHPCE. The CHB proportions range 0.0001 to 0.01--so CEU is effectively the proportion that is not YRI, and we can just use CEU as the variable here. (use CEU for consistency with other analyses).

```{r}
flares <- fread("raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")

dgels <- lapply(dgels,FUN=function(d){
    tmpsamp <- as.data.table(d$samples,keep.rownames=T)
    tmpsamp <- merge.data.table(tmpsamp,flares,by.x="brnum",by.y="ID")
    tmpsamp <- as.data.frame(tmpsamp,row.names=tmpsamp$rn)[rownames(d$samples),]
    tmpsamp$rn <- NULL
    d$samples <- tmpsamp
    return(d)
})
```


### DE TEST 1 ### CEU proportion, WITHOUT best pfc rin as a covar
```{r}
setDTthreads(1,restore_after_fork = FALSE)
clus <- parallel::makeCluster(9,type = "PSOCK")
fulldgelres.anc.norin <- parallel::clusterApply(x=dgels,cl = clus,fun=function(X){
    
    library(data.table)
    library(limma)
    library(edgeR)
    
    # prevent resource competition, though we aren't explicitly using DT at all here:

    dge <- X
    des <- model.matrix(~0+e4copies+Sex+CEU+Age, data = dge$samples)
    
    # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
    
    # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
    v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$brnum),adaptive.span = T,sample.weights = T)
    cont <- makeContrasts(globAnc=CEU,levels=des)
    v.swt.fit <- contrasts.fit(v.swt,contrasts=cont)
    v.swt.fit.e <- eBayes(v.swt.fit)
    v.swt.e.tt.ga <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
    
    return(v.swt.e.tt.ga)
})
parallel::stopCluster(clus)

names(fulldgelres.anc.norin) <- names(dgels)
```

### DE TEST 2 ### WITH CEU proportion, model WITH best pfc RIN as a covar
```{r}
setDTthreads(1,restore_after_fork = FALSE)
clus <- parallel::makeCluster(9,type = "PSOCK")
fulldgelres.anc.rin <- parallel::clusterApply(x=dgels,cl = clus,fun=function(X){
    
    library(data.table)
    library(limma)
    library(edgeR)
    
    # prevent resource competition, though we aren't explicitly using DT at all here:

    dge <- X
    des <- model.matrix(~0+e4copies+Sex+CEU+Age+Rin, data = dge$samples)
    
    # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
    
    # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
    v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$brnum),adaptive.span = T,sample.weights = T)
    cont <- makeContrasts(localAnc=CEU,levels=des)
    v.swt.fit <- contrasts.fit(v.swt,contrasts=cont)
    v.swt.fit.e <- eBayes(v.swt.fit)
    v.swt.e.tt.ga <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
    
    return(v.swt.e.tt.ga)
})
parallel::stopCluster(clus)

names(fulldgelres.anc.rin) <- names(dgels)
```


```{r}
concatres <- list(CEU_noRin_DE=fulldgelres.anc.norin,CEU_wRin_DE=fulldgelres.anc.rin)

saveRDS(concatres,"processed-data/11_DEanalyses/05a-GlobalCEUprop_DE_with_e4alleleCountModel_toptabs_testsWithandWithout_RIN.RDS")
```


## tabulate how many CEU models picked up DE per cluster at FDR<0.2
```{r}
tabres <- mapply(a=fulldgelres.anc.norin,ar=fulldgelres.anc.rin,SIMPLIFY=FALSE,FUN=function(a,ar){

    ancs <- list(a,ar)
    names(ancs) <- c("E4_wAncestry_noRin","E4_wAncestry_wRin")
    
    ancsf <- mapply(x=ancs,n=names(ancs),SIMPLIFY=FALSE,FUN=function(x,n){
        x <- as.data.table(x)[adj.P.Val<0.2,gene_name]
        rettab <- cbind(x,n)
        rettab <- as.data.table(rettab)

        if(length(x)==0){
            return(NULL)
        } else
        {
            setnames(rettab,c("gene_name","DEtest_FDRunder0.2"))
            return(rettab)
        }
    })
    ancsf <- rbindlist(ancsf,fill=T)
    # account for cases where no fdr sig genes in any of the models:
    if(nrow(ancsf)==0){
        ancsf <- as.data.table(cbind(gene_name=NA,DEtest_FDRunder0.2=NA))
    } 
    # 
    ancsf[,Nsigmodels:=.N,by="gene_name"]

    ancsnom <- mapply(x=ancs,n=names(ancs),SIMPLIFY=FALSE,FUN=function(x,n){
        x <- as.data.table(x)[P.Value<0.05,gene_name]
        rettab <- cbind(x,n)
        rettab <- as.data.table(rettab)

        if(length(x)==0){
            return(NULL)
        } else
        {
            setnames(rettab,c("gene_name","DEtest_unadjP.05"))
            return(rettab)
        }
    })
    ancsnom <- rbindlist(ancsnom,fill=T)
    ancsnom[,Nsigmodels:=.N,by="gene_name"]
    
    retlist <- list(ancsf=ancsf,ancsnom=ancsnom)
    return(retlist)
})


### then, concatenate these within variable of interest and sig cutoff

ancfdrs <- rbindlist(mapply(x=tabres,n=names(tabres),SIMPLIFY=FALSE,FUN=function(x,n){as.data.table(cbind(x$ancsf,n))}),fill=T)
setnames(ancfdrs,c("DEtest_FDRunder0.2","Nsigmodels","n"),c("DEtest_FDRunder0.2_forCluster","Nsigmodels_forCluster","louvCluster"))
ancfdrs <- ancfdrs[,.(gene_name,louvCluster,DEtest_FDRunder0.2_forCluster,Nsigmodels_forCluster)]

ancunadjs <- rbindlist(mapply(x=tabres,n=names(tabres),SIMPLIFY=FALSE,FUN=function(x,n){as.data.table(cbind(x$ancsnom,n))}),fill=T)
setnames(ancunadjs,c("DEtest_unadjP.05","Nsigmodels","n"),c("DEtest_unadjP.05_forCluster","Nsigmodels_forCluster","louvCluster"))
ancunadjs <- ancunadjs[,.(gene_name,louvCluster,DEtest_unadjP.05_forCluster,Nsigmodels_forCluster)]
```

save these last two tables
```{r}
fwrite(ancfdrs,"processed-data/11_DEanalyses/05b-GlobalCEUprop_DE_with_e4alleleCountModel_FDR0.2_byGene_byNmodelsOf2max_byCluster.txt",sep='\t',quote=F,row.names=F,col.names=T)

fwrite(ancunadjs,"processed-data/11_DEanalyses/05c-GlobalCEUprop_DE_with_e4alleleCountModel_unadjP05_byGene_byNmodelsOf2max_byCluster.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

```{r}
rm(list=ls())
gc(full=T)
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.2 (2024-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.40.0 parallelly_1.42.0   colorout_1.3-0.2    edgeR_4.4.1        
[5] limma_3.62.1        data.table_1.16.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] digest_0.6.37     codetools_0.2-20  fastmap_1.2.0     xfun_0.50        
 [5] lattice_0.22-6    knitr_1.49        htmltools_0.5.8.1 rmarkdown_2.29   
 [9] cli_3.6.3         grid_4.4.2        statmod_1.5.0     compiler_4.4.2   
[13] rprojroot_2.0.4   here_1.0.1        rstudioapi_0.17.1 tools_4.4.2      
[17] evaluate_1.0.3    yaml_2.3.10       locfit_1.5-9.10  
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.2 (2024-10-31)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package      * version  date (UTC) lib source
    BiocParallel * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 P  cli            3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
    codetools      0.2-20   2024-03-31 [1] CRAN (R 4.4.2)
    colorout     * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    data.table   * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
    digest         0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
    edgeR        * 4.4.1    2024-12-02 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate       1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
    fastmap        1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
    here           1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
    htmltools      0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
    knitr          1.49     2024-11-08 [1] CRAN (R 4.4.1)
    lattice        0.22-6   2024-03-20 [1] CRAN (R 4.4.2)
    limma        * 3.62.1   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    locfit         1.5-9.10 2024-06-24 [1] CRAN (R 4.4.0)
    parallelly   * 1.42.0   2025-01-30 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.4    2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown      2.29     2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot      2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi     0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
    sessioninfo    1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
    statmod        1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
    xfun           0.50     2025-01-07 [1] CRAN (R 4.4.1)
    yaml           2.3.10   2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
