---
title: "01-allgeneGLMPCA_louvainCluster_spotplots"
output: html_document
date: "2024-12-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(Polychrome)
library(escheR)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load QC'ed spe, clustering results; subset to louvain clusterings only (leiden clusterings all returned 100s-1000s of clusters)
```{r}
lc2 <- readRDS("processed-data/05_QC/04-lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed.RDS")

SNNlouvlist <- readRDS("processed-data/07_Clustering/01-GLMPCAallgene_SNN10-15-20_leidenAndLouvain_res05-1-2.RDS")
## all clustering assignments are in columns with names reflecting the clustering method, SNN k value, and resolution parameter. each list entry is the results for one SNN k value so we can merge the columns into one big table for simplicity here.
SNNlouvs <- merge.data.table(SNNlouvlist[[1]],SNNlouvlist[[2]],by="rn")
SNNlouvs <- merge.data.table(SNNlouvs,SNNlouvlist[[3]],by="rn")

## just use the louvain clusterings
keepclus <- c("rn",grep(names(SNNlouvs),pattern="_louv_",value=T))
SNNlouvs <- SNNlouvs[,..keepclus]
SNNlouvs <- DataFrame(SNNlouvs,row.names=SNNlouvs$rn)[colnames(lc2),]
SNNlouvs$rn <- NULL

rm(keepclus,SNNlouvlist)
```

### define a function to extract a legend from a ggplot (so we have one legend per output pdf); generate a gridded PDF for each clustering result
```{r}
# helper function to extract a legend and pipe back to a diff plot: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 

extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}


lapply(colnames(SNNlouvs),FUN=function(clus){
    colidx <- which(colnames(SNNlouvs)==clus)
    nclusts <- length(unique(SNNlouvs[, colidx]))
    colLabels(lc2) <- factor(SNNlouvs[, colidx],levels=paste0("X",c(1:nclusts)))
    pal <- palette36.colors(n = nclusts+2)[c(3:(nclusts+2))]
    # next line is needed for consistency in color-cluster assignments between libdspatial and the generated legend below
    names(pal) <- paste0("X",c(1:nclusts))
    
    ## create a unified legend spanning all possible cluster values (regardless of presence in a given sample)
    leg.tmp <- ggplot(as.data.table(colData(lc2))[,.(label,array_row,array_col)],aes(x=array_row,y=array_col,col=label))+geom_point()+
        scale_color_manual(values=as.character(pal))+
        guides(color=guide_legend(ncol=2,override.aes=list(size=3)))+
        theme(
                    legend.text = element_text(size = 10),
                    legend.title = element_text(size = 10, hjust = 0.5),
                    legend.spacing.y = unit(0, "cm"),
                    legend.spacing.x = unit(0.1, "cm"))
    leg.tmp2 <- extract_legend(leg.tmp)
    dev.off()
    ### make spotplots
    plt <-
        spatialLIBD::vis_grid_clus(
            lc2,
            clustervar = "label",
            pdf_file = NULL,
            colors = pal,
            spatial = F,
            auto_crop = T,
            return_plots = T,
            point_size = 1.85,
            sort_clust = F
        )
    
    ### adjust font sizes, drop individual plot legends
    plt <- lapply(
        plt,
        FUN = function(x) {
            x +
                theme(
                    plot.title = element_text(size = 11, hjust = 0.5),
                    strip.text = element_text(size = 10),
                )+
                guides(fill="none")
        }
    )
    
    # save
    pdf(
        paste0(
            "plots/07_Clustering/01-spotplots/01-allgeneGLMPCA_",
            clus,
            ".pdf"
        ),
        width = 26,
        height = 90
    )
    ### arrangeGrob() on the spotplot list --> 3 x 3 of plots from that list only, constituting the "first column" of the parent; leg.tmp2 is an additional thing being plotted in the second column of the parent. width of the plot grid is set to 24, width of the second column (legend only) set to 2.
    do.call("grid.arrange",list(arrangeGrob(grobs=plt,ncol=3),leg.tmp2,ncol=2,widths=c(24,2)))
    dev.off()
    
})
```

### some preliminary annotations of glmpca SNN k=10 louvain resolution 1 markers looks promising for some neuropeptide-rich cells surrounding LC. so load the rough annotations and subset to the pops of interest to plot the same way.

```{r}
snn10louv1 <- SNNlouvs$snn10_louv_res1
annot <- fread("processed-data/07_Clustering/03-glmpca_snn10_louv1_tentativeannotsbytop10lfcmks.txt")
annot <- unique(annot[,.(clus,label)])

snn10louv1 <- as.data.table(cbind(colnames(lc2),snn10louv1))
snn10louv1 <- merge.data.table(snn10louv1,annot,by.x="snn10louv1",by.y="clus")
snn10louv1 <- DataFrame(snn10louv1,row.names=snn10louv1$V1)[colnames(lc2),]

colLabels(lc2) <- snn10louv1$label

lc2sub <- lc2[,lc2$label %in% c("Inh.SST.CCK.HTR2C","LC.1","5HT.1","Exc.17A6.RBP4.TAC1.HTR2C","Unsure.2.PCDHGC5.AVP")]
lc2sub$label <- factor(lc2sub$label,levels=c("Inh.SST.CCK.HTR2C","LC.1","5HT.1","Exc.17A6.RBP4.TAC1.HTR2C","Unsure.2.PCDHGC5.AVP"))

### write plot
nclusts <- length(unique(lc2sub$label))
pal <- palette36.colors(n = nclusts+2)[c(3:(nclusts+2))]
# next line is needed for consistency in color-cluster assignments between libdspatial and the generated legend below
names(pal) <- levels(lc2sub$label)

## create a unified legend spanning all possible cluster values (regardless of presence in a given sample)
leg.tmp <- ggplot(as.data.table(colData(lc2sub))[,.(label,array_row,array_col)],aes(x=array_row,y=array_col,col=label))+geom_point()+
  scale_color_manual(values=pal)+
  guides(color=guide_legend(ncol=2,override.aes=list(size=3)))+
  theme(
              legend.text = element_text(size = 10),
              legend.title = element_text(size = 10, hjust = 0.5),
              legend.spacing.y = unit(0, "cm"),
              legend.spacing.x = unit(0.1, "cm"))
leg.tmp2 <- extract_legend(leg.tmp)
dev.off()
### make spotplots
s <- unique(lc2sub$sample_id)

plt <- lapply(s,FUN=function(s){
   cursub <- lc2sub[,lc2sub$sample_id==s]
   
   mke <- make_escheR(cursub)
   mke <- mke |> add_fill("label",size=2.5,point_size = 2.5)
   mke <- mke+ggtitle(s)
   return(mke)
})

# pull legend from a plot, then wipe legends from individual plots
leg.tmp2 <- extract_legend(plt[[1]])
dev.off()

### adjust font sizes, drop individual plot legends
plt <- lapply(
  plt,
  FUN = function(x) {
      x +
          theme(
              plot.title = element_text(size = 14, hjust = 0.5),
              strip.text = element_text(size = 10),
          )+
          guides(fill="none")
  }
)


# save
pdf(
      "plots/07_Clustering/01-spotplots/01-allgeneGLMPCA_SNN10_louvain1_clusofinterestonly.pdf",
  width = 26,
  height = 90,
  onefile=FALSE
)
### arrangeGrob() on the spotplot list --> 3 x 3 of plots from that list only, constituting the "first column" of the parent; leg.tmp2 is an additional thing being plotted in the second column of the parent. width of the plot grid is set to 24, width of the second column (legend only) set to 2.
do.call("grid.arrange",list(arrangeGrob(grobs=plt,ncol=3),leg.tmp2,ncol=2,widths=c(24,2)))
dev.off()

```


reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] Polychrome_1.5.1            BiocParallel_1.38.0        
 [3] parallelly_1.38.0           colorout_1.3-0.2           
 [5] spatialLIBD_1.16.2          SpatialExperiment_1.14.0   
 [7] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
 [9] Biobase_2.64.0              GenomicRanges_1.56.1       
[11] GenomeInfoDb_1.40.1         IRanges_2.38.1             
[13] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[15] MatrixGenerics_1.16.0       matrixStats_1.3.0          
[17] gridExtra_2.3               ggplot2_3.5.1              
[19] data.table_1.15.4           rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0             bitops_1.0-7             
  [4] filelock_1.0.3            fields_16.2               tibble_3.2.1             
  [7] XML_3.99-0.17             lifecycle_1.0.4           edgeR_4.2.0              
 [10] doParallel_1.0.17         rprojroot_2.0.4           lattice_0.22-6           
 [13] magrittr_2.0.3            limma_3.60.3              plotly_4.10.4            
 [16] sass_0.4.9                rmarkdown_2.27            jquerylib_0.1.4          
 [19] yaml_2.3.9                httpuv_1.6.15             spam_2.10-0              
 [22] sessioninfo_1.2.2         cowplot_1.1.3             DBI_1.2.3                
 [25] RColorBrewer_1.1-3        golem_0.4.1               maps_3.4.2               
 [28] abind_1.4-5               zlibbioc_1.50.0           purrr_1.0.2              
 [31] RCurl_1.98-1.14           rappdirs_0.3.3            GenomeInfoDbData_1.2.12  
 [34] ggrepel_0.9.5             irlba_2.3.5.1             DelayedMatrixStats_1.26.0
 [37] codetools_0.2-20          DelayedArray_0.30.1       DT_0.33                  
 [40] scuttle_1.14.0            tidyselect_1.2.1          farver_2.1.2             
 [43] UCSC.utils_1.0.0          ScaledMatrix_1.12.0       viridis_0.6.5            
 [46] shinyWidgets_0.8.6        BiocFileCache_2.12.0      GenomicAlignments_1.40.0 
 [49] jsonlite_1.8.8            BiocNeighbors_1.22.0      scater_1.32.0            
 [52] iterators_1.0.14          foreach_1.5.2             tools_4.4.1              
 [55] Rcpp_1.0.13               glue_1.7.0                SparseArray_1.4.8        
 [58] xfun_0.45                 here_1.0.1                dplyr_1.1.4              
 [61] withr_3.0.0               BiocManager_1.30.23       fastmap_1.2.0            
 [64] fansi_1.0.6               digest_0.6.36             rsvd_1.0.5               
 [67] R6_2.5.1                  mime_0.12                 colorspace_2.1-0         
 [70] RSQLite_2.3.7             config_0.3.2              utf8_1.2.4               
 [73] tidyr_1.3.1               generics_0.1.3            rtracklayer_1.64.0       
 [76] httr_1.4.7                htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [79] scatterplot3d_0.3-44      pkgconfig_2.0.3           gtable_0.3.5             
 [82] blob_1.2.4                XVector_0.44.0            htmltools_0.5.8.1        
 [85] dotCall64_1.1-1           scales_1.3.0              png_0.1-8                
 [88] attempt_0.3.1             knitr_1.48                rstudioapi_0.16.0        
 [91] rjson_0.2.21              curl_5.2.1                cachem_1.1.0             
 [94] BiocVersion_3.19.1        vipor_0.4.7               AnnotationDbi_1.66.0     
 [97] restfulr_0.0.15           pillar_1.9.0              grid_4.4.1               
[100] vctrs_0.6.5               promises_1.3.0            BiocSingular_1.20.0      
[103] dbplyr_2.5.0              beachmat_2.20.0           xtable_1.8-4             
[106] beeswarm_0.4.0            paletteer_1.6.0           evaluate_0.24.0          
[109] magick_2.8.3              cli_3.6.3                 locfit_1.5-9.10          
[112] compiler_4.4.1            Rsamtools_2.20.0          crayon_1.5.3             
[115] labeling_0.4.3            rematch2_2.1.2            ggbeeswarm_0.7.2         
[118] viridisLite_0.4.2         munsell_0.5.1             Biostrings_2.72.1        
[121] lazyeval_0.2.2            Matrix_1.7-0              ExperimentHub_2.12.0     
[124] benchmarkme_1.0.8         sparseMatrixStats_1.16.0  bit64_4.0.5              
[127] KEGGREST_1.44.1           statmod_1.5.0             shiny_1.8.1.1            
[130] AnnotationHub_3.12.0      memoise_2.0.1             bslib_0.7.0              
[133] benchmarkmeData_1.0.4     bit_4.0.5                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-17
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular           1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings             2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bslib                  0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64              1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   knitr                  1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome           * 1.5.1     2022-05-03 [1] CRAN (R 4.4.0)
   promises               1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                  1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   S4Arrays               1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scatterplot3d          0.3-44    2023-05-05 [1] CRAN (R 4.4.0)
   scuttle                1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spam                   2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
