---
title: "02-louvainClusts_on_allgeneGLMPCA"
output: html_document
date: "2024-11-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(HDF5Array)
library(ggrastr)
library(Polychrome)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

# only use louvain clusterings (leiden returned 100s to 1000s)
```{r}
lcglmpca <- HDF5Array("processed-data/06_FeatSelection_Dimred_Harmony/03b_lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed_allgeneGLMPCA.h5",name="GLM-PCA")

## get the first 4 pcs, only ones we're plotting
lcglmpca <- as.data.table(lcglmpca[,c("PC1","PC2","PC3","PC4")],keep.rownames=T)

SNNlouvlist <- readRDS("processed-data/07_Clustering/01-GLMPCAallgene_SNN10-15-20_leidenAndLouvain_res05-1-2.RDS")
## all clustering assignments are in columns with names reflecting the clustering method, SNN k value, and resolution parameter. each list entry is the results for one SNN k value so we can merge the columns into one big table for simplicity here.
SNNlouvs <- merge.data.table(SNNlouvlist[[1]],SNNlouvlist[[2]],by="rn")
SNNlouvs <- merge.data.table(SNNlouvs,SNNlouvlist[[3]],by="rn")

## just use the louvain clusterings
keepclus <- c("rn",grep(names(SNNlouvs),pattern="_louv_",value=T))
SNNlouvs <- SNNlouvs[,..keepclus]

rm(keepclus,SNNlouvlist)
```

## define helper function to extract a single legend - from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

## plot cluster assignments onto dims 1-2, 1-3, 1-4, 2-3, 2-4, 3-4
```{r}
lapply(colnames(SNNlouvs)[2:ncol(SNNlouvs)],FUN=function(clus){
    nclus <- length(unique(SNNlouvs[,get(clus)]))
    
    pal <- pal <- palette36.colors(n = nclus+2)[c(3:(nclus+2))]
    # next line is needed for consistency in color-cluster assignments between libdspatial and the generated legend below
    names(pal) <- paste0("X",c(1:nclus))
    
    # append cluster assignments to lcglmpca
    lcglmpca <- merge.data.table(lcglmpca,SNNlouvs[,.(rn,get(clus))],by="rn")
    setnames(lcglmpca,6,"clus")
    ## fetch a legend with this color palette to place next to all PC plots
    leg.tmp <- ggplot(lcglmpca,aes(x=PC1,y=PC2,col=clus))+
        geom_point()+
        scale_color_manual(values=as.character(pal))+
        guides(color=guide_legend(ncol=2,override.aes=list(size=3)))+
        theme(
                    legend.text = element_text(size = 10),
                    legend.title = element_text(size = 10, hjust = 0.5),
                    legend.spacing.y = unit(0, "cm"),
                    legend.spacing.x = unit(0.1, "cm"))
    leg.tmp2 <- extract_legend(leg.tmp)
    dev.off()
    
    
    ## now generate the six pc pair plots
    colcombs <- list(c("PC1","PC2"),
                     c("PC1","PC3"),
                     c("PC1","PC4"),
                     c("PC2","PC3"),
                     c("PC2","PC4"),
                     c("PC3","PC4"))
    p <- lapply(colcombs,FUN=function(cc){
        ggplot(lcglmpca,aes_string(x=cc[1],y=cc[2],color="clus"))+
            rasterize(geom_point(size=0.1),dpi=300,dev="cairo_png")+
            guides(color="none")+
            ggtitle(paste0(clus,":\n",cc[1]," vs ",cc[2]))+
            theme(title = element_text(size=10))
    })
    
    pdf(paste0("plots/07_Clustering/02-GLMPCA/02-",clus,"_on_PCs.pdf"),height=13,width=21)
        do.call("grid.arrange",list(arrangeGrob(grobs=p,ncol=3),leg.tmp2,ncol=2,widths=c(19,2)))
    dev.off()
})

```

reprod info:
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] Polychrome_1.5.1      BiocParallel_1.38.0   parallelly_1.38.0    
 [4] colorout_1.3-0.2      ggrastr_1.0.2         HDF5Array_1.32.0     
 [7] rhdf5_2.48.0          DelayedArray_0.30.1   SparseArray_1.4.8    
[10] S4Arrays_1.4.1        abind_1.4-5           IRanges_2.38.1       
[13] S4Vectors_0.42.1      MatrixGenerics_1.16.0 matrixStats_1.3.0    
[16] BiocGenerics_0.50.0   Matrix_1.7-0          gridExtra_2.3        
[19] ggplot2_3.5.1         data.table_1.15.4     rlang_1.1.4          

loaded via a namespace (and not attached):
 [1] utf8_1.2.4           generics_0.1.3       lattice_0.22-6      
 [4] digest_0.6.36        magrittr_2.0.3       evaluate_0.24.0     
 [7] grid_4.4.1           fastmap_1.2.0        rprojroot_2.0.4     
[10] fansi_1.0.6          scales_1.3.0         codetools_0.2-20    
[13] cli_3.6.3            crayon_1.5.3         scatterplot3d_0.3-44
[16] XVector_0.44.0       munsell_0.5.1        withr_3.0.0         
[19] yaml_2.3.9           ggbeeswarm_0.7.2     parallel_4.4.1      
[22] tools_4.4.1          dplyr_1.1.4          colorspace_2.1-0    
[25] Rhdf5lib_1.26.0      here_1.0.1           png_0.1-8           
[28] vctrs_0.6.5          R6_2.5.1             lifecycle_1.0.4     
[31] zlibbioc_1.50.0      vipor_0.4.7          beeswarm_0.4.0      
[34] pkgconfig_2.0.3      pillar_1.9.0         gtable_0.3.5        
[37] glue_1.7.0           xfun_0.45            tibble_3.2.1        
[40] tidyselect_1.2.1     rstudioapi_0.16.0    knitr_1.48          
[43] rhdf5filters_1.16.0  farver_2.1.2         htmltools_0.5.8.1   
[46] labeling_0.4.3       rmarkdown_2.27       compiler_4.4.1      
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-11-29
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package        * version date (UTC) lib source
   abind          * 1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   beeswarm         0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   BiocGenerics   * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel   * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli              3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools        0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout       * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace       2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon           1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table     * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   DelayedArray   * 0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   digest           0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr            1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate         0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi            1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver           2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap          1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics         0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggbeeswarm       0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2        * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr        * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   glue             1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra      * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable           0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array      * 1.32.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here             1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools        0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   IRanges        * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   knitr            1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling         0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lattice          0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle        1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr         2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix         * 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats    * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   munsell          0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly     * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar           1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig        2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   png              0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome     * 1.5.1   2022-05-03 [1] CRAN (R 4.4.0)
   R6               2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   rhdf5          * 2.48.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters     1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib         1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P rlang          * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown        2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot        2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi       0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Arrays       * 1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors      * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales           1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   scatterplot3d    0.3-44  2023-05-05 [1] CRAN (R 4.4.0)
   sessioninfo      1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   SparseArray    * 1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble           3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect       1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8             1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs            0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor            0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   withr            3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun             0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector          0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml             2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc         1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
