---
title: "03-Boxplot_top_APOE4_allelecount_DEGs"
author: "Bernie Mulvey"
date: "2025-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 10,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

get the LCNM+/- DE results and the LC parent cluster DE results
```{r}
nmsub <- readRDS("processed-data/11_DEanalyses/04a-E4andSexDE_with_e4alleleCountModel_toptabs_testsWithandWithout_RINandorAncestry_by_LCadjAndLCNMadj.RDS")

## the first four slots are adjacency subclusters which dont include LC
nmsub <- nmsub[5:8]
nmsub2 <- mapply(s=nmsub,n=names(nmsub),SIMPLIFY=FALSE,FUN=function(s,n){
    retlist <- list(s$LC.1_NMpos,s$LC.1_NMneg)
    names(retlist) <- c("LC.1_NMpos","LC.1_NMneg")
    
    retlist <- lapply(retlist,FUN=function(x){
        y <- rbindlist(x,fill=T,idcol="de_var")
        y[,de_model:=n]
        return(y)
    })
    rettab <- rbindlist(retlist,fill=T,idcol="clust")
    return(rettab)
})
nmsub2 <- rbindlist(nmsub2,fill=T)

## get parent cluster DE results
alllc <- readRDS("processed-data/11_DEanalyses/03a-E4andSexDE_with_e4alleleCountModel_toptabs_testsWithandWithout_RINandorAncestry.RDS")

alllc <- mapply(s=alllc,n=names(alllc),SIMPLIFY=FALSE,FUN=function(s,n){
    y <- s$LC.1
    y <- rbindlist(y,fill=T,idcol="de_var")
    y[,de_model:=n]
    y[,clust:="LC.1"]
    return(y)
})
alllc <- rbindlist(alllc,fill=T)

## get columns in same order and concat
ordnames <- names(nmsub2)
alllc <- alllc[,..ordnames]
plttab <- rbind(nmsub2,alllc)
```

```{r}
clpair <- list(c("LC.1_NMpos","LC.1_NMneg"),c("LC.1_NMpos","LC.1"),c("LC.1_NMneg","LC.1"))

plttab2 <- plttab[,.(clust,de_var,de_model,chr,gene_id,t)]

plts <- lapply(clpair,FUN=function(x){
    y <- copy(plttab2)[clust==x[[1]]&!(chr %in% c("chrX","chrY"))]
    z <- copy(plttab2)[clust==x[[2]]&!(chr %in% c("chrX","chrY"))]
    y[,chr:=NULL]
    y[,clust:=NULL]
    y[,de_model:=gsub(de_model,pattern="_LCNMadj",replacement="")]

    z[,chr:=NULL]
    z[,clust:=NULL]
    z[,de_model:=gsub(de_model,pattern="_LCNMadj",replacement="")]
    
    setnames(y,"t",paste0(x[[1]],"_t"))
    setnames(z,"t",paste0(x[[2]],"_t"))
    y <- merge.data.table(y,z,by=c("de_var","de_model","gene_id"))
    
    p <- ggplot(y,aes_string(x=paste0(x[1],"_t"),y=paste0(x[2],"_t")))+
        geom_point(size=0.4,stroke=0.1)+
        facet_wrap(~de_var+de_model)+
        geom_abline(intercept=0,slope=1,col="red",linetype="dashed")+
        xlab(x[[1]])+
        ylab(x[[2]])+
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
        
    return(p)
    
})


png("plots/11_DEanalyses/04_LCNMfrxn_DEs_vs_eachother_andWholeLC.png",height=1000*length(plts),width=1000)
do.call(grid.arrange,c(plts,ncol=1))
dev.off()
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
RR version 4.4.2 (2024-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils    
[5] methods   base     

other attached packages:
[1] colorout_1.3-0.2  gridExtra_2.3    
[3] ggtext_0.1.2      ggplot2_3.5.1    
[5] data.table_1.16.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] Matrix_1.7-1        gtable_0.3.6       
 [3] gam_1.22-5          dplyr_1.1.4        
 [5] compiler_4.4.2      crayon_1.5.3       
 [7] tidyselect_1.2.1    Rcpp_1.0.14        
 [9] Biobase_2.66.0      xml2_1.3.6         
[11] siggenes_1.80.0     splines_4.4.2      
[13] scales_1.3.0        lattice_0.22-6     
[15] here_1.0.1          R6_2.5.1           
[17] labeling_0.4.3      generics_0.1.3     
[19] knitr_1.49          MASS_7.3-61        
[21] BiocGenerics_0.52.0 iterators_1.0.14   
[23] tibble_3.2.1        munsell_0.5.1      
[25] rprojroot_2.0.4     pillar_1.10.1      
[27] xfun_0.50           multtest_2.62.0    
[29] cli_3.6.3           withr_3.0.2        
[31] magrittr_2.0.3      foreach_1.5.2      
[33] grid_4.4.2          gridtext_0.1.5     
[35] scrime_1.3.5        lifecycle_1.0.4    
[37] vctrs_0.6.5         evaluate_1.0.3     
[39] glue_1.8.0          farver_2.1.2       
[41] codetools_0.2-20    stats4_4.4.2       
[43] survival_3.7-0      colorspace_2.1-1   
[45] tools_4.4.2         pkgconfig_2.0.3    
> sessioninfo::session_info()
─ Session info ────────────────────────────────
 setting  value
 version  R version 4.4.2 (2024-10-31)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-21
 pandoc   NA

─ Packages ────────────────────────────────────
 !  package      * version date (UTC) lib source
    Biobase        2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics   0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 P  cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
    codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.2)
    colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace     2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    crayon         1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table   * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
    dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate       1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver         2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    foreach        1.5.2   2022-02-02 [1] CRAN (R 4.4.0)
    gam            1.22-5  2024-09-12 [1] CRAN (R 4.4.1)
    generics       0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
    ggplot2      * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
    ggtext       * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue           1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext       0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable         0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    iterators      1.0.14  2022-02-05 [1] CRAN (R 4.4.0)
    knitr          1.49    2024-11-08 [1] CRAN (R 4.4.1)
    labeling       0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lattice        0.22-6  2024-03-20 [1] CRAN (R 4.4.2)
    lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    MASS           7.3-61  2024-06-13 [1] CRAN (R 4.4.2)
    Matrix         1.7-1   2024-10-18 [1] CRAN (R 4.4.2)
    multtest       2.62.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    munsell        0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
    pillar         1.10.1  2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6             2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
    Rcpp           1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
 VP rlang        * 1.1.4   2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    scales         1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
    scrime         1.3.5   2018-12-01 [1] CRAN (R 4.4.0)
    sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
    siggenes       1.80.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    survival       3.7-0   2024-06-05 [1] CRAN (R 4.4.2)
    tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    withr          3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun           0.50    2025-01-07 [1] CRAN (R 4.4.1)
    xml2           1.3.6   2023-12-04 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────