---
title: "06c-init_louvainCluster_RMSD"
output: html_document
date: "2026-02-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
r <- readRDS("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/06c_2-SNNk10_HDG_SVG_2575_louvain1_clustRMSDs.RDS")

anno <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")

names(r) <- anno[match(names(r),clusid),anno]
r <- as.data.table(r,keep.rownames=T)
setnames(r,2,"RMSD")

n <- "snnHARMONYlmbna_HDG_SVG_2575_louv_res1"

pdf(paste0("plots/16_sensitivityanalysis_reproc_wo3donors/Sens1_featureSelection_dimred_harmony_clustering/06c_2_",n,"_rmsds.pdf"),width=10,height=6)
ggplot(r,aes(x=rn,y=RMSD))+
    geom_col()+
    xlab("Cluster")+
    theme(axis.text.x=element_text(size=8,angle=90,hjust=1,vjust=0.5))
dev.off()


```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-3      ggrastr_1.0.2       Polychrome_1.5.4    ggbeeswarm_0.7.3   
[5] gridExtra_2.3       ggplot2_4.0.2       data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] gtable_0.3.6         dplyr_1.2.0          compiler_4.5.2      
 [4] tidyselect_1.2.1     dichromat_2.0-0.1    png_0.1-8           
 [7] scales_1.4.0         yaml_2.3.12          fastmap_1.2.0       
[10] here_1.0.2           R6_2.6.1             labeling_0.4.3      
[13] generics_0.1.4       knitr_1.51           BiocGenerics_0.56.0 
[16] tibble_3.3.1         rprojroot_2.1.1      pillar_1.11.1       
[19] RColorBrewer_1.1-3   rlang_1.1.7          xfun_0.56           
[22] S7_0.2.1             otel_0.2.0           cli_3.6.5           
[25] withr_3.0.2          magrittr_2.0.4       digest_0.6.39       
[28] grid_4.5.2           rstudioapi_0.18.0    beeswarm_0.4.0      
[31] lifecycle_1.0.5      vipor_0.4.7          S4Vectors_0.48.0    
[34] vctrs_0.7.1          scatterplot3d_0.3-44 evaluate_1.0.5      
[37] glue_1.8.0           farver_2.1.2         stats4_4.5.2        
[40] rsconnect_1.7.0      colorspace_2.1-2     rmarkdown_2.30      
[43] tools_4.5.2          pkgconfig_2.0.3      htmltools_0.5.9     
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-17
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package       * version  date (UTC) lib source
 beeswarm        0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 BiocGenerics    0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli             3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 colorout      * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace      2.1-2    2025-09-22 [1] CRAN (R 4.5.0)
 data.table    * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 dichromat       2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest          0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr           1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate        1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver          2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap         1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics        0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 ggbeeswarm    * 0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2       * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr       * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 glue            1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra     * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gtable          0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here            1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools       0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 knitr           1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling        0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lifecycle       1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magrittr        2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 otel            0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 pillar          1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig       2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 png             0.1-8    2022-11-29 [1] CRAN (R 4.5.0)
 Polychrome    * 1.5.4    2025-04-06 [1] CRAN (R 4.5.0)
 R6              2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer    1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 rlang           1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown       2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot       2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect       1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi      0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Vectors       0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7              0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales          1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scatterplot3d   0.3-44   2023-05-05 [1] CRAN (R 4.5.0)
 sessioninfo     1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 tibble          3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect      1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs           0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor           0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr           3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun            0.56     2026-01-18 [1] CRAN (R 4.5.2)
 yaml            2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
