---
title: "06d-init_louvainClusts_on_dimreds"
output: html_document
date: "2026-02-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(ggrastr)
library(Polychrome)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 11), axis.text.y = element_text(size = 10), axis.title.y = element_text(size =11), plot.title = element_text(size = 13,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=9), legend.title = element_text(size=10,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

# only use louvain clusterings (leiden returned 100s to 1000s)
```{r}
dimrs <- readRDS("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/02_02-PCA_harmo_UMAP_matrices_HDG_SVG_2575only.RDS")

## we want the uncorrected dimred (slot 1) and HARMONYd (slot 3) dimred from these two lists.
alldimr <- c(dimrs[c(1,3)])

## to save RAM, subset each of these to the first three PCs
## get the first 3 pcs, only ones we're plotting
alldimr <- lapply(alldimr,function(adr){
   keepn <- colnames(adr)[1:3]
   y <- as.data.table(adr,keep.rownames=T)
   keepn <- c("rn",keepn)
   y <- y[,..keepn]
   # standardize col names for plotting 
   setnames(y,c(2:4),paste0("Dimred_component_",c(1:3)))
   return(y)
})


### get clusterings
louvs <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(louvs,2,"clusid")

## get annotations list, append
anno <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")

louvs <- merge.data.table(louvs,anno[,.(clusid,anno)],by="clusid")

## clean up
rm(anno)
```

## define helper function to extract a single legend - from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

####################
#### Full dimRs ####
####################

## plot cluster assignments onto dims 1-2, 1-3, 2-3
```{r}
lapply(names(alldimr),FUN=function(adrn){
   ## the clusterings corresponding to non-glm PCAs don't contain the string PCA at all, so we need to remove 'PCA' from those list entry names for this to work properly
   keepcols <- c("rn","anno")
   tmplouvs <- copy(louvs)[,..keepcols]
   
   curdr <- copy(alldimr[[adrn]])
   
   clus <- "anno"
   nclus <- length(unique(tmplouvs[,get(clus)]))
   pal <- palette36.colors(n = nclus+2)[c(3:(nclus+2))]
      
      # next line is needed for consistency in color-cluster assignments between libdspatial and the generated legend below
   names(pal) <- unique(tmplouvs$anno)
   
   pltdat <- merge.data.table(curdr,tmplouvs[,.(rn,anno)],by="rn")

   ## make shared legend
   leg.tmp <- ggplot(pltdat,aes(x=Dimred_component_1,y=Dimred_component_2,col=get(clus)))+
      geom_point(size=0.2,stroke = 0)+
      scale_color_manual(values=as.character(pal))+
      guides(color=guide_legend(ncol=2,override.aes=list(size=3)))+
      theme(
         legend.text = element_text(size = 10),
         legend.title = element_text(size = 10, hjust = 0.5),
         legend.spacing.y = unit(0, "cm"),
         legend.spacing.x = unit(0.1, "cm"))
   leg.tmp2 <- extract_legend(leg.tmp)
   dev.off()
      
## now generate the 3 pc pair plots
colcombs <- list(c("Dimred_component_1","Dimred_component_2"),
                 c("Dimred_component_1","Dimred_component_3"),
                 c("Dimred_component_2","Dimred_component_3"))

   p <- lapply(colcombs,FUN=function(cc){
      ggplot(pltdat,aes_string(x=cc[1],y=cc[2],color=clus))+
         scale_color_manual(values=pal)+
         rasterize(geom_point(size=0.4,stroke=0),dpi=300,dev="cairo_png")+
         guides(color="none")+
         ggtitle(paste0(clus,":\non ",adrn))+
         theme(title = element_text(size=9))
   })
   ## append legend into the fourth slot of the list so we can have a 2x2 grid of three plots and a legend.  
   p[[length(p)+1]] <- leg.tmp2
   
   pdf(paste0("plots/16_sensitivityanalysis_reproc_wo3donors/Sens1_featureSelection_dimred_harmony_clustering/06d_2_clusts_on",adrn,"_on_PCA_and_Harmo.pdf"),height=13,width=21)
   #do.call("grid.arrange",list(arrangeGrob(grobs=p,ncol=3),leg.tmp2,ncol=2,widths=c(19,2)))
   do.call("grid.arrange",c(p,ncol=2))
   dev.off()
})


```

################
#### UMAPS ####
################
## we also want to do the UMAPs in the same manner as these two things above.
```{r}
## we want the uncorrected and corrected UMAPs (slots 2 and 4)
allumap <- dimrs[c(2,4)]
allumap <- lapply(allumap,function(x){as.data.table(x,keep.rownames=T)})
```

```{r}
lapply(names(allumap),FUN=function(adrn){
   ## the clusterings corresponding to non-glm PCAs don't contain the string PCA at all, so we need to remove 'PCA' from those list entry names for this to work properly.
   
   cluscols <- "anno"
   keepcols <- c("rn",cluscols)
   
   tmplouvs <- copy(louvs)[,..keepcols]
   
   curdr <- copy(allumap[[adrn]])
   
   clus <- "anno"
   nclus <- length(unique(tmplouvs[,get(clus)]))
   pal <- palette36.colors(n = nclus+2)[c(3:(nclus+2))]
      
   # next line is needed for consistency in color-cluster assignments between libdspatial and the generated legend below
   names(pal) <- unique(tmplouvs$anno)
   
   pltdat <- merge.data.table(curdr,tmplouvs[,.(rn,anno)],by="rn")

   p <- ggplot(pltdat,aes_string(x="UMAP1",y="UMAP2",color=clus))+
         scale_color_manual(values=pal)+
         rasterize(geom_point(size=0.4,stroke=0),dpi=300,dev="cairo_png")+
         ggtitle(paste0(clus,":\non ",adrn))+
         guides(color=guide_legend(override.aes=list(size=2)))+
         theme(title = element_text(size=9))
      
      pdf(paste0("plots/07_2_featureSelection_dimred_harmony_clustering/06d_2_clusters_on_",adrn,".pdf"),height=9,width=12)
      print(p)
      dev.off()
})

```

reprod info:
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.44.0 parallelly_1.46.1   colorout_1.3-3      Polychrome_1.5.4   
[5] ggrastr_1.0.2       gridExtra_2.3       ggplot2_4.0.2       data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] gtable_0.3.6         dplyr_1.2.0          compiler_4.5.2      
 [4] tidyselect_1.2.1     ggbeeswarm_0.7.3     parallel_4.5.2      
 [7] dichromat_2.0-0.1    png_0.1-8            scales_1.4.0        
[10] yaml_2.3.12          fastmap_1.2.0        here_1.0.2          
[13] R6_2.6.1             labeling_0.4.3       generics_0.1.4      
[16] knitr_1.51           tibble_3.3.1         rprojroot_2.1.1     
[19] pillar_1.11.1        RColorBrewer_1.1-3   rlang_1.1.7         
[22] xfun_0.56            S7_0.2.1             otel_0.2.0          
[25] cli_3.6.5            withr_3.0.2          magrittr_2.0.4      
[28] digest_0.6.39        grid_4.5.2           rstudioapi_0.18.0   
[31] beeswarm_0.4.0       lifecycle_1.0.5      vipor_0.4.7         
[34] vctrs_0.7.1          scatterplot3d_0.3-44 evaluate_1.0.5      
[37] glue_1.8.0           farver_2.1.2         codetools_0.2-20    
[40] rsconnect_1.7.0      colorspace_2.1-2     rmarkdown_2.30      
[43] tools_4.5.2          pkgconfig_2.0.3      htmltools_0.5.9     
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-17
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package       * version  date (UTC) lib source
 beeswarm        0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 BiocParallel  * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli             3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools       0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout      * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 colorspace      2.1-2    2025-09-22 [1] CRAN (R 4.5.0)
 data.table    * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 dichromat       2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest          0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr           1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate        1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver          2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap         1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics        0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 ggbeeswarm      0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2       * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrastr       * 1.0.2    2023-06-01 [1] CRAN (R 4.5.0)
 glue            1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra     * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gtable          0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here            1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools       0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 knitr           1.51     2025-12-20 [1] CRAN (R 4.5.2)
 labeling        0.4.3    2023-08-29 [1] CRAN (R 4.5.0)
 lifecycle       1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magrittr        2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 otel            0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly    * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar          1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig       2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 png             0.1-8    2022-11-29 [1] CRAN (R 4.5.0)
 Polychrome    * 1.5.4    2025-04-06 [1] CRAN (R 4.5.0)
 R6              2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer    1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 rlang           1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown       2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot       2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect       1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi      0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S7              0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 scales          1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scatterplot3d   0.3-44   2023-05-05 [1] CRAN (R 4.5.0)
 sessioninfo     1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 tibble          3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect      1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs           0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor           0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 withr           3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun            0.56     2026-01-18 [1] CRAN (R 4.5.2)
 yaml            2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
