---
title: "02-Initial_PCAs"
author: "Bernie Mulvey"
date: "2026-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(scran)
library(scater)
library(harmony)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### load SPE and feature sets; subset to just the one we decided on originally with all samples, 
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

feats <- readRDS("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/01d_2_LCsampsonly_HiDevG_HVG_SVG_andmixed_featsets.RDS")
feats <- feats$HDG_SVG_2575
```

run pca, HARMONY, and umaps thereupon; extract dimred mats and save separately
```{r}
# standard PCA
set.seed(42)
lc3 <- scater::runPCA(lc3,subset_row=feats,name="PCA_HDG_SVG_2575")

# standard UMAPs (which require PCA first)
set.seed(42)
lc3 <- scater::runUMAP(lc3,dimred="PCA_HDG_SVG_2575",name="UMAP_HDG_SVG_2575")
    
    
# harmony runs -- we need to temporarily rename the PCA for the feature set to "PCA" for runharmony to accept it
reducedDimNames(lc3)[which(reducedDimNames(lc3)=="PCA_HDG_SVG_2575")] <- "PCA"
    
# data-based initialization of lambda (development ver of Harmony; requires explicit lambda=NULL)
set.seed(42)
lc3 <- RunHarmony(lc3,lambda=NULL,group.by.vars="sample_id",reduction.save="HARMONYlmbna_HDG_SVG_2575")
    
# restore the feature-set-specific name for the uncorrected PCA
reducedDimNames(lc3)[which(reducedDimNames(lc3)=="PCA")] <- "PCA_HDG_SVG_2575"
    
    
# UMAP of data-based initialization of lambda
set.seed(42)
lc3 <- scater::runUMAP(lc3,dimred="HARMONYlmbna_HDG_SVG_2575",name="HARMONYlmbna_HDG_SVG_2575_UMAP")

## check that we have 4 reduced dims on the object
stopifnot(length(reducedDimNames(lc3))==4)

```

### pull out and save the reducedDims; check prop variance explained by the top 50 PCs
```{r}
rdm <- reducedDims(lc3)
rdm <- as.list(rdm)
# already removed 10x ones:
# rdm <- rdm[grep(names(rdm),pattern="^10x",value=T,invert=T)]

ans <- sum(attr(rdm[[1]],"percentVar"))
# [1] 30.28718

saveRDS(rdm,"processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/02_02-PCA_harmo_UMAP_matrices_HDG_SVG_2575only.RDS")
```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] colorout_1.3-3              harmony_1.2.4              
 [5] Rcpp_1.1.1                  scater_1.38.0              
 [7] scran_1.38.0                scuttle_1.20.0             
 [9] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
[11] SummarizedExperiment_1.40.0 Biobase_2.70.0             
[13] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[15] IRanges_2.44.0              S4Vectors_0.48.0           
[17] BiocGenerics_0.56.0         generics_0.1.4             
[19] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[21] gridExtra_2.3               ggtext_0.1.2               
[23] ggplot2_4.0.2               data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1    viridisLite_0.4.2   dplyr_1.2.0         vipor_0.4.7        
 [5] farver_2.1.2        viridis_0.6.5       S7_0.2.1            fastmap_1.2.0      
 [9] bluster_1.20.0      digest_0.6.39       rsvd_1.0.5          lifecycle_1.0.5    
[13] cluster_2.1.8.1     statmod_1.5.1       magrittr_2.0.4      compiler_4.5.2     
[17] rlang_1.1.7         tools_4.5.2         yaml_2.3.12         igraph_2.2.1       
[21] knitr_1.51          S4Arrays_1.10.1     dqrng_0.4.1         here_1.0.2         
[25] DelayedArray_0.36.0 xml2_1.5.2          RColorBrewer_1.1-3  rsconnect_1.7.0    
[29] abind_1.4-8         withr_3.0.2         grid_4.5.2          beachmat_2.26.0    
[33] edgeR_4.8.2         scales_1.4.0        dichromat_2.0-0.1   cli_3.6.5          
[37] rmarkdown_2.30      otel_0.2.0          metapod_1.18.0      RSpectra_0.16-2    
[41] rstudioapi_0.18.0   rjson_0.2.23        ggbeeswarm_0.7.3    parallel_4.5.2     
[45] XVector_0.50.0      vctrs_0.7.1         Matrix_1.7-4        BiocSingular_1.26.0
[49] BiocNeighbors_2.4.0 ggrepel_0.9.6       irlba_2.3.7         beeswarm_0.4.0     
[53] magick_2.9.0        locfit_1.5-9.12     limma_3.66.0        glue_1.8.0         
[57] codetools_0.2-20    uwot_0.2.4          cowplot_1.2.0       RcppAnnoy_0.0.23   
[61] gtable_0.3.6        ScaledMatrix_1.18.0 tibble_3.3.1        pillar_1.11.1      
[65] htmltools_0.5.9     R6_2.6.1            rprojroot_2.1.1     evaluate_1.0.5     
[69] lattice_0.22-7      RhpcBLASctl_0.23-42 gridtext_0.1.5      SparseArray_1.10.8 
[73] xfun_0.56           pkgconfig_2.0.3    
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-12
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beachmat               2.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bluster                1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 cluster                2.1.8.1  2025-03-12 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 cowplot                1.2.0    2025-07-07 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 dqrng                  0.4.1    2024-05-28 [1] CRAN (R 4.5.0)
 edgeR                  4.8.2    2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 harmony              * 1.2.4    2025-10-10 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 igraph                 2.2.1    2025-10-27 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7    2026-01-30 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 limma                  3.66.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 locfit                 1.5-9.12 2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 metapod                1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                 * 1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 RcppAnnoy              0.0.23   2026-01-12 [1] CRAN (R 4.5.2)
 RhpcBLASctl            0.23-42  2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 RSpectra               0.16-2   2024-07-18 [1] CRAN (R 4.5.0)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scater               * 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 scran                * 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scuttle              * 1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 statmod                1.5.1    2025-10-09 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 uwot                   0.2.4    2025-11-10 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5    2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.5.0)
 withr                  3.0.2    2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 xml2                   1.5.2    2026-01-17 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
