---
title: "07b_2-Compare_louvain_LCs_to_manualLCannots"
author: "Bernie Mulvey"
date: "2026-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(dreamlet) # for pseudobulking
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load counts only object, manual LC annots, louvain annots
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

manlcs <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/07_2_LCsampsonly_manual_expression_defined_LC.txt")

louvs <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(louvs,2,"clusid")
```

## load annots of LC cluster and get pseudobulk expression for each
```{r}
anno <- fread("processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno[,.(clusid,anno)],by="clusid")
```

```{r}
louvs[anno!="LC",anno:="Other"]
    
tmplouv <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]
colLabels(lc3) <- tmplouv$anno
    
lcpb <- aggregateToPseudoBulk(lc3,sample_id = "sample_id",cluster_id = "label")
lcpb <- processAssays(lcpb,formula=~1,assays = "LC",min.cells = 5,min.count=1,min.samples = 20,normalize.method = "TMM")
lcpb <- as.matrix(lcpb$LC$E)
```

## get pseudobulk expression for the two manual annotations
```{r}
manpbs <- lapply(names(manlcs)[2:ncol(manlcs)],function(m){
    curmname <- c("rn",m)
    tmpman <- manlcs[,..curmname]
    setnames(tmpman,2,"manan")
    
    tmpman[manan %in% grep(manan,pattern="_LC",value=T),manan:="LC"]
    tmpman[manan!="LC",manan:="Other"]
    
    tmpman <- DataFrame(tmpman,row.names=tmpman$rn)[colnames(lc3),]
    colLabels(lc3) <- tmpman$manan
    
    manpb <- aggregateToPseudoBulk(lc3,sample_id = "sample_id",cluster_id = "label")
    manpb <- processAssays(manpb,formula=~1,assays = "LC",min.cells = 5,min.count=1,min.samples = 20,normalize.method = "TMM")
    manpb <- as.matrix(manpb$LC$E)
    return(manpb)
})
names(manpbs) <- names(manlcs)[2:ncol(manlcs)]
```

## compare the pseudobulk expression of the manual annotations to the louvain annotations
```{r}
## pair each sample and its genes --> correlate --> repeat for all sample pairs in each manual annotation
ans <- mapply(p=manpbs,n=names(manpbs),SIMPLIFY = FALSE,FUN=function(p,n){
    keepg <- intersect(rownames(lcpb),rownames(p))
    subans <- lapply(colnames(lcpb)[colnames(lcpb) %in% colnames(p)],FUN = function(s){
        ans <- cor.test(x=lcpb[keepg,s],y=p[keepg,s],method="spearman",use="pairwise.complete.obs")
        return(ans$estimate)
    })
    stopifnot(is.list(subans))
    names(subans) <- colnames(lcpb)[colnames(lcpb) %in% colnames(p)]
    retans <- as.data.table(subans)
    retans[,louvclus:="LC_hdgsvg2575"]
    retans[,manclus:=n]
    return(retans)
})
```

convert to tabular format for each louvain cluster
```{r}
ans2 <- rbindlist(ans,fill=T)
ans2 <- melt.data.table(ans2,id.vars = c("louvclus","manclus"))
setnames(ans2,c(3,4),c("sample","pbulk_xpr_spearmancor"))
```

save
```{r}
saveRDS(ans2,"processed-data/07_2_LCsampsonly_featureSelection_dimred_harmony_clustering/07b_2-manualLCdefs_vs_louvainLC.RDS")
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] dreamlet_1.8.0              variancePartition_1.40.1   
 [3] limma_3.66.0                SpatialExperiment_1.20.0   
 [5] SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1       
 [9] Seqinfo_1.0.0               IRanges_2.44.0             
[11] S4Vectors_0.48.0            BiocGenerics_0.56.0        
[13] generics_0.1.4              MatrixGenerics_1.22.0      
[15] matrixStats_1.5.0           BiocParallel_1.44.0        
[17] parallelly_1.46.1           colorout_1.3-3             
[19] gridExtra_2.3               ggtext_0.1.2               
[21] ggplot2_4.0.2               data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] splines_4.5.2             bitops_1.0-9              tibble_3.3.1             
  [4] graph_1.88.1              XML_3.99-0.20             lifecycle_1.0.5          
  [7] Rdpack_2.6.5              mixsqp_0.3-54             edgeR_4.8.2              
 [10] rprojroot_2.1.1           lattice_0.22-7            MASS_7.3-65              
 [13] backports_1.5.0           magrittr_2.0.4            metafor_4.8-0            
 [16] rmarkdown_2.30            yaml_2.3.12               otel_0.2.0               
 [19] DBI_1.2.3                 minqa_1.2.8               RColorBrewer_1.1-3       
 [22] abind_1.4-8               EnvStats_3.1.0            purrr_1.2.1              
 [25] msigdbr_25.1.1            rmeta_3.0                 RCurl_1.98-1.17          
 [28] ggrepel_0.9.6             pbkrtest_0.5.5            irlba_2.3.7              
 [31] annotate_1.88.0           DelayedMatrixStats_1.32.0 codetools_0.2-20         
 [34] DelayedArray_0.36.0       xml2_1.5.2                tidyselect_1.2.1         
 [37] farver_2.1.2              lme4_1.1-38               mathjaxr_2.0-0           
 [40] iterators_1.0.14          tools_4.5.2               progress_1.2.3           
 [43] Rcpp_1.1.1                zenith_1.12.0             glue_1.8.0               
 [46] SparseArray_1.10.8        xfun_0.56                 here_1.0.2               
 [49] dplyr_1.2.0               withr_3.0.2               numDeriv_2016.8-1.1      
 [52] fastmap_1.2.0             boot_1.3-32               caTools_1.18.3           
 [55] digest_0.6.39             truncnorm_1.0-9           R6_2.6.1                 
 [58] scattermore_1.2           gtools_3.9.5              dichromat_2.0-0.1        
 [61] RSQLite_2.4.5             RhpcBLASctl_0.23-42       tidyr_1.3.2              
 [64] corpcor_1.6.10            prettyunits_1.2.0         httr_1.4.7               
 [67] S4Arrays_1.10.1           pkgconfig_2.0.3           gtable_0.3.6             
 [70] rsconnect_1.7.0           blob_1.3.0                S7_0.2.1                 
 [73] XVector_0.50.0            remaCor_0.0.20            htmltools_0.5.9          
 [76] zigg_0.0.2                GSEABase_1.72.0           scales_1.4.0             
 [79] png_0.1-8                 fANCOVA_0.6-1             reformulas_0.4.4         
 [82] ashr_2.2-63               knitr_1.51                rstudioapi_0.18.0        
 [85] reshape2_1.4.5            rjson_0.2.23              curl_7.0.0               
 [88] nlme_3.1-168              nloptr_2.2.1              cachem_1.1.0             
 [91] stringr_1.6.0             KernSmooth_2.23-26        parallel_4.5.2           
 [94] vipor_0.4.7               metadat_1.4-0             AnnotationDbi_1.72.0     
 [97] pillar_1.11.1             grid_4.5.2                vctrs_0.7.1              
[100] gplots_3.3.0              mashr_0.2.79              xtable_1.8-4             
[103] beeswarm_0.4.0            Rgraphviz_2.54.0          evaluate_1.0.5           
[106] KEGGgraph_1.70.0          invgamma_1.2              magick_2.9.0             
[109] mvtnorm_1.3-3             cli_3.6.5                 locfit_1.5-9.12          
[112] compiler_4.5.2            rlang_1.1.7               crayon_1.5.3             
[115] SQUAREM_2021.1            plyr_1.8.9                ggbeeswarm_0.7.3         
[118] stringi_1.8.7             babelgene_22.9            assertthat_0.2.1         
[121] lmerTest_3.2-0            Biostrings_2.78.0         aod_1.3.3                
[124] Matrix_1.7-4              hms_1.1.4                 sparseMatrixStats_1.22.0 
[127] bit64_4.6.0-1             KEGGREST_1.50.0           statmod_1.5.1            
[130] rbibutils_2.4.1           Rfast_2.1.5.2             gridtext_0.1.5           
[133] broom_1.0.12              memoise_2.0.1             RcppParallel_5.1.11-1    
[136] bit_4.6.0                 EnrichmentBrowser_2.40.0 
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-17
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [1] CRAN (R 4.5.0)
 annotate               1.88.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 AnnotationDbi          1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 aod                    1.3.3      2023-12-13 [1] CRAN (R 4.5.0)
 ashr                   2.2-63     2023-08-21 [1] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [1] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [1] CRAN (R 4.5.0)
 backports              1.5.0      2024-05-23 [1] CRAN (R 4.5.0)
 beeswarm               0.4.0      2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [1] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [1] CRAN (R 4.5.0)
 bitops                 1.0-9      2024-10-03 [1] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [1] CRAN (R 4.5.2)
 boot                   1.3-32     2025-08-29 [1] CRAN (R 4.5.2)
 broom                  1.0.12     2026-01-27 [1] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [1] CRAN (R 4.5.0)
 caTools                1.18.3     2024-09-04 [1] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 corpcor                1.6.10     2021-09-16 [1] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [1] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [1] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [1] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 DelayedMatrixStats     1.32.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0      2026-02-03 [1] CRAN (R 4.5.2)
 dreamlet             * 1.8.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 edgeR                  4.8.2      2025-12-23 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 EnrichmentBrowser      2.40.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 EnvStats               3.1.0      2025-04-24 [1] CRAN (R 4.5.0)
 evaluate               1.0.5      2025-08-27 [1] CRAN (R 4.5.0)
 fANCOVA                0.6-1      2020-11-13 [1] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [1] CRAN (R 4.5.0)
 ggtext               * 0.1.2      2022-09-16 [1] CRAN (R 4.5.0)
 glue                   1.8.0      2024-09-30 [1] CRAN (R 4.5.0)
 gplots                 3.3.0      2025-11-30 [1] CRAN (R 4.5.2)
 graph                  1.88.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 gridExtra            * 2.3        2017-09-09 [1] CRAN (R 4.5.0)
 gridtext               0.1.5      2022-09-16 [1] CRAN (R 4.5.0)
 GSEABase               1.72.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 gtable                 0.3.6      2024-10-25 [1] CRAN (R 4.5.0)
 gtools                 3.9.5      2023-11-20 [1] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [1] CRAN (R 4.5.0)
 hms                    1.1.4      2025-10-17 [1] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [1] CRAN (R 4.5.2)
 httr                   1.4.7      2023-08-15 [1] CRAN (R 4.5.0)
 invgamma               1.2        2025-07-02 [1] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [1] CRAN (R 4.5.2)
 iterators              1.0.14     2022-02-05 [1] CRAN (R 4.5.0)
 KEGGgraph              1.70.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KEGGREST               1.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 KernSmooth             2.23-26    2025-01-01 [1] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5      2026-01-08 [1] CRAN (R 4.5.2)
 limma                * 3.66.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 lme4                   1.1-38     2025-12-02 [1] CRAN (R 4.5.2)
 lmerTest               3.2-0      2026-01-13 [1] CRAN (R 4.5.2)
 locfit                 1.5-9.12   2025-03-05 [1] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [1] CRAN (R 4.5.0)
 mashr                  0.2.79     2023-10-18 [1] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [1] CRAN (R 4.5.2)
 mathjaxr               2.0-0      2025-12-01 [1] CRAN (R 4.5.2)
 Matrix                 1.7-4      2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [1] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [1] CRAN (R 4.5.0)
 metadat                1.4-0      2025-02-04 [1] CRAN (R 4.5.0)
 metafor                4.8-0      2025-01-28 [1] CRAN (R 4.5.0)
 minqa                  1.2.8      2024-08-17 [1] CRAN (R 4.5.0)
 mixsqp                 0.3-54     2023-12-20 [1] CRAN (R 4.5.0)
 msigdbr                25.1.1     2025-07-21 [1] CRAN (R 4.5.2)
 mvtnorm                1.3-3      2025-01-10 [1] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [1] CRAN (R 4.5.2)
 nloptr                 2.2.1      2025-03-17 [1] CRAN (R 4.5.0)
 numDeriv               2016.8-1.1 2019-06-06 [1] CRAN (R 4.5.0)
 otel                   0.2.0      2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1     2026-01-08 [1] CRAN (R 4.5.2)
 pbkrtest               0.5.5      2025-07-18 [1] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [1] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [1] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [1] CRAN (R 4.5.0)
 prettyunits            1.2.0      2023-09-24 [1] CRAN (R 4.5.0)
 progress               1.2.3      2023-12-06 [1] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [1] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [1] CRAN (R 4.5.0)
 rbibutils              2.4.1      2026-01-21 [1] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [1] CRAN (R 4.5.2)
 RcppParallel           5.1.11-1   2025-08-27 [1] CRAN (R 4.5.0)
 RCurl                  1.98-1.17  2025-03-22 [1] CRAN (R 4.5.0)
 Rdpack                 2.6.5      2026-01-23 [1] CRAN (R 4.5.2)
 reformulas             0.4.4      2026-02-02 [1] CRAN (R 4.5.2)
 remaCor                0.0.20     2025-08-20 [1] CRAN (R 4.5.0)
 reshape2               1.4.5      2025-11-12 [1] CRAN (R 4.5.0)
 Rfast                  2.1.5.2    2025-10-10 [1] CRAN (R 4.5.0)
 Rgraphviz              2.54.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 RhpcBLASctl            0.23-42    2023-02-11 [1] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7      2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30       2025-09-28 [1] CRAN (R 4.5.0)
 rmeta                  3.0        2018-03-20 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0      2025-12-06 [1] CRAN (R 4.5.2)
 RSQLite                2.4.5      2025-11-30 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1     2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [1] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [1] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [1] CRAN (R 4.5.0)
 Seqinfo              * 1.0.0      2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 sparseMatrixStats      1.22.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 SpatialExperiment    * 1.20.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SQUAREM                2021.1     2021-01-13 [1] CRAN (R 4.5.0)
 statmod                1.5.1      2025-10-09 [1] CRAN (R 4.5.0)
 stringi                1.8.7      2025-03-27 [1] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [1] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1      2026-01-11 [1] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [1] CRAN (R 4.5.0)
 truncnorm              1.0-9      2023-03-20 [1] CRAN (R 4.5.0)
 variancePartition    * 1.40.1     2025-12-11 [1] Bioconductor 3.22 (R 4.5.2)
 vctrs                  0.7.1      2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [1] CRAN (R 4.5.0)
 withr                  3.0.2      2024-10-28 [1] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [1] CRAN (R 4.5.2)
 XML                    3.99-0.20  2025-11-08 [1] CRAN (R 4.5.0)
 xml2                   1.5.2      2026-01-17 [1] CRAN (R 4.5.2)
 xtable                 1.8-4      2019-04-21 [1] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [1] CRAN (R 4.5.2)
 zenith                 1.12.0     2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 zigg                   0.0.2      2025-02-07 [1] CRAN (R 4.5.0)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
