---
title: "08_2-LCsampsonly-Filter all TF-target GSEA results to expressed TFs"
author: "Bernie Mulvey"
date: "2026-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load TFtarg NM GSEA result sets and the SPE (which we need for filtering on expression)
```{r}
lcnmdegsea <- readRDS("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens4_DEanalyses_removedsampsAndFinalNMseg/05_2-LCNMposVsNeg_msigdb_and_tftarg_GSEAs.RDS")
lcnmdegsea <- lcnmdegsea[c("tftarg_signed","tftarg_unsigned")]
names(lcnmdegsea) <- paste0("LCNM_",names(lcnmdegsea))
## add an NM subcluster column for these ones
lcnmdegsea$LCNM_tftarg_signed[dir=="LCNM_pos_up",subanno:="LC_NMpos"]
lcnmdegsea$LCNM_tftarg_signed[dir=="LCNM_neg_up",subanno:="LC_NMneg"]
lcnmdegsea$LCNM_tftarg_unsigned[,anno:="LC"] # because we're looking for enrichment in DE regardless of direction

gseas <- lcnmdegsea
rm(lcnmdegsea)
```

SPE, append log counts on the fly, add cluster labels and NM cluster labels; append FLARE ancestry estimates
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

louvs <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(louvs,2,"clusid")

anno <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno[,.(clusid,anno)],by="clusid")

lcnmsub <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens3_subclustering_and_NMF/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnmsub <- lcnmsub[,.(rn,NMseg_subclus)]
lcnmsub[!(NMseg_subclus %in% c("LC_NMneg","LC_NMpos")),NMseg_subclus:=NA]
louvs <- merge.data.table(louvs,lcnmsub,by="rn")
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$anno
lc3$lcnm <- louvs$NMseg_subclus

## append YRI estimates and make categorical predominant ancestry grouping for corresponding DE analyses
flares <- fread("raw-data/ancestryAndHaplotypes/FLARE/global_ancestry_estimates.txt")
flares[ID=="Br6119",ID:="Br6119(re-dis)"]
stopifnot(nrow(flares[ID %in% lc3$brnum])==30)

tmpcd <- as.data.table(colData(lc3),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,flares[,.(ID,YRI)],by.x="brnum",by.y="ID")
stopifnot(nrow(tmpcd)==ncol(lc3))
tmpcd[,predomAncest:=NA]
tmpcd[,predomAncest:=ifelse(YRI<0.25,yes="CEU",no=predomAncest)]
tmpcd[,predomAncest:=ifelse(YRI>0.75,yes="YRI",no=predomAncest)]
colData(lc3) <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lc3),]

rm(lcnmsub,louvs,anno,tmpcd,flares)
```

# drop duplicated gene ids/symbols; set rownames of spe to gene symbol, since those are the identifiers we're stuck with for TF-targ sets
```{r}
dropg <- as.data.table(rowData(lc3),keep.rownames=T)[,.(gene_id,gene_name)]
dropg1 <- dropg[duplicated(gene_name)|duplicated(gene_id),] 

lc3 <- lc3[!(rownames(lc3)%in%dropg1$gene_id),]
rownames(lc3) <- rowData(lc3)$gene_name
lc3 <- lc3[!(rownames(lc3)%in%dropg1$gene_name),]

#### separate out the TF symbol from the full set name so we can filter the SPE accordingly
gseas <- lapply(gseas,FUN=function(x){
    
    y <- x[!(pathway %in% grep(pathway,pattern="Rummagene",value=T))]
    y[,tf:=apply(.SD,MARGIN = 1,FUN=function(h){strsplit(h,"_",1)[[1]][1]}),.SDcols="pathway"]
    ### rummagene TFs are named differently:
    z <- x[pathway %in% grep(pathway,pattern="Rummagene",value=T)]
    # reverse the output of a strsplit of a strsplit o.0 to get the string immediately preceding "_Rummagene_.."
    z[,tf:=apply(.SD,MARGIN=1,FUN=function(p){
        rev.default(strsplit(
            strsplit(p,"_Rummagene_transcription_factors",-1)[[1]][1],
            "_")[[1]]
        )[1]
    }),.SDcols="pathway"]
    #z[,tf:=gsub(pathway,pattern=".*-.*_(.*)_Rummagene_transcription_factors)",replacement="\\1")]
    fix <- rbind(z,y)
    # make all upper case
    fix[,tf:=toupper(tf)]
    return(fix)
})


# filter SPE to TFs that were tested in GSEA
keepgenes <- unique(unlist(lapply(gseas,FUN=function(x){x$tf})))
keepgenes <- keepgenes[keepgenes %in% rownames(lc3)|keepgenes %in% toupper(rownames(lc3))|toupper(keepgenes) %in% rownames(lc3)|toupper(keepgenes) %in% toupper(rownames(lc3))]
lc3 <- lc3[keepgenes,]

rm(keepgenes,dropg,dropg1)
gc(full=T)
```

### for NM+/- DE only: extract and filter to TFs detected in the cluster at >1% of spots in in one or more of the approp. DE groups {≥ 25% of samples with ≥ 1% of spots with TF detectable AND a mean % spots detected of >=0.75% for the entire group}
```{r}
## logcounts per spot in long format for TFs in the gsea results
lc3.lct <- as.data.table(as.data.frame(as.matrix(logcounts(lc3))),keep.rownames=T)
lc3.lct <- melt.data.table(lc3.lct,id.vars="rn",value.name="logcounts")
setnames(lc3.lct,"rn","tf")

## append (eventually) pertinent metadata
subcd <- as.data.table(colData(lc3),keep.rownames=T)[,.(rn,Sex,APOE,sample_id,label,lcnm,YRI,predomAncest)]
subcd[APOE=="E4/ E4",APOE:="E4hom"]
subcd[APOE=="E2/ E2",APOE:="E2hom"]
subcd[APOE=="E3/ E4",APOE:="E4het"]
subcd[APOE=="E2/ E3",APOE:="E2het"]
subcd[,APOEsimple:=ifelse(APOE %in% c("E4hom","E4het"),yes="E4",no="E2")]
lc3.lct <- merge.data.table(lc3.lct,subcd,by.x="variable",by.y="rn")

rm(subcd,lc3)
gc(full=T)

# make a co-list indicating which DE test (sex, APOE discrete, APOE simple, or categorized YRI) based on the corresponding metadata columns
degroups <- c("lcnm","lcnm") # NMpos vs. neg signed and unsigned

# and another indicating which cluster column to use
cluscol <- c("none","none")

stopifnot(length(degroups)==length(cluscol)&length(gseas)==length(degroups))

lcnms.filt <- mapply(R=gseas,N=names(gseas),G=degroups,Clus=cluscol,SIMPLIFY=FALSE,FUN=function(R,N,G,Clus){
    ## exclude spots not assigned anything in the given grouping (is.na(G))
    ## for example, spots from samples that aren't at either extreme of the YRI proportion estimate and thus not assigned either CEU or YRI as predomAncest
    groupwise <- lapply(unique(lc3.lct[!is.na(G),get(G)]),FUN=function(S){
            curdat <- lc3.lct[get(G)==S]
            
            if (Clus=="none") {
                # in this case, we're grouping by a subclustering column so no need to further subset here
                curdat <- curdat
            } else if (length(grep(N,pattern="nmf",value=T)>0)){
                curclus <- "LC"
                curdat <- curdat[label=="LC"]
            } else {
                curclus <- unique(R$anno)
                curdat <- curdat[get(Clus)==curclus]
            }
            
            curdat[,Nspot:=.N/length(unique(tf)),by="sample_id"] # each spot will have n rows, one row per TF, so we can get the # of spots by dividing
            
            ## filter to spots from current cluster with N≥10 spots (the same limit we used for DE)
            curdat <- curdat[Nspot>10]
            
            # this is a clunky approach, but the TOTAL %age of spots expressing the TF in question (ie logcounts>0) for that sample will fill in for any rows where logcounts are >0. we can then get a table of unique combinations of sample X TF x %age of spots
            curdat[logcounts>0,pctxpr:=.N/Nspot,by=c("tf","sample_id")]
            # set NAs in pctxpr to 0
            curdat[is.na(pctxpr),pctxpr:=0]
            # get each unique pair of tf-sample-pctxpr, where we fetch the max row value of pctxpr for a tf-sample pair (which will be >0 and = the % of spots expressing in all rows where the gene was detected for that sample)
            tffilt <- curdat[,max(pctxpr),by=c("tf","sample_id")]
            # calculate mean % spots with TF detectable across all cluster spots in the group 
            # and calculate number of samples with over 1% of cluster-spots with TF detectable. each tf-sample pair is a unique row now, so take the mean across rows with a given TF value for group-spots-detected mean, or tabulate the number of TF occurrences with the pctxpr filter to get the number of passing rows
            setnames(tffilt,"V1","pctxpr")
            tffilt[,meanxpr:=mean(pctxpr),by="tf"]
            tffilt[pctxpr>0.01,Nxpr:=.N,by="tf"]
            # na rows returned in Nxpr will be meaningless in this result, since we just will subset to max Nxpr analogous to what we did above, resulting in a table of X tfs with their mean + spot %age and number of samples over threshold. set na values to 0 to avoid errors relating to NAs here.
            tffilt[is.na(Nxpr),Nxpr:=0]
            tffilt2 <- tffilt[,max(Nxpr),by="tf"]
            setnames(tffilt2,"V1","Nxpr")
            ## add baack the mean expression to the number of threshold-passing samples by merging parts of the old data ttable to the above result
            tffilt2 <- merge.data.table(tffilt2,unique(tffilt[,.(tf,meanxpr)],by="tf"))
            
            # FINALLY  get the TFs meeting criteria (expressed in at least 0.5% of spots of cluster for at least 25% of samples from the given group) back and filter the result table to them
            minsamp <- floor(0.25*length(unique(curdat$sample_id)))
            keeptfs <- tffilt2[Nxpr>=minsamp&meanxpr>0.0075,tf]
            # return qualifying TFs for this group
            return(keeptfs)
        })
    names(groupwise) <- unique(lc3.lct[,get(G)])
    
    filtres <- copy(R[tf %in% unique(unlist(groupwise))|
                          tf %in% toupper(unique(unlist(groupwise)))|
                          toupper(tf) %in% unique(unlist(groupwise))|
                          toupper(tf) %in% toupper(unique(unlist(groupwise)))])
    
    i<-1
    for (i in c(1:length(groupwise))){
        filtres[,newcol:=ifelse(tf %in% groupwise[[i]],yes="yes",no="no")]
        setnames(filtres,"newcol",paste0("TF_",names(groupwise)[i],"_Expr"))
    }
    return(filtres)
})

## clear out everything and repeat from scratch for sex DE and APO DE TF-gene analyses, where the thresholds will be different
rm(list=grep(ls(),pattern="lcnms.filt",value=T,invert=T))
gc(full=T)
```

# sex and E4vE2 GSEAs filtering:
```{r}
gsea.tf <- readRDS("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens4_DEanalyses_removedsampsAndFinalNMseg/03bii_2-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEA_TFtargs.RDS")
gsea.tf <- gsea.tf[grep(names(gsea.tf),pattern="Sex|E4vE2",value=T)]

gsea.tf.sex <- rbindlist(gsea.tf[grep(names(gsea.tf),pattern="Sex",value=T)],idcol="params")
gsea.tf.sex[,Signed:="Yes"]

gsea.tf.apo <- rbindlist(gsea.tf[grep(names(gsea.tf),pattern="E4vE2",value=T)],idcol="params")
gsea.tf.apo[,Signed:="Yes"]

gsea.tf.us <- readRDS("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens4_DEanalyses_removedsampsAndFinalNMseg/03bii_2-MulticlusterDE_OneInterxnPerModel_dropMixClusts_GSEA_TFtargs_UNsigned.RDS")
gsea.tf.us <- gsea.tf.us[grep(names(gsea.tf),pattern="Sex|E4vE2",value=T)]

gsea.tf.us.sex <- rbindlist(gsea.tf.us[grep(names(gsea.tf.us),pattern="Sex",value=T)],idcol="params")
gsea.tf.us.sex[,Signed:="No"]

gsea.tf.us.apo <- rbindlist(gsea.tf.us[grep(names(gsea.tf.us),pattern="E4vE2",value=T)],idcol="params")
gsea.tf.us.apo[,Signed:="No"]

gsea.sex <- rbind(gsea.tf.sex,gsea.tf.us.sex)
gsea.apo <- rbind(gsea.tf.apo,gsea.tf.us.apo)

gsea.sex[,Domain:=gsub(params,pattern="(.*)_Sex",replacement="\\1")]
gsea.sex[,Comparison:="Male vs Female"]
gsea.sex[,params:=NULL]
gsea.sex[,`Geneset Source`:="Enrichr (as used in Mulvey 2025, Cell Rep.)"]

gsea.apo[,Domain:=gsub(params,pattern="(.*)_E4vE2",replacement="\\1")]
gsea.apo[,Comparison:="E4 vs E2"]
gsea.apo[,params:=NULL]
gsea.apo[,`Geneset Source`:="Enrichr (as used in Mulvey 2025, Cell Rep.)"]


### extract TF symbol from the original enrichr term
gseas <- list(gsea.apo,gsea.sex)
names(gseas) <- c("E4vE2","Sex")

gseas <- lapply(gseas,function(x){
  y <- x[!(pathway %in% grep(pathway,pattern="Rummagene",value=T))]
  y[,TF:=apply(.SD,MARGIN = 1,FUN=function(h){strsplit(h,"_",1)[[1]][1]}),.SDcols="pathway"]
  ### rummagene TFs are named differently:
  z <- x[pathway %in% grep(pathway,pattern="Rummagene",value=T)]
  # reverse the output of a strsplit of a strsplit o.0 to get the string immediately preceding "_Rummagene_.."
  z[,TF:=apply(.SD,MARGIN=1,FUN=function(p){
      rev.default(strsplit(
          strsplit(p,"_Rummagene_transcription_factors",-1)[[1]][1],
          "_")[[1]]
      )[1]
    }),.SDcols="pathway"]
  fix <- rbind(z,y)
  
  # make all upper case
  fix[,TF:=toupper(TF)]
  
  return(fix)
})


# desired order:  "Geneset Source" "Domain"         "Comparison"     "Signed"     
# "TF"        "dir"            "pval"           "padj"          
# "NES"            "ES"             "leadingEdge"    "size"       "log2err"
# "pathway" (full enrichr term)

keepn <- names(gseas[[1]])[c(14,12,13,11,15,10,2,3,6,5,8,7,4,1)]
gseas[[1]] <- gseas[[1]][,..keepn]
gseas[[2]] <- gseas[[2]][,..keepn]

## cleanup
rm(gsea.apo,gsea.sex,gsea.tf,gsea.tf.us,gsea.tf.sex,gsea.tf.apo,gsea.tf.us.apo,gsea.tf.us.sex)
```

## now find expressed TFs in our spatial domains: at least 4% of spots in at least N tissue sections, where N is half the size of the smallest E4orE2, Sex, or Ancestry group; OR, TFs expressed on >2% of spots on average across all samples. (that way, if something wacky like a sex-and-ancestry-specific TF expression phenomenon is happening but robustly enough, we'll still catch it. a test pass for ACHE_SERT_Npep suggests this is very few genes--5 of 12k passing).
```{r}
lc3 <- readRDS("processed-data/06-LCposDonorsOnly_countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

## add logcounts
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

## add domain labels
louvs <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/04_2-LConly_HDG_SVG_2575_Louv1_clustering.txt")
setnames(louvs,2,"clusid")

anno <- fread("processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens1_LCsampsonly_featureSelection_dimred_harmony_clustering/08_2-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno[,.(clusid,anno)],by="clusid")
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

## we want the anno column, which matches the domains as listed in the gsea tables
lc3$domain <- louvs$anno

## subset to those five domains
lc3 <- lc3[,lc3$domain %in% c("ACHE_SERT_LC_WM","ACHE_SERT_Npep","Astro_Vasc_Mural","LC","Oligo_1")]

## find the smallest domain-group for our expressed threshold
domgrp <- as.data.table(colData(lc3))[,.(sample_id,Sex,APOE,Ancestry,domain)]
domgrp[,E4E2:=ifelse(APOE %in% grep(APOE,pattern="E4",value=T),"E4","E2")]
domgrp <- unique(domgrp)
domgrp[,.N,by=c("domain","Sex")][N==min(N),N] # 31
domgrp[,.N,by=c("domain","E4E2")][N==min(N),N] # 24
domgrp[,.N,by=c("domain","Ancestry")][N==min(N),N] # 20
# so, half of the lowest domain-condition group (20) --> require at least 10 tissue sections passing our expression threshold, which will be at least 4% of spots with nonzero counts

# make sure rownames are gene symbols before continuing
rownames(lc3) <- rowData(lc3)$gene_name

# drop any genes that have fewer than 10 counts for efficiency (impossible to have 10 samples with a domain and nonzero expression if total raw counts is under 10)
keepg <- rowSums(counts(lc3))>9 # only drops 6576 genes but we'll do it
lc3 <- lc3[keepg,]

## cleanup for parallel run by domain:
rm(keepg,anno,louvs)
gc(full=T)

### determine genes expressed on a per-domain basis (remember, at least 10 samples with >4% of spots in a domain expressing, OR 2% of spots in domain expressing on average)
doms <- unique(lc3$domain)

setDTthreads(1,restore_after_fork=FALSE)
domxprgs <- mclapply(doms,mc.cores=5,FUN=function(D){
    setDTthreads(1,restore_after_fork=FALSE)
    lct <- as.data.table(as.data.frame(as.matrix(logcounts(lc3[,lc3$domain==D]))),keep.rownames=T)
    setnames(lct,"rn","symb")
    lct <- melt(lct,id.vars="symb",value.name="logcounts")
    ## the key thing happening above is that a row is created for EVERY gene for every spot, so the number of genes we end up with here is a constant we can divide by to get the number of spots in each sample
    
    ### get sample ID from the unique spot identifiers
    lct[,samp:=gsub(variable,pattern="^.*_(.*_.*)$",replacement="\\1")]
    lct[,Nspot:=.N/length(unique(symb)),by="samp"]
            
    # this is a clunky approach, but the TOTAL %age of spots expressing the TF in question (ie logcounts>0) for that sample will fill in for ALL of the rows where logcounts are >0 in this next step. we can then get a table of unique combinations of sample X TF x %age of spots
    lct[logcounts>0,pctxpr:=.N/Nspot,by=c("symb","samp")]
    # set NAs in pctxpr to 0
    lct[is.na(pctxpr),pctxpr:=0]
    # get each unique pair of tf-sample-pctxpr, where we fetch the max row value of pctxpr for a tf-sample pair (which will be >0 and = the % of spots expressing in any row where the gene was detected for that sample)
    lctfilt <- lct[,max(pctxpr),by=c("symb","samp")]
    # calculate mean % spots expressing dataset wide and number of samples with expression over 4%. each tf occurs once per sample, so take the mean across rows with a given TF value for samplewise means, or tabulate the number of TF occurrences with the pctxpr filter to get the number of passing rows
    setnames(lctfilt,"V1","pctxpr")
    lctfilt[,meanxpr:=mean(pctxpr),by="symb"]
    lctfilt[pctxpr>0.04,Nxpr:=.N,by="symb"]
    # na rows returned in Nxpr will be meaningless in this result, since we just will subset to max Nxpr analogous to what we did above, resulting in a table of X tfs with their mean + spot %age and number of samples over threshold. set na values to 0 to avoid errors relating to NAs here.
    lctfilt[is.na(Nxpr),Nxpr:=0]
    lctfilt2 <- lctfilt[,max(Nxpr),by="symb"]
    setnames(lctfilt2,"V1","Nxpr")
    ## add baack the mean expression to the number of threshold-passing samples by merging parts of the old data ttable to the above result
    lctfilt2 <- merge.data.table(lctfilt2,unique(lctfilt[,.(symb,meanxpr)],by="symb"))

    # FINALLY  get the genes meeting criteria (expressed in at least 4% of spots of domain for at least 12 samples, or >2% of spots on average across ALL samples with the domain) back and filter the result table to them
    
    xprg <- lctfilt2[Nxpr>9|meanxpr>0.02,symb]
    # return qualifying gene symbols
    return(xprg)
})
names(domxprgs) <- doms
rm(lc3)
gc(full=T)
```

## now filter our gsea tables to TF-domain pairs that passed the expression threshold above.
```{r}
gseas2 <- lapply(gseas,function(x){
  z <- mapply(d=domxprgs,n=names(domxprgs),SIMPLIFY=FALSE,function(d,n){
    y <- x[Domain==n & TF %in% d]
    return(y)
  })
  ret <- rbindlist(z)
  return(ret)
})

## append the NM DE
gseas2 <- c(gseas2,lcnms.filt)

## fix something unintended in the NM tables
gseas2$LCNM_tftarg_signed[,TF_NA_Expr:=NULL]
gseas2$LCNM_tftarg_unsigned[,TF_NA_Expr:=NULL]


saveRDS(gseas2,"processed-data/16_sensitivityanalysis_reproc_wo3donors/Sens4_DEanalyses_removedsampsAndFinalNMseg/08_2-LCNM_E4vE2_SexDE_TFtarg_GSEAs_filterToExpressed.RDS")
```

```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1          
 [3] default_1.0.0               colorout_1.3-3             
 [5] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [9] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[11] IRanges_2.44.0              S4Vectors_0.48.0           
[13] BiocGenerics_0.56.0         generics_0.1.4             
[15] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[17] data.table_1.18.2.1        

loaded via a namespace (and not attached):
 [1] gtable_0.3.6        beeswarm_0.4.0      rjson_0.2.23        xfun_0.56          
 [5] ggplot2_4.0.2       ggrepel_0.9.6       lattice_0.22-7      vctrs_0.7.1        
 [9] tools_4.5.2         tibble_3.3.1        pkgconfig_2.0.3     BiocNeighbors_2.4.0
[13] Matrix_1.7-4        RColorBrewer_1.1-3  S7_0.2.1            lifecycle_1.0.5    
[17] compiler_4.5.2      farver_2.1.2        codetools_0.2-20    vipor_0.4.7        
[21] htmltools_0.5.9     yaml_2.3.12         pillar_1.11.1       rsconnect_1.7.0    
[25] DelayedArray_0.36.0 viridis_0.6.5       magick_2.9.0        abind_1.4-8        
[29] tidyselect_1.2.1    rsvd_1.0.5          digest_0.6.39       BiocSingular_1.26.0
[33] dplyr_1.2.0         rprojroot_2.1.1     fastmap_1.2.0       grid_4.5.2         
[37] here_1.0.2          cli_3.6.5           SparseArray_1.10.8  scater_1.38.0      
[41] magrittr_2.0.4      S4Arrays_1.10.1     dichromat_2.0-0.1   scales_1.4.0       
[45] ggbeeswarm_0.7.3    rmarkdown_2.30      XVector_0.50.0      otel_0.2.0         
[49] gridExtra_2.3       ScaledMatrix_1.18.0 beachmat_2.26.0     evaluate_1.0.5     
[53] knitr_1.51          viridisLite_0.4.3   irlba_2.3.7         rlang_1.1.7        
[57] Rcpp_1.1.1          glue_1.8.0          scuttle_1.20.0      rstudioapi_0.18.0  
[61] R6_2.6.1           
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-02-22
 rstudio  2026.01.0+392 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 beachmat               2.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.5.0)
 Biobase              * 2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 default              * 1.0.0    2017-08-07 [1] CRAN (R 4.5.0)
 DelayedArray           0.36.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.5.0)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 dplyr                  1.2.0    2026-02-03 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 farver                 2.1.2    2024-05-13 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics             * 0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3    2025-11-29 [1] CRAN (R 4.5.2)
 ggplot2                4.0.2    2026-02-03 [1] CRAN (R 4.5.2)
 ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.5.0)
 glue                   1.8.0    2024-09-30 [1] CRAN (R 4.5.0)
 gridExtra              2.3      2017-09-09 [1] CRAN (R 4.5.0)
 gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.5.0)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges              * 2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7    2026-01-30 [1] CRAN (R 4.5.2)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.46.1   2026-01-08 [1] CRAN (R 4.5.2)
 pillar                 1.11.1   2025-09-17 [1] CRAN (R 4.5.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.5.0)
 R6                     2.6.1    2025-02-15 [1] CRAN (R 4.5.0)
 RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.5.0)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.5.0)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1    2025-11-14 [1] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0    2025-04-24 [1] CRAN (R 4.5.0)
 scater                 1.38.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 scuttle                1.20.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment * 1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 tibble                 3.3.1    2026-01-11 [1] CRAN (R 4.5.2)
 tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.5.0)
 vctrs                  0.7.1    2026-01-23 [1] CRAN (R 4.5.2)
 vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.5.0)
 viridis                0.6.5    2024-01-29 [1] CRAN (R 4.5.0)
 viridisLite            0.4.3    2026-02-04 [1] CRAN (R 4.5.2)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────
