---
title: "05-LMM_spotlogcounts_to_NMmetrics"
author: "Bernie Mulvey"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
require(multcomp)
require(car)
require(broom)
require(lme4)
require(parallel)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

read in spe, subset to NM+ cells in LC cluster
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

## calculate log counts
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

## read in nm subclusterings
lcnm <- fread("processed-data/10_subclustering/04-LCsubclus-by-NM_otherclusts-by-LCorLCNMadjacency.txt")
lcnm <- lcnm[,.(rn,NMseg_subclus)]
lcnm <- DataFrame(lcnm,row.names=lcnm$rn)[colnames(lc3),]

## assign
colLabels(lc3) <- lcnm$NMseg_subclus

# subset to LC NM+, make rownames symbols
lcnm <- lc3[,lc3$label=="LC.1_NMpos"]
rownames(lcnm) <- rowData(lcnm)$gene_name
rm(lc3)
gc()

### get metadata -- only intensity and proportion for NM measures are continuous and appropriate for an LMM. (the N nm bodies and centroid signal count are both counts, and would require GLMMs). we can still inspect top genes from the continuous measures plotting across all four metrics with plots later.

lcnm.cd <- as.data.table(colData(lcnm),keep.rownames=T)
lcnm.cd <- lcnm.cd[,.(rn,Intensity_NM,Prop_NM,Age,Sex,APOE,brnum,sample_id)]
## add column for E4 copies
lcnm.cd[,e4copies:=0]
lcnm.cd[APOE=="E3/ E4",e4copies:=1]
lcnm.cd[APOE=="E4/ E4",e4copies:=2]
lcnm.cd[,APOE:=NULL]

keepg <- rowSums(counts(lcnm))>0

lcnm.lct <- logcounts(lcnm[keepg,])
lcnm.lct <- as.data.frame(as.matrix(lcnm.lct))
```

## clean up for modeling run
```{r}
rm(lcnm,keepg)
gc(full=T)
```

iterate over each gene and metric
```{r}
setDTthreads(1,restore_after_fork=FALSE)

nmcors <- lapply("Intensity_NM","Prop_NM"),function(nm){
    results <- mclapply(c(1:nrow(lcnm.lct)),mc.cores=8,FUN=function(i){
        setDTthreads(1,restore_after_fork=FALSE)
        xpr <- lcnm.lct[i,]
        xpr <- t(xpr)
        xpr <- as.data.table(xpr,keep.rownames=T)
        xpr <- merge.data.table(xpr,lcnm.cd,by="rn")
        
        curg <- colnames(xpr)[2]
        
        setnames(xpr,2,"val")
        setnames(xpr,nm,"NMvar")
        # fit lmm
        lmm <- lmer(val ~ NMvar + Age + Sex + e4copies + (1|brnum) + (1|brnum:sample_id),data=xpr)
        
        an <- as.data.table(Anova(lmm),keep.rownames=T)
        ## for some reason Sex comes out as Sex here but "SexM" for the fixef. so to make sure we keep the sex coefficients, change Sex to SexM here
        an[rn=="Sex",rn:="SexM"]
        
        ## get model coefficients
        fixfx <- as.data.table(fixef(lmm),keep.rownames=T)
        setnames(fixfx,c("rn","coef"))
        fixfx <- fixfx[rn!="(Intercept)"]
        
        ## append model coefficients
        
        an <- merge.data.table(an,fixfx,by="rn")
        
        an[rn=="NMvar",rn:=nm]
        an[,gene:=curg]
        
        return(an)
    })
    names(results) <- rownames(lcnm.lct)
    return(results)
})
names(nmcors) <- c("N_NMbodies","CentroidSignalCount_NM","Intensity_NM","Prop_NM")

### a handful of these return error messages indicating one of the variables is constant. https://stats.stackexchange.com/questions/366784/how-to-get-around-the-glmer-warning-downdated-vtv-is-not-positive-definite

## only retain entries that are tabular results
nmcors2 <- lapply(nmcors,FUN=function(x){
    x[unlist(lapply(x,is.data.table))]
})
nmcors2 <- lapply(nmcors2,FUN=function(x){
    y <- rbindlist(x,idcol="gene_name")
    setnames(y,"Pr(>Chisq)","p")
    return(y)
})

saveRDS(nmcors2,"processed-data/Xenium_prioritization/05-LCNMspots_geneLogcounts_LMMed_by_NMmetric.RDS")
```

## rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.2 (2024-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] parallelly_1.42.0           colorout_1.3-0.2           
 [3] lme4_1.1-35.5               Matrix_1.7-1               
 [5] broom_1.0.7                 car_3.1-3                  
 [7] carData_3.0-5               multcomp_1.4-26            
 [9] TH.data_1.1-2               MASS_7.3-61                
[11] survival_3.7-0              mvtnorm_1.3-2              
[13] SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1
[15] SummarizedExperiment_1.36.0 Biobase_2.66.0             
[17] GenomicRanges_1.58.0        GenomeInfoDb_1.42.1        
[19] IRanges_2.40.1              S4Vectors_0.44.0           
[21] BiocGenerics_0.52.0         MatrixGenerics_1.18.0      
[23] matrixStats_1.5.0           data.table_1.16.4          
[25] rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] gridExtra_2.3           sandwich_3.1-1          magrittr_2.0.3         
 [4] scater_1.34.0           compiler_4.4.2          vctrs_0.6.5            
 [7] pkgconfig_2.0.3         crayon_1.5.3            fastmap_1.2.0          
[10] backports_1.5.0         magick_2.8.5            XVector_0.46.0         
[13] scuttle_1.16.0          rmarkdown_2.29          sessioninfo_1.2.2      
[16] UCSC.utils_1.2.0        nloptr_2.1.1            ggbeeswarm_0.7.2       
[19] purrr_1.0.2             xfun_0.50               zlibbioc_1.52.0        
[22] beachmat_2.22.0         jsonlite_1.8.9          DelayedArray_0.32.0    
[25] BiocParallel_1.40.0     irlba_2.3.5.1           R6_2.5.1               
[28] boot_1.3-31             Rcpp_1.0.14             knitr_1.49             
[31] zoo_1.8-12              splines_4.4.2           tidyselect_1.2.1       
[34] rstudioapi_0.17.1       abind_1.4-8             yaml_2.3.10            
[37] viridis_0.6.5           codetools_0.2-20        lattice_0.22-6         
[40] tibble_3.2.1            evaluate_1.0.3          pillar_1.10.1          
[43] generics_0.1.3          rprojroot_2.0.4         ggplot2_3.5.1          
[46] munsell_0.5.1           scales_1.3.0            minqa_1.2.8            
[49] glue_1.8.0              tools_4.4.2             BiocNeighbors_2.0.1    
[52] ScaledMatrix_1.14.0     grid_4.4.2              tidyr_1.3.1            
[55] colorspace_2.1-1        nlme_3.1-166            GenomeInfoDbData_1.2.13
[58] beeswarm_0.4.0          BiocSingular_1.22.0     vipor_0.4.7            
[61] Formula_1.2-5           cli_3.6.3               rsvd_1.0.5             
[64] viridisLite_0.4.2       S4Arrays_1.6.0          dplyr_1.1.4            
[67] gtable_0.3.6            digest_0.6.37           SparseArray_1.6.0      
[70] ggrepel_0.9.6           rjson_0.2.23            htmltools_0.5.8.1      
[73] lifecycle_1.0.4         httr_1.4.7              here_1.0.1             
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.2 (2024-10-31)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-28
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package              * version  date (UTC) lib source
    abind                  1.4-8    2024-09-12 [1] CRAN (R 4.4.1)
    backports              1.5.0    2024-05-23 [1] CRAN (R 4.4.0)
    beachmat               2.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocNeighbors          2.0.1    2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel           1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    boot                   1.3-31   2024-08-28 [1] CRAN (R 4.4.2)
    broom                * 1.0.7    2024-09-26 [1] CRAN (R 4.4.1)
    car                  * 3.1-3    2024-09-27 [1] CRAN (R 4.4.1)
    carData              * 3.0-5    2022-01-06 [1] CRAN (R 4.4.0)
 P  cli                    3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
    codetools              0.2-20   2024-03-31 [1] CRAN (R 4.4.2)
    colorout             * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1    2024-07-26 [1] CRAN (R 4.4.0)
    crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
    DelayedArray           0.32.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    digest                 0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
    fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
    Formula                1.2-5    2023-02-24 [1] CRAN (R 4.4.0)
    generics               0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13   2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2    2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2                3.5.1    2024-04-23 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra              2.3      2017-09-09 [1] CRAN (R 4.4.0)
    gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7    2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1   2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1  2022-10-03 [1] CRAN (R 4.4.0)
    jsonlite               1.8.9    2024-09-20 [1] CRAN (R 4.4.1)
    knitr                  1.49     2024-11-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6   2024-03-20 [1] CRAN (R 4.4.2)
    lifecycle              1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
    lme4                 * 1.1-35.5 2024-07-03 [1] CRAN (R 4.4.1)
    magick                 2.8.5    2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
    MASS                 * 7.3-61   2024-06-13 [1] CRAN (R 4.4.2)
    Matrix               * 1.7-1    2024-10-18 [1] CRAN (R 4.4.2)
    MatrixGenerics       * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.4.1)
    minqa                  1.2.8    2024-08-17 [1] CRAN (R 4.4.0)
    multcomp             * 1.4-26   2024-07-18 [1] CRAN (R 4.4.0)
    munsell                0.5.1    2024-04-01 [1] CRAN (R 4.4.0)
    mvtnorm              * 1.3-2    2024-11-04 [1] CRAN (R 4.4.1)
    nlme                   3.1-166  2024-08-14 [1] CRAN (R 4.4.2)
    nloptr                 2.1.1    2024-06-25 [1] CRAN (R 4.4.0)
    parallelly           * 1.42.0   2025-01-30 [1] CRAN (R 4.4.1)
    pillar                 1.10.1   2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
    purrr                  1.0.2    2023-08-10 [1] CRAN (R 4.4.0)
    R6                     2.5.1    2021-08-19 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14   2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4    2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown              2.29     2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.4.0)
    S4Arrays               1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sandwich               3.1-1    2024-09-15 [1] CRAN (R 4.4.1)
    ScaledMatrix           1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0    2023-11-28 [1] CRAN (R 4.4.0)
    scater                 1.34.0   2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
    scuttle                1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
    SingleCellExperiment * 1.28.1   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SummarizedExperiment * 1.36.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    survival             * 3.7-0    2024-06-05 [1] CRAN (R 4.4.2)
    TH.data              * 1.1-2    2023-04-17 [1] CRAN (R 4.4.0)
    tibble                 3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1    2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5    2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.4.0)
    xfun                   0.50     2025-01-07 [1] CRAN (R 4.4.1)
    XVector                0.46.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10   2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    zoo                    1.8-12   2023-04-13 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
