---
title: "06b-Tabulate_xpr_LuskinPeriLC-snseq_subclus_mks"
author: "Bernie Mulvey"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##
```

read in spe
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

## calculate log counts
lc3 <- scater::computeLibraryFactors(lc3)
lc3 <- scater::logNormCounts(lc3)

## read in clustering and annots
louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

anno <- fread("processed-data/08_validitycheck_25hdg75svg_louv1/10-25hdg75svg_louv1_annots.txt")
louvs <- merge.data.table(louvs,anno,by="clusid")
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

## assign
colLabels(lc3) <- louvs$anno

## rownames to symbols
rownames(lc3) <- rowData(lc3)$gene_name
```

fetch table of mouse LC/peri-LC subcluster markers noted for examination
```{r}
mslc <- fread("processed-data/Xenium_prioritization/06a-luskin_snseq_LC_GABA_subclust_mks_to_check.txt")

lc3 <- lc3[mslc$gene_name,]
```

## tabulate prop of each cluster with counts > 0
```{r}
xprprops <- lapply(rownames(lc3),FUN=function(g){
    ans <- lapply(unique(lc3$label),FUN=function(a){
        subans <- sum(counts(lc3[g,lc3$label==a])>0)/ncol(lc3[g,lc3$label==a])
        return(subans)
    })
    names(ans) <- unique(lc3$label)
    return(ans)
})
names(xprprops) <- rownames(lc3)

anstab <- lapply(xprprops,FUN=function(x){
    y <- cbind(as.numeric(x),names(x))
    y <- as.data.table(y)
    setnames(y,c("prop_nonzero","cluster"))
    return(y)
})
anstab <- rbindlist(anstab,idcol="gene")

# save
fwrite(anstab,"processed-data/Xenium_prioritization/06b-ProportionSpots_per_clus_w_LuskinSubclusMks.txt",sep='\t',quote=F,row.names=F)
```

## rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.2 (2024-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] colorout_1.3-0.2            SpatialExperiment_1.16.0   
 [3] SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
 [5] Biobase_2.66.0              GenomicRanges_1.58.0       
 [7] GenomeInfoDb_1.42.1         IRanges_2.40.1             
 [9] S4Vectors_0.44.0            BiocGenerics_0.52.0        
[11] MatrixGenerics_1.18.0       matrixStats_1.5.0          
[13] data.table_1.16.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] beeswarm_0.4.0          gtable_0.3.6            rjson_0.2.23           
 [4] xfun_0.50               ggplot2_3.5.1           ggrepel_0.9.6          
 [7] lattice_0.22-6          generics_0.1.3          vctrs_0.6.5            
[10] tools_4.4.2             parallel_4.4.2          tibble_3.2.1           
[13] pkgconfig_2.0.3         BiocNeighbors_2.0.1     Matrix_1.7-1           
[16] lifecycle_1.0.4         GenomeInfoDbData_1.2.13 compiler_4.4.2         
[19] munsell_0.5.1           codetools_0.2-20        vipor_0.4.7            
[22] htmltools_0.5.8.1       yaml_2.3.10             pillar_1.10.1          
[25] crayon_1.5.3            BiocParallel_1.40.0     DelayedArray_0.32.0    
[28] viridis_0.6.5           magick_2.8.5            abind_1.4-8            
[31] tidyselect_1.2.1        rsvd_1.0.5              digest_0.6.37          
[34] BiocSingular_1.22.0     dplyr_1.1.4             rprojroot_2.0.4        
[37] fastmap_1.2.0           grid_4.4.2              here_1.0.1             
[40] colorspace_2.1-1        cli_3.6.3               SparseArray_1.6.0      
[43] scater_1.34.0           magrittr_2.0.3          S4Arrays_1.6.0         
[46] UCSC.utils_1.2.0        scales_1.3.0            ggbeeswarm_0.7.2       
[49] rmarkdown_2.29          XVector_0.46.0          httr_1.4.7             
[52] gridExtra_2.3           ScaledMatrix_1.14.0     beachmat_2.22.0        
[55] evaluate_1.0.3          knitr_1.49              viridisLite_0.4.2      
[58] irlba_2.3.5.1           Rcpp_1.0.14             glue_1.8.0             
[61] scuttle_1.16.0          rstudioapi_0.17.1       jsonlite_1.8.9         
[64] R6_2.5.1                zlibbioc_1.52.0        
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.2 (2024-10-31)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-03-02
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package              * version date (UTC) lib source
    abind                  1.4-8   2024-09-12 [1] CRAN (R 4.4.1)
    beachmat               2.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocNeighbors          2.0.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel           1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 P  cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
    codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.2)
    colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.16.4  2024-12-06 [1] CRAN (R 4.4.1)
    DelayedArray           0.32.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    digest                 0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate               1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1  2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13  2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2                3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6   2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra              2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gtable                 0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1  2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
    jsonlite               1.8.9   2024-09-20 [1] CRAN (R 4.4.1)
    knitr                  1.49    2024-11-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.2)
    lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magick                 2.8.5   2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    Matrix                 1.7-1   2024-10-18 [1] CRAN (R 4.4.2)
    MatrixGenerics       * 1.18.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0   2025-01-07 [1] CRAN (R 4.4.1)
    munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
    pillar                 1.10.1  2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23  2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4   2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rmarkdown              2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi             0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
    S4Arrays               1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ScaledMatrix           1.14.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
    scater                 1.34.0  2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
    scuttle                1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
    SingleCellExperiment * 1.28.1  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SummarizedExperiment * 1.36.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
    xfun                   0.50    2025-01-07 [1] CRAN (R 4.4.1)
    XVector                0.46.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10  2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0  2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────
