---
title: "01-get_featSets_for_subclustering"
author: "Bernie Mulvey"
date: "2025-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(scater)
library(scran)
library(scry) # get null residuals
# library(SpatialFeatureExperiment)

## rstudio (but not positron) GUI tweaks
# require(colorout)
# ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

# load sectionwise, filtered, lognormed SPE; append the initial cluster labels (25HDG-75SVG louvain res 1)
```{r}
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")

louvs <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/04-Initial_louvLeid_res0.5-1-2_clusterings.RDS")
louvs <- louvs <- louvs$HARMONYlmbna_HDG_SVG_2575[,.(rn,snnHARMONYlmbna_HDG_SVG_2575_louv_res1)]
setnames(louvs,2,"clusid")

autoan <- readRDS("processed-data/07_featureSelection_dimred_harmony_clustering/05b-init_clusterings_autoannoLCrapheAndGABAergic.RDS")
autoan <- autoan$snnHARMONYlmbna_HDG_SVG_2575_louv_res1

louvs <- merge.data.table(louvs,autoan,by="clusid",all.x=T)
louvs <- DataFrame(louvs,row.names=louvs$rn)[colnames(lc3),]

colLabels(lc3) <- louvs$autoanno
```

## make an SPE for each cluster individually
```{r}
clussub <- lapply(unique(colLabels(lc3)),function(x){ 
    sub <- lc3[,colLabels(lc3)==x]
    
    # drop genes that have 0 counts for the cluster in question
    sub <- sub[rowSums(counts(sub))>0,]

    # calculate the lognorm counts
    sub <- computeLibraryFactors(sub)
    sub <- logNormCounts(sub)
    return(sub)
})
names(clussub) <- unique(colLabels(lc3))

## we don't need the full spe now
rm(lc3,louvs,autoan)
gc(full=T)
```

get HVGs per cluster
```{r}
subhvgs <- lapply(clussub,function(x){
    lv <- modelGeneVar(x)
    
    ## get all variable genes for now
    hvg <- scran::getTopHVGs(lv,prop=1.0)

    hvgtab <- cbind(hvg,seq(1,length(hvg),1))
    colnames(hvgtab) <- c("gene_id","HVG_rank")
    hvgtab <- merge.data.table(as.data.table(hvgtab),as.data.table(rowData(x))[,.(gene_id,gene_name)],by="gene_id")
    hvgtab[,HVG_rank:=as.numeric(HVG_rank)]
    return(hvgtab)
})

## get a 5%ile, 10%ile, 20%ile set of HVGs for each subcluster
subhvgsets <- lapply(subhvgs,function(x){
    x[,HVG_rank:=as.numeric(HVG_rank)]
    hvg5 <- x[HVG_rank<=nrow(x)*0.05,]
    hvg10 <- x[HVG_rank<=nrow(x)*0.10,]
    hvg20 <- x[HVG_rank<=nrow(x)*0.20,]
    retlist <- list(hvg5,hvg10,hvg20)
    names(retlist) <- c("hvg5","hvg10","hvg20")
    return(retlist)
})
```

## save feature sets. (no need to save the spe by cluster list -- very fast to make, but big)
```{r}
saveRDS(subhvgsets,"processed-data/10_subclustering/01-HVG5-10-20_sets_per_cluster.RDS")
```

## repro inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.2 (2024-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0        
 [2] parallelly_1.42.0          
 [3] colorout_1.3-0.2           
 [4] scry_1.18.0                
 [5] scran_1.34.0               
 [6] scater_1.34.0              
 [7] scuttle_1.16.0             
 [8] gridExtra_2.3              
 [9] ggtext_0.1.2               
[10] ggplot2_3.5.1              
[11] data.table_1.16.4          
[12] SpatialExperiment_1.16.0   
[13] SingleCellExperiment_1.28.1
[14] SummarizedExperiment_1.36.0
[15] Biobase_2.66.0             
[16] GenomicRanges_1.58.0       
[17] GenomeInfoDb_1.42.1        
[18] IRanges_2.40.1             
[19] S4Vectors_0.44.0           
[20] BiocGenerics_0.52.0        
[21] MatrixGenerics_1.18.0      
[22] matrixStats_1.5.0          
[23] rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1       
 [2] viridisLite_0.4.2      
 [3] dplyr_1.1.4            
 [4] vipor_0.4.7            
 [5] viridis_0.6.5          
 [6] bluster_1.16.0         
 [7] rsvd_1.0.5             
 [8] lifecycle_1.0.4        
 [9] cluster_2.1.6          
[10] statmod_1.5.0          
[11] magrittr_2.0.3         
[12] compiler_4.4.2         
[13] tools_4.4.2            
[14] igraph_2.1.4           
[15] knitr_1.49             
[16] S4Arrays_1.6.0         
[17] dqrng_0.4.1            
[18] here_1.0.1             
[19] DelayedArray_0.32.0    
[20] xml2_1.3.6             
[21] abind_1.4-8            
[22] withr_3.0.2            
[23] grid_4.4.2             
[24] beachmat_2.22.0        
[25] colorspace_2.1-1       
[26] edgeR_4.4.1            
[27] scales_1.3.0           
[28] cli_3.6.3              
[29] crayon_1.5.3           
[30] generics_0.1.3         
[31] metapod_1.14.0         
[32] httr_1.4.7             
[33] rjson_0.2.23           
[34] ggbeeswarm_0.7.2       
[35] zlibbioc_1.52.0        
[36] parallel_4.4.2         
[37] XVector_0.46.0         
[38] vctrs_0.6.5            
[39] Matrix_1.7-1           
[40] jsonlite_1.8.9         
[41] BiocSingular_1.22.0    
[42] BiocNeighbors_2.0.1    
[43] ggrepel_0.9.6          
[44] irlba_2.3.5.1          
[45] beeswarm_0.4.0         
[46] magick_2.8.5           
[47] locfit_1.5-9.10        
[48] limma_3.62.1           
[49] glue_1.8.0             
[50] codetools_0.2-20       
[51] gtable_0.3.6           
[52] UCSC.utils_1.2.0       
[53] ScaledMatrix_1.14.0    
[54] munsell_0.5.1          
[55] tibble_3.2.1           
[56] pillar_1.10.1          
[57] GenomeInfoDbData_1.2.13
[58] R6_2.5.1               
[59] rprojroot_2.0.4        
[60] evaluate_1.0.3         
[61] lattice_0.22-6         
[62] gridtext_0.1.5         
[63] Rcpp_1.0.14            
[64] SparseArray_1.6.0      
[65] xfun_0.50              
[66] pkgconfig_2.0.3        
> sessioninfo::session_info()
─ Session info ────────────────────────────────
 setting  value
 version  R version 4.4.2 (2024-10-31)
 os       macOS Sonoma 14.7.1
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-02-06
 pandoc   NA

─ Packages ────────────────────────────────────
 !  package              * version  date (UTC) lib source
    abind                  1.4-8    2024-09-12 [1] CRAN (R 4.4.1)
    beachmat               2.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocNeighbors          2.0.1    2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel         * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    bluster                1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
 P  cli                    3.6.3    2024-06-21 [2] CRAN (R 4.4.0)
    cluster                2.1.6    2023-12-01 [1] CRAN (R 4.4.2)
    codetools              0.2-20   2024-03-31 [1] CRAN (R 4.4.2)
    colorout             * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
    colorspace             2.1-1    2024-07-26 [1] CRAN (R 4.4.0)
    crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.4.0)
    data.table           * 1.16.4   2024-12-06 [1] CRAN (R 4.4.1)
    DelayedArray           0.32.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dplyr                  1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
    dqrng                  0.4.1    2024-05-28 [1] CRAN (R 4.4.0)
    edgeR                  4.4.1    2024-12-02 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate               1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
    generics               0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
    GenomeInfoDb         * 1.42.1   2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13   2024-12-12 [1] Bioconductor
    GenomicRanges        * 1.58.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2    2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2              * 3.5.1    2024-04-23 [1] CRAN (R 4.4.0)
    ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.4.1)
    ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.4.0)
    glue                   1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.4.0)
    gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.4.0)
    gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
    httr                   1.4.7    2023-08-15 [1] CRAN (R 4.4.0)
    igraph                 2.1.4    2025-01-23 [1] CRAN (R 4.4.1)
    IRanges              * 2.40.1   2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1  2022-10-03 [1] CRAN (R 4.4.0)
    jsonlite               1.8.9    2024-09-20 [1] CRAN (R 4.4.1)
    knitr                  1.49     2024-11-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6   2024-03-20 [1] CRAN (R 4.4.2)
    lifecycle              1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
    limma                  3.62.1   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    locfit                 1.5-9.10 2024-06-24 [1] CRAN (R 4.4.0)
    magick                 2.8.5    2024-09-20 [1] CRAN (R 4.4.1)
    magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
    Matrix                 1.7-1    2024-10-18 [1] CRAN (R 4.4.2)
    MatrixGenerics       * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.4.1)
    metapod                1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    munsell                0.5.1    2024-04-01 [1] CRAN (R 4.4.0)
    parallelly           * 1.42.0   2025-01-30 [1] CRAN (R 4.4.1)
    pillar                 1.10.1   2025-01-07 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
    R6                     2.5.1    2021-08-19 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14   2025-01-12 [1] CRAN (R 4.4.1)
    rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.4    2025-01-17 [2] CRAN (R 4.4.1) (on disk 1.1.5)
    rprojroot              2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
    rsvd                   1.0.5    2021-04-16 [1] CRAN (R 4.4.0)
    S4Arrays               1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ScaledMatrix           1.14.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.3.0    2023-11-28 [1] CRAN (R 4.4.0)
    scater               * 1.34.0   2024-10-29 [1] Bioconductor 3.20 (R 4.4.1)
    scran                * 1.34.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scry                 * 1.18.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scuttle              * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.2    2021-12-06 [1] CRAN (R 4.4.0)
    SingleCellExperiment * 1.28.1   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    SparseArray            1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    SpatialExperiment    * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    statmod                1.5.0    2023-01-06 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5    2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.4.0)
    withr                  3.0.2    2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.50     2025-01-07 [1] CRAN (R 4.4.1)
    xml2                   1.3.6    2023-12-04 [1] CRAN (R 4.4.0)
    XVector                0.46.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    zlibbioc               1.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────