---
title: "09cV2-Enrichr_LCnmfFactorSpecificGenes"
author: "Bernie Mulvey"
date: "2026-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(enrichR)
library(SpatialExperiment)
library(biomaRt)
library(stringr)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```



we'll run all of the factors with specific genes from both unfilterd and pre-filtered NMF runs and see what there is to see. 

```{r}
unfilt <- fread("processed-data/10_subclustering/07thru10_ver2_53factorNMF/09aV2-LCNMF_factors_and_factorSpecGenes_min25specGenes.txt")

filt <- fread("processed-data/10_subclustering/07thru10_ver2_53factorNMF/09bV2-LCNMFLogCt18Filt_factors_and_factorSpecGenes_min25specGenes.txt")

### we actually want the full NMF matrix and we need the SPE rowData to convert to entrez symbols
lc3 <- readRDS("processed-data/06-countsOnly_noImgData_QCed_SPE_split_to_tissSections.RDS")
rddt <- as.data.table(rowData(lc3))
rm(lc3)

biomart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL",dataset="hsapiens_gene_ensembl")
lut <- biomaRt::getBM(attributes=c("ensembl_gene_id",
                                   "entrezgene_id",
                                   "entrezgene_accession",
                                   "entrezgene_description"),
                      mart=biomart,
                      filter="ensembl_gene_id",
                      values=unique(rddt$gene_id)) # uh ok "accession" = symbol

lut <- as.data.table(lut)
rddt <- merge.data.table(rddt[,.(gene_id,chr)],lut[,.(ensembl_gene_id,entrezgene_accession)],by.x="gene_id",by.y="ensembl_gene_id")
rm(biomart)

## append these to the factor spec genes
unfilt <- merge.data.table(unfilt,rddt[,.(gene_id,entrezgene_accession)],by="gene_id")

filt <- merge.data.table(filt,rddt[,.(gene_id,entrezgene_accession)],by="gene_id")
### this appears to have added some rows, so drop duplicated genes
dupg <- unfilt[,.N,by=c("factor","gene_id")][N>1,gene_id]
unfilt <- unfilt[!(gene_id %in% dupg)]

dupg <- filt[,.N,by=c("factor","gene_id")][N>1,gene_id]
filt <- filt[!(gene_id %in% dupg)]

## temporarily load the NMF matrices to get the background genes for filt and unfilt
nmf <- readRDS("processed-data/10_subclustering/07thru10_ver2_53factorNMF/07-lcnmf53_tol1e-9.RDS")
nmfbg <- rddt[gene_id %in% rownames(nmf$w)&entrezgene_accession!="",unique(entrezgene_accession)]
rm(nmf)

nmffilt <- readRDS("processed-data/10_subclustering/07thru10_ver2_53factorNMF/07b-lcnmf53_tol1e-9_rowSumLogctFilt18.RDS")
filtbg <- rddt[gene_id %in% rownames(nmffilt$w)&entrezgene_accession!="",unique(entrezgene_accession)]
rm(nmffilt)

### clean up
rm(lut,dupg,rddt)
```

### enrichr setup
## rummagenes don't work here unfortunately.
```{r}
checkdbs <- enrichR::listEnrichrDbs()
enrichr_dbs <- grep(checkdbs$libraryName,pattern="Aging|Allen|ARCHS4|Azimuth|BioPlex|CellMarker|ChEA_2022|CORUM|DepMap_CRISPR_GeneDependency_CellLines_2023|Descartes|DGIdb_Drug_Targets_2024|Disease_Perturbations|Disease_Signatures|Drug_Perturbations|DrugMatrix|Elsevier_Pathway|ENCODE_Histone_Modifications_2015|ENCODE_TF_ChIP-seq_2015|Roadmap|FANTOM6|Gene_Perturbations|GO_Biological_Process_2025|GO_Cellular_Component_2025|GO_Molecular_Function_2025|HMS_LINCS|HuBMAP|huMAP|IDG_Drug_Targs_2022|Jensen_DISEASES_Curated_2025|KEGG_2021_Human|Kinase_Perturbations|KOMP2|L1000|Ligand_Perturbations|lncHUB|MAGNET|MGI_Mammalian_Phenotype_Level_4_2024|NIBR_DRUGseq_2025|PerturbAtlas|PFOCR_Pathways_2023|PPI_Hub_Proteins|Proteomics|Reactome_Pathways_2024|SILAC|SubCell|SynGO_2024|Tabula_Muris|Tabula_Sapiens|TF_Perturbations|TF-LOF_Expression|TG_GATES_2020|The_Kinase_Library_2024|Tissue_Protein_Expression|Transcription_Factor_PPIs|TRANSFAC_and_JASPAR|TRRUST|WikiPathways_2024_Human|^DSigDB",value=T)

## filter out the ARCHS4 predictions re: rare disease, NIH PIs--interesting but not useful here
enrichr_dbs <- grep(enrichr_dbs,pattern="AutoRIF|GeneRIF|from_GEO",value=T,invert=T)

rm(checkdbs)
```

Interestingly, enrichr allows for the gene sets to have a "membership value of 0-1" in the second column. we COULD test this later on by adjusting the w matrix of the factor specific-genes to range 0-1 by subsetting to a table of those genes and that factor's values, then dt[,LCnmfX_scaled:=(LCnmfX-min(LCnmfX))/(max(LCnmfX)-min(LCnmfX))]

but for now, let's just feed the genes in as a straight-up vector (against the whole set of genes fed into the NMF run.)
```{r}
## need to run serially, otherwise server will start blocking requests and/or gene sets requested will get mixed up. plus other unexpected errors
unfilt.res <- list()
i <- 1
for (i in c(1:length(unique(unfilt$factor)))){
    f <- unique(unfilt$factor)[i]
    tmpunfilt <- copy(unfilt[factor==f])
    
    cur_enrichr_dbs <- enrichr_dbs
    
    ## drop any dbs that throw weird factor-geneset-specific errors
    if(f=="LCnmf9"|f=="LCnmf10"|f=="LCnmf2"|f=="LCnmf1"|f=="LCnmf7"|f=="LCnmf16"|f=="LCnmf17"){
        cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="PerturbAtlas_MouseGenePerturbationSigs"]
    } else if (f=="LCnmf31"){
      cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="Elsevier_Pathway_Collection"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="CellMarker_2024"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="PerturbAtlas_MouseGenePerturbationSigs"]
    } else if (f=="LCnmf53"){
        cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="CellMarker_2024"]
    } else if (f=="LCnmf3"){
      cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="HMS_LINCS_KinomeScan"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="Tabula_Sapiens"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="PerturbAtlas_MouseGenePerturbationSigs"]
    } else if (f=="LCnmf8"){
      cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="HMS_LINCS_KinomeScan"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="PerturbAtlas_MouseGenePerturbationSigs"]
    } else if (f=="LCnmf5"){
      cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="HMS_LINCS_KinomeScan"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="Tabula_Sapiens"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="SynGO_2024"]
    } else if (f=="LCnmf22"){
      cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="HMS_LINCS_KinomeScan"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="CellMarker_2024"]
    } else {
      cur_enrichr_dbs <- enrichr_dbs
    }
    
    factlist <- enrichr(genes=tmpunfilt[factor==f,entrezgene_accession],
                        background = nmfbg,
                        databases=cur_enrichr_dbs)
    
  
    
    factlist.dt <- rbindlist(factlist,idcol="Enrichr_database")
    factlist.dt[,LCnmf_unfilt_factor:=f]
    
    unfilt.res[[i]] <- factlist.dt
    names(unfilt.res)[i] <- f
    
    rm(f,factlist.dt,factlist,tmpunfilt,cur_enrichr_dbs)
    gc(full=T)
}
rm(i)

#concat and save
unfilt.res2 <- rbindlist(unfilt.res)
# tabulate how many genes were hit in the set
unfilt.res2[,N_genes:=apply(.SD,MARGIN=1,simplify=FALSE,FUN=function(x){
  max(c(1,length(str_split_fixed(x,";",n=Inf)))) # if no separator, 1 gene matched
}),.SDcols="Genes"]
unfilt.res2[,N_genes:=unlist(N_genes)] ## fix weirdness with apply within d.t.

# get rid of deprecated columns
unfilt.res2[,Old.P.value:=NULL]
unfilt.res2[,Old.Adjusted.P.value:=NULL]

saveRDS(unfilt.res2,"processed-data/10_subclustering/07thru10_ver2_53factorNMF/09cV2-LCnmfUNfilt_factorSpecGenes_enrichr_results.RDS")
###
```

cleanup
```{r}
rm(unfilt,unfilt.res,unfilt.res2,nmfbg)
gc(full=T)
```

### pre-filtered-genes NMF factors
```{r}
filt.res <- list()
i<-1
for (i in c(1:length(unique(filt$factor)))){
    f <- unique(filt$factor)[i]
    tmpfilt <- copy(filt[factor==f])
    
    cur_enrichr_dbs <- enrichr_dbs

    if(f=="LCnmf1"|f=="LCnmf3"|f=="LCnmf10"|f=="LCnmf4"|f=="LCnmf12"|f=="LCnmf2"|f=="LCnmf13"|f=="LCnmf14"|f=="LCnmf6"|f=="LCnmf20"){
        cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="PerturbAtlas_MouseGenePerturbationSigs"]
    
    } else if (f=="LCnmf30"){
        cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="Elsevier_Pathway_Collection"]
        cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="PerturbAtlas_MouseGenePerturbationSigs"]
        
    } else if (f=="LCnmf5"){
      cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="HMS_LINCS_KinomeScan"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="Tabula_Sapiens"]
      cur_enrichr_dbs <- cur_enrichr_dbs[cur_enrichr_dbs!="DGIdb_Drug_Targets_2024"]
      
    } else if (f=="LCnmf24"){
      cur_enrichr_dbs <- enrichr_dbs[enrichr_dbs!="HMS_LINCS_KinomeScan"]
      
    } else {
        cur_enrichr_dbs <- enrichr_dbs
    }
    # 
    factlist <- enrichr(genes=tmpfilt[factor==f,entrezgene_accession],
                        background = filtbg,
                        databases=cur_enrichr_dbs)
    
    factlist.dt <- rbindlist(factlist,idcol="Enrichr_database")
    factlist.dt[,LCnmf_filt_factor:=f]
    
    filt.res[[i]] <- factlist.dt
    names(filt.res)[i] <- f
    
    rm(f,factlist.dt,factlist,tmpfilt)
}
rm(i)

# concat and save
filt.res2 <- rbindlist(filt.res)
# tabulate how many genes were hit in the set
filt.res2[,N_genes:=apply(.SD,MARGIN=1,simplify=FALSE,FUN=function(x){
  max(c(1,length(str_split_fixed(x,";",n=Inf)))) # if no separator, 1 gene matched
}),.SDcols="Genes"]
filt.res2[,N_genes:=unlist(N_genes)] ## fix weirdness with apply within d.t.
# get rid of deprecated columns
filt.res2[,Old.P.value:=NULL]
filt.res2[,Old.Adjusted.P.value:=NULL]
saveRDS(filt.res2,"processed-data/10_subclustering/07thru10_ver2_53factorNMF/09cV2-LCnmfFilt_factorSpecGenes_enrichr_results.RDS")
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.44.0         parallelly_1.46.1           colorout_1.3-3             
 [4] msigdbr_25.1.1              fgsea_1.36.2                clusterProfiler_4.18.4     
 [7] Matrix_1.7-4                data.table_1.18.2.1         RcppML_0.5.6               
[10] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0 SummarizedExperiment_1.40.0
[13] Biobase_2.70.0              GenomicRanges_1.62.1        Seqinfo_1.0.0              
[16] IRanges_2.44.0              S4Vectors_0.48.0            BiocGenerics_0.56.0        
[19] generics_0.1.4              MatrixGenerics_1.22.0       matrixStats_1.5.0          

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3      jsonlite_2.0.0          rstudioapi_0.18.0      
  [4] tidydr_0.0.6            magrittr_2.0.4          ggtangle_0.1.1         
  [7] ggbeeswarm_0.7.3        magick_2.9.0            farver_2.1.2           
 [10] rmarkdown_2.30          fs_1.6.6                vctrs_0.7.1            
 [13] memoise_2.0.1           ggtree_4.0.4            htmltools_0.5.9        
 [16] S4Arrays_1.10.1         curl_7.0.0              BiocNeighbors_2.4.0    
 [19] SparseArray_1.10.8      gridGraphics_0.5-1      htmlwidgets_1.6.4      
 [22] plyr_1.8.9              cachem_1.1.0            igraph_2.2.1           
 [25] lifecycle_1.0.5         pkgconfig_2.0.3         gson_0.1.0             
 [28] rsvd_1.0.5              R6_2.6.1                fastmap_1.2.0          
 [31] digest_0.6.39           aplot_0.2.9             enrichplot_1.30.4      
 [34] ggnewscale_0.5.2        patchwork_1.3.2         AnnotationDbi_1.72.0   
 [37] rprojroot_2.1.1         scater_1.38.0           irlba_2.3.7            
 [40] RSQLite_2.4.5           beachmat_2.26.0         polyclip_1.10-7        
 [43] httr_1.4.7              abind_1.4-8             compiler_4.5.2         
 [46] here_1.0.2              withr_3.0.2             fontquiver_0.2.1       
 [49] bit64_4.6.0-1           S7_0.2.1                viridis_0.6.5          
 [52] DBI_1.2.3               ggforce_0.5.0           R.utils_2.13.0         
 [55] MASS_7.3-65             rappdirs_0.3.4          DelayedArray_0.36.0    
 [58] sessioninfo_1.2.3       rjson_0.2.23            tools_4.5.2            
 [61] vipor_0.4.7             otel_0.2.0              scatterpie_0.2.6       
 [64] beeswarm_0.4.0          ape_5.8-1               R.oo_1.27.1            
 [67] glue_1.8.0              nlme_3.1-168            GOSemSim_2.36.0        
 [70] grid_4.5.2              cluster_2.1.8.1         reshape2_1.4.5         
 [73] gtable_0.3.6            R.methodsS3_1.8.2       tidyr_1.3.2            
 [76] BiocSingular_1.26.0     ScaledMatrix_1.18.0     XVector_0.50.0         
 [79] ggrepel_0.9.6           pillar_1.11.1           stringr_1.6.0          
 [82] babelgene_22.9          yulab.utils_0.2.4       splines_4.5.2          
 [85] tweenr_2.0.3            dplyr_1.2.0             treeio_1.34.0          
 [88] lattice_0.22-7          bit_4.6.0               tidyselect_1.2.1       
 [91] fontLiberation_0.1.0    GO.db_3.22.0            Biostrings_2.78.0      
 [94] scuttle_1.20.0          knitr_1.51              fontBitstreamVera_0.1.1
 [97] gridExtra_2.3           xfun_0.56               stringi_1.8.7          
[100] lazyeval_0.2.2          ggfun_0.2.0             yaml_2.3.12            
[103] evaluate_1.0.5          codetools_0.2-20        gdtools_0.4.4          
[106] tibble_3.3.1            qvalue_2.42.0           ggplotify_0.1.3        
[109] cli_3.6.5               systemfonts_1.3.1       dichromat_2.0-0.1      
[112] Rcpp_1.1.1              png_0.1-8               assertthat_0.2.1       
[115] ggplot2_4.0.2           blob_1.3.0              DOSE_4.4.0             
[118] viridisLite_0.4.3       tidytree_0.4.7          ggiraph_0.9.3          
[121] scales_1.4.0            purrr_1.2.1             crayon_1.5.3           
[124] rlang_1.1.7.9000        cowplot_1.2.0           fastmatch_1.1-8        
[127] KEGGREST_1.50.0        
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-22
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ────────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 AnnotationDbi          1.72.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 ape                    5.8-1      2024-12-16 [2] CRAN (R 4.5.0)
 aplot                  0.2.9      2025-09-12 [2] CRAN (R 4.5.0)
 assertthat             0.2.1      2019-03-21 [2] CRAN (R 4.5.0)
 babelgene              22.9       2022-09-29 [2] CRAN (R 4.5.0)
 beachmat               2.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 beeswarm               0.4.0      2021-06-01 [2] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocNeighbors          2.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocSingular           1.26.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Biostrings             2.78.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 bit                    4.6.0      2025-03-06 [2] CRAN (R 4.5.0)
 bit64                  4.6.0-1    2025-01-16 [2] CRAN (R 4.5.0)
 blob                   1.3.0      2026-01-14 [2] CRAN (R 4.5.2)
 cachem                 1.1.0      2024-05-16 [2] CRAN (R 4.5.0)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 cluster                2.1.8.1    2025-03-12 [2] CRAN (R 4.5.0)
 clusterProfiler      * 4.18.4     2025-12-15 [2] Bioconductor 3.22 (R 4.5.2)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 cowplot                1.2.0      2025-07-07 [2] CRAN (R 4.5.0)
 crayon                 1.5.3      2024-06-20 [2] CRAN (R 4.5.0)
 curl                   7.0.0      2025-08-19 [2] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DBI                    1.2.3      2024-06-02 [2] CRAN (R 4.5.0)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 DOSE                   4.4.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 dplyr                  1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 enrichplot             1.30.4     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 fastmatch              1.1-8      2026-01-17 [2] CRAN (R 4.5.2)
 fgsea                * 1.36.2     2026-01-05 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 fontBitstreamVera      0.1.1      2017-02-01 [2] CRAN (R 4.5.0)
 fontLiberation         0.1.0      2016-10-15 [2] CRAN (R 4.5.0)
 fontquiver             0.2.1      2017-02-01 [2] CRAN (R 4.5.0)
 fs                     1.6.6      2025-04-12 [2] CRAN (R 4.5.0)
 gdtools                0.4.4      2025-10-06 [2] CRAN (R 4.5.0)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 ggbeeswarm             0.7.3      2025-11-29 [2] CRAN (R 4.5.2)
 ggforce                0.5.0      2025-06-18 [2] CRAN (R 4.5.0)
 ggfun                  0.2.0      2025-07-15 [2] CRAN (R 4.5.0)
 ggiraph                0.9.3      2026-01-19 [2] CRAN (R 4.5.2)
 ggnewscale             0.5.2      2025-06-20 [2] CRAN (R 4.5.0)
 ggplot2                4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggplotify              0.1.3      2025-09-20 [2] CRAN (R 4.5.0)
 ggrepel                0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 ggtangle               0.1.1      2026-01-16 [2] CRAN (R 4.5.2)
 ggtree                 4.0.4      2026-01-05 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 GO.db                  3.22.0     2026-01-31 [2] Bioconductor
 GOSemSim               2.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 gridExtra              2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gridGraphics           0.5-1      2020-12-13 [2] CRAN (R 4.5.0)
 gson                   0.1.0      2023-03-07 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 htmlwidgets            1.6.4      2023-12-06 [2] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [2] CRAN (R 4.5.0)
 igraph                 2.2.1      2025-10-27 [2] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 jsonlite               2.0.0      2025-03-27 [2] CRAN (R 4.5.0)
 KEGGREST               1.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lazyeval               0.2.2      2019-03-15 [2] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [2] CRAN (R 4.5.2)
 Matrix               * 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 memoise                2.0.1      2021-11-26 [2] CRAN (R 4.5.0)
 msigdbr              * 25.1.1     2025-07-21 [2] CRAN (R 4.5.2)
 nlme                   3.1-168    2025-03-31 [2] CRAN (R 4.5.2)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 parallelly           * 1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 patchwork              1.3.2      2025-08-25 [2] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 plyr                   1.8.9      2023-10-02 [2] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [2] CRAN (R 4.5.0)
 polyclip               1.10-7     2024-07-23 [2] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 qvalue                 2.42.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 R.methodsS3            1.8.2      2022-06-13 [2] CRAN (R 4.5.0)
 R.oo                   1.27.1     2025-05-02 [2] CRAN (R 4.5.0)
 R.utils                2.13.0     2025-02-24 [2] CRAN (R 4.5.0)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 rappdirs               0.3.4      2026-01-17 [2] CRAN (R 4.5.2)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RcppML               * 0.5.6      2026-02-07 [2] Github (zdebruine/RcppML@5449a5b)
 reshape2               1.4.5      2025-11-12 [2] CRAN (R 4.5.0)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 RSQLite                2.4.5      2025-11-30 [2] CRAN (R 4.5.2)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 rsvd                   1.0.5      2021-04-16 [2] CRAN (R 4.5.0)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 ScaledMatrix           1.18.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scater                 1.38.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 scatterpie             0.2.6      2025-09-12 [2] CRAN (R 4.5.0)
 scuttle                1.20.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 stringi                1.8.7      2025-03-27 [2] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [2] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 systemfonts            1.3.1      2025-10-01 [2] CRAN (R 4.5.0)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidydr                 0.0.6      2025-07-25 [2] CRAN (R 4.5.0)
 tidyr                  1.3.2      2025-12-19 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 tidytree               0.4.7      2026-01-08 [2] CRAN (R 4.5.2)
 treeio                 1.34.0     2025-10-30 [2] Bioconductor 3.22 (R 4.5.1)
 tweenr                 2.0.3      2024-02-26 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 vipor                  0.4.7      2023-12-18 [2] CRAN (R 4.5.0)
 viridis                0.6.5      2024-01-29 [2] CRAN (R 4.5.0)
 viridisLite            0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 withr                  3.0.2      2024-10-28 [2] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)
 yulab.utils            0.2.4      2026-02-02 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

───────────────────────────────────────────────────────────────────────────────────────
