---
title: "10f-speLCnmfNOfilt_projectToWeberDivecha_snSeq"
author: "Bernie Mulvey"
date: "2026-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(RcppML)
library(SpatialExperiment)
library(Matrix) # otherwise rcppml gives error 'no item called "package:Matrix" on the search list'
library(dplyr)
library(ggplot2)
library(WeberDivechaLCdata)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
lcsn <- WeberDivechaLCdata_singleNucleus()

colnames(colData(lcsn))
## several clusterings
# label-30, numerical
# label_inhibitory (14 subclusters)
# label_merged (10 clusters annotated)
# supervised_NE,unsupervised_NE,unsupervised_5HT (all T/F)
# we'll use label_merged here. label_inhibitory could be interesting for NMF on the other neuronal clusters some other time though.
```

## load filtered LC NMF output
```{r}
lcnmf <- readRDS("processed-data/10_subclustering/07thru10_ver2_53factorNMF/07-lcnmf53_tol1e-9.RDS")
```

```{r}
## drop multimatches
dropg <- as.data.table(table(rownames(lcsn)))[N>1] # none
keepg <- rownames(lcnmf$w)[rownames(lcnmf$w) %in% rownames(lcsn)] 

## subset SCE and w accordingly
lcsn <- lcsn[keepg,]
lcnmf$w <- lcnmf$w[keepg,]

stopifnot(nrow(lcnmf$w)==nrow(lcsn))
```

## use rcppml::project (pretty straightforward but erik's script for posterity and the L2 parameter used). also, the argument order has changed for project relative to this code, i think. nonetheless:
https://github.com/LieberInstitute/spatial_hpc/blob/main/code/NMF/project_nmf_patterns.R

Note that Erik's code implies there are rownames on the nmf w matrix, which there aren't. so we also need to re-load the full mouse SPE to determine which genes were there to start with (i.e. are the rows of the w matrix) and which remained after ortholog filtering above.
```{r}
setDTthreads(1,restore_after_fork = FALSE)

set.seed(42)
options(RcppML.threads = 10)
lffprojinto.webdiv <- RcppML::project(data=logcounts(lcsn),w=as.matrix(lcnmf$w),L2=0.00005)
lffprojinto.webdiv <- t(lffprojinto.webdiv)


stopifnot(nrow(lffprojinto.webdiv)==ncol(lcsn))
rownames(lffprojinto.webdiv) <- colnames(lcsn)
colnames(lffprojinto.webdiv) <- paste0("V",c(1:53))
## append to sce
colData(lcsn) <- cbind(colData(lcsn),lffprojinto.webdiv)
colnames(colData(lcsn))[c(17:ncol(colData(lcsn)))] <- paste0("LCspeFiltNMF_",c(1:53))

usefulname <- colnames(colData(lcsn))[17:ncol(colData(lcsn))]

seed3 = as.matrix(colData(lcsn)[,usefulname])
seed3 = seed3>0
d3 = cbind.data.frame(webdivClus=as.data.frame(colData(lcsn))[,"label_merged"],
                      seed3) %>% 
  group_by(webdivClus) %>% add_tally(name="total") %>%
  group_by(webdivClus, total) %>%
  summarise_at(usefulname, sum) %>%
  tidyr::pivot_longer(usefulname, values_to="n", names_to="nmf") %>%
  mutate(prop=n/total)

seed4 = as.matrix(colData(lcsn)[,usefulname])
seed4 = apply(seed4, 2, scale)
d4 = cbind.data.frame(webdivClus=as.data.frame(colData(lcsn))[,"label_merged"],
                      seed4) %>% 
  group_by(webdivClus) %>%
  summarise_at(usefulname, mean) %>% 
  tidyr::pivot_longer(usefulname, values_to="scaled.avg", names_to="nmf")


pltdat <- left_join(d3[,c("webdivClus","nmf","prop")], 
                    d4[,c("webdivClus","nmf","scaled.avg")])

# to d.t.
pltdat <- as.data.table(pltdat)
pltdat[,nmf:=factor(nmf,usefulname)]

#reorder clusters to put LC neurons first
lcfirst <- c("NE","5HT","excitatory","inhibitory","neurons_ambiguous","astrocytes","macrophages_microglia","OPCs","oligodendrocytes","endothelial_mural")
pltdat[,webdivClus:=factor(webdivClus,rev(lcfirst))]
```


```{r}
pltdat <- pltdat[!is.na(scaled.avg)]

range(pltdat$scaled.avg) # -1.81, 6.4
quantile(pltdat$scaled.avg,c(0.02,0.98)) # -1.08, 3.63

png("plots/10_subclustering/07thru10_ver2_53factorNMF/10f-speLCnmfNOfilt_allfactors_projectToWeberDivechaSnseq.png",height=1000,width=8000)
ggplot(pltdat, aes(x=nmf, y=webdivClus, size=prop, color=scaled.avg))+
  geom_count()+
  theme_bw()+
  scale_size(range=c(0,3))+
  scale_color_gradient2(low = "darkblue",mid="white",high="red",midpoint = 0,na.value = NA)+
  labs(x="Spatial LC NMF Factor",y="Weber-Divecha Merged Cluster")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5,size=10),
        axis.text.y=element_text(size=10),
        axis.title.x=element_text(size=12),
        axis.title.y=element_text(size=12))
dev.off()

# 10 astro, 16 macro/micro but NOT mural; 31 mural/macro/micro, others NE or nothing per se

## also plot for the factors with factor-specific genes that we analyzed in GSEA; this will give us a better idea as to whether those factors are LC neuron-driven or something else.
factspec <- fread("processed-data/10_subclustering/07thru10_ver2_53factorNMF/09aV2-LCNMF_factors_and_factorSpecGenes_min25specGenes.txt")
gsfacts <- unique(factspec$factor)
gsfacts <- gsub(gsfacts,pattern="LCnmf(.*)$",replacement="\\1")
gsfacts <- paste0("LCspeFiltNMF_",gsfacts)

png("plots/10_subclustering/07thru10_ver2_53factorNMF/10f-speLCnmfNOfilt_GSEAdFactors_projectToWeberDivechaSnseq.png",height=1000,width=1500)
ggplot(pltdat[nmf %in% gsfacts], aes(x=nmf, y=webdivClus, size=prop, color=scaled.avg))+
  geom_count()+
  theme_bw()+
  scale_size(range=c(0,5))+
  scale_color_gradient2(low = "darkblue",mid="white",high="red",midpoint = 0,,na.value = NA)+
  labs(x="Spatial LC NMF Factor",y="Weber-Divecha Merged Cluster")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5,size=10),
        axis.text.y=element_text(size=10),
        axis.title.x=element_text(size=12),
        axis.title.y=element_text(size=12))
dev.off()

## as stated above.
```


rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] colorout_1.3-3              ggplot2_4.0.2              
 [3] dplyr_1.2.0                 Matrix_1.7-4               
 [5] SpatialExperiment_1.20.0    SingleCellExperiment_1.32.0
 [7] SummarizedExperiment_1.40.0 Biobase_2.70.0             
 [9] GenomicRanges_1.62.1        Seqinfo_1.0.0              
[11] IRanges_2.44.0              S4Vectors_0.48.0           
[13] BiocGenerics_0.56.0         generics_0.1.4             
[15] MatrixGenerics_1.22.0       matrixStats_1.5.0          
[17] Seurat_5.4.0                SeuratObject_5.3.0         
[19] sp_2.2-0                    RcppML_0.5.6               
[21] data.table_1.18.2.1        

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3     rstudioapi_0.18.0      jsonlite_2.0.0        
  [4] magrittr_2.0.4         magick_2.9.0           spatstat.utils_3.2-1  
  [7] farver_2.1.2           rmarkdown_2.30         vctrs_0.7.1           
 [10] ROCR_1.0-12            spatstat.explore_3.7-0 htmltools_0.5.9       
 [13] S4Arrays_1.10.1        SparseArray_1.10.8     sctransform_0.4.3     
 [16] parallelly_1.46.1      KernSmooth_2.23-26     htmlwidgets_1.6.4     
 [19] ica_1.0-3              plyr_1.8.9             plotly_4.12.0         
 [22] zoo_1.8-15             igraph_2.2.1           mime_0.13             
 [25] lifecycle_1.0.5        pkgconfig_2.0.3        R6_2.6.1              
 [28] fastmap_1.2.0          fitdistrplus_1.2-6     future_1.69.0         
 [31] shiny_1.12.1           digest_0.6.39          patchwork_1.3.2       
 [34] rprojroot_2.1.1        tensor_1.5.1           RSpectra_0.16-2       
 [37] irlba_2.3.7            labeling_0.4.3         progressr_0.18.0      
 [40] spatstat.sparse_3.1-0  httr_1.4.7             polyclip_1.10-7       
 [43] abind_1.4-8            compiler_4.5.2         here_1.0.2            
 [46] withr_3.0.2            S7_0.2.1               fastDummies_1.7.5     
 [49] MASS_7.3-65            DelayedArray_0.36.0    rjson_0.2.23          
 [52] tools_4.5.2            lmtest_0.9-40          otel_0.2.0            
 [55] httpuv_1.6.16          future.apply_1.20.1    goftest_1.2-3         
 [58] glue_1.8.0             nlme_3.1-168           promises_1.5.0        
 [61] grid_4.5.2             Rtsne_0.17             cluster_2.1.8.1       
 [64] reshape2_1.4.5         gtable_0.3.6           spatstat.data_3.1-9   
 [67] tidyr_1.3.2            XVector_0.50.0         spatstat.geom_3.7-0   
 [70] RcppAnnoy_0.0.23       ggrepel_0.9.6          RANN_2.6.2            
 [73] pillar_1.11.1          stringr_1.6.0          spam_2.11-3           
 [76] RcppHNSW_0.6.0         later_1.4.5            splines_4.5.2         
 [79] lattice_0.22-7         survival_3.8-6         deldir_2.0-4          
 [82] tidyselect_1.2.1       miniUI_0.1.2           pbapply_1.7-4         
 [85] knitr_1.51             gridExtra_2.3          scattermore_1.2       
 [88] xfun_0.56              stringi_1.8.7          lazyeval_0.2.2        
 [91] yaml_2.3.12            evaluate_1.0.5         codetools_0.2-20      
 [94] tibble_3.3.1           cli_3.6.5              uwot_0.2.4            
 [97] xtable_1.8-4           reticulate_1.44.1      dichromat_2.0-0.1     
[100] Rcpp_1.1.1             globals_0.19.0         spatstat.random_3.4-4 
[103] png_0.1-8              spatstat.univar_3.1-6  parallel_4.5.2        
[106] dotCall64_1.2          listenv_0.10.0         viridisLite_0.4.3     
[109] scales_1.4.0           ggridges_0.5.7         purrr_1.2.1           
[112] rlang_1.1.7.9000       cowplot_1.2.0         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2026-03-24
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ──────────────────────────────────────────────────────────────────────────
 package              * version    date (UTC) lib source
 abind                  1.4-8      2024-09-12 [2] CRAN (R 4.5.0)
 Biobase              * 2.70.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics         * 0.56.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5      2025-04-23 [2] CRAN (R 4.5.0)
 cluster                2.1.8.1    2025-03-12 [2] CRAN (R 4.5.0)
 codetools              0.2-20     2024-03-31 [2] CRAN (R 4.5.2)
 colorout             * 1.3-3      2026-01-31 [2] Github (jalvesaq/colorout@64863bb)
 cowplot                1.2.0      2025-07-07 [2] CRAN (R 4.5.0)
 data.table           * 1.18.2.1   2026-01-27 [2] CRAN (R 4.5.2)
 DelayedArray           0.36.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 deldir                 2.0-4      2024-02-28 [2] CRAN (R 4.5.0)
 dichromat              2.0-0.1    2022-05-02 [2] CRAN (R 4.5.0)
 digest                 0.6.39     2025-11-19 [2] CRAN (R 4.5.2)
 dotCall64              1.2        2024-10-04 [2] CRAN (R 4.5.0)
 dplyr                * 1.2.0      2026-02-03 [2] CRAN (R 4.5.2)
 evaluate               1.0.5      2025-08-27 [2] CRAN (R 4.5.0)
 farver                 2.1.2      2024-05-13 [2] CRAN (R 4.5.0)
 fastDummies            1.7.5      2025-01-20 [2] CRAN (R 4.5.0)
 fastmap                1.2.0      2024-05-15 [2] CRAN (R 4.5.0)
 fitdistrplus           1.2-6      2026-01-24 [2] CRAN (R 4.5.2)
 future                 1.69.0     2026-01-16 [2] CRAN (R 4.5.2)
 future.apply           1.20.1     2025-12-09 [2] CRAN (R 4.5.2)
 generics             * 0.1.4      2025-05-09 [2] CRAN (R 4.5.0)
 GenomicRanges        * 1.62.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 ggplot2              * 4.0.2      2026-02-03 [2] CRAN (R 4.5.2)
 ggrepel                0.9.6      2024-09-07 [2] CRAN (R 4.5.0)
 ggridges               0.5.7      2025-08-27 [2] CRAN (R 4.5.0)
 globals                0.19.0     2026-02-02 [2] CRAN (R 4.5.2)
 glue                   1.8.0      2024-09-30 [2] CRAN (R 4.5.0)
 goftest                1.2-3      2021-10-07 [2] CRAN (R 4.5.0)
 gridExtra              2.3        2017-09-09 [2] CRAN (R 4.5.0)
 gtable                 0.3.6      2024-10-25 [2] CRAN (R 4.5.0)
 here                   1.0.2      2025-09-15 [2] CRAN (R 4.5.0)
 htmltools              0.5.9      2025-12-04 [2] CRAN (R 4.5.2)
 htmlwidgets            1.6.4      2023-12-06 [2] CRAN (R 4.5.0)
 httpuv                 1.6.16     2025-04-16 [2] CRAN (R 4.5.0)
 httr                   1.4.7      2023-08-15 [2] CRAN (R 4.5.0)
 ica                    1.0-3      2022-07-08 [2] CRAN (R 4.5.0)
 igraph                 2.2.1      2025-10-27 [2] CRAN (R 4.5.0)
 IRanges              * 2.44.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 irlba                  2.3.7      2026-01-30 [2] CRAN (R 4.5.2)
 jsonlite               2.0.0      2025-03-27 [2] CRAN (R 4.5.0)
 KernSmooth             2.23-26    2025-01-01 [2] CRAN (R 4.5.2)
 knitr                  1.51       2025-12-20 [2] CRAN (R 4.5.2)
 labeling               0.4.3      2023-08-29 [2] CRAN (R 4.5.0)
 later                  1.4.5      2026-01-08 [2] CRAN (R 4.5.2)
 lattice                0.22-7     2025-04-02 [2] CRAN (R 4.5.2)
 lazyeval               0.2.2      2019-03-15 [2] CRAN (R 4.5.0)
 lifecycle              1.0.5      2026-01-08 [2] CRAN (R 4.5.2)
 listenv                0.10.0     2025-11-02 [2] CRAN (R 4.5.0)
 lmtest                 0.9-40     2022-03-21 [2] CRAN (R 4.5.0)
 magick                 2.9.0      2025-09-08 [2] CRAN (R 4.5.0)
 magrittr               2.0.4      2025-09-12 [2] CRAN (R 4.5.0)
 MASS                   7.3-65     2025-02-28 [2] CRAN (R 4.5.2)
 Matrix               * 1.7-4      2025-08-28 [2] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics       * 1.22.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 matrixStats          * 1.5.0      2025-01-07 [2] CRAN (R 4.5.0)
 mime                   0.13       2025-03-17 [2] CRAN (R 4.5.0)
 miniUI                 0.1.2      2025-04-17 [2] CRAN (R 4.5.0)
 nlme                   3.1-168    2025-03-31 [2] CRAN (R 4.5.2)
 otel                   0.2.0      2025-08-29 [2] CRAN (R 4.5.0)
 parallelly             1.46.1     2026-01-08 [2] CRAN (R 4.5.2)
 patchwork              1.3.2      2025-08-25 [2] CRAN (R 4.5.0)
 pbapply                1.7-4      2025-07-20 [2] CRAN (R 4.5.0)
 pillar                 1.11.1     2025-09-17 [2] CRAN (R 4.5.0)
 pkgconfig              2.0.3      2019-09-22 [2] CRAN (R 4.5.0)
 plotly                 4.12.0     2026-01-24 [2] CRAN (R 4.5.2)
 plyr                   1.8.9      2023-10-02 [2] CRAN (R 4.5.0)
 png                    0.1-8      2022-11-29 [2] CRAN (R 4.5.0)
 polyclip               1.10-7     2024-07-23 [2] CRAN (R 4.5.0)
 progressr              0.18.0     2025-11-06 [2] CRAN (R 4.5.0)
 promises               1.5.0      2025-11-01 [2] CRAN (R 4.5.0)
 purrr                  1.2.1      2026-01-09 [2] CRAN (R 4.5.2)
 R6                     2.6.1      2025-02-15 [2] CRAN (R 4.5.0)
 RANN                   2.6.2      2024-08-25 [2] CRAN (R 4.5.0)
 RColorBrewer           1.1-3      2022-04-03 [2] CRAN (R 4.5.0)
 Rcpp                   1.1.1      2026-01-10 [2] CRAN (R 4.5.2)
 RcppAnnoy              0.0.23     2026-01-12 [2] CRAN (R 4.5.2)
 RcppHNSW               0.6.0      2024-02-04 [2] CRAN (R 4.5.0)
 RcppML               * 0.5.6      2026-02-07 [2] Github (zdebruine/RcppML@5449a5b)
 reshape2               1.4.5      2025-11-12 [2] CRAN (R 4.5.0)
 reticulate             1.44.1     2025-11-14 [2] CRAN (R 4.5.2)
 rjson                  0.2.23     2024-09-16 [2] CRAN (R 4.5.0)
 rlang                  1.1.7.9000 2026-02-04 [1] Github (r-lib/rlang@bae1fe1)
 rmarkdown              2.30       2025-09-28 [2] CRAN (R 4.5.0)
 ROCR                   1.0-12     2026-01-23 [2] CRAN (R 4.5.2)
 rprojroot              2.1.1      2025-08-26 [2] CRAN (R 4.5.0)
 RSpectra               0.16-2     2024-07-18 [2] CRAN (R 4.5.0)
 rstudioapi             0.18.0     2026-01-16 [2] CRAN (R 4.5.2)
 Rtsne                  0.17       2023-12-07 [2] CRAN (R 4.5.0)
 S4Arrays               1.10.1     2025-12-08 [2] Bioconductor 3.22 (R 4.5.2)
 S4Vectors            * 0.48.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 S7                     0.2.1      2025-11-14 [2] CRAN (R 4.5.2)
 scales                 1.4.0      2025-04-24 [2] CRAN (R 4.5.0)
 scattermore            1.2        2023-06-12 [2] CRAN (R 4.5.0)
 sctransform            0.4.3      2026-01-10 [2] CRAN (R 4.5.2)
 Seqinfo              * 1.0.0      2025-10-29 [2] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3      2025-02-05 [2] CRAN (R 4.5.0)
 Seurat               * 5.4.0      2025-12-14 [2] CRAN (R 4.5.2)
 SeuratObject         * 5.3.0      2025-12-12 [2] CRAN (R 4.5.2)
 shiny                  1.12.1     2025-12-09 [2] CRAN (R 4.5.2)
 SingleCellExperiment * 1.32.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 sp                   * 2.2-0      2025-02-01 [2] CRAN (R 4.5.0)
 spam                   2.11-3     2026-01-08 [2] CRAN (R 4.5.2)
 SparseArray            1.10.8     2025-12-18 [2] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment    * 1.20.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 spatstat.data          3.1-9      2025-10-18 [2] CRAN (R 4.5.0)
 spatstat.explore       3.7-0      2026-01-22 [2] CRAN (R 4.5.2)
 spatstat.geom          3.7-0      2026-01-20 [2] CRAN (R 4.5.2)
 spatstat.random        3.4-4      2026-01-21 [2] CRAN (R 4.5.2)
 spatstat.sparse        3.1-0      2024-06-21 [2] CRAN (R 4.5.0)
 spatstat.univar        3.1-6      2026-01-17 [2] CRAN (R 4.5.2)
 spatstat.utils         3.2-1      2026-01-10 [2] CRAN (R 4.5.2)
 stringi                1.8.7      2025-03-27 [2] CRAN (R 4.5.0)
 stringr                1.6.0      2025-11-04 [2] CRAN (R 4.5.0)
 SummarizedExperiment * 1.40.0     2025-10-29 [2] https://bioc-release.r-universe.dev (R 4.5.2)
 survival               3.8-6      2026-01-16 [2] CRAN (R 4.5.2)
 tensor                 1.5.1      2025-06-17 [2] CRAN (R 4.5.0)
 tibble                 3.3.1      2026-01-11 [2] CRAN (R 4.5.2)
 tidyr                  1.3.2      2025-12-19 [2] CRAN (R 4.5.2)
 tidyselect             1.2.1      2024-03-11 [2] CRAN (R 4.5.0)
 uwot                   0.2.4      2025-11-10 [2] CRAN (R 4.5.0)
 vctrs                  0.7.1      2026-01-23 [2] CRAN (R 4.5.2)
 viridisLite            0.4.3      2026-02-04 [2] CRAN (R 4.5.2)
 withr                  3.0.2      2024-10-28 [2] CRAN (R 4.5.0)
 xfun                   0.56       2026-01-18 [2] CRAN (R 4.5.2)
 xtable                 1.8-4      2019-04-21 [2] CRAN (R 4.5.0)
 XVector                0.50.0     2025-10-29 [2] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12     2025-12-10 [2] CRAN (R 4.5.2)
 zoo                    1.8-15     2025-12-15 [2] CRAN (R 4.5.2)

 [1] /Users/bmulvey/Library/R/arm64/4.5/library
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 * ── Packages attached to the search path.

─────────────────────────────────────────────────────────────────────────────────────
