---
title: "12-Markers_of_WebDivecha_inhibitory4"
author: "Bernie Mulvey"
date: "2025-08-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(SingleCellExperiment)
# devtools::install_github("dmcable/spacexr", build_vignettes = FALSE, force = TRUE)
library(spatialLIBD)
library(WeberDivechaLCdata)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

## load pre-annotated webdiv snseq and regen the collapsed-original labels
```{r} 
lcsn <- WeberDivechaLCdata_singleNucleus()

## determine what hierarchical clustering levels there are
tmpcd <- as.data.table(colData(lcsn),keep.rownames=T)
checklabs <- unique(tmpcd[,.(label,label_merged,label_inhibitory)])
## so the inhibitories are re-clusterings (multiple label_inhibitory per label)
View(checklabs[label_merged!="inhibitory"])
## the other 23 labels (of 30) have single label merged assignments.
## so we will make a cluster hierarchy as follows: cluster_merged > label > label or inhibitory subcluster

# to ensure there's enough LCNE cells for RCTD to work with, change label_merged and then label itself to NE for any cell classified as supervised NE OR unsup. NE
tmpcd[,label_merged:=as.character(label_merged)]
tmpcd[unsupervised_NE==TRUE,label_merged:="NE"]
tmpcd[supervised_NE==TRUE,label_merged:="NE"]

## then, make label a character column of merged_label (parent) and label (granular clust #)
tmpcd[,label:=as.character(label)]
tmpcd[,label:=paste0(label_merged,"_",label)]
## for NE neurons, disregard the granular cluster of origin since we're integrating several definition types now
tmpcd[label_merged=="NE",label:="NE"]
## finally, add info about inhibitory subcluster; ONLY for inhibitory merged clusters
tmpcd[,sublabel:=label]
tmpcd[label_merged=="inhibitory"&!is.na(label_inhibitory),sublabel:=paste0(sublabel,
                                                "_sub",label_inhibitory)]
## for sublabels, we will need to filter out sublabels with fewer than RCTD's suggested minimum of 25 cells (see loop below).

stopifnot(length(unique(tmpcd$label))>length(unique(tmpcd$label_merged)))
stopifnot(length(unique(tmpcd$sublabel))>length(unique(tmpcd$label)))

colData(lcsn) <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(lcsn),]

rm(tmpcd,checklabs)
```

## subset to inhibitory sublabels and run spatialLIBD registrationwrapper for markers
```{r}
lcsn <- lcsn[,lcsn$label %in% grep(unique(lcsn$label),pattern="inhibitory",value=T)]
dim(lcsn)
# 2725 cells

lcsnmk <- registration_wrapper(lcsn,var_registration="label",var_sample_id="Sample",gene_ensembl = "gene_id",gene_name = "gene_name",min_ncells = 10)

lcsnmk2 <- as.data.table(lcsnmk$enrichment)

## filter to collapsed 'merged_original' label of interest, inhibitory 4 
keepn <- c("ensembl","gene",grep(names(lcsnmk2),pattern="inhibitory_4",value=T))
lcsnmk2 <- lcsnmk2[,..keepn]
setnames(lcsnmk2,c(3:6),c("tstat","pval","fdr","logFC"))

fwrite(lcsnmk2,"processed-data/10_subclustering/12-WebDiv_inhibitory_4_mks_vs_other_inhibitorys.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

repinf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices
[5] utils     datasets  methods   base     

other attached packages:
 [1] BiocParallel_1.40.0        
 [2] parallelly_1.43.0          
 [3] WeberDivechaLCdata_1.8.0   
 [4] ExperimentHub_2.14.0       
 [5] AnnotationHub_3.14.0       
 [6] BiocFileCache_2.14.0       
 [7] dbplyr_2.5.0               
 [8] spatialLIBD_1.18.0         
 [9] SpatialExperiment_1.16.0   
[10] SingleCellExperiment_1.28.1
[11] SummarizedExperiment_1.36.0
[12] Biobase_2.66.0             
[13] GenomicRanges_1.58.0       
[14] GenomeInfoDb_1.42.3        
[15] IRanges_2.40.1             
[16] S4Vectors_0.44.0           
[17] BiocGenerics_0.52.0        
[18] MatrixGenerics_1.18.1      
[19] matrixStats_1.5.0          
[20] data.table_1.17.4          
[21] rlang_1.1.5                

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3      
  [2] jsonlite_2.0.0          
  [3] magrittr_2.0.3          
  [4] ggbeeswarm_0.7.2        
  [5] magick_2.8.6            
  [6] farver_2.1.2            
  [7] BiocIO_1.16.0           
  [8] fields_16.3.1           
  [9] zlibbioc_1.52.0         
 [10] vctrs_0.6.5             
 [11] memoise_2.0.1           
 [12] config_0.3.2            
 [13] Rsamtools_2.22.0        
 [14] paletteer_1.6.0         
 [15] RCurl_1.98-1.17         
 [16] benchmarkme_1.0.8       
 [17] htmltools_0.5.8.1       
 [18] S4Arrays_1.6.0          
 [19] curl_6.2.2              
 [20] BiocNeighbors_2.0.1     
 [21] SparseArray_1.6.2       
 [22] sass_0.4.10             
 [23] bslib_0.9.0             
 [24] htmlwidgets_1.6.4       
 [25] plotly_4.10.4           
 [26] cachem_1.1.0            
 [27] GenomicAlignments_1.42.0
 [28] mime_0.13               
 [29] lifecycle_1.0.4         
 [30] iterators_1.0.14        
 [31] pkgconfig_2.0.3         
 [32] rsvd_1.0.5              
 [33] Matrix_1.7-3            
 [34] R6_2.6.1                
 [35] fastmap_1.2.0           
 [36] GenomeInfoDbData_1.2.13 
 [37] shiny_1.10.0            
 [38] digest_0.6.37           
 [39] rematch2_2.1.2          
 [40] AnnotationDbi_1.68.0    
 [41] rprojroot_2.0.4         
 [42] scater_1.34.1           
 [43] irlba_2.3.5.1           
 [44] RSQLite_2.3.9           
 [45] beachmat_2.22.0         
 [46] filelock_1.0.3          
 [47] httr_1.4.7              
 [48] abind_1.4-8             
 [49] compiler_4.4.3          
 [50] here_1.0.1              
 [51] withr_3.0.2             
 [52] bit64_4.6.0-1           
 [53] doParallel_1.0.17       
 [54] attempt_0.3.1           
 [55] viridis_0.6.5           
 [56] DBI_1.2.3               
 [57] maps_3.4.3              
 [58] sessioninfo_1.2.3       
 [59] rappdirs_0.3.3          
 [60] DelayedArray_0.32.0     
 [61] rjson_0.2.23            
 [62] tools_4.4.3             
 [63] vipor_0.4.7             
 [64] beeswarm_0.4.0          
 [65] httpuv_1.6.16           
 [66] glue_1.8.0              
 [67] restfulr_0.0.15         
 [68] promises_1.3.2          
 [69] grid_4.4.3              
 [70] generics_0.1.4          
 [71] gtable_0.3.6            
 [72] tidyr_1.3.1             
 [73] ScaledMatrix_1.14.0     
 [74] BiocSingular_1.22.0     
 [75] XVector_0.46.0          
 [76] ggrepel_0.9.6           
 [77] BiocVersion_3.20.0      
 [78] foreach_1.5.2           
 [79] pillar_1.10.2           
 [80] spam_2.11-1             
 [81] limma_3.62.2            
 [82] later_1.4.2             
 [83] benchmarkmeData_1.0.4   
 [84] dplyr_1.1.4             
 [85] lattice_0.22-6          
 [86] rtracklayer_1.66.0      
 [87] bit_4.6.0               
 [88] tidyselect_1.2.1        
 [89] locfit_1.5-9.12         
 [90] Biostrings_2.74.1       
 [91] scuttle_1.16.0          
 [92] knitr_1.50              
 [93] gridExtra_2.3           
 [94] edgeR_4.4.2             
 [95] xfun_0.52               
 [96] statmod_1.5.0           
 [97] DT_0.33                 
 [98] UCSC.utils_1.2.0        
 [99] lazyeval_0.2.2          
[100] yaml_2.3.10             
[101] shinyWidgets_0.9.0      
[102] evaluate_1.0.3          
[103] codetools_0.2-20        
[104] tibble_3.2.1            
[105] BiocManager_1.30.25     
[106] cli_3.6.4               
[107] xtable_1.8-4            
[108] jquerylib_0.1.4         
[109] golem_0.5.1             
[110] dichromat_2.0-0.1       
[111] Rcpp_1.0.14             
[112] png_0.1-8               
[113] XML_3.99-0.18           
[114] parallel_4.4.3          
[115] ggplot2_3.5.2           
[116] blob_1.2.4              
[117] dotCall64_1.2           
[118] bitops_1.0-9            
[119] viridisLite_0.4.2       
[120] scales_1.4.0            
[121] purrr_1.0.4             
[122] crayon_1.5.3            
[123] cowplot_1.1.3           
[124] KEGGREST_1.46.0         
> sessioninfo::session_info()
─ Session info ─────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-08-05
 pandoc   NA
 quarto   1.7.32 @ /Applications/Positron.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────
 !  package              * version   date (UTC) lib source
    abind                  1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
    AnnotationDbi          1.68.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    AnnotationHub        * 3.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
    beachmat               2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
    benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
    benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
    Biobase              * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocFileCache        * 2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocGenerics         * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocIO                 1.16.0    2024-12-12 [1] Github (Bioconductor/BiocIO@8e80624)
    BiocManager            1.30.25   2024-08-28 [1] CRAN (R 4.4.1)
    BiocNeighbors          2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
    BiocParallel         * 1.40.0    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    BiocSingular           1.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    BiocVersion            3.20.0    2024-05-04 [1] Bioconductor 3.20 (R 4.4.0)
    Biostrings             2.74.1    2024-12-16 [1] Bioconductor 3.20 (R 4.4.2)
    bit                    4.6.0     2025-03-06 [1] CRAN (R 4.4.1)
    bit64                  4.6.0-1   2025-01-16 [1] CRAN (R 4.4.1)
    bitops                 1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
    blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
    bslib                  0.9.0     2025-01-30 [1] CRAN (R 4.4.1)
    cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 VP cli                    3.6.4     2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.3)
    config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
    cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
    crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
    curl                   6.2.2     2025-03-24 [1] CRAN (R 4.4.1)
    data.table           * 1.17.4    2025-05-26 [1] CRAN (R 4.4.3)
    DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
    dbplyr               * 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
    DelayedArray           0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    dichromat              2.0-0.1   2022-05-02 [1] CRAN (R 4.4.0)
    digest                 0.6.37    2024-08-19 [1] CRAN (R 4.4.1)
    doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
    dotCall64              1.2       2024-10-04 [1] CRAN (R 4.4.1)
    dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
    DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
    edgeR                  4.4.2     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    evaluate               1.0.3     2025-01-10 [1] CRAN (R 4.4.1)
    ExperimentHub        * 2.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
    fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
    fields                 16.3.1    2025-03-08 [1] CRAN (R 4.4.1)
    filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
    foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
    generics               0.1.4     2025-05-09 [1] CRAN (R 4.4.1)
    GenomeInfoDb         * 1.42.3    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
    GenomeInfoDbData       1.2.13    2024-12-12 [1] Bioconductor
    GenomicAlignments      1.42.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    GenomicRanges        * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2                3.5.2     2025-04-09 [1] CRAN (R 4.4.1)
    ggrepel                0.9.6     2024-09-07 [1] CRAN (R 4.4.1)
    glue                   1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
    golem                  0.5.1     2024-08-27 [1] CRAN (R 4.4.1)
    gridExtra              2.3       2017-09-09 [1] CRAN (R 4.4.0)
    gtable                 0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
    here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
    htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
    htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
    httpuv                 1.6.16    2025-04-16 [1] CRAN (R 4.4.1)
    httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
    IRanges              * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
    irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
    iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
    jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
    jsonlite               2.0.0     2025-03-27 [1] CRAN (R 4.4.1)
    KEGGREST               1.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    knitr                  1.50      2025-03-16 [1] CRAN (R 4.4.2)
    later                  1.4.2     2025-04-08 [1] CRAN (R 4.4.1)
    lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.3)
    lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
    lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
    limma                  3.62.2    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    locfit                 1.5-9.12  2025-03-05 [1] CRAN (R 4.4.1)
    magick                 2.8.6     2025-03-23 [1] CRAN (R 4.4.1)
    magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
    maps                   3.4.3     2025-05-26 [1] CRAN (R 4.4.1)
    Matrix                 1.7-3     2025-03-11 [1] CRAN (R 4.4.1)
    MatrixGenerics       * 1.18.1    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
    matrixStats          * 1.5.0     2025-01-07 [1] CRAN (R 4.4.1)
    memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
    mime                   0.13      2025-03-17 [1] CRAN (R 4.4.1)
    paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
    parallelly           * 1.43.0    2025-03-24 [1] CRAN (R 4.4.1)
    pillar                 1.10.2    2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
    plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
    png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
    promises               1.3.2     2024-11-28 [1] CRAN (R 4.4.1)
    purrr                  1.0.4     2025-02-05 [1] CRAN (R 4.4.1)
    R6                     2.6.1     2025-02-15 [1] CRAN (R 4.4.1)
    rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
    RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp                   1.0.14    2025-01-12 [1] CRAN (R 4.4.1)
    RCurl                  1.98-1.17 2025-03-22 [1] CRAN (R 4.4.1)
    rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
    restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
    rjson                  0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 VP rlang                * 1.1.5     2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
    Rsamtools              2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    RSQLite                2.3.9     2024-12-03 [1] CRAN (R 4.4.1)
    rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
    rtracklayer            1.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Arrays               1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    S4Vectors            * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sass                   0.4.10    2025-04-11 [1] CRAN (R 4.4.1)
    ScaledMatrix           1.14.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    scales                 1.4.0     2025-04-24 [1] CRAN (R 4.4.1)
    scater                 1.34.1    2025-03-03 [1] Bioconductor 3.20 (R 4.4.2)
    scuttle                1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    sessioninfo            1.2.3     2025-02-05 [1] CRAN (R 4.4.1)
    shiny                  1.10.0    2024-12-14 [1] CRAN (R 4.4.1)
    shinyWidgets           0.9.0     2025-02-21 [1] CRAN (R 4.4.1)
    SingleCellExperiment * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
    spam                   2.11-1    2025-01-20 [1] CRAN (R 4.4.1)
    SparseArray            1.6.2     2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
    SpatialExperiment    * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    spatialLIBD          * 1.18.0    2024-11-07 [1] Bioconductor 3.20 (R 4.4.3)
    statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
    SummarizedExperiment * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
    tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
    tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
    UCSC.utils             1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
    vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
    viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
    WeberDivechaLCdata   * 1.8.0     2024-10-31 [1] Bioconductor 3.20 (R 4.4.1)
    withr                  3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
    xfun                   0.52      2025-04-02 [1] CRAN (R 4.4.1)
    XML                    3.99-0.18 2025-01-01 [1] CRAN (R 4.4.1)
    xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
    XVector                0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
    yaml                   2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
    zlibbioc               1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────
