---
title: "Rostrocaudal_Info_Condense_for_SupTab1"
author: "Bernie Mulvey"
date: "06-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)

## >= and <= comparators that account for floating point error
greaterOrEqual <- function(x,y) {
    precision <- sqrt(.Machine$double.eps)
    (x >= y) | (abs(x-y) <= precision)
}

lessOrEqual <- function(x,y){
    precision <- sqrt(.Machine$double.eps)
    (x <= y) | (abs(x-y) <= precision)
}

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


read in TSV with QC cutting info
```{r}
roca <- fread("plots/manuscript/SuppTabs/TabDelims/LC_sectioning_log_5_31_26.txt")

## get samples where TH/NM was not present at the cutting face of block
fromoutside <- roca[freeze_thaw_round==1&(NM_at_start=="N"|TH_at_start=="N")]
## 4 right hemisphere blocks ^
fromoutside[,.(brnum,cutting_from_to)]
#     brnum cutting_from_to
#    <char>          <char>
# 1: Br5276             R-C
# 2: Br6222             R-C
# 3: Br5517             R-C
# 4: Br5712             C-R

roca[,Rend_Cend_orInterior:="Interior"]
roca[brnum %in% c("Br5276", "Br6222", "Br5517")&hemisphere=="Right",Rend_Cend_orInterior:="Rostral_End"]
roca[brnum %in% c("Br5712")&hemisphere=="Right",Rend_Cend_orInterior:="Caudal_End"]

## drop QC sections now, not necess for sup dat
roca <- roca[sample_type!="QC"]
## make brnums unique by hemisphere for adjusting these. Br5854 left was not designated 5854-L here
roca[brnum=="Br5854"&hemisphere=="Left",blockid:="Br5854-L"]

## change so that sequencing of cutting 1, 2, etc only covers omic samples
roca[,brnum:=gsub(blockid,pattern="-",replacement="")]
roca[,sampling:=frank(freeze_thaw_round, ties.method = "dense"), by = brnum]
roca[sample_type=="Xenium",sample_id:=brnum] # used as sample identifier for practicality

roca2 <- roca[,.(brnum, hemisphere, cutting_from_to, sampling, Rend_Cend_orInterior, sample_type, sample_id)]
setnames(roca2,"sample_id","sample_identifier(Vis-sample_id__Xen-brnum)")
fwrite(roca2,"plots/manuscript/SuppTabs/ST1sheet-RostrocaudalSectioningInfo.txt",sep='\t',quote=F,row.names=F)
```


seshinf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.5.2 (2025-10-31)
Platform: aarch64-apple-darwin20
Running under: macOS Tahoe 26.3.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.1

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] BiocParallel_1.44.0 parallelly_1.47.0   colorout_1.3-3      data.table_1.18.2.1

loaded via a namespace (and not attached):
 [1] Matrix_1.7-4                rjson_0.2.23                compiler_4.5.2             
 [4] Rcpp_1.1.1                  SpatialExperiment_1.20.0    SummarizedExperiment_1.40.0
 [7] Biobase_2.70.0              GenomicRanges_1.62.1        magick_2.9.0               
[10] parallel_4.5.2              IRanges_2.44.0              Seqinfo_1.0.0              
[13] yaml_2.3.12                 fastmap_1.2.0               lattice_0.22-7             
[16] here_1.0.2                  XVector_0.50.0              S4Arrays_1.10.1            
[19] generics_0.1.4              knitr_1.51                  BiocGenerics_0.56.0        
[22] DelayedArray_0.36.1         MatrixGenerics_1.22.0       rprojroot_2.1.1            
[25] rlang_1.1.7                 xfun_0.56                   otel_0.2.0                 
[28] SparseArray_1.10.8          cli_3.6.5                   magrittr_2.0.4             
[31] digest_0.6.39               grid_4.5.2                  rstudioapi_0.18.0          
[34] lifecycle_1.0.5             S4Vectors_0.48.1            SingleCellExperiment_1.32.0
[37] evaluate_1.0.5              codetools_0.2-20            abind_1.4-8                
[40] rsconnect_1.7.0             stats4_4.5.2                rmarkdown_2.30             
[43] matrixStats_1.5.0           tools_4.5.2                 htmltools_0.5.9            
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.5.2 (2025-10-31)
 os       macOS Tahoe 26.3.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2026-06-01
 rstudio  2026.01.1+403 Apple Blossom (desktop)
 pandoc   3.6.3 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.8.25 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────────────
 package              * version  date (UTC) lib source
 abind                  1.4-8    2024-09-12 [1] CRAN (R 4.5.0)
 Biobase                2.70.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocGenerics           0.56.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 BiocParallel         * 1.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 cli                    3.6.5    2025-04-23 [1] CRAN (R 4.5.0)
 codetools              0.2-20   2024-03-31 [1] CRAN (R 4.5.2)
 colorout             * 1.3-3    2026-01-31 [1] Github (jalvesaq/colorout@64863bb)
 data.table           * 1.18.2.1 2026-01-27 [1] CRAN (R 4.5.2)
 DelayedArray           0.36.1   2026-03-31 [1] Bioconductor 3.22 (R 4.5.2)
 digest                 0.6.39   2025-11-19 [1] CRAN (R 4.5.2)
 evaluate               1.0.5    2025-08-27 [1] CRAN (R 4.5.0)
 fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.5.0)
 generics               0.1.4    2025-05-09 [1] CRAN (R 4.5.0)
 GenomicRanges          1.62.1   2025-12-08 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 here                   1.0.2    2025-09-15 [1] CRAN (R 4.5.0)
 htmltools              0.5.9    2025-12-04 [1] CRAN (R 4.5.2)
 IRanges                2.44.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 knitr                  1.51     2025-12-20 [1] CRAN (R 4.5.2)
 lattice                0.22-7   2025-04-02 [1] CRAN (R 4.5.2)
 lifecycle              1.0.5    2026-01-08 [1] CRAN (R 4.5.2)
 magick                 2.9.0    2025-09-08 [1] CRAN (R 4.5.0)
 magrittr               2.0.4    2025-09-12 [1] CRAN (R 4.5.0)
 Matrix                 1.7-4    2025-08-28 [1] local (/Users/bmulvey/Desktop/Matrix)
 MatrixGenerics         1.22.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 matrixStats            1.5.0    2025-01-07 [1] CRAN (R 4.5.0)
 otel                   0.2.0    2025-08-29 [1] CRAN (R 4.5.0)
 parallelly           * 1.47.0   2026-04-17 [1] CRAN (R 4.5.2)
 Rcpp                   1.1.1    2026-01-10 [1] CRAN (R 4.5.2)
 rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.5.0)
 rlang                  1.1.7    2026-01-09 [1] CRAN (R 4.5.2)
 rmarkdown              2.30     2025-09-28 [1] CRAN (R 4.5.0)
 rprojroot              2.1.1    2025-08-26 [1] CRAN (R 4.5.0)
 rsconnect              1.7.0    2025-12-06 [1] CRAN (R 4.5.2)
 rstudioapi             0.18.0   2026-01-16 [1] CRAN (R 4.5.2)
 S4Arrays               1.10.1   2025-12-08 [1] Bioconductor 3.22 (R 4.5.2)
 S4Vectors              0.48.1   2026-04-04 [1] https://bioc-release.r-universe.dev (R 4.5.3)
 Seqinfo                1.0.0    2025-10-29 [1] Bioconductor 3.22 (R 4.5.2)
 sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.5.0)
 SingleCellExperiment   1.32.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SparseArray            1.10.8   2025-12-18 [1] Bioconductor 3.22 (R 4.5.2)
 SpatialExperiment      1.20.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 SummarizedExperiment   1.40.0   2025-10-29 [1] https://bioc-release.r-universe.dev (R 4.5.2)
 xfun                   0.56     2026-01-18 [1] CRAN (R 4.5.2)
 XVector                0.50.0   2025-10-29 [1] Bioconductor 3.22 (R 4.5.1)
 yaml                   2.3.12   2025-12-10 [1] CRAN (R 4.5.2)

 [1] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.5/library
 * ── Packages attached to the search path.

────────────────────────────────────────────────────────────────────────────────────────────
