---
title: "02-Nonspatial clustering QC"
output: html_document
date: "2024-09-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(bluster)
library(HDF5Array)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

# Silhouette Widths
## load glmpca matrix, clustering results from the same matrix (the two inputs to bluster for sil. widths)
```{r}
lcglmpca <- HDF5Array("processed-data/06_FeatSelection_Dimred_Harmony/03b_lffLC_noEdges_noLocOutliers_remainder1pctileUMIorGene-removed_allgeneGLMPCA.h5",name="GLM-PCA")

SNNlouvlist <- readRDS("processed-data/06_FeatSelection_Dimred_Harmony/03c_GLMPCAallgene_SNN10-15-20_leidenAndLouvain_res05-1-2.RDS")
## all clustering assignments are in columns with names reflecting the clustering method, SNN k value, and resolution parameter. each list entry is the results for one SNN k value so we can merge the columns into one big table for simplicity here.

SNNlouvs <- merge.data.table(SNNlouvlist[[1]],SNNlouvlist[[2]],by="rn")
SNNlouvs <- merge.data.table(SNNlouvs,SNNlouvlist[[3]],by="rn")

## are any of these like, clearly wacked out results?
SNNlouvs[,lapply(.SD,FUN=function(x){length(unique(x))}),.SDcols=c(2:ncol(SNNlouvs))]

## uhhh, yes, leiden clusterings all had 400-5100 clusters o.0
## lets just use the louvain clusterings
keepclus <- c("rn",grep(names(SNNlouvs),pattern="_louv_",value=T))
SNNlouvs <- SNNlouvs[,..keepclus]

rm(keepclus,SNNlouvlist)
```

### calculate silhouette withs from the cluster assignments and the corresponding dimensionality reduction matrix (here GLMPCA of all genes)
```{r}
## get the clustering column names (everything except "rn" in keepclus)
keepclus <- keepclus[2:length(keepclus)]

## reorder the rows of the pca matrix to match the data.table input. (way harder to do this vice versa)
lcglmpca <- lcglmpca[SNNlouvs$rn,]
stopifnot(sum(rownames(lcglmpca)==SNNlouvs$rn)==nrow(lcglmpca))

## run silhouette width approximation for each clustering result and the underlying glmpca mat
silres <- mclapply(X=keepclus,mc.cores=5,FUN=function(clusts){
    getcols <- c("rn",clusts)
  clusts <- copy(SNNlouvs[,..getcols])
  clusts <- as.data.frame(clusts,row.names = clusts$rn)
  clusts <- clusts[rownames(lcglmpca),]
  clusts$rn <- NULL
  names(clusts)[1] <- "curclus"
  silwidths <- approxSilhouette(x = lcglmpca,clusters = clusts$curclus)
  return(silwidths)
})
names(silres) <- keepclus

### save results
saveRDS(silwidths,"processed-data/07_Clustering/02-SNNk10-15-20_louvain_allgeneGLMPCA_silwidths.RDS")

rm(silwidths)
```

### cluster purity in bluster (same inputs)
```{r}
## run cluster purity in bluster for each clustering result and the underlying glmpca mat
clustpurity <- mclapply(X=keepclus,mc.cores=5,FUN=function(clusts){
    getcols <- c("rn",clusts)
    clusts <- copy(SNNlouvs[,..getcols])
    clusts <- as.data.frame(clusts,row.names = clusts$rn)
    clusts <- clusts[rownames(lcglmpca),]
    clusts$rn <- NULL
    names(clusts)[1] <- "curclus"
    clustpurity <- neighborPurity(x = lcglmpca,clusters = clusts$curclus)
    return(clustpurity)
})

names(clustpurity) <- keepclus

### save results
saveRDS(clustpurity,"processed-data/07_Clustering/02-SNNk10-15-20_louvain_allgeneGLMPCA_clustpurity.RDS")
```

cluster RMSDs
```{r}
crmsd <- mclapply(X=keepclus,mc.cores=5,FUN=function(clusts){
    getcols <- c("rn",clusts)
    clusts <- copy(SNNlouvs[,..getcols])
    clusts <- as.data.frame(clusts,row.names = clusts$rn)
    clusts <- clusts[rownames(lcglmpca),]
    clusts$rn <- NULL
    names(clusts)[1] <- "curclus"
    cRMSD <- clusterRMSD(x = lcglmpca,clusters = clusts$curclus)
    return(cRMSD)
})
names(crmsd) <- keepclus

saveRDS(crmsd,"processed-data/07_Clustering/02-SNNk10-15-20_louvain_allgeneGLMPCA_clustRMSDs.RDS")
```

## cluster modularity requires the spot-spot distance matrix from the SNN graph upstream of clustering, which we did not save and so cannot execute.

interpretation notes from OSCA re: purity and silhouette widths: "The main difference between the cluster purity and silhouette width is that the former ignores the intra-cluster variance. This provides a simpler interpretation of cluster separation; a low silhouette width may still occur in well-separated clusters if the internal heterogeneity is high, while no such complication exists for the cluster purity. Comparing these two metrics can give us an indication of which clusters are heterogeneous, well-separated, or both."

# reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] HDF5Array_1.32.0            rhdf5_2.48.0               
 [3] DelayedArray_0.30.1         SparseArray_1.4.8          
 [5] S4Arrays_1.4.1              abind_1.4-5                
 [7] Matrix_1.7-0                BiocParallel_1.38.0        
 [9] parallelly_1.38.0           colorout_1.3-0.2           
[11] bluster_1.14.0              SpatialExperiment_1.14.0   
[13] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
[15] Biobase_2.64.0              GenomicRanges_1.56.1       
[17] GenomeInfoDb_1.40.1         IRanges_2.38.1             
[19] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[21] MatrixGenerics_1.16.0       matrixStats_1.3.0          
[23] gridExtra_2.3               ggplot2_3.5.1              
[25] data.table_1.15.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] gtable_0.3.5            rjson_0.2.21            xfun_0.45              
 [4] lattice_0.22-6          rhdf5filters_1.16.0     vctrs_0.6.5            
 [7] tools_4.4.1             generics_0.1.3          tibble_3.2.1           
[10] fansi_1.0.6             cluster_2.1.6           pkgconfig_2.0.3        
[13] BiocNeighbors_1.22.0    lifecycle_1.0.4         GenomeInfoDbData_1.2.12
[16] compiler_4.4.1          munsell_0.5.1           codetools_0.2-20       
[19] htmltools_0.5.8.1       yaml_2.3.9              pillar_1.9.0           
[22] crayon_1.5.3            sessioninfo_1.2.2       magick_2.8.3           
[25] tidyselect_1.2.1        digest_0.6.36           dplyr_1.1.4            
[28] rprojroot_2.0.4         fastmap_1.2.0           grid_4.4.1             
[31] here_1.0.1              colorspace_2.1-0        cli_3.6.3              
[34] magrittr_2.0.3          utf8_1.2.4              withr_3.0.0            
[37] scales_1.3.0            UCSC.utils_1.0.0        rmarkdown_2.27         
[40] XVector_0.44.0          httr_1.4.7              igraph_2.0.3           
[43] evaluate_0.24.0         knitr_1.48              Rcpp_1.0.13            
[46] glue_1.7.0              rstudioapi_0.16.0       jsonlite_1.8.8         
[49] Rhdf5lib_1.26.0         R6_2.5.1                zlibbioc_1.50.0        
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-16
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package              * version date (UTC) lib source
   abind                * 1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics         * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors          1.22.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bluster              * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   cluster                2.1.6   2023-12-01 [1] CRAN (R 4.4.1)
   codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   DelayedArray         * 0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate               0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi                  1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12  2024-05-01 [1] Bioconductor
   GenomicRanges        * 1.56.1  2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue                   1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array            * 1.32.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   igraph                 2.0.3   2024-03-13 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   jsonlite               1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr                  1.48    2024-07-07 [1] CRAN (R 4.4.1)
   lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magick                 2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix               * 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   rhdf5                * 2.48.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters           1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib               1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                  0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Arrays             * 1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SparseArray          * 1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SummarizedExperiment * 1.34.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr                  3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector                0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
